### 2019-05-14 manuscript/main.tex
added manuscript draft

\documentclass{article}

\usepackage{amsmath,epsfig,bm, amssymb,graphicx,algorithm,algorithmic,color,
  bbm}
\usepackage{hyperref}

\begin{document}

\title{Network-based GWAS on familial breast cancer: method comparison and new biomarkers}
\author{H\'ector Climente-Gonz\'alez, Christine Lonjou, Chlo\'e-Agathe Azencott}

\maketitle

\section{Methods}

\subsection{GENESIS}

GENESIS is a GWAS study of familial breast cancer on the French population. The
index cases are patients with a breast cancer affected sister and no BRCA1/2
mutations \cite{sinilnikova_genesis:_2016}. The controls are unaffected
colleagues and/or friends of the cases. In total, the GENESIS dataset consists
of 2\,566 samples, of which 1\,293 are cases and 1\,273 are controls. The genotyping
platform was iCOGS, a custom Illumina array designed
to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}. It contains ~211\,155
SNPs previously associated with cancer susceptibility, survival, or other
cancer-related traits; and in selected genes and pathways.

We used the age of the patients as covariate. Specifically, the
age at diagnosis for cases, and the age at the interview for controls.

\subsection{Preprocessing and quality control}

A PCA revealed no differential population structure between cases and controls.

28 samples with ?? were removed. A subset of 20 duplicated SNPs in FGFR2 were also removed.
Following the suggested protocol in ??, we LD pruned the dataset with $R^2$ 0.75.

We imputed missing genotypes using IMPUTE2 \cite{howie_flexible_2009}. The reference panel were the
genotypes from the 1000 Genomes Project Phase 3 \cite{the_1000_genomes_project_consortium_global_2015}. We selected 503 samples, the
number of samples of European origin in the reference panel. After imputation, we discarded
SNPs with a minor allele frequency lower than 0.1\%, not in Hardy - Weinberg
equilibrium (P-value \textless 0.001), and/or those with missing values on more than 10\% of the samples. In addition, we removed the samples with more than 10\% missing genotypes. The final dataset included 1\,279 controls and 1\,298 cases, genotyped over 128\,595 SNPs. 
\subsection{SNP- and gene-level association scores}

\subsection{High-weight module discovery algorithms}

\subsubsection{Node scores}

We used PLINK v1.90 \cite{chang_second-generation_2015} to fit a per-SNP
logistic regression, using the age as covariate. The assumed model of
inheritance was additive. Then, we computed a P-value of the null hypothesis of
the coefficient being different than 0. We used VEGAS2v2 to compute the
gene-level association score \cite{mishra_vegas2:_2015}. In order to map a SNP
to a gene we relied on their physical position, being within 50 kb of the
borders of the gene.

\subsubsection{Network}

We used the physical interactions from BioGRID version 3.5.171
\cite{oughtred_biogrid_2019} to get protein-protein interactions when required.
Out of the six methods tested, five use a protein-protein interaction network.

\subsubsection{Methods used}

Finding the highest-scoring, densely interconnected subnetwork on a graph is an
open problem in the field. Hence, several solutions have been proposed to the
problem. In this paper, we apply 5 methods designed to explore the
protein-protein interaction network, and one method, SConES, which explores
SNP-networks.

\noindent {\bf Heinz} \cite{dittrich_identifying_2008}

\noindent {\bf Hierarchichal HotNet} \cite{reyna_hierarchical_2018}

\noindent {\bf jActiveModules}

\noindent {\bf LEAN} \cite{gwinner_network-based_2016}

\noindent {\bf SConES} \cite{azencott_efficient_2013}

\noindent {\bf SigMod}

\subsection{Validation of selected biomarkers}

\subsubsection{Classification accuracy of selected biomarkers}

To evaluate the quality of the SNPs selected by the different methods, we used
the performance of a classifier trained exclusively on those SNPs. The
underlying assumption is that SNPs that are good biomarkers will be good
predictiors of the phenotype of unseen samples. To that end, we followed a
5-fold cross-validation schema, where over 5 iterations 80\% of the data is used
to select SNPs and train the classifier, and 20\% is used to compute and
evaluate the predictions.

Specifically we fitted an L1-penalized logistic regression over the age and the selected SNPs, using the outcome (case/control) as
target variable. When the method retrieved a list of genes (all of them except
SConES), all the SNPs mapped to any of those genes were used.

\subsubsection{Comparison to external dataset}

\subsection{Code availability}

The implementation of the different network-based methods to GWAS data is
available at \url{https://github.com/hclimente/gwas-tools}. The code necessary
to reproduce all the analyses in this article is available at \url{https://github.com/hclimente/genewa}.

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2019-10-20 manuscript/manuscript.tex
finished first draft of the manuscript

% Created 2019-10-19 Sat 12:19
% Intended LaTeX compiler: pdflatex
\documentclass[twocolumn, 10pt]{article}
\usepackage{amsmath,amssymb}
\usepackage[format=hang,labelfont=bf,textfont=small,singlelinecheck=false,justification=raggedright,margin={20pt,20pt},figurename=Fig.]{caption}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{threeparttable}
\newcommand{\mean}[1]{$\overline{\mbox{#1}}$}
\newcommand{\median}[1]{$\hat{\mbox{#1}}$}

\makeatletter
\def\affiliation#1{\gdef\@affiliation{#1}}
\def\abstract#1{\gdef\@abstract{#1}}
\def\graphabst#1{\gdef\@graphabst{#1}}
\def\keywords#1{\gdef\@keywords{#1}}
\def\corresp#1{\gdef\@corresp{#1}}
\def\bioauthor#1#2{\centerline{\textbf{#1}}\par #2}
\newcommand{\MakeTitle}{
  \newpage
  \begin{center}
    \Large \@title\par
    \large \@author
  \end{center}
  \@affiliation\par
  \@abstract\par
  \vskip 1em
}
\makeatother

\title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
\author{Héctor Climente-González$^{1,2,3}$,Christine Lonjou$^{1,2,3}$,Fabienne Lesueur$^{1,2,3}$,GENESIS
  investigators,Nadine Andrieu$^{1,2,3}$,Chloé-Agathe Azencott$^{3,1,2}$}
\affiliation{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France; \\
  $^{2}$INSERM, U900, F-75005 Paris, France; \\
  $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France; \\
}
\hypersetup{
 pdfauthor={Héctor Climente-González, Chloé-Agathe Azencott},
 pdftitle={Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 26.2 (Org mode 9.2.3)}, 
 pdflang={English}}
\begin{document}

\onecolumn
\MakeTitle
\begin{abstract}

Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building. It does so by jointly considering the statistical association between a gene and a phenotype, obtained experimentally, and the biological context of each gene, represented as a network. In this work we study the utility of six network methods to discover new biomarkers in GWAS
data by searching subnetworks highly associated to a phenotype. We interrogate a familial breast cancer GWAS focused on BRCA1/2 negative French women. By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility. We
perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods. Interestingly, a combination solution subnetworks provided a concise subnetwork of 93 genes, enriched in known familial breast cancer susceptibility genes (BABAM1, BLM, CASP8, FGFR2, and TOX3, Fisher's exact test
P-value = $7.8 \times 10^{-5}$ ) and more central than average. Additionally, it includes subnetworks of mechanisms related to cancer, like protein folding (HSPA1A, HSPA1B, and HSPA1L) or mitocondrial ribosomes (MRPS30, MRPS31, MRPS18B). We also observed a general disregulation in the neighborhood of COPS5, a gene related to multiple hallmarks of cancer.

\end{abstract}

\section{Introduction}
\label{sec:orgbee581e}

In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \cite{bush_chapter_2012}. To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering the disease of interest, and the controls from the general population. Then, a per-SNP statistical association test is conducted (e.g. \(\chi\)\textsuperscript{2}). Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort. Once the risk SNPs have been discovered, they can be used for diagnosis, guide the treatment, and deepen our understanding of the disease.

GWAS have successfully identified thousands of variants underlying many common diseases \cite{buniello_nhgri-ebi_2019}. However, the experimental setting also presents intrinsic challenges. Some of them stem from the high-dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped. This limits the statistical power of the experiment, as only variants with large and moderate effects can be detected. And it is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}. Additionally, to avoid false positives, a conservative multiple test correction is applied, typically Bonferroni. However, Bonferroni is known to be overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}. Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood. The fact that SNPs in non-coding regions often top the results of GWAS illustrates well this challenge. On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}. The omnigenic model \cite{boyle_expanded_2017} explains this by the dense functional inter-relatedness between genes, influencing each others' behavior, which allows alterations in most genes to impact the ``core'' genes involved in a disease. A comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.

In this regard many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}. As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships between them are the edges that link them. These functional relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their gene products establish a protein-protein interaction. Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent biomolecules \cite{barabasi_network_2011}. In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011} \cite{pinero_uncovering_2016}. This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes. Additionally, as genes that contribute to a disease tend to participate in similar biological functions, guilt-by-association strategies have proved effective at identifying disease genes \cite{huang_systematic_2018}.

Network-based, biomarker discovery methods exploit the guilt-by-association strategy to identify disease genes on GWAS data \cite{azencott_network-guided_2016}. In essence, each SNP has a measure of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge. Then, the problem becomes finding a functionally-related set of genes that is highly associated with the disease. Different solutions have been proposed to this problem, often stemming from divergent different mathematical frameworks and considerations of what the optimal solution looks like. Some methods strongly constrain the problem to certain kinds of subnetworks. Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on star subnetworks, i.e. instances were both a gene and its direct interactors are associated with the disease. Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on interconnected genes with high association with the disease. However, they differ in their tolerance to the inclusion of lowly associated nodes, and the possible number of disconnected subnetworks in the solution. Lastly, other methods also consider the topology of the network, favoring solutions that are densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.

In this work, we analyze the effectiveness of these six methods to discover new biomarkers on GWAS data. We focus on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a homogeneous dataset of familial breast cancer in the French population. After following a classical GWAS approach, we use these network-based methods to recover additional familial breast cancer biomarkers. Some of them are known, while others are specific to this dataset. Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus network of predisposition to familial breast cancer.

\section{Methods}
\label{sec:org9e7ca64}
\subsection{GENESIS}
\label{sec:org920eb00}

GENESIS is a GWAS study of familial breast cancer on the French population \cite{sinilnikova_genesis:_2016}. The index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had sister with breast cancer, and negative for BRCA1 and BRCA2 mutations. The controls are unaffected colleagues and/or friends of the cases, born around the year of birth of the corresponding case (\textpm{} 3 years). We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases. The genotyping platform was iCOGS, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}. It contains 210\,918 SNPs previously associated with cancer susceptibility, survival, or other cancer-related traits; and in selected genes and pathways.

\subsection{Preprocessing and quality control}
\label{sec:orgd26c48b}

We discarded SNPs with a minor allele frequency lower than 0.1\%, not in Hardy - Weinberg equilibrium (P-value \textless 0.001), and those with missing values on more than 10\% of the samples. A subset of 20 duplicated SNPs in FGFR2 were also removed. In addition, we removed the samples with more than 10\% missing genotypes, and an additional 28 samples with TODO. The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.

We looked for population structure that could create confounding associations. A PCA revealed no differential population structure between cases and controls (Supplementary Figure \ref{sfig:pcs}). Independently, we did not find evidence of genomic inflation (\(\lambda\) = 1.05) either, further confirming the absence of confounding population structure.

\subsection{High-score subnetwork search algorithms}
\label{sec:org6ced6c9}
\subsubsection{SNP and gene association}
\label{methods:node_score}
To measure association between a genotype and the phenotype, we performed a per-SNP 1df \(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}. Then, we used VEGAS2v2 to compute the gene-level association score \cite{mishra_vegas2:_2015} from the SNP P-values. In order to map SNPs to genes we used their overlap on the genome: all SNPs located within the boundaries of a gene, \textpm{} 50 kb, were mapped to that gene. To compute the gene association we used the 10\% of SNPs with lowest P-values. We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP. Out of those, we focused exclusively on the 32\,767 that could be mapped to an HGNC symbol. Out of the SNPs 197\,083 remaining after quality control, 164\,037 mapped to at least one of these genes.

We use such mapping to compare the outputs of methods who produce SNP- to those that produce gene-lists, and vice versa. In the former case, we consider any gene that can be mapped to any of the selected SNPs as selected as well. In the latter, we consider all the SNPs that can be mapped to that gene as selected by the method.

\subsubsection{Mathematical notation}
\label{methods:notation}
In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$. $V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$. Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges. When referring to a subnetwork S, $V_{S}$ is the set of nodes in S and $E_{S}$ is the set of edges in S. A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.

On top of a weight, nodes have other properties provided by the topology of the graph. In this article we focus on two: degree centrality, and betweenness centrality. The degree centrality, or degree, is the number of edges that a node has. The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.

In addition, we use several matrices that describe different properties of a graph. The described matrices are square, and have as many rows and columns as nodes are in the network. The element $(i,j)$ represents a  selected relationship between $v_{i}$ and $v_{j}$. The \emph{adjacency matrix} $W_{G}$ contains a 1 when the corresponding nodes are connected through an edge, and 0 otherwise; the diagonal is zero. The \emph{degree matrix} $D_{G}$ is a diagonal matrix which contains the degree of the different nodes. Lastly, the \emph{Laplacian matrix} $L_{G}$ is defined as $L_{G} = D_{G} - W_{G}$.

\subsubsection{Methods used}
\label{methods:methods}
\begin{table}[htbp]
\caption{\label{tab:method_comparison}
Summary of the differences between the studied algorithms.}
\centering
\begin{tabular}{l|lllllll}
Method & Field & Nodes & Exhaustive & Solution & Components & Input & Scoring\\
\hline
dmGWAS & GWAS & Gene & No & - & 1 & Summary & -log\textsubscript{10}(P)\\
heinz & Omics & Gene & Yes & - & 1 & Summary & BUM\\
HotNet2 & Omics & Gene & Yes & Modular & \(\ge\) 1 & Summary & Local FDR\\
LEAN & Omics & Gene & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P)\\
SConES & GWAS & SNP & Yes & Modular & \(\ge\) 1 & Genotypes & \(\chi\)\textsuperscript{2}\\
SigMod & GWAS & Gene & Yes & Modular & 1 & Summary & -log\textsubscript{10}(P)\\
\end{tabular}
\end{table}

Beyond the assumption that genes that contribute to the same function will be nearby in the protein-protein interaction network (PPIN), they might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.). That is not the only choice to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, what is the best network to use, etc. In consequence, multiple solutions have been proposed. In this article, we examine six of them: five that explore the protein-protein interaction network, and one which explores SNP networks. We selected methods that were open source, had an implementation available, and an accessible documentation. Their main differences are summarized in Table \ref{tab:method_comparison}.

\begin{description}
\item[{dmGWAS}] dmGWAS searches the subgraph with the highest local density in low P-values \cite{jia_dmgwas:_2011}. To that end it searches candidate subnetwork solutions using a greedy procedure involving the following steps:

\begin{enumerate}
\item Select a seed node.
\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph as

\begin{equation*} 
Z_m = \frac{\sum z_i}{\sqrt{k}}
\end{equation*}

where \emph{k} is the number of genes in the subgraph, z\textsubscript{i} = \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}), and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
\item Identify neighboring nodes i.e. nodes at distance \(\le\) \emph{d}. We set d = 2.
\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than Z\textsubscript{m} \texttimes{} (1 + r). In our experiments, we set r = 0.1.
\item Repeat 2-4 until no increment Z\textsubscript{m} \texttimes{} (1 + r) is possible.
\end{enumerate}

Lastly, the module's Z-score is normalized as

\begin{equation*}
Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)}
\end{equation*}

where Z\textsubscript{m}(\(\pi\)) represent a vector containing 100000 random subsets of the same number of genes.

We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}. We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
\end{description}

\begin{description}
\item[{heinz}] The goal of heinz is to identify the highest-scored connected subgraph on the network \cite{dittrich_identifying_2008}. The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive value when there is. This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired FDR. Thus formulated, the problem is NP-complete. To solve it efficiently it is re-casted as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S):

\begin{equation*}
p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
\end{equation*}

were p(v) = w(v) - w' is the \emph{profit} of adding a node, c(e) = w' is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$. All three are positive quantities. heinz implements the algorithm from \cite{ljubic_algorithmic_2006}, which in practice is often fast and optimal, neither is guaranteed. We used BioNet's implementation of heinz, available on Bioconductor \cite{beisser_bionet:_2010,heinz}.

\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}. To that end, it considers both the local topology of the network and the scores of the nodes. The former is captured by an insulated heat diffusion process: at the beginning, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}). This process continues until equilibrium is reached, and results in a similarity matrix F. F is used to compute the similarity matrix E that accounts also for similarities in node scores as

\begin{equation*} 
E = F \operatorname{diag}(w(V)), 
\end{equation*}

where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal. We scored the nodes as in \cite{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be. In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}. To obtain densely connected subnetworks, HotNet2 prunes E, only preserving edges such that w(E) \textgreater{} \(\delta\). Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores. HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\). Both parameters are set automatically by the algorithm, and are robust \cite{leiserson_pan-cancer_2015}. HotNet2 is implemented in Python \cite{hotnet2}.

\item[{LEAN}] LEAN searches disregulated ``star'' gene subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}. By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node). For a particular subnetwork of size \emph{m}, the P-values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}. Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P-values lower or equal to p\textsubscript{k} under the null hypothesis. The minimum of these \emph{k} P-values is the score of the subnetwork. This score is transformed into a P-value through an empirical distribution obtained via a subsampling scheme, where sets of \emph{m} genes are selected randomly, and their score computed. Lastly, P-values are corrected for multiple testing through a Benjamini-Hochberg correction. We used the implementation of LEAN from the LEANR R package \cite{leanr}.
\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem

\begin{equation}
\label{eq:scones}
\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
\end{equation}

where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model. Given two hyperparameters, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure. We used the version on SConES implemented in the R package martini \cite{martini}. As in \cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds. Note that the solution to the above problem can consist of several connected subnetworks which are disconnected from each other. In this case, the selected hyperparameters were \(\eta\) = 3.51, \(\lambda\) = 210.29 for SConES GS; \(\eta\) = 3.51, \(\lambda\) = 97.61 for SConES GM; and \(\eta\) = 3.51, \(\lambda\) = 45.31 for SConES GI.
\end{description}

\begin{description}
\item[{SigMod}] SigMod aims at identifying the most densely connected gene subnetwork that is most strongly associated to the phenotype \cite{liu_sigmod:_2017}. It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the Laplacian matrix rather than the adjacency matrix (Section \ref{methods:notation}), to quantify solutions containing many edges.

\begin{equation*}
\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
\end{equation*}

As SConES, this optimization problem can also be solved by a graph min-cut approach.

SigMod presents three important differences with SConES. First it is designed for gene-gene networks. Second, by replacing the Laplacian by the adjacency matrix, it favors subnetworks containing many edges. SConES, instead, penalizes connections between the selected selected and unselected nodes. Third, it returns a single connected subnetwork, which it achieves by exploring a grid of hyperparameters and processing their respective solutions. Specifically, for the range of \(\lambda\) = \(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}. Such a large change implies that the network is strongly interconnected. This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other. A score is assigned to each candidate solution by summing their node scores and normalizing by size. The candidate solution with the highest standardized score is the chosen solution. SigMod is implemented in an R package \cite{sigmod}.
\end{description}

\subsubsection{Gene-gene network}
\label{methods:gene_network}
Out of the six methods tested, five use a gene-gene interaction network (Section \ref{methods:methods}). Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), for practical reasons we focused on a PPIN, as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it. We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}. Unless specified otherwise, we used only interactions coming from high-throughput experiments to avoid biasing the topology of the network by well-studied genes with more known interactions on average. Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of HGNC symbols. The scoring function for the nodes changed from method to method (Section \ref{methods:methods}).

Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies. In that case, out of the 179\,332 interactions in HINT, we mapped 173\,797 to a pair of HGNC symbols.

\subsubsection{SNP networks}
\label{methods:snp_network}
SConES \cite{azencott_efficient_2013} is the only of the studied methods designed to handle SNP networks. As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs. The authors suggested three ways of building these networks: connecting the those SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists (``GI network''). We focused on the GI network, as it is the network that fits better the scope of this article. However, at different stages of the manuscript we also used GS and GM. For the GM network, we used the mapping described in Section \ref{methods:node_score}. For the GI network, we used the PPI as described in Section \ref{methods:gene_network}. For all three networks the node score used is the association of the individual SNPs with the phenotype; specifically, we used the 1 d.f. \(\chi\)\textsuperscript{2}.

\subsubsection{Consensus network}
\label{methods:consensus}
The different high-weight subnetwork discovery algorithms make different assumptions on the properties of the solutions, and employ different strategies to find them. Hence, combining the outcome of the different approaches might provide a more complete outlook on the specific alterations on the GENESIS dataset. We built such consensus network by retaining the nodes that were selected by at least two of the methods. We combined the results of 6 methods: dmGWAS, heinz, HotNet2, LEAN, SConES GI, and SigMod.

\subsection{Evaluation of methods}
\label{sec:orgc968928}
\subsubsection{Classification accuracy of selected biomarkers}
\label{methods:comparison}
A desirable solution is one that is sparse, while offering a good predictive power on unseen samples. We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression trained exclusively on those SNPs to predict the outcome (case/control). The L1 penalty helps to account for LD to reduce the number of SNPs included in the model (size of the active set), while improving the generalization of the classifier. The value of the regularization parameter, which controls both the magnitude and the sparsity of the coefficients, was set by cross-validation. To that end, we used the different network-methods on a random subset of 80\% of the samples. On this same subset we trained our classifier exclusively on the SNPs selected by a particular method. When the method retrieved a list of genes (all of them except SConES), all the SNPs mapped to any of those genes were used. Then we evaluated performance of the classifier on the remaining 20\% of the dataset. We repeated this procedure 5 times to estimate the average and the deviation of the different performance measures. The different performance measures we used where: size of the solution, size of the active set, specificity, and sensitivity. The size of the active set provides an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.

Additionally, for each of the methods, we evaluated their stability and their runtime. The stability of an algorithm is its sensitivity to small changes of the input, and is measured using the Pearson's correlation between different runs as suggested in \cite{nogueira_measuring_2016}. To obtain a baseline, we also performed the procedure using all the SNPs. Lastly, another desirable property is that the method retrieves a good candidate causal subnetwork. In consequence, we compared the outcome of each of the methods to the consensus subnetwork of all the solutions (Section \ref{methods:consensus}).

\subsubsection{Biological relevance of the genes}
\label{methods:bcac}
An alternative way to validate the results is comparing our results to an external dataset. For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}. Across this article we refer to these genes as \emph{familial breast cancer genes}.

Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015}. BCAC is one of the largest efforts in GWAS, genotyping over 120\,000 women of European ancestry. As opposed to GENESIS, samples were not selected based on family history, and hence is enriched in sporadic breast cancers. Another difference is that BCAC is a relatively heterogeneous study on a pan-European sample, while GENESIS is a homogeneous dataset focused on the French population. Despite these differences, there should be shared genetic architecture. On top of that, that overlap should become larger when the results are aggregated at the gene level. For that purpose, we computed the gene association as in Section \ref{methods:node_score}. iCOGS array was used for genotyping in BCAC \cite{sakoda_turning_2013}, the same array as for GENESIS \cite{sinilnikova_genesis:_2016}. Although imputed data is available, we used exclusively the SNPs available on GENESIS after quality control to make the results comparable.

\subsection{Code availability}
\label{sec:org1d907c0}

This work required developing computational pipelines for several GWAS analyses, such physically mapping SNPs to genes, computing gene scores, and performing six different network analyses. For each of those processes, a streamlined, project-agnostic pipeline with a clear interface was created. They are compiled in the following GitHub repository: \url{https://github.com/hclimente/gwas-tools}. The code that applies these pipelines to the GENESIS project, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}. Although the GENESIS data is not publicly available, the published code should work on any other GWAS dataset. All the produced gene subnetworks were deposited on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.

\section{Results}
\label{sec:orgc718f04}
\subsection{FGFR2 is strongly associated with familial breast cancer}
\label{results:conventional}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/figure_1.png}
\caption{\label{fig:snp_gene_manhattan}
Association in GENESIS. The red line represents the Bonferroni threshold. \textbf{(A)} SNP association, measured from the outcome of a 1df \(\chi\)\textsuperscript{2} allelic test. Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated. The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}. \textbf{(B)} Gene association, measured by P-value of VEGAS2v2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P-values. The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.}
\end{figure}

We conducted association analyses in the GENESIS dataset at both the SNP and the gene levels (Section \ref{methods:node_score}). Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Figure \ref{fig:snp_gene_manhattan}A). The former overlaps with gene FGFR2; the latter with CASC16, and it is located near the protein-coding gene TOX3. Variants in both FGFR2 and TOX3 were related to breast cancer susceptibility in other cohorts negative for BRCA1/2 \cite{rinella_genetic_2013}. Only the peak in chromosome 10 replicated in the gene-level analysis, with FGFR2 just above the threshold of significance (Figure \ref{fig:snp_gene_manhattan}B).

These results show the overlap in the genetic architecture of the disease between the French population and other cohorts, especially at the gene level. In addition, there are other SNPs whose P-values, although higher than the conventional threshold of significance, show a strong association with familial breast cancer. The most prominent of such regions are 3p24 and 8q24, both of which  which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}. This motivates exploring network methods, which trade statistical significance for biological relevance.

\subsection{Network methods successfully identify genes linked to breast cancer}
\label{results:separate_networks}
\begin{table}[htbp]
  \begin{threeparttable}
\caption{\label{tab:gene_solutions}
Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
\centering
\begin{tabular}{lrrrrr}
Network & \# genes & \# edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & \(\rho\)\textsubscript{consensus}\\
\hline
HINT HT & 13\,619 & 142\,541 & 16\,706 & 0.46 & 0.066\\
\hline
Consensus & 55 & 117 & 74\,062 & 0.0051 & 1\\
dmGWAS & 194 & 450 & 49\,115 & 0.19 & 0.41\\
heinz & 4 & 3 & 113\,633 & 0.0012 & 0.21\\
HotNet2 & 440 & 374 & 7\,739 & 0.048 & 0.31\\
LEAN & 0 & 0 & - & - & -\\
SConES GI & 0 (1) & 0 & - & - & -\\
SigMod & 142 & 249 & 92\,603 & 0.0083 & 0.73\\
\end{tabular}
\begin{tablenotes}
\emph{\# genes}: number of genes selected out of those that are part of the PPIN; in parentheses, the total number of genes. \mean{Betweenness}: mean betweenness of the selected genes in the PPIN. \median{P}\textsubscript{gene}: median P-value of the selected genes. \(\rho\)\textsubscript{consensus}: Pearson's correlation between the subnetwork and the consensus network.
\end{tablenotes}
\end{threeparttable}
\end{table}

We applied six network methods to the GENESIS dataset (Section \ref{methods:methods}), obtaining six solutions (Supplementary Figure \ref{sfig:overview_solutions}, Supplementary Files 1 and 2): one for each of the five gene-based methods (Section \ref{methods:gene_network}), and one for SConES GI (Section \ref{methods:snp_network}). The solutions are very heterogeneous (Table \ref{tab:gene_solutions} and Supplementary table \ref{tab:snp_solutions}): none of the subnetworks examined by LEAN is significant (adjusted P-value < 0.05), while HotNet2 produced the largest solution subnetwork with 440 genes. SConES GI failed to recover genes in the PPIN, but it recovered one genomic region mapped to RNA gene RNU6-420P. All solution subnetworks except LEAN's are, on average, more strongly associated to familial breast cancer than the whole PPIN (median P-values $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{sfig:manhattan_solutions}). This exemplifies the trade-off between statistical significance and biological relevance. However, there are nuances between solutions: heinz strongly favored highly associated genes, while dmGWAS is less conservative (median gene P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select lowly associated genes.

The solution subnetworks present other desirable properties. First, four of the methods succeeded at recovering genes involved in the disease (Supplementary Figure \ref{sfig:disgenet_comparison}), as their subnetworks were enriched in familial breast cancer genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03). We also compared the outcome of the network methods to the association tests conducted on the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015} (Supplementary Figure \ref{sfig:bcac_comparison}). Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC. This confirms the capability of network methods to find the same signal as in more powered studies by leveraging on prior knowledge. Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in three solutions (dmGWAS, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}). This agrees with the notion that disease genes are more central than other, non-essential genes \cite{pinero_uncovering_2016}. We observe that this conclusion holds in this disease, as familial breast cancer genes have higher betweenness centrality than others (one-tailed Wilcoxon rank-sum test P-value = 2.64 \texttimes{} 10\textsuperscript{-5}, Supplementary Figure \ref{sfig:consensus_stats}C). Interestingly, SConES' selected SNPs are also more central than the average SNP (Supplementary table \ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than unrelated SNPs. However, very central nodes are also more likely to be connecting a random pair of nodes, making then more likely to be selected by the examined methods. Hence, further work is needed draw conclusions.

As the solutions were quite different from each other it is hard to draw joint conclusions. The 4-gene solution selected by heinz includes the familial breast cancer gene TOX3, in region 16q12. By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or else. In fact, SConES GI, which adds to GM the interactions between genes, retrieves 4 subnetworks in intergenic regions, and 1 overlapping an RNA gene (RNU6-420P). SigMod, despite being related to SConES, produces a vastly different, large solution. On top of recovering three familial breast cancer genes, a keratin-based region of its subnetwork affects the cytoskeleton (\emph{structural constituent of cytoskeleton}, GO enrichment's adjusted P-value = 9.10 \texttimes{} 10\textsuperscript{-4}), a potentially novel susceptibility mechanisms to cancer. Interestingly, dmGWAS solution is also related to cytoskeleton (\emph{tubulin binding}, GO enrichment's adjusted P-value = 0.031). But, additionally, it includes a submodule of proteins related to \emph{unfolded protein binding} (GO enrichment's adjusted P-value = 0.045), which was related to cancer susceptibility \cite{calderwood_heat_2016}. Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes. The second largest subnetwork (13 nodes), contains two familial breast cancer genes: CASP8 and BLM.

\subsection{heinz retrieves a small, highly informative set of biomarkers in a fast and stable fashion}
\label{sec:orgb834a56}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/figure_4.pdf}
\caption{\label{fig:benchmark}
Comparison of network-based GWAS methods on GENESIS. Each method was run 5 times of a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}). \textbf{(A)} Number of SNPs selected by each method and number of SNPs on the active set used by the Lasso classifier. Points are the average over the 5 runs; lines represent the standard error of the mean. A grey diagonal line with slope 1 is added for comparison. For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs. \textbf{(B)} Sensitivity and specificity on test set of the L1-penalized logistic regression trained on the features selected by each of the methods. In addition, the performance of the classifier trained on all SNPs is displayed. Points are the average over the 5 runs; lines represent the standard error of the mean. \textbf{(C)} Pairwise Pearson's correlations of the solutions used by different methods. A Pearson's correlation of 1 means the two solutions are the same. A Pearson's correlation of 0 means that there is no SNP in common between the two solutions. \textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP). For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) which took VEGAS2v2 on average to compute the gene scores from SNP summary statistics.}
\end{figure}

As the methods produced such different results, we compared their solutions in a 5-fold subsampling setting (Section \ref{methods:comparison}). Specifically, we measured the following properties (Figure \ref{fig:benchmark}): (i) size of the solution subnetwork; (ii) sensitivity and specificity of an L1-penalized logistic regression on the selected SNPs; (iii) stability; and (iv) computational runtime.

Both solution size and active set of SNPs selected by Lasso varies greatly between the different methods (Figure \ref{fig:benchmark}A). heinz has the smallest solutions, with an average of 182 selected our of which 5.6\% (10.2) are selected by Lasso. The largest solutions come from SConES GI (6256.6 SNPs), and dmGWAS (4255.0 SNPs). Interestingly, heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%), although it is  high for all the methods (> 86\%). This suggests methods are selecting informative SNPs on average.

The sensitivity and specificity of the classifier on the testing data informs us about the usefulness of the selected SNPs as patient classification (Figure \ref{fig:benchmark}B). All classifiers' sensitivities were in the 0.42 -- 0.51 range; the specificities, between 0.54 and 0.62. On average, SigMod had the highest sensitivity (0.51); dmGWAS, the highest specificity (0.52). Both heinz and SigMod had on average better sensitivity than the classifier trained on all the SNPs, but none had superior specificity. However, the differences are negligible, well within the 95\% confidence interval.

Another desirable quality of an algorithm is stability (Section \ref{methods:comparison}). Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure \ref{fig:benchmark}C). The other methods displayed similarly low stabilities.

In terms of computational runtime, the fastest method was heinz (Figure \ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds. The slowest method was dmGWAS (1 day and 17 hours on average) followed by SConES GI (1 day and 4.32 hours on average). Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours. Considering that, it took 3 days and 2.4 hours on average for dmGWAS' to produce results.

\subsection{No solution is perfect}
\label{sec:orgc8f05c9}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/figure_2.pdf}
\caption{\label{fig:issues}
Drawbacks confronted when using network guided methods. \textbf{(A)} dmGWAS solution subnetwork. Genes with a P-value < 0.1 are highlighted in red. \textbf{(B)} Centrality degree and -log\textsubscript{10} of the VEGAS P-value for the nodes in SigMod solution subnetwork. \textbf{(C)} Genomic regions where either SConES GS, GM or GI select SNPs.}
\end{figure}

In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different (Supplementary Figure \ref{sfig:pearson_methods}A). That is due to the particulars of the methods, and directly or indirectly, they provide information about the dataset. For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.

In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P-value for the subnetwork (Table \ref{tab:gene_solutions}). Due to this parsimonious and highly associated solution, it was the best method to select a set of good biomarkers for classification. (Figure \ref{fig:benchmark}B). Its conservativeness stems from its preprocessing step, which models the gene p-values as a mixture model of a beta and a uniform distribution, controlled by an FDR parameter. Due to the limited signal at the gene level in this dataset (Figure \ref{fig:snp_gene_manhattan}B), only 36 of them are retain a positive score after applying the BUM model (Section \ref{methods:methods}). Hence, heinz's solution subnetwork consists only of 4 genes, which does not provide much insight of the biology of cancer. Importantly, it ignores genes that are strongly associated to cancer in this dataset like FGFR2.

On the other end of the spectrum, we have large solutions provided by dmGWAS, HotNet2, and SigMod. dmGWAS' subnetwork is the least associated subnetwork on average. This is due to the greedy framework it uses, which considers all nodes at distance 2 of the examined, and accepts weakly associated genes if they are linked to another, strongly associated one. This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure \ref{fig:issues}A). SigMod displays the same tendency, as the most central genes are the least associated to the disease (Figure \ref{fig:issues}B). This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses. In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks. Lastly, SigMod misses some of the most strongly associated, familial breast cancer genes in the dataset, like FGFR2 and TOX3.

By virtue of using a SNP subnetwork, SConES analyzes each SNP in their context. It therefore selects SNPs in genes none of whose interactors are associated to the disease, as well as SNPs in non-coding regions or in non-interacting genes. In fact, due to linkage disequilibrium, such genes are favored by SConES, as selecting an SNPs in an LD block which overlaps with a gene favors selecting the rest of the gene. This might explain why the GS and GM networks, heavily affected by linkage disequilibrium, produce similar results (Supplementary Figure \ref{sfig:pearson_methods}B). On the other hand, SConES penalizes selecting SNPs and not their neighbors. This makes it conservative regarding SNPs with many interactions, for instance those mapped to hubs in the PPIN. For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure \ref{fig:issues}C). In fact SConES GS and SConES GM select regions related to breast cancer, like 16q12 (TOX3, Section \ref{results:conventional}), 3p24 (SLC4A7/NEK10 \cite{ahmed_newly_2009}), 5p12 (FGF10, MRPS30 \cite{quigley_5p12_2014}), and 10q26 (FGFR2, Section \ref{results:conventional}). On top of that only SConES GS selects region 8q24 (POU5F1B \cite{breyer_expressed_2014}). We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN due to the same cause: the absence of joint association of a module. Although in the case of SConES other hyperparameters could lead to a more informative solution (lower \(\lambda\)), finding the appropriate configuration is hard. In addition, the iCOGS platform is not a real GWAS experiment: the genome is not unbiasedly surveyed, some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied - hurting the accuracy with which the GS network captures the genome structure.

\subsection{Aggregating solutions provides insights into the biology of cancer}
\label{results:consensus}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/figure_3.pdf}
\caption{\label{fig:consensus}
Consensus subnetwork on GENESIS (Section \ref{methods:consensus}). Each node is represented by a pie chart, which accounts the methods that selected it. The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known familial breast cancer genes (colored in pink).}
\end{figure}

To leverage on the strengths of each of the methods and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section \ref{methods:consensus}). The consensus subnetwork (Figure \ref{fig:consensus}) contains 93 genes and is enriched in familial breast cancer genes (Fisher's exact test P-value = 7.8 \texttimes{} 10\textsuperscript{-5}). Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure \ref{sfig:consensus_stats}A). Encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure \ref{sfig:consensus_stats}B). Globally, a GO enrichment shows the involvement of two cellular processes: unfolded protein binding, and structural constituent of cytoskeleton (adjusted P-values of 0.001, 0.001, respectively), which were already observed in different solutions (Section \ref{results:separate_networks}). Remarkably, many of the selected genes are related to mitochondrial translation. For instance, MRPS30 (VEGAS P-value = 0.001), encodes a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}. Albeit disconnected from MRPS30, the consensus network includes a 2-node subnetwork composed of two mitochondrial ribosomal protein (MRPS31 - VEGAS P-value = 7.67 \texttimes{} 10\textsuperscript{-3} - and MRPS18B - VEGAS P-value = 7.92 \texttimes{} 10\textsuperscript{-3}), which suggests an involvement of mitochondrial ribosomes in carcinogenesis \cite{required}.

We also examined the topological properties of the nodes. The genes in the consensus network have higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value = 4.29 \texttimes{} 10\textsuperscript{-18}). Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value = 0.57). Centrality, however, is weakly anti-correlated with association to the disease (Pearson correlation coefficient = -0.26, Supplementary Figure \ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes. In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}. For instance, the most central node in the consensus network is COPS5 (Supplementary Figure \ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}. Despite its lack of association in GENESIS (VEGAS P-value = 0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value = 0.006).

The consensus subnetwork is not completely connected: out of the 93 genes, the largest connected subnetwork includes only 49. A GO enrichment analysis showed that this component is related to three major cellular processes: unfolded protein binding, structural constituent of cytoskeleton, and poly(U) RNA binding (adjusted P-values of 0.01, 0.04, and 0.04, respectively). We found support in the literature of the involvement of each of these functions in the development of cancer, as discussed next. The consensus network also contains a protein directly involved in caspase-mediated apoptosis, CASP8 (VEGAS P-value = 1.95 \texttimes{} 10\textsuperscript{-4}). This is related to the enriched activity, \emph{unfolded protein binding}, which inhibits caspase-dependent apoptosis, raising the chances of developing cancer \cite{calderwood_heat_2016}. It involves three Hsp70 chaperones of the consensus subnetwork: HSPA1A, HSPA1B, and HSPA1L. They all are closely encoded in 6p21. In fact out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association. HSPA1A was the most strongly associated one (VEGAS P-value = 8.37 \texttimes{} 10\textsuperscript{-4}). Remarkably, 14 of the 93 genes are in subnetworks of size 1 (isolated) or two, as they do not have a consistently altered neighborhood. One of them is the familial breast cancer FGFR2 (Section \ref{results:conventional}). Another one is SLC4A7 (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{5}), a gene encoding a sodium bicarbonate cotransporter. In the past the genomic region containing both SLC4A7 and nearby gene NEK10 (VEGAS P-value = 1.56 \texttimes{} 10\textsuperscript{-5}) were associated with familial breast cancer \cite{ahmed_newly_2009}. NEK10 is a gene that might be involved in cell-cycle control, but it is absent from the PPIN and hence it could not be studied by gene methods. Despite that, the fact that both dmGWAS, HotNet2 and SigMod link SLC4A7 in their different subnetwork supports the notion that this gene is the responsible for cancer susceptibility.

\subsection{Limitations of network analyses}
\label{sec:org73465bb}

The strength of network-based analyses comes from leveraging prior knowledge to boost discovery. In consequence, they falter in front of understudied genes, especially those not in the network. Out of the 32767 genes that we can map the genotyped SNPs, 60.7\% (19887) are not in the protein-protein interaction network. The majority of those (14660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure \ref{sfig:biotypes_excluded}). The importance of these genes, like CASC16, is highlighted in Section \ref{results:conventional}. Among the excluded protein-coding genes we find genes like NEK10 (P-value 1.6 \texttimes{} 10\textsuperscript{-5}) or POU5F1B, both linked to breast cancer susceptibility \cite{ahmed_newly_2009}. However, on average protein coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value = 2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47). As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions. As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely. It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in NEK10 and CASC16. Lastly, all the methods rely heavily on how SNPs are mapped to genes. In Section \ref{results:conventional} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.

As not all databases compile the same interactions, the choice of the PPIN determines the final output. In this work we used exclusively interactions from HINT from high-throughput experiments. This responds to concerns of some authors about biases introduced by adding interactions coming from targeted studies in the literature \cite{cai_broker_2010,das_hint:_2012}. It is a ``rich getting richer'' phenomenon, where popular genes have a higher proportion of their interactions described. On the other hand, one study found that the best predictor of the performance of a network for disease gene discovery is the size of the network \cite{huang_systematic_2018}. This would support using the largest amount of interactions. To clarify their impact on this study, we compared the impact of using only physical interactions from high-throughput experiment versus interactions from both high-throughput and the literature (Section \ref{methods:gene_network}. For most of the methods using a larger network did not greatly impact the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure \ref{sfig:lc_ht_comparison}).

\section{Discussion}
\label{sec:orgbfa6dab}

In this article we evaluate the viability of systems biology approaches to GWAS, and examine a GWAS dataset on familial breast cancer focused on BRCA1/2 negative French women. Systems biology addresses two of the largest GWAS issues: interpretability and an overly conservative statistical framework that hinders discovery. This is achieved by considering the biological context of each of the genes and SNPs. Based on divergent considerations of what the desired set of biomarkers is, several methods for network-guided biomarker discovery have been proposed. In this article we reviewed the performance of six of them on GWAS. Despite their differences, most of them produced a relevant subset of biomarkers, recovering known familial breast cancer genes. We also discuss the limitations of such analyses, related to the lack of known interactions around some genes. A crucial step for the gene based methods is the computation of the gene score. In the manuscript we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations. However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), other algorithms might have more statistical power.

The network methods we studied differ in what the optimal solution subnetwork looks like. On the one hand, SConES and heinz prefer small highly associated solutions. On the other hand, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context. While the former provide a reduced set of biomarkers, the latter deepen our understanding of the disease and provide biological hypotheses. They are not exempt of limitations. dmGWAS and SigMod's solution's size require further analyses, which risk oversimplifying their richness. Also, incautious practitioners might be misled by some genes, which are very central in the solution subnetworks, while being weakly associated. Nonetheless, they are pushed into the solution by their privileged topological properties. On the other end, conservative solutions, like SConES GI and heinz might not shed much light on the etiology of the disease.

To overcome the problems posed by the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork. We use a straightforward aggregation to generate it, including any node that was recovered by at least two methods. The resulting network is a synthesis of the altered mechanism: is smaller than the largest solutions (SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz). The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering familial breast cancer genes as well as genome regions associated to cancer susceptibility. However, thanks to its small size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.

In order to produce the consensus network, we had to face the different interfaces, preprocessing steps, and unexpected behaviors of the various methods. To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation. They are available on GitHub: \url{https://github.com/hclimente/gwas-tools}. Importantly, those methods that had a permissive license were compiled into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.

\section{Acknowledgments}
\label{sec:org7847920}

H.C-G. was funded by the European Union’s Horizon 2020 research and innovation program under the Marie Skłodowska-Curie [666003].

\bibliographystyle{ieeetr}
\bibliography{bibliography}

\clearpage
\setcounter{figure}{0}
\setcounter{section}{0}
\setcounter{table}{0}

\section*{Supplementary materials}
\label{sec:org7e6740f}

\begin{table}[htbp]
\begin{threeparttable}
\caption{\label{tab:snp_solutions}
Summary statistics on the results of SConES on the three SNP-SNP interaction networks. The first row within each block contains the summary statistics on the whole network.}
\centering
\begin{tabular}{lrrllr}
Network & SNPs & Edges & Subnetworks & \mean{Betweenness} & \median{P}\textsubscript{SNP}\\
\hline
GS & 197\,083 & 197\,060 & - & 2.03 \texttimes{} 10\textsuperscript{7} & 0.49\\
SConES GS & 1\,590 & 1\,585 & 5 & 2.52 \texttimes{} 10\textsuperscript{7} & 0.023\\
\hline
GM & 197\,083 & 6\,442\,446 & - & 3.99 \texttimes{} 10\textsuperscript{6} & 0.49\\
SConES GM & 1\,692 & 177\,611 & 5 & 4.40 \texttimes{} 10\textsuperscript{6} & 0.055\\
\hline
GI & 197\,083 & 28\,733\,720 & - & 1.46 \texttimes{} 10\textsuperscript{6} & 0.49\\
SConES GI & 408 & 539 & 5 & 9.33 \texttimes{} 10\textsuperscript{6} & 0.076\\
\end{tabular}
\begin{tablenotes}
\mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network; \median{P}\textsubscript{SNP}: median P-value of the selected SNPs.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{table}[htbp]
\begin{threeparttable}
  \caption{\label{tab:scones_gene_solutions}
Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
\centering
\begin{tabular}{lrrrlr}
Network & Genes & Edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & $\rho_{consensus}$\\
\hline
SConES GS & 5 & 0 & 9805 & 2.7 \texttimes{} 10\textsuperscript{-5} & 0.19\\
SConES GM & 28 & 2 & 4267 & 0.067 & 0.12\\
\end{tabular}
\begin{tablenotes}
\mean{Betweenness}: mean betweenness of the selected genes in the full network; \median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson's correlation with the consensus network.
\end{tablenotes}
\end{threeparttable}
\end{table}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
\caption{\label{sfig:pcs}
GENESIS shows no differential population structure between cases and controls. \textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS. Cases are colored in green, controls in orange.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_2.pdf}
\caption{\label{sfig:overview_solutions}
Overview of the subnetworks produced by the different network methods. \textbf{(dmGWAS, heinz, HotNet2, LEAN, and SigMod)} contain gene subnetworks; \textbf{(SConES GI)}, SNP subnetworks.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_3.png}
\caption{\label{sfig:manhattan_solutions}
Manhattan plots showing the biomolecules selected by each method. In \textbf{(Consensus, dmGWAS, heinz, HotNet2, and SigMod)} datapoints are genes; in \textbf{(SConES GI)}, SNPs. LEAN was excluded, as it did not select any gene.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_4.pdf}
\caption{\label{sfig:bcac_comparison}
Proportion of the Bonferroni significant biomolecules (in either the GENESIS or the BCAC datasets) selected by each of the methods on the GENESIS data. \textbf{(Consensus, dmGWAS, heinz, HotNet2, and SigMod)} involve significant genes, only among those present in the protein-protein interaction network. \textbf{(SConES GS, GM and GI)} involve significant SNPs. LEAN was excluded, as it did not select any gene. The presented network methods recover a higher proportion of significant genes than of non-significant genes in both datasets, despite their lack of significance in GENESIS.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_5.pdf}
\caption{\label{sfig:disgenet_comparison}
:Proportion of the selected genes by each of the methods on the GENESIS data that is a known familial breast cancer gene (Section \ref{methods:bcac}). Only genes present in the protein-protein interaction network were considered. LEAN was excluded, as it did not select any gene. The presented network methods recover a higher proportion of familial breast cancer genes than of other genes, despite their lack of significance in GENESIS.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
\caption{\label{sfig:pearson_methods}
Pearson's correlation between the different solution subnetworks. \textbf{(A)} Correlation between selected SNPs. \textbf{(B)} Correlation between selected genes. In general, the solutions display a very low overlap.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
\caption{\label{sfig:consensus_names}
Consensus subnetwork on GENESIS (Section \ref{methods:consensus}). \textbf{(A)} Each node is represented by a pie chart, which accounts the methods that selected it. The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known familial breast cancer genes (colored in pink). This panel is equivalent to Figure \ref{fig:consensus}. \textbf{(B)} The name of every gene is indicated.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_8.pdf}
\caption{\label{sfig:consensus_stats}
Genes on the consensus network. Familial breast cancer genes are colored in pink; the rest are colored in grey. \textbf{(A)} Number of methods selecting every gene in the subnetwork. \textbf{(B)} VEGAS P-values of association of the genes, with regards to the number of methods that selected them. \textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network. To improve visualization, we removed outliers. \textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS P-value of the genes in the consensus network. The blue line represents a fitted generalized linear model.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_9.pdf}
\caption{\label{sfig:biotypes_excluded}
Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/sfigure_10.pdf}
\caption{\label{sfig:lc_ht_comparison}
Comparison of benchmark on high-throughput interactions to benchmark on both high-throughput and literature curated interactions. Grey lines represent no change between the benchmarks (1 for ratios, 0 for differences). \textbf{(A)} Ratios of the selected features between both benchmarks and of the active set. \textbf{(B)} Shifts in sensitivity and specificity. \textbf{(C)} Shift in Pearson's correlation between benchmarks. \textbf{(D)} Ratio between the runtimes of the benchmarks. For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119980 seconds (1 day and 9.33 hours) which took VEGAS2v2 on average to compute the gene scores from SNP summary statistics. In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}
\end{figure}
\end{document}

### 2019-11-16 manuscript/manuscript.tex
updated with fabienne's comments

@@
 \author{Héctor Climente-González$^{1,2,3}$,Christine Lonjou$^{1,2,3}$,Fabienne Lesueur$^{1,2,3}$,GENESIS
-  investigators,Nadine Andrieu$^{1,2,3}$,Chloé-Agathe Azencott$^{3,1,2}$}
+  investigators,Dominique Stoppa-Lyonnet$^{4,5,6}$,Nadine Andrieu$^{1,2,3}$,Chloé-Agathe Azencott$^{3,1,2}$}
 \affiliation{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France; \\
   $^{2}$INSERM, U900, F-75005 Paris, France; \\
-  $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France; \\
+  $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
+  $^4$Service de Génétique, Institut Curie, F-75005 Paris, France;
+  $^5$INSERM, U830, F-75005 Paris, France;
+  $^6$Université Paris Descartes;
+
 }
@@
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
-It does so by jointly considering the statistical association between a gene and a phenotype, obtained experimentally, and the biological context of each gene, represented as a network.
-In this work we study the utility of six network methods to discover new biomarkers in GWAS
-data by searching subnetworks highly associated to a phenotype.
-We interrogate a familial breast cancer GWAS focused on BRCA1/2 negative French women.
+It does so by jointly considering the statistical association between a gene and a phenotype, and the biological context of each gene, represented as a network.
+In this work we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks highly associated to a phenotype.
+We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
 By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-We
-perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-Interestingly, a combination solution subnetworks provided a concise subnetwork of 93 genes, enriched in known familial breast cancer susceptibility genes (BABAM1, BLM, CASP8, FGFR2, and TOX3, Fisher's exact test
-P-value = $7.8 \times 10^{-5}$ ) and more central than average.
-Additionally, it includes subnetworks of mechanisms related to cancer, like protein folding (HSPA1A, HSPA1B, and HSPA1L) or mitocondrial ribosomes (MRPS30, MRPS31, MRPS18B).
-We also observed a general disregulation in the neighborhood of COPS5, a gene related to multiple hallmarks of cancer.
+We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
+Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value = $7.8 \times 10^{-5}$ ) and more central than average.
+Additionally, it includes subnetworks of mechanisms related to cancer, like protein folding (\emph{HSPA1A}, \emph{HSPA1B}, and \emph{HSPA1L}) or mitochondrial ribosomes (\emph{MRPS30}, \emph{MRPS31}, \emph{MRPS18B}).
+We also observed a general dysregulation in the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
 
@@
 Then, a per-SNP statistical association test is conducted (e.g.
-\(\chi\)\textsuperscript{2}).
+logistic regression).
 Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
-Once the risk SNPs have been discovered, they can be used for diagnosis, guide the treatment, and deepen our understanding of the disease.
+Once the risk SNPs have been discovered, they can be used for risk assessment, and to deepen our understanding of the disease.
 
@@
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
-The fact that SNPs in non-coding regions often top the results of GWAS illustrates well this challenge.
 On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The omnigenic model \cite{boyle_expanded_2017} explains this by the dense functional inter-relatedness between genes, influencing each others' behavior, which allows alterations in most genes to impact the ``core'' genes involved in a disease.
+The omnigenic model \cite{boyle_expanded_2017} explains this by the dense functional inter-relatedness between genes, influencing each other's behavior, which allows alterations in most genes to impact the ``core'' genes involved in a disease.
 A comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
@@
 In this work, we analyze the effectiveness of these six methods to discover new biomarkers on GWAS data.
-We focus on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a homogeneous dataset of familial breast cancer in the French population.
+We focus on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study of familial breast cancer conducted in the French population.
 After following a classical GWAS approach, we use these network-based methods to recover additional familial breast cancer biomarkers.
@@
 
-GENESIS is a GWAS study of familial breast cancer on the French population \cite{sinilnikova_genesis:_2016}.
-The index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had sister with breast cancer, and negative for BRCA1 and BRCA2 mutations.
-The controls are unaffected colleagues and/or friends of the cases, born around the year of birth of the corresponding case (\textpm{} 3 years).
+The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
+Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for BRCA1 and BRCA2 pathogenic variants.
+Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of the corresponding case (\textpm{} 3 years).
 We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
-The genotyping platform was iCOGS, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
-It contains 210\,918 SNPs previously associated with cancer susceptibility, survival, or other cancer-related traits; and in selected genes and pathways.
+The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
+It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as functional candidate variants in selected genes and pathways.
 
@@
 
-We discarded SNPs with a minor allele frequency lower than 0.1\%, not in Hardy - Weinberg equilibrium (P-value \textless 0.001), and those with missing values on more than 10\% of the samples.
-A subset of 20 duplicated SNPs in FGFR2 were also removed.
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy - Weinberg equilibrium in controls (P-value \textless 0.001), and those missing on more than 10\% of the samples.
+A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
 In addition, we removed the samples with more than 10\% missing genotypes, and an additional 28 samples with TODO.
@@
 Then, we used VEGAS2v2 to compute the gene-level association score \cite{mishra_vegas2:_2015} from the SNP P-values.
-In order to map SNPs to genes we used their overlap on the genome: all SNPs located within the boundaries of a gene, \textpm{} 50 kb, were mapped to that gene.
-To compute the gene association we used the 10\% of SNPs with lowest P-values.
+In order to map SNPs to genes we used their overlap on the genome: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
+To compute the gene association we used the 10\% of SNPs linked to the gene with lowest P-values.
 We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
@@
 The described matrices are square, and have as many rows and columns as nodes are in the network.
-The element $(i,j)$ represents a  selected relationship between $v_{i}$ and $v_{j}$.
-The \emph{adjacency matrix} $W_{G}$ contains a 1 when the corresponding nodes are connected through an edge, and 0 otherwise; the diagonal is zero.
-The \emph{degree matrix} $D_{G}$ is a diagonal matrix which contains the degree of the different nodes.
-Lastly, the \emph{Laplacian matrix} $L_{G}$ is defined as $L_{G} = D_{G} - W_{G}$.
+The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
+The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected through an edge, and 0 otherwise; the diagonal is zero.
+The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
+Lastly, the \emph{Laplacian matrix} $L_G$ is defined as $L_G = D_G - W_G$.
 
@@
 \label{methods:methods}
+
 \begin{table}[htbp]
+  \begin{threeparttable}
     \caption{\label{tab:method_comparison}
@@
     \begin{tabular}{l|lllllll}
-Method & Field & Nodes & Exhaustive & Solution & Components & Input & Scoring\\
+      Method & Field & Nodes & Exhaustive & Solution & Components & Input & Scoring & Reference\\
       \hline
-dmGWAS & GWAS & Gene & No & - & 1 & Summary & -log\textsubscript{10}(P)\\
-heinz & Omics & Gene & Yes & - & 1 & Summary & BUM\\
-HotNet2 & Omics & Gene & Yes & Modular & \(\ge\) 1 & Summary & Local FDR\\
-LEAN & Omics & Gene & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P)\\
-SConES & GWAS & SNP & Yes & Modular & \(\ge\) 1 & Genotypes & \(\chi\)\textsuperscript{2}\\
-SigMod & GWAS & Gene & Yes & Modular & 1 & Summary & -log\textsubscript{10}(P)\\
+      dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
+      heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
+      HotNet2 & Omics & Genes & Yes & Modular & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
+      LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
+      SConES & GWAS & SNPs & Yes & Modular & \(\ge\) 1 & Genotypes & \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
+      SigMod & GWAS & Genes & Yes & Modular & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
     \end{tabular}
+    \begin{tablenotes}
+      /Field/: field in which the algorithm was developed.
+/Nodes/, the type of network, either gene (protein-protein interaction network usually) or a SNP network.
+/Exhaustive/: whether all the possible solutions given the selected hyperparameters are explored.
+/Solution/: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.
+/Components/: number of connected subnetworks in the solution.
+/Input/: genotype data or GWAS summary statistics.
+/Scoring/: how SNP/gene P-values are transformed into node scores.
+    \end{tablenotes}
+  \end{threeparttable}
 \end{table}
@@
 \item[{dmGWAS}] dmGWAS searches the subgraph with the highest local density in low P-values \cite{jia_dmgwas:_2011}.
-To that end it searches candidate subnetwork solutions using a greedy procedure involving the following steps:
+To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
 
@@
 As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs.
-The authors suggested three ways of building these networks: connecting the those SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists (``GI network'').
+The authors suggested three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists (``GI network'').
 We focused on the GI network, as it is the network that fits better the scope of this article.
@@
 On this same subset we trained our classifier exclusively on the SNPs selected by a particular method.
-When the method retrieved a list of genes (all of them except SConES), all the SNPs mapped to any of those genes were used.
+When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes were used.
 Then we evaluated performance of the classifier on the remaining 20\% of the dataset.
 We repeated this procedure 5 times to estimate the average and the deviation of the different performance measures.
-The different performance measures we used where: size of the solution, size of the active set, specificity, and sensitivity.
+The different performance measures we used were: size of the solution, size of the active set, specificity, and sensitivity.
 The size of the active set provides an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
@@
 For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}.
-Across this article we refer to these genes as \emph{familial breast cancer genes}.
+Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
 
 Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015}.
-BCAC is one of the largest efforts in GWAS, genotyping over 120\,000 women of European ancestry.
+BCAC has conducted one of the largest efforts in GWAS, involving over 120\,000 women of European ancestry.
 As opposed to GENESIS, samples were not selected based on family history, and hence is enriched in sporadic breast cancers.
@@
 The code that applies these pipelines to the GENESIS project, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
-Although the GENESIS data is not publicly available, the published code should work on any other GWAS dataset.
+Although the GENESIS dataset is not publicly available, the published code should work on any other GWAS dataset.
 All the produced gene subnetworks were deposited on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
@@
 \label{sec:orgc718f04}
-\subsection{FGFR2 is strongly associated with familial breast cancer}
+\subsection{A conventional GWAS shows that FGFR2 is strongly associated with familial breast cancer}
 \label{results:conventional}
@@
 Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Figure \ref{fig:snp_gene_manhattan}A).
-The former overlaps with gene FGFR2; the latter with CASC16, and it is located near the protein-coding gene TOX3.
-Variants in both FGFR2 and TOX3 were related to breast cancer susceptibility in other cohorts negative for BRCA1/2 \cite{rinella_genetic_2013}.
-Only the peak in chromosome 10 replicated in the gene-level analysis, with FGFR2 just above the threshold of significance (Figure \ref{fig:snp_gene_manhattan}B).
+The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
+Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other cohorts negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
+Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Figure \ref{fig:snp_gene_manhattan}B).
 
-These results show the overlap in the genetic architecture of the disease between the French population and other cohorts, especially at the gene level.
+These results show the overlap in the genetic architecture of the disease between the studied French population sample and other populations, especially at the gene level.
 In addition, there are other SNPs whose P-values, although higher than the conventional threshold of significance, show a strong association with familial breast cancer.
-The most prominent of such regions are 3p24 and 8q24, both of which  which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
+The most prominent of such regions are 3p24 and 8q24, both of which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
 This motivates exploring network methods, which trade statistical significance for biological relevance.
 
-\subsection{Network methods successfully identify genes linked to breast cancer}
+\subsection{Network methods successfully identify genes associated with breast cancer}
 \label{results:separate_networks}
@@
 \begin{tablenotes}
-\emph{\# genes}: number of genes selected out of those that are part of the PPIN; in parentheses, the total number of genes.
+\emph{\# genes}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
 \mean{Betweenness}: mean betweenness of the selected genes in the PPIN.
@@
 SConES GI failed to recover genes in the PPIN, but it recovered one genomic region mapped to RNA gene RNU6-420P.
-All solution subnetworks except LEAN's are, on average, more strongly associated to familial breast cancer than the whole PPIN (median P-values $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{sfig:manhattan_solutions}).
+All solution subnetworks except LEAN's are, on average, more strongly associated to breast cancer than the whole PPIN (median P-values $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{sfig:manhattan_solutions}).
 This exemplifies the trade-off between statistical significance and biological relevance.
@@
 The solution subnetworks present other desirable properties.
-First, four of the methods succeeded at recovering genes involved in the disease (Supplementary Figure \ref{sfig:disgenet_comparison}), as their subnetworks were enriched in familial breast cancer genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
-We also compared the outcome of the network methods to the association tests conducted on the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015} (Supplementary Figure \ref{sfig:bcac_comparison}).
+First, four of the methods succeeded at recovering genes involved in the disease (Supplementary Figure \ref{sfig:disgenet_comparison}), as their subnetworks were enriched in breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
+We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015} (Supplementary Figure \ref{sfig:bcac_comparison}).
 Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC.
@@
 This agrees with the notion that disease genes are more central than other, non-essential genes \cite{pinero_uncovering_2016}.
-We observe that this conclusion holds in this disease, as familial breast cancer genes have higher betweenness centrality than others (one-tailed Wilcoxon rank-sum test P-value = 2.64 \texttimes{} 10\textsuperscript{-5}, Supplementary Figure \ref{sfig:consensus_stats}C).
+We observe that this conclusion holds in this disease, as known breast cancer susceptibility genes have higher betweenness centrality than others (one-tailed Wilcoxon rank-sum test P-value = 2.64 \texttimes{} 10\textsuperscript{-5}, Supplementary Figure \ref{sfig:consensus_stats}C).
 Interestingly, SConES' selected SNPs are also more central than the average SNP (Supplementary table \ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than unrelated SNPs.
@@
 As the solutions were quite different from each other it is hard to draw joint conclusions.
-The 4-gene solution selected by heinz includes the familial breast cancer gene TOX3, in region 16q12.
-By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or else.
-In fact, SConES GI, which adds to GM the interactions between genes, retrieves 4 subnetworks in intergenic regions, and 1 overlapping an RNA gene (RNU6-420P).
+The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
+By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or not.
+In fact, SConES GI, which adds to GM the interactions between genes, retrieves 4 subnetworks in intergenic regions, and 1 overlapping an RNA gene (\emph{RNU6-420P}).
 SigMod, despite being related to SConES, produces a vastly different, large solution.
-On top of recovering three familial breast cancer genes, a keratin-based region of its subnetwork affects the cytoskeleton (\emph{structural constituent of cytoskeleton}, GO enrichment's adjusted P-value = 9.10 \texttimes{} 10\textsuperscript{-4}), a potentially novel susceptibility mechanisms to cancer.
+On top of recovering three breast cancer susceptibility genes, a keratin-based region of its subnetwork affects the cytoskeleton (\emph{structural constituent of cytoskeleton}, GO enrichment's adjusted P-value = 9.10 \texttimes{} 10\textsuperscript{-4}), a potentially novel susceptibility mechanism for cancer susceptibility.
 Interestingly, dmGWAS solution is also related to cytoskeleton (\emph{tubulin binding}, GO enrichment's adjusted P-value = 0.031).
-But, additionally, it includes a submodule of proteins related to \emph{unfolded protein binding} (GO enrichment's adjusted P-value = 0.045), which was related to cancer susceptibility \cite{calderwood_heat_2016}.
+But, additionally, it includes a submodule of proteins related to \emph{unfolded protein binding} (GO enrichment's adjusted P-value = 0.045), which has been previously related to cancer susceptibility \cite{calderwood_heat_2016}.
 Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
-The second largest subnetwork (13 nodes), contains two familial breast cancer genes: CASP8 and BLM.
+The second largest subnetwork (13 nodes), contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
 
@@
 
-As the methods produced such different results, we compared their solutions in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+As the employed methods produced such different results, we compared their solutions in a 5-fold subsampling setting (Section \ref{methods:comparison}).
 Specifically, we measured the following properties (Figure \ref{fig:benchmark}): (i) size of the solution subnetwork; (ii) sensitivity and specificity of an L1-penalized logistic regression on the selected SNPs; (iii) stability; and (iv) computational runtime.
@@
 Both solution size and active set of SNPs selected by Lasso varies greatly between the different methods (Figure \ref{fig:benchmark}A).
-heinz has the smallest solutions, with an average of 182 selected our of which 5.6\% (10.2) are selected by Lasso.
+heinz has the smallest solutions, with an average of 182 selected SNPs, out of which 5.6\% (10.2) are selected by Lasso.
 The largest solutions come from SConES GI (6256.6 SNPs), and dmGWAS (4255.0 SNPs).
@@
 
-The sensitivity and specificity of the classifier on the testing data informs us about the usefulness of the selected SNPs as patient classification (Figure \ref{fig:benchmark}B).
+To determine whether the selected SNPs could be used for patient classification we computed the sensitivity and the specificity of the classifier on the testing data (Figure \ref{fig:benchmark}B).
 All classifiers' sensitivities were in the 0.42 -- 0.51 range; the specificities, between 0.54 and 0.62.
@@
 Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure \ref{fig:benchmark}C).
-The other methods displayed similarly low stabilities.
+Conversely, the other methods displayed similarly low stabilities.
 
@@
 In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different (Supplementary Figure \ref{sfig:pearson_methods}A).
-That is due to the particulars of the methods, and directly or indirectly, they provide information about the dataset.
+That is due to the particularities of the methods which directly or indirectly provide information about the dataset.
 For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
@@
 Hence, heinz's solution subnetwork consists only of 4 genes, which does not provide much insight of the biology of cancer.
-Importantly, it ignores genes that are strongly associated to cancer in this dataset like FGFR2.
+Importantly, it ignores genes that are strongly associated to cancer in this dataset like \emph{FGFR2}.
 
@@
 In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
-Lastly, SigMod misses some of the most strongly associated, familial breast cancer genes in the dataset, like FGFR2 and TOX3.
+Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
 
@@
 It therefore selects SNPs in genes none of whose interactors are associated to the disease, as well as SNPs in non-coding regions or in non-interacting genes.
-In fact, due to linkage disequilibrium, such genes are favored by SConES, as selecting an SNPs in an LD block which overlaps with a gene favors selecting the rest of the gene.
+In fact, due to linkage disequilibrium, such genes are favored by SConES, as selecting SNPs in an LD block which overlaps with a gene favors selecting the rest of the gene.
 This might explain why the GS and GM networks, heavily affected by linkage disequilibrium, produce similar results (Supplementary Figure \ref{sfig:pearson_methods}B).
@@
 For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure \ref{fig:issues}C).
-In fact SConES GS and SConES GM select regions related to breast cancer, like 16q12 (TOX3, Section \ref{results:conventional}), 3p24 (SLC4A7/NEK10 \cite{ahmed_newly_2009}), 5p12 (FGF10, MRPS30 \cite{quigley_5p12_2014}), and 10q26 (FGFR2, Section \ref{results:conventional}).
-On top of that only SConES GS selects region 8q24 (POU5F1B \cite{breyer_expressed_2014}).
+In fact SConES GS and SConES GM select regions related to breast cancer, like 16q12 (\emph{TOX3}, Section \ref{results:conventional}), 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), and 10q26 (\emph{FGFR2}, Section \ref{results:conventional}).
+On top of that only SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
 We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN due to the same cause: the absence of joint association of a module.
-Although in the case of SConES other hyperparameters could lead to a more informative solution (lower \(\lambda\)), finding the appropriate configuration is hard.
-In addition, the iCOGS platform is not a real GWAS experiment: the genome is not unbiasedly surveyed, some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied - hurting the accuracy with which the GS network captures the genome structure.
+Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
+lower \(\lambda\), Section \ref{methods:methods}), it is unclear what is the best strategy to find them.
+In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hurting the accuracy with which the GS network captures the genome structure.
 
@@
 Each node is represented by a pie chart, which accounts the methods that selected it.
-The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known familial breast cancer genes (colored in pink).}
+The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).}
 \end{figure}
@@
 To leverage on the strengths of each of the methods and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section \ref{methods:consensus}).
-The consensus subnetwork (Figure \ref{fig:consensus}) contains 93 genes and is enriched in familial breast cancer genes (Fisher's exact test P-value = 7.8 \texttimes{} 10\textsuperscript{-5}).
+The consensus subnetwork (Figure \ref{fig:consensus}) contains 93 genes and is enriched in breast cancer susceptibility genes (Fisher's exact test P-value = 7.8 \texttimes{} 10\textsuperscript{-5}).
 Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure \ref{sfig:consensus_stats}A).
@@
 In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is COPS5 (Supplementary Figure \ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+For instance, the most central node in the consensus network is \emph{COPS5} (Supplementary Figure \ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
 Despite its lack of association in GENESIS (VEGAS P-value = 0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value = 0.006).
@@
 We found support in the literature of the involvement of each of these functions in the development of cancer, as discussed next.
-The consensus network also contains a protein directly involved in caspase-mediated apoptosis, CASP8 (VEGAS P-value = 1.95 \texttimes{} 10\textsuperscript{-4}).
+The consensus network also contains a protein directly involved in caspase-mediated apoptosis, \emph{CASP8} (VEGAS P-value = 1.95 \texttimes{} 10\textsuperscript{-4}).
 This is related to the enriched activity, \emph{unfolded protein binding}, which inhibits caspase-dependent apoptosis, raising the chances of developing cancer \cite{calderwood_heat_2016}.
 It involves three Hsp70 chaperones of the consensus subnetwork: HSPA1A, HSPA1B, and HSPA1L.
-They all are closely encoded in 6p21.
-In fact out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
+They genes encoding these proteins are all near each other at 6p21.
+In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
 HSPA1A was the most strongly associated one (VEGAS P-value = 8.37 \texttimes{} 10\textsuperscript{-4}).
-Remarkably, 14 of the 93 genes are in subnetworks of size 1 (isolated) or two, as they do not have a consistently altered neighborhood.
-One of them is the familial breast cancer FGFR2 (Section \ref{results:conventional}).
-Another one is SLC4A7 (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{5}), a gene encoding a sodium bicarbonate cotransporter.
-In the past the genomic region containing both SLC4A7 and nearby gene NEK10 (VEGAS P-value = 1.56 \texttimes{} 10\textsuperscript{-5}) were associated with familial breast cancer \cite{ahmed_newly_2009}.
-NEK10 is a gene that might be involved in cell-cycle control, but it is absent from the PPIN and hence it could not be studied by gene methods.
-Despite that, the fact that both dmGWAS, HotNet2 and SigMod link SLC4A7 in their different subnetwork supports the notion that this gene is the responsible for cancer susceptibility.
-
-\subsection{Limitations of network analyses}
-\label{sec:org73465bb}
-
-The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
-In consequence, they falter in front of understudied genes, especially those not in the network.
-Out of the 32767 genes that we can map the genotyped SNPs, 60.7\% (19887) are not in the protein-protein interaction network.
-The majority of those (14660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure \ref{sfig:biotypes_excluded}).
-The importance of these genes, like CASC16, is highlighted in Section \ref{results:conventional}.
-Among the excluded protein-coding genes we find genes like NEK10 (P-value 1.6 \texttimes{} 10\textsuperscript{-5}) or POU5F1B, both linked to breast cancer susceptibility \cite{ahmed_newly_2009}.
-However, on average protein coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value = 2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
-As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
-As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
-It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in NEK10 and CASC16.
-Lastly, all the methods rely heavily on how SNPs are mapped to genes.
-In Section \ref{results:conventional} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
-
-As not all databases compile the same interactions, the choice of the PPIN determines the final output.
-In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns of some authors about biases introduced by adding interactions coming from targeted studies in the literature \cite{cai_broker_2010,das_hint:_2012}.
-It is a ``rich getting richer'' phenomenon, where popular genes have a higher proportion of their interactions described.
-On the other hand, one study found that the best predictor of the performance of a network for disease gene discovery is the size of the network \cite{huang_systematic_2018}.
-This would support using the largest amount of interactions.
-To clarify their impact on this study, we compared the impact of using only physical interactions from high-throughput experiment versus interactions from both high-throughput and the literature (Section \ref{methods:gene_network}.
-For most of the methods using a larger network did not greatly impact the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure \ref{sfig:lc_ht_comparison}).
+Remarkably, 14 of the 93 genes are in subnetworks of size 1 (isolated) or 2, as they do not have a consistently altered neighborhood.
+One of them is the well-known breast cancer susceptibility gene \emph{FGFR2} (Section \ref{results:conventional}).
+Another one is the also well-known \emph{SLC4A7} gene (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{5}), which encodes a sodium bicarbonate cotransporter.
+The genomic region containing both \emph{SLC4A7} and nearby gene \emph{NEK10} (VEGAS P-value = 1.56 \texttimes{} 10\textsuperscript{-5}) have been consistently associated with breast cancer susceptibility \cite{ahmed_newly_2009}.
+\emph{NEK10} is a gene that might be involved in cell-cycle control, but it is absent from the PPIN and hence it could not be studied by gene methods.
+Despite that, the fact that both dmGWAS, HotNet2 and SigMod link \emph{SLC4A7} in their different subnetwork supports the notion that this gene is the responsible for breast cancer susceptibility.
 
@@
 Based on divergent considerations of what the desired set of biomarkers is, several methods for network-guided biomarker discovery have been proposed.
-In this article we reviewed the performance of six of them on GWAS.
+We reviewed the performance of six of them on GWAS.
 Despite their differences, most of them produced a relevant subset of biomarkers, recovering known familial breast cancer genes.
@@
 We use a straightforward aggregation to generate it, including any node that was recovered by at least two methods.
-The resulting network is a synthesis of the altered mechanism: is smaller than the largest solutions (SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
-The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering familial breast cancer genes as well as genome regions associated to cancer susceptibility.
+The resulting network is a synthesis of the altered mechanism: it is smaller than the largest solutions (SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
+The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
 However, thanks to its small size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.
+
+The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
+In consequence, they show their shortcomings in front of understudied genes, especially those not in the network.
+Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
+The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure \ref{sfig:biotypes_excluded}).
+The importance of these genes, like \emph{CASC16}, is highlighted in Section \ref{results:conventional}.
+Among the excluded protein-coding genes we find genes like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}) or \emph{POU5F1B}, both linked to breast cancer susceptibility \cite{ahmed_newly_2009}.
+However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value = 2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
+As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
+As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
+It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
+Lastly, all the methods rely heavily on how SNPs are mapped to genes.
+In Section \ref{results:conventional} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+
+As not all databases compile the same interactions, the choice of the PPIN determines the final output.
+In this work we used exclusively interactions from HINT from high-throughput experiments.
+This responds to concerns of some authors about biases introduced by adding interactions coming from targeted studies in the literature \cite{cai_broker_2010,das_hint:_2012} where a ``rich getting richer'' phenomenon is observed: popular genes have a higher proportion of their interactions described.
+On the other hand, one study found that the best predictor of the performance of a network for disease gene discovery is the size of the network \cite{huang_systematic_2018}.
+This also supports using the largest amount of interactions.
+To clarify their impact on this study, we compared the impact of using only physical interactions from high-throughput experiment versus interactions from both high-throughput and the literature (Section \ref{methods:gene_network}).
+We conclude that for most of the methods a larger network did not greatly impact the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure \ref{sfig:lc_ht_comparison}).
 
@@
 
-\section{Acknowledgments}
+\section{Funding and acknowledgments}
 \label{sec:org7847920}
 
-H.C-G. was funded by the European Union’s Horizon 2020 research and innovation program under the Marie Skłodowska-Curie [666003].
+This project was supported by funding from Agence Nationale de la Recherche (ANR-18-CE45-0021-01).
+H.C-G.
+was funded by the European Union’s Horizon 2020 research and innovation program under the Marie Skłodowska-Curie [666003].
+Financial support for GENESIS resource and genotyping was provided by the Ligue Nationale contre le Cancer (grants PRE05/DSL, PRE07/DSL, PRE11/NA), the French National Institute of Cancer (INCa grant No b2008-029/LL-LC) and the comprehensive cancer center SiRIC, (Site de Recherche Intégrée sur le Cancer: Grant INCa-DGOS-4654).
+
+GENESIS (GENE SISters) is a French national study sponsored by UNICANCER (Sinilnikova et al.
+BMC Cancer 2016).
+We wish to thank the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie : Séverine Eon-Marchais, M.
+Marcou, D.
+Le Gal, L.
+Toulemonde, J.
+Beauvallet, N.
+Mebirouk, E.
+Cavaciuti), the biological resource centre (S.
+Mazoyer, F.
+Damiola, L.
+Barjhoux, C.
+Verny-Pierre, V.
+Sornin) and all the GENESIS collaborating cancer clinics.
 
@@
 Manhattan plots showing the biomolecules selected by each method.
-In \textbf{(Consensus, dmGWAS, heinz, HotNet2, and SigMod)} datapoints are genes; in \textbf{(SConES GI)}, SNPs.
+In \textbf{(Consensus, dmGWAS, heinz, HotNet2, and SigMod)} datapoints are genes; in \textbf{(SConES GS, GM, and GI)}, SNPs.
 LEAN was excluded, as it did not select any gene.}
@@
 \textbf{(SConES GS, GM and GI)} involve significant SNPs.
-LEAN was excluded, as it did not select any gene.
+LEAN is not displayed as it did not select any gene.
 The presented network methods recover a higher proportion of significant genes than of non-significant genes in both datasets, despite their lack of significance in GENESIS.}
@@
 \caption{\label{sfig:disgenet_comparison}
-:Proportion of the selected genes by each of the methods on the GENESIS data that is a known familial breast cancer gene (Section \ref{methods:bcac}).
+:Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene (Section \ref{methods:bcac}).
 Only genes present in the protein-protein interaction network were considered.
-LEAN was excluded, as it did not select any gene.
-The presented network methods recover a higher proportion of familial breast cancer genes than of other genes, despite their lack of significance in GENESIS.}
+LEAN is not displayed as it did not select any gene.
+The presented network methods recover a higher proportion of breast cancer susceptibility genes than of other genes, despite their lack of significance in GENESIS.}
 \end{figure}
@@
 \textbf{(A)} Each node is represented by a pie chart, which accounts the methods that selected it.
-The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known familial breast cancer genes (colored in pink).
+The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).
 This panel is equivalent to Figure \ref{fig:consensus}.
@@
 Genes on the consensus network.
-Familial breast cancer genes are colored in pink; the rest are colored in grey.
+Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
 \textbf{(A)} Number of methods selecting every gene in the subnetwork.

### 2020-03-02 manuscript/manuscript.tex
update manuscript

@@
-% Created 2019-10-19 Sat 12:19
-% Intended LaTeX compiler: pdflatex
 \documentclass[twocolumn, 10pt]{article}
@@
 \usepackage{threeparttable}
+\usepackage[numbers]{natbib}
 \newcommand{\mean}[1]{$\overline{\mbox{#1}}$}
@@
 \title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
-\author{Héctor Climente-González$^{1,2,3}$,Christine Lonjou$^{1,2,3}$,Fabienne Lesueur$^{1,2,3}$,GENESIS
-  investigators,Dominique Stoppa-Lyonnet$^{4,5,6}$,Nadine Andrieu$^{1,2,3}$,Chloé-Agathe Azencott$^{3,1,2}$}
-\affiliation{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France; \\
-  $^{2}$INSERM, U900, F-75005 Paris, France; \\
+\author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, GENESIS study collaborators, Dominique Stoppa-Lyonnet$^{4,5,6}$, Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
+\affiliation{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
+  $^{2}$INSERM, U900, F-75005 Paris, France;\\
   $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
-  $^4$Service de Génétique, Institut Curie, F-75005 Paris, France;
-  $^5$INSERM, U830, F-75005 Paris, France;
-  $^6$Université Paris Descartes;
-
+  $^{4}$Service de Génétique, Institut Curie, F-75005 Paris, France;\\
+  $^{5}$INSERM, U830, F-75005 Paris, France;\\
+  $^{6}$Université Paris Descartes;\\
 }
 \hypersetup{
- pdfauthor={Héctor Climente-González, Chloé-Agathe Azencott},
+  pdfauthor={Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, GENESIS investigators, Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott},
   pdftitle={Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer},
@@
   pdfsubject={},
- pdfcreator={Emacs 26.2 (Org mode 9.2.3)}, 
   pdflang={English}}
@@
 \MakeTitle
+
 \begin{abstract}
-
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
-It does so by jointly considering the statistical association between a gene and a phenotype, and the biological context of each gene, represented as a network.
-In this work we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks highly associated to a phenotype.
+It does so by jointly considering the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
+In this work, we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks highly associated to this phenotype.
 We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
-By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
@@
 We also observed a general dysregulation in the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-
+By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:orgbee581e}
 
@@
 On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The omnigenic model \cite{boyle_expanded_2017} explains this by the dense functional inter-relatedness between genes, influencing each other's behavior, which allows alterations in most genes to impact the ``core'' genes involved in a disease.
-A comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
+The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the ``core'' genes directly involved in a disease's mechanism.
+Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
 
 In this regard many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
-As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships between them are the edges that link them.
-These functional relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their gene products establish a protein-protein interaction.
-Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent biomolecules \cite{barabasi_network_2011}.
+As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and between them are the edges that link them.
+These functional relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction.
+Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}.
 In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011} \cite{pinero_uncovering_2016}.
@@
 We focus on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study of familial breast cancer conducted in the French population.
-After following a classical GWAS approach, we use these network-based methods to recover additional familial breast cancer biomarkers.
-Some of them are known, while others are specific to this dataset.
+After a classical GWAS approach, we use these network-based methods to recover additional breast cancer biomarkers.
 Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus network of predisposition to familial breast cancer.
@@
 \section{Methods}
-\label{sec:org9e7ca64}
 \subsection{GENESIS}
-\label{sec:org920eb00}
 
 The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
-Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for BRCA1 and BRCA2 pathogenic variants.
+Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
 Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of the corresponding case (\textpm{} 3 years).
@@
 \subsection{Preprocessing and quality control}
-\label{sec:orgd26c48b}
 
@@
 A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
-In addition, we removed the samples with more than 10\% missing genotypes, and an additional 28 samples with TODO.
+In addition, we removed the samples with more than 10\% missing genotypes.
+After control for relatedness, 17 additional samples were removed (6 for sample identity error, 6 false ``friend/control'' having family link with other samples, 3 ``friend/control'' having a high relatedness score).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with CLIS, 1 with \emph{BRCA1/2} mutation detected in the family, 1 with molecular diagnosis not received).
 The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
@@
 We looked for population structure that could create confounding associations.
-A PCA revealed no differential population structure between cases and controls (Supplementary Figure \ref{sfig:pcs}).
+A PCA revealed no visual differential population structure between cases and controls (Supplementary Figure \ref{sfig:pcs}).
 Independently, we did not find evidence of genomic inflation (\(\lambda\) = 1.05) either, further confirming the absence of confounding population structure.
@@
 \subsection{High-score subnetwork search algorithms}
-\label{sec:org6ced6c9}
 \subsubsection{SNP and gene association}
 \label{methods:node_score}
-To measure association between a genotype and the phenotype, we performed a per-SNP 1df \(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
+To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
 Then, we used VEGAS2v2 to compute the gene-level association score \cite{mishra_vegas2:_2015} from the SNP P-values.
@@
 We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
-Out of those, we focused exclusively on the 32\,767 that could be mapped to an HGNC symbol.
-Out of the SNPs 197\,083 remaining after quality control, 164\,037 mapped to at least one of these genes.
+Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
+Out of the SNPs 197\,083 remaining after quality control, 164\,037 were mapped to at least one of these genes.
 
-We use such mapping to compare the outputs of methods who produce SNP- to those that produce gene-lists, and vice versa.
+We use such mapping to compare the outputs of methods that produce SNP- to those that produce gene-lists, and vice versa.
 In the former case, we consider any gene that can be mapped to any of the selected SNPs as selected as well.
@@
 The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
-The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected through an edge, and 0 otherwise; the diagonal is zero.
+The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; the diagonal is zero.
 The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
@@
 \begin{table}[htbp]
+  \caption{\label{tab:method_comparison} Summary of the differences between the studied algorithms.}
+  \centering
   \begin{threeparttable}
-    \caption{\label{tab:method_comparison}
-      Summary of the differences between the studied algorithms.}
-    \centering
-    \begin{tabular}{l|lllllll}
+    \begin{tabular}{l|llllllll}
       Method & Field & Nodes & Exhaustive & Solution & Components & Input & Scoring & Reference\\
@@
       heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
-      HotNet2 & Omics & Genes & Yes & Modular & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
+      HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
       LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
-      SConES & GWAS & SNPs & Yes & Modular & \(\ge\) 1 & Genotypes & \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
-      SigMod & GWAS & Genes & Yes & Modular & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
+      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
+      SigMod & GWAS & Genes & Yes & Module & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
     \end{tabular}
     \begin{tablenotes}
-      /Field/: field in which the algorithm was developed.
-/Nodes/, the type of network, either gene (protein-protein interaction network usually) or a SNP network.
-/Exhaustive/: whether all the possible solutions given the selected hyperparameters are explored.
-/Solution/: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.
-/Components/: number of connected subnetworks in the solution.
-/Input/: genotype data or GWAS summary statistics.
-/Scoring/: how SNP/gene P-values are transformed into node scores.
+      \textbf{\emph{Field}}: field in which the algorithm was developed.\\
+      \textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network.\\
+      \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.\\
+      \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.\\
+      \textbf{\emph{Components}}: number of connected subnetworks in the solution.\\
+      \textbf{\emph{Input}}: genotype data or GWAS summary statistics.\\
+      \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.\\
+      \textbf{\emph{Reference}}: original publication featuring the algorithm.
     \end{tablenotes}
@@
 In consequence, multiple solutions have been proposed.
-In this article, we examine six of them: five that explore the protein-protein interaction network, and one which explores SNP networks.
+In this article, we examine six of them: five that explore the PPIN, and one which explores SNP networks.
 We selected methods that were open source, had an implementation available, and an accessible documentation.
@@
 \item Select a seed node.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph as
+\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph $S$ as
 
 \begin{equation*} 
-Z_m = \frac{\sum z_i}{\sqrt{k}}
+Z_m = \frac{\sum_{i \in S} z_i}{\sqrt{k}}
 \end{equation*}
 
-where \emph{k} is the number of genes in the subgraph, z\textsubscript{i} = \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}), and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+where \emph{k} is the number of genes in the subgraph; z\textsubscript{i} is the score of gene $i$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
 \item Identify neighboring nodes i.e. nodes at distance \(\le\) \emph{d}. We set d = 2.
@@
 
-where Z\textsubscript{m}(\(\pi\)) represent a vector containing 100000 random subsets of the same number of genes.
+where Z\textsubscript{m}(\(\pi\)) represent a vector containing 100\,000 random subsets of the same number of genes.
 
@@
 All three are positive quantities.
-heinz implements the algorithm from \cite{ljubic_algorithmic_2006}, which in practice is often fast and optimal, neither is guaranteed.
+heinz implements the algorithm from \citet{ljubic_algorithmic_2006}, which in practice is often fast and optimal, neither is guaranteed.
 We used BioNet's implementation of heinz, available on Bioconductor \cite{beisser_bionet:_2010,heinz}.
@@
 where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-We scored the nodes as in \cite{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
+We scored the nodes as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
 In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
@@
 We used the version on SConES implemented in the R package martini \cite{martini}.
-As in \cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
+As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
 Note that the solution to the above problem can consist of several connected subnetworks which are disconnected from each other.
@@
 \item[{SigMod}] SigMod aims at identifying the most densely connected gene subnetwork that is most strongly associated to the phenotype \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the Laplacian matrix rather than the adjacency matrix (Section \ref{methods:notation}), to quantify solutions containing many edges.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the adjacency matrix rather than the Laplacian matrix (Section \ref{methods:notation}), to quantify solutions containing many edges.
 
@@
 First it is designed for gene-gene networks.
-Second, by replacing the Laplacian by the adjacency matrix, it favors subnetworks containing many edges.
-SConES, instead, penalizes connections between the selected selected and unselected nodes.
+Second, by replacing the adjacency by the Laplacian matrix, it favors subnetworks containing many edges.
+SConES, instead, penalizes connections between the selected and unselected nodes.
 Third, it returns a single connected subnetwork, which it achieves by exploring a grid of hyperparameters and processing their respective solutions.
@@
 
+\begin{description}
+\item[{Consensus}] In addition, we built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
+\end{description}
+
 \subsubsection{Gene-gene network}
@@
 Unless specified otherwise, we used only interactions coming from high-throughput experiments to avoid biasing the topology of the network by well-studied genes with more known interactions on average.
-Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of HGNC symbols.
+Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols.
 The scoring function for the nodes changed from method to method (Section \ref{methods:methods}).
@@
 Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
-In that case, out of the 179\,332 interactions in HINT, we mapped 173\,797 to a pair of HGNC symbols.
+In that case, out of the 179\,332 interactions in HINT, we mapped 173\,797 to a pair of gene symbols.
 
@@
 We focused on the GI network, as it is the network that fits better the scope of this article.
-However, at different stages of the manuscript we also used GS and GM.
+However, at different stages of this work we also used GS and GM.
 For the GM network, we used the mapping described in Section \ref{methods:node_score}.
 For the GI network, we used the PPI as described in Section \ref{methods:gene_network}.
-For all three networks the node score used is the association of the individual SNPs with the phenotype; specifically, we used the 1 d.f.
-\(\chi\)\textsuperscript{2}.
-
-\subsubsection{Consensus network}
-\label{methods:consensus}
-The different high-weight subnetwork discovery algorithms make different assumptions on the properties of the solutions, and employ different strategies to find them.
-Hence, combining the outcome of the different approaches might provide a more complete outlook on the specific alterations on the GENESIS dataset.
-We built such consensus network by retaining the nodes that were selected by at least two of the methods.
-We combined the results of 6 methods: dmGWAS, heinz, HotNet2, LEAN, SConES GI, and SigMod.
+For all three networks the node score used is the association of the individual SNPs with the phenotype; specifically, we used the 1~d.f.~\(\chi\)\textsuperscript{2}.
 
 \subsection{Evaluation of methods}
-\label{sec:orgc968928}
 \subsubsection{Classification accuracy of selected biomarkers}
@@
 On this same subset we trained our classifier exclusively on the SNPs selected by a particular method.
-When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes were used.
+When the method retrieved a list of genes (all of them except SConES), all the SNPs mapped to any of those genes were used.
 Then we evaluated performance of the classifier on the remaining 20\% of the dataset.
 We repeated this procedure 5 times to estimate the average and the deviation of the different performance measures.
-The different performance measures we used were: size of the solution, size of the active set, specificity, and sensitivity.
+The performance measures we used were: size of the solution, size of the active set, specificity, and sensitivity.
 The size of the active set provides an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
@@
 To obtain a baseline, we also performed the procedure using all the SNPs.
-Lastly, another desirable property is that the method retrieves a good candidate causal subnetwork.
-In consequence, we compared the outcome of each of the methods to the consensus subnetwork of all the solutions (Section \ref{methods:consensus}).
 
@@
 \label{methods:bcac}
-An alternative way to validate the results is comparing our results to an external dataset.
+An alternative way to evaluate the results is comparing our results to an external dataset.
 For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}.
@@
 BCAC has conducted one of the largest efforts in GWAS, involving over 120\,000 women of European ancestry.
-As opposed to GENESIS, samples were not selected based on family history, and hence is enriched in sporadic breast cancers.
+As opposed to GENESIS, samples were not selected based on family history, and hence the study is enriched in sporadic breast cancers.
 Another difference is that BCAC is a relatively heterogeneous study on a pan-European sample, while GENESIS is a homogeneous dataset focused on the French population.
 Despite these differences, there should be shared genetic architecture.
-On top of that, that overlap should become larger when the results are aggregated at the gene level.
+On top of that, that overlap should become larger at the gene level.
 For that purpose, we computed the gene association as in Section \ref{methods:node_score}.
-iCOGS array was used for genotyping in BCAC \cite{sakoda_turning_2013}, the same array as for GENESIS \cite{sinilnikova_genesis:_2016}.
-Although imputed data is available, we used exclusively the SNPs available on GENESIS after quality control to make the results comparable.
+The iCOGS array was used for genotyping in BCAC \cite{sakoda_turning_2013}, the same array as for GENESIS \cite{sinilnikova_genesis:_2016}.
+Although imputed data are available, we used exclusively the SNPs available on GENESIS after quality control to make the results comparable.
 
 \subsection{Code availability}
-\label{sec:org1d907c0}
 
@@
 \section{Results}
-\label{sec:orgc718f04}
-\subsection{A conventional GWAS shows that FGFR2 is strongly associated with familial breast cancer}
+
+\subsection{A conventional GWAS confirms that FGFR2 is strongly associated with familial breast cancer}
 \label{results:conventional}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/figure_1.png}
-\caption{\label{fig:snp_gene_manhattan}
-Association in GENESIS.
-The red line represents the Bonferroni threshold.
-\textbf{(A)} SNP association, measured from the outcome of a 1df \(\chi\)\textsuperscript{2} allelic test.
-Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
-The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
-\textbf{(B)} Gene association, measured by P-value of VEGAS2v2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P-values.
-The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.}
-\end{figure}
 
 We conducted association analyses in the GENESIS dataset at both the SNP and the gene levels (Section \ref{methods:node_score}).
-Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Figure \ref{fig:snp_gene_manhattan}A).
+Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure \ref{sfig:snp_gene_manhattan}A).
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
 Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other cohorts negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
-Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Figure \ref{fig:snp_gene_manhattan}B).
-
-These results show the overlap in the genetic architecture of the disease between the studied French population sample and other populations, especially at the gene level.
+Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Supplementary Figure \ref{sfig:snp_gene_manhattan}B).
+These results show an overlap in the genetic architecture of the disease between the studied French population sample and other populations, especially at the gene level.
 In addition, there are other SNPs whose P-values, although higher than the conventional threshold of significance, show a strong association with familial breast cancer.
 The most prominent of such regions are 3p24 and 8q24, both of which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-This motivates exploring network methods, which trade statistical significance for biological relevance.
+We applied an L1-penalized logistic regression (Section \ref{methods:comparison}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+Unconstrained by considerations about statistical significance, it selected 100 SNPs, from all the aforementioned regions and others (Supplementary Figure \ref{sfig:snp_gene_manhattan}C).
+Yet, it is unclear why those SNPs were selected, as emphasized by the low association of some of them, which further complicates the biological interpretation.
+In this regard, and in opposition to what would be predicted by the omnigenic model, the genes to which this SNPs map to (Section \ref{required}) are not interconnected in the HINT protein-protein interaction network (Section \ref{required}).
+This motivates exploring network methods which, by trading statistical significance for biological relevance, search susceptibility subnetworks.
+In fact, such methods provided comparable classification accuracies to L1-penalized logistic regression (Figure \ref{fig:benchmark}D), while providing more interpretable solutions.
 
@@
 \begin{table}[htbp]
+  \centering
   \begin{threeparttable}
@@
 \begin{tablenotes}
-\emph{\# genes}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
-\mean{Betweenness}: mean betweenness of the selected genes in the PPIN.
-\median{P}\textsubscript{gene}: median P-value of the selected genes.
-\(\rho\)\textsubscript{consensus}: Pearson's correlation between the subnetwork and the consensus network.
+  \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.\\
+  \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.\\
+  \textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.\\
+  \textbf{\(\rho\)\textsubscript{consensus}}: Pearson's correlation between the subnetwork and the consensus network.
 \end{tablenotes}
@@
 
-We applied six network methods to the GENESIS dataset (Section \ref{methods:methods}), obtaining six solutions (Supplementary Figure \ref{sfig:overview_solutions}, Supplementary Files 1 and 2): one for each of the five gene-based methods (Section \ref{methods:gene_network}), and one for SConES GI (Section \ref{methods:snp_network}).
-The solutions are very heterogeneous (Table \ref{tab:gene_solutions} and Supplementary table \ref{tab:snp_solutions}): none of the subnetworks examined by LEAN is significant (adjusted P-value < 0.05), while HotNet2 produced the largest solution subnetwork with 440 genes.
-SConES GI failed to recover genes in the PPIN, but it recovered one genomic region mapped to RNA gene RNU6-420P.
-All solution subnetworks except LEAN's are, on average, more strongly associated to breast cancer than the whole PPIN (median P-values $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{sfig:manhattan_solutions}).
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.9\linewidth]{./figures/figure_1.png}
+  \caption{\label{fig:solution_overview} TODO.}
+\end{figure}
+
+We applied six network methods to the GENESIS dataset (Section \ref{methods:methods}).
+As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained five solutions (Supplementary Figure \ref{fig:solution_overview}, Supplementary Files 1 and 2): one for each of the remaining four gene-based methods (Section \ref{methods:gene_network}), and one for SConES GI (Section \ref{methods:snp_network}).
+The solutions are very heterogeneous (Table \ref{tab:gene_solutions} and Supplementary table \ref{tab:snp_solutions}), making hard to draw joint conclusions.
+HotNet2 produced the largest solution subnetwork with 440 genes.
+SConES GI failed to recover genes in the PPIN, but it recovered one genomic region mapped to RNA gene \emph{RNU6-420P}.
+All solution subnetworks except LEAN's are, on average, more strongly associated to breast cancer than the whole PPIN (median P-values $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
@@
 The solution subnetworks present other desirable properties.
-First, four of the methods succeeded at recovering genes involved in the disease (Supplementary Figure \ref{sfig:disgenet_comparison}), as their subnetworks were enriched in breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
-We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015} (Supplementary Figure \ref{sfig:bcac_comparison}).
+First, four of the methods succeeded at recovering genes involved in the disease (Supplementary Figure \ref{fig:solution_overview}D), as their subnetworks were enriched in breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
+We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015} (Supplementary Figure \ref{fig:solution_overview}D).
 Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC.
@@
 
-\subsection{heinz retrieves a small, highly informative set of biomarkers in a fast and stable fashion}
-\label{sec:orgb834a56}
+\subsection{A case study: the consensus network}
+\label{results:consensus}
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.5\linewidth]{./figures/figure_3.pdf}
+  \caption{\label{fig:consensus}
+    Consensus subnetwork on GENESIS (Section \ref{methods:methods}).
+Each node is represented by a pie chart, which shows the methods that selected it.
+The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).}
+\end{figure}
+
+The heterogeneity in the solutions suggested, and the shared encouraging properties suggested that each method was capturing different susceptibility mechanisms.
+Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure \ref{sfig:consensus_stats}A).
+Encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure \ref{sfig:consensus_stats}B).
+To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section \ref{methods:methods}).
+This subnetwork (Figure \ref{fig:consensus}) contains 93 genes and shares the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes (Fisher's exact test P-value = 7.8 \texttimes{} 10\textsuperscript{-5}), higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value = 4.29 \texttimes{} 10\textsuperscript{-18}) and TODO pathways.
+Topologically, the consensus network can be divided into a large connected component composed of 49 genes, and multiple smaller subnetworks.
+Among the latter, 14 of the 93 genes are in subnetworks of size 1 (isolated) or 2, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves (TODO TEST).
+Some of them are known breast cancer susceptibility gene like \emph{FGFR2} (Section \ref{results:conventional}) and \emph{SLC4A7} (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{-5}, \citet{ahmed_newly_2009}).
+
+TODO Yet, there are other novel functions: Globally, a GO enrichment shows the involvement of two cellular processes: unfolded protein binding, and structural constituent of cytoskeleton (adjusted P-values of 0.001, 0.001, respectively), which were already observed in different solutions (Section \ref{results:separate_networks}).
+Remarkably, many of the selected genes are related to mitochondrial translation.
+For instance, MRPS30 (VEGAS P-value = 0.001), encodes a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
+Albeit disconnected from MRPS30, the consensus network includes a 2-node subnetwork composed of two mitochondrial ribosomal protein (MRPS31 - VEGAS P-value = 7.67 \texttimes{} 10\textsuperscript{-3} - and MRPS18B - VEGAS P-value = 7.92 \texttimes{} 10\textsuperscript{-3}), which suggests an involvement of mitochondrial ribosomes in carcinogenesis \cite{required}.
+
+We also examined the topological properties of the nodes.
+The genes in the consensus network have higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value = 4.29 \texttimes{} 10\textsuperscript{-18}).
+Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value = 0.57).
+Centrality, however, is weakly anti-correlated with association to the disease (Pearson correlation coefficient = -0.26, Supplementary Figure \ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
+In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
+For instance, the most central node in the consensus network is \emph{COPS5} (Supplementary Figure \ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+Despite its lack of association in GENESIS (VEGAS P-value = 0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value = 0.006).
+
+% The consensus subnetwork is not completely connected: out of the 93 genes, the largest connected subnetwork includes only 49.
+A GO enrichment analysis showed that this component is related to three major cellular processes: unfolded protein binding, structural constituent of cytoskeleton, and poly(U) RNA binding (adjusted P-values of 0.01, 0.04, and 0.04, respectively).
+We found support in the literature of the involvement of each of these functions in the development of cancer, as discussed next.
+The consensus network also contains a protein directly involved in caspase-mediated apoptosis, \emph{CASP8} (VEGAS P-value = 1.95 \texttimes{} 10\textsuperscript{-4}).
+This is related to the enriched activity, \emph{unfolded protein binding}, which inhibits caspase-dependent apoptosis, raising the chances of developing cancer \cite{calderwood_heat_2016}.
+It involves three Hsp70 chaperones of the consensus subnetwork: HSPA1A, HSPA1B, and HSPA1L.
+The genes encoding these proteins are all near each other at 6p21.
+In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
+\emph{HSPA1A} was the most strongly associated one (VEGAS P-value = 8.37 \texttimes{} 10\textsuperscript{-4}).
+
+\subsection{Methods are comparably stable, and produce similarly good predictors}
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/figure_4.pdf}
+\includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
 \caption{\label{fig:benchmark}
@@
 
-As the employed methods produced such different results, we compared their solutions in a 5-fold subsampling setting (Section \ref{methods:comparison}).
-Specifically, we measured the following properties (Figure \ref{fig:benchmark}): (i) size of the solution subnetwork; (ii) sensitivity and specificity of an L1-penalized logistic regression on the selected SNPs; (iii) stability; and (iv) computational runtime.
+We compared the six methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+Specifically, we measured five properties (Figure \ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression on the selected SNPs; stability; and computational runtime.
 
 Both solution size and active set of SNPs selected by Lasso varies greatly between the different methods (Figure \ref{fig:benchmark}A).
-heinz has the smallest solutions, with an average of 182 selected SNPs, out of which 5.6\% (10.2) are selected by Lasso.
-The largest solutions come from SConES GI (6256.6 SNPs), and dmGWAS (4255.0 SNPs).
+Heinz produced the smallest solutions, with an average of 182 selected SNPs.
+The largest solutions come from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
 Interestingly, heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%), although it is  high for all the methods (> 86\%).
@@
 
-Another desirable quality of an algorithm is stability (Section \ref{methods:comparison}).
+Another desirable quality of an algorithm is stability of the solution with regards to different inputs (Section \ref{methods:comparison}).
 Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure \ref{fig:benchmark}C).
@@
 
-In terms of computational runtime, the fastest method was heinz (Figure \ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds.
-The slowest method was dmGWAS (1 day and 17 hours on average) followed by SConES GI (1 day and 4.32 hours on average).
-Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
-Considering that, it took 3 days and 2.4 hours on average for dmGWAS' to produce results.
+In terms of computational runtime, the fastest method was heinz (Figure \ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds, and HotNet2 was slowest (3 days and 14 hours on average).
+Including the time required to compute the gene scores slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
+Considering that, it took 5 days on average for HotNet2 to produce results.
 
-\subsection{No solution is perfect}
+\subsection{The consensus overcomes the problems of the individual solutions}
 \label{sec:orgc8f05c9}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/figure_2.pdf}
+\includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
 \caption{\label{fig:issues}
@@
 Genes with a P-value < 0.1 are highlighted in red.
-\textbf{(B)} Centrality degree and -log\textsubscript{10} of the VEGAS P-value for the nodes in SigMod solution subnetwork.
+\textbf{(B)} Centrality degree and -log\textsubscript{10} of the VEGAS2v2 P-value for the nodes in SigMod solution subnetwork.
 \textbf{(C)} Genomic regions where either SConES GS, GM or GI select SNPs.}
@@
 (Figure \ref{fig:benchmark}B).
-Its conservativeness stems from its preprocessing step, which models the gene p-values as a mixture model of a beta and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Figure \ref{fig:snp_gene_manhattan}B), only 36 of them are retain a positive score after applying the BUM model (Section \ref{methods:methods}).
+Its conservativeness stems from its preprocessing step, which models the gene P-values as a mixture model of a beta and a uniform distribution, controlled by an FDR parameter.
+Due to the limited signal at the gene level in this dataset (Figure \ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after applying the BUM model (Section \ref{methods:methods}).
 Hence, heinz's solution subnetwork consists only of 4 genes, which does not provide much insight of the biology of cancer.
@@
 dmGWAS' subnetwork is the least associated subnetwork on average.
-This is due to the greedy framework it uses, which considers all nodes at distance 2 of the examined, and accepts weakly associated genes if they are linked to another, strongly associated one.
+This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
+This framework considers all nodes at distance 2 of the examined, and accepts weakly associated genes if they are linked to another, strongly associated one.
 This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure \ref{fig:issues}A).
@@
 On top of that only SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
-We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN due to the same cause: the absence of joint association of a module.
+We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a module.
 Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
@@
 
-\subsection{Aggregating solutions provides insights into the biology of cancer}
-\label{results:consensus}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/figure_3.pdf}
-\caption{\label{fig:consensus}
-  Consensus subnetwork on GENESIS (Section \ref{methods:consensus}).
-Each node is represented by a pie chart, which accounts the methods that selected it.
-The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).}
-\end{figure}
+\subsection{Network methods boost biomarker discovery}
 
-To leverage on the strengths of each of the methods and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section \ref{methods:consensus}).
-The consensus subnetwork (Figure \ref{fig:consensus}) contains 93 genes and is enriched in breast cancer susceptibility genes (Fisher's exact test P-value = 7.8 \texttimes{} 10\textsuperscript{-5}).
-Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure \ref{sfig:consensus_stats}A).
-Encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure \ref{sfig:consensus_stats}B).
-Globally, a GO enrichment shows the involvement of two cellular processes: unfolded protein binding, and structural constituent of cytoskeleton (adjusted P-values of 0.001, 0.001, respectively), which were already observed in different solutions (Section \ref{results:separate_networks}).
-Remarkably, many of the selected genes are related to mitochondrial translation.
-For instance, MRPS30 (VEGAS P-value = 0.001), encodes a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
-Albeit disconnected from MRPS30, the consensus network includes a 2-node subnetwork composed of two mitochondrial ribosomal protein (MRPS31 - VEGAS P-value = 7.67 \texttimes{} 10\textsuperscript{-3} - and MRPS18B - VEGAS P-value = 7.92 \texttimes{} 10\textsuperscript{-3}), which suggests an involvement of mitochondrial ribosomes in carcinogenesis \cite{required}.
-
-We also examined the topological properties of the nodes.
-The genes in the consensus network have higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value = 4.29 \texttimes{} 10\textsuperscript{-18}).
-Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value = 0.57).
-Centrality, however, is weakly anti-correlated with association to the disease (Pearson correlation coefficient = -0.26, Supplementary Figure \ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
-In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is \emph{COPS5} (Supplementary Figure \ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS (VEGAS P-value = 0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value = 0.006).
-
-The consensus subnetwork is not completely connected: out of the 93 genes, the largest connected subnetwork includes only 49.
-A GO enrichment analysis showed that this component is related to three major cellular processes: unfolded protein binding, structural constituent of cytoskeleton, and poly(U) RNA binding (adjusted P-values of 0.01, 0.04, and 0.04, respectively).
-We found support in the literature of the involvement of each of these functions in the development of cancer, as discussed next.
-The consensus network also contains a protein directly involved in caspase-mediated apoptosis, \emph{CASP8} (VEGAS P-value = 1.95 \texttimes{} 10\textsuperscript{-4}).
-This is related to the enriched activity, \emph{unfolded protein binding}, which inhibits caspase-dependent apoptosis, raising the chances of developing cancer \cite{calderwood_heat_2016}.
-It involves three Hsp70 chaperones of the consensus subnetwork: HSPA1A, HSPA1B, and HSPA1L.
-They genes encoding these proteins are all near each other at 6p21.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
-HSPA1A was the most strongly associated one (VEGAS P-value = 8.37 \texttimes{} 10\textsuperscript{-4}).
-Remarkably, 14 of the 93 genes are in subnetworks of size 1 (isolated) or 2, as they do not have a consistently altered neighborhood.
-One of them is the well-known breast cancer susceptibility gene \emph{FGFR2} (Section \ref{results:conventional}).
-Another one is the also well-known \emph{SLC4A7} gene (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{5}), which encodes a sodium bicarbonate cotransporter.
-The genomic region containing both \emph{SLC4A7} and nearby gene \emph{NEK10} (VEGAS P-value = 1.56 \texttimes{} 10\textsuperscript{-5}) have been consistently associated with breast cancer susceptibility \cite{ahmed_newly_2009}.
-\emph{NEK10} is a gene that might be involved in cell-cycle control, but it is absent from the PPIN and hence it could not be studied by gene methods.
-Despite that, the fact that both dmGWAS, HotNet2 and SigMod link \emph{SLC4A7} in their different subnetwork supports the notion that this gene is the responsible for breast cancer susceptibility.
+We compared the results of different network methods to the European cohort of BCAC, the largest GWAS to date in breast cancer (Section \ref{methods:bcac}).
+This comparison is pertinent, since despite caveats we expect a significant shared genetic factors with GENESIS, and since it is a much larger study (XX times more samples, 13 million SNPs through imputation).
+We conducted a study at both the SNP and the gene levels, equivalent to the one shown in Section \ref{results:conventional} for GENESIS.
+Despite the low overlap using a conventional approach (TODO), the solutions provided by the different network approaches overlap significantly (Fisher's exact test P-value $\leq X \times 10^{-X}$).
+This illustrates the ability of network guided GWAS to boost discovery.
 
 \section{Discussion}
-\label{sec:orgbfa6dab}
 
-In this article we evaluate the viability of systems biology approaches to GWAS, and examine a GWAS dataset on familial breast cancer focused on BRCA1/2 negative French women.
-Systems biology addresses two of the largest GWAS issues: interpretability and an overly conservative statistical framework that hinders discovery.
+In this article we evaluate the viability of a systems biology take on genetic studies by examining a GWAS dataset on familial breast cancer.
+Such an approach addresses two of the largest issues with GWAS: interpretability and an overly conservative statistical framework that hinders discovery.
 This is achieved by considering the biological context of each of the genes and SNPs.
@@
 A crucial step for the gene based methods is the computation of the gene score.
-In the manuscript we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), other algorithms might have more statistical power.
+In this work we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms might have more statistical power.
 
-The network methods we studied differ in what the optimal solution subnetwork looks like.
-On the one hand, SConES and heinz prefer small highly associated solutions.
-On the other hand, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While the former provide a reduced set of biomarkers, the latter deepen our understanding of the disease and provide biological hypotheses.
-They are not exempt of limitations.
-dmGWAS and SigMod's solution's size require further analyses, which risk oversimplifying their richness.
-Also, incautious practitioners might be misled by some genes, which are very central in the solution subnetworks, while being weakly associated.
-Nonetheless, they are pushed into the solution by their privileged topological properties.
-On the other end, conservative solutions, like SConES GI and heinz might not shed much light on the etiology of the disease.
+The network methods we studied differ in what the optimal solution subnetwork looks like, which acted as a double-edged sword.
+In one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
+On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
+While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which risk oversimplifying their richness.
+HotNet2 balances both approaches at the expense of producing a constellation of many, highly associated, small subnetworks.
+Despite their differences, all the solutions produced comparable discrimination capabilities on a linear classifier trained on them.
+On the negative side, the methods were also remarkably unstable, and hence likely to produce very different solutions in face of slightly different inputs.
 
@@
 The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
-However, thanks to its small size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.
+However, thanks to its smaller size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.
 
@@
 The importance of these genes, like \emph{CASC16}, is highlighted in Section \ref{results:conventional}.
-Among the excluded protein-coding genes we find genes like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}) or \emph{POU5F1B}, both linked to breast cancer susceptibility \cite{ahmed_newly_2009}.
+Among the excluded protein-coding genes we find genes like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, both linked to breast cancer susceptibility \cite{ahmed_newly_2009}.
 However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value = 2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
@@
 This responds to concerns of some authors about biases introduced by adding interactions coming from targeted studies in the literature \cite{cai_broker_2010,das_hint:_2012} where a ``rich getting richer'' phenomenon is observed: popular genes have a higher proportion of their interactions described.
-On the other hand, one study found that the best predictor of the performance of a network for disease gene discovery is the size of the network \cite{huang_systematic_2018}.
+On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network.
 This also supports using the largest amount of interactions.
@@
 
-\section{Funding and acknowledgments}
-\label{sec:org7847920}
+\section*{Funding and acknowledgments}
 
@@
 
-GENESIS (GENE SISters) is a French national study sponsored by UNICANCER (Sinilnikova et al.
+GENESIS (GENE SISters) is a French national study promoted by UNICANCER (Sinilnikova et al.
 BMC Cancer 2016).
@@
 Verny-Pierre, V.
-Sornin) and all the GENESIS collaborating cancer clinics.
+Sornin) and all the GENESIS collaborating cancer clinics clinics (Clinique Sainte Catherine, Avignon: H.
+Dreyfus; Hôpital Saint Jacques, Besançon: M-A.
+Collonge-Rame; Institut Bergonié, Bordeaux: M.Longy, A.
+Floquet, E.
+Barouk-Simonet; CHU, Brest: S.
+Audebert; Centre François Baclesse, Caen: P.
+Berthet; Hôpital Dieu, Chambéry: S.
+Fert-Ferrer; Centre Jean Perrin, Clermont-Ferrand: Y-J.
+Bignon; Hôpital Pasteur, Colmar: J-M.
+Limacher; Hôpital d’Enfants CHU – Centre Georges François Leclerc, Dijon: L.
+Faivre-Olivier; CHU, Fort de France: O.
+Bera; CHU Albert Michallon, Grenoble: D.
+Leroux; Hôpital Flaubert, Le Havre: V.
+Layet; Centre Oscar Lambret, Lille: P.
+Vennin†, C.
+Adenis; Hôpital Jeanne de Flandre, Lille: S.
+Lejeune-Dumoulin, S.
+Manouvier-Hanu; CHRU Dupuytren, Limoges: L.
+Venat-Bouvet; Centre Léon Bérard, Lyon: C.
+Lasset, V.
+Bonadona; Hôpital Edouard Herriot, Lyon: S.
+Giraud; Institut Paoli-Calmettes, Marseille: F.
+Eisinger, L.
+Huiart; Centre Val d’Aurelle – Paul Lamarque, Montpellier: I.
+Coupier; CHU Arnaud de Villeneuve, Montpellier: I.
+Coupier, P.
+Pujol; Centre René Gauducheau, Nantes: C.
+Delnatte; Centre Catherine de Sienne, Nantes: A.
+Lortholary; Centre Antoine Lacassagne, Nice: M.
+Frénay, V.
+Mari; Hôpital Caremeau, Nîmes: J.
+Chiesa; Réseau Oncogénétique Poitou Charente, Niort: P.
+Gesta; Institut Curie, Paris: D.
+Stoppa-Lyonnet, M.
+Gauthier-Villars, B.
+Buecher, A.
+de Pauw, C.
+Abadie, M.
+Belotti; Hôpital Saint-Louis, Paris: O.
+Cohen-Haguenauer; Centre Viggo-Petersen, Paris: F.
+Cornélis; Hôpital Tenon, Paris: A.
+Fajac; GH Pitié Salpétrière et Hôpital Beaujon, Paris: C.
+Colas, F.
+Soubrier, P.
+Hammel, A.
+Fajac; Institut Jean Godinot, Reims: C.
+Penet, T.
+D.
+Nguyen; Polyclinique Courlancy, Reims: L.
+Demange†, C.
+Penet; Centre Eugène Marquis, Rennes: C.
+Dugast; Centre Henri Becquerel, Rouen: A.
+Chevrier, T.
+Frebourg, J.
+Tinat, I.
+Tennevet, A.
+Rossi; Hôpital René Huguenin/Institut Curie, Saint Cloud: C.
+Noguès, L.
+Demange†, E.
+Mouret-Fourme; CHU, Saint-Etienne: F.
+Prieur; Centre Paul Strauss, Strasbourg: J-P.
+Fricker, H.
+Schuster; Hôpital Civil, Strasbourg: O.
+Caron, C.
+Maugard; Institut Claudius Regaud, Toulouse: L.
+Gladieff, V.
+Feillel; Hôpital Bretonneau, Tours: I.
+Mortemousque; Centre Alexis Vautrin, Vandoeuvre-les-Nancy: E.
+Luporsi; Hôpital de Bravois, Vandoeuvre-les-Nancy: P.
+Jonveaux; Gustave Roussy, Villejuif: A.
+Chompret†, O.
+Caron).
 
-\bibliographystyle{ieeetr}
+\bibliographystyle{plainnat}
 \bibliography{bibliography}
@@
 \section*{Supplementary materials}
-\label{sec:org7e6740f}
 
@@
 \hline
-SConES GS & 5 & 0 & 9805 & 2.7 \texttimes{} 10\textsuperscript{-5} & 0.19\\
-SConES GM & 28 & 2 & 4267 & 0.067 & 0.12\\
+SConES GS & 5 & 0 & 9\,805 & 2.7 \texttimes{} 10\textsuperscript{-5} & 0.19\\
+SConES GM & 28 & 2 & 4\,267 & 0.067 & 0.12\\
 \end{tabular}
@@
 \includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
-\caption{\label{sfig:pcs}
-GENESIS shows no differential population structure between cases and controls.
+\caption{\label{sfig:pcs} GENESIS shows no differential population structure between cases and controls.
 \textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
@@
   \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_2.pdf}
-\caption{\label{sfig:overview_solutions}
-Overview of the subnetworks produced by the different network methods. \textbf{(dmGWAS, heinz, HotNet2, LEAN, and SigMod)} contain gene subnetworks; \textbf{(SConES GI)}, SNP subnetworks.}
+  \includegraphics[width=.9\linewidth]{./figures/sfigure_2.png}
+  \caption{\label{sfig:snp_gene_manhattan} Association in GENESIS.
+The red line represents the Bonferroni threshold.
+\textbf{(A)} SNP association, measured from the outcome of a 1 d.f.
+$\chi^2$ allelic test.
+Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
+The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
+\textbf{(B)} Gene association, measured by P-value of VEGAS2v2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P-values.
+The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
+\textbf{(C)} SNP association as in panel (A).
+The SNPs in black are selected by a L1-penalized logistic regression (Section \ref{methods:comparison}, $\lambda = 0.03$).}
 \end{figure}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_3.png}
-\caption{\label{sfig:manhattan_solutions}
-Manhattan plots showing the biomolecules selected by each method.
-In \textbf{(Consensus, dmGWAS, heinz, HotNet2, and SigMod)} datapoints are genes; in \textbf{(SConES GS, GM, and GI)}, SNPs.
-LEAN was excluded, as it did not select any gene.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_4.pdf}
-\caption{\label{sfig:bcac_comparison}
-Proportion of the Bonferroni significant biomolecules (in either the GENESIS or the BCAC datasets) selected by each of the methods on the GENESIS data.
-\textbf{(Consensus, dmGWAS, heinz, HotNet2, and SigMod)} involve significant genes, only among those present in the protein-protein interaction network.
-\textbf{(SConES GS, GM and GI)} involve significant SNPs.
-LEAN is not displayed as it did not select any gene.
-The presented network methods recover a higher proportion of significant genes than of non-significant genes in both datasets, despite their lack of significance in GENESIS.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_5.pdf}
-\caption{\label{sfig:disgenet_comparison}
-:Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene (Section \ref{methods:bcac}).
-Only genes present in the protein-protein interaction network were considered.
-LEAN is not displayed as it did not select any gene.
-The presented network methods recover a higher proportion of breast cancer susceptibility genes than of other genes, despite their lack of significance in GENESIS.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
+\includegraphics[width=.9\linewidth]{./figures/sfigure_3.pdf}
 \caption{\label{sfig:pearson_methods}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
+\includegraphics[width=.9\linewidth]{./figures/sfigure_4.pdf}
 \caption{\label{sfig:consensus_names}
-Consensus subnetwork on GENESIS (Section \ref{methods:consensus}).
-\textbf{(A)} Each node is represented by a pie chart, which accounts the methods that selected it.
+Consensus subnetwork on GENESIS (Section \ref{methods:methods}).
+\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
 The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).
-This panel is equivalent to Figure \ref{fig:consensus}.
-\textbf{(B)} The name of every gene is indicated.}
+This panel is identical to Figure \ref{fig:consensus}.
+\textbf{(B)} Manhattan plot showing the genes included in the subnetwork.
+\textbf{(C)} Every gene name is indicated.
+\textbf{(D,E)} Proportion of the Bonferroni significant genes (in GENESIS and BCAC, respectively) included in the consensus network.
+\textbf{F} Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene, as in Figure \ref{fig:solution_overview}D}.
 \end{figure}
@@
 \centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_8.pdf}
+\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_5.pdf}
 \caption{\label{sfig:consensus_stats}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_9.pdf}
+\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
 \caption{\label{sfig:biotypes_excluded}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_10.pdf}
+\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
 \caption{\label{sfig:lc_ht_comparison}
@@
 \textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119980 seconds (1 day and 9.33 hours) which took VEGAS2v2 on average to compute the gene scores from SNP summary statistics.
+For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) which took VEGAS2v2 on average to compute the gene scores from SNP summary statistics.
 In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}
 \end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.9\linewidth]{./figures/sfigure_8.png}
+  \caption{\label{sfig:scones_gsm} TODO.}
+\end{figure}
+
 \end{document}

### 2020-03-02 manuscript/manuscript.tex
removed all latex errors

@@
 \title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
-\author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, GENESIS study collaborators, Dominique Stoppa-Lyonnet$^{4,5,6}$, Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
-\affiliation{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
+\author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$,\\
+GENESIS Study collaborators, Dominique Stoppa-Lyonnet$^{4,5,6}$,\\
+Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
+\date{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
   $^{2}$INSERM, U900, F-75005 Paris, France;\\
@@
   $^{5}$INSERM, U830, F-75005 Paris, France;\\
-  $^{6}$Université Paris Descartes;\\
+  $^{6}$Université Paris Descartes.\\
 }
 \hypersetup{
-  pdfauthor={Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, GENESIS investigators, Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott},
+  pdfauthor={Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, GENESIS Study collaborators, Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott},
   pdftitle={Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer},
@@
 \onecolumn
-\MakeTitle
+\maketitle
 
-\begin{abstract}
+\Large\textbf{Abstract}
+
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
@@
 By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-\end{abstract}
 
@@
     \begin{tablenotes}
-      \textbf{\emph{Field}}: field in which the algorithm was developed.\\
-      \textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network.\\
-      \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.\\
-      \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.\\
-      \textbf{\emph{Components}}: number of connected subnetworks in the solution.\\
-      \textbf{\emph{Input}}: genotype data or GWAS summary statistics.\\
-      \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.\\
-      \textbf{\emph{Reference}}: original publication featuring the algorithm.
+      \item \textbf{\emph{Field}}: field in which the algorithm was developed.\\
+      \item \textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network.\\
+      \item \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.\\
+      \item \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.\\
+      \item \textbf{\emph{Components}}: number of connected subnetworks in the solution.\\
+      \item \textbf{\emph{Input}}: genotype data or GWAS summary statistics.\\
+      \item \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.\\
+      \item \textbf{\emph{Reference}}: original publication featuring the algorithm.
     \end{tablenotes}
@@
 \begin{tablenotes}
-  \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.\\
-  \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.\\
-  \textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.\\
-  \textbf{\(\rho\)\textsubscript{consensus}}: Pearson's correlation between the subnetwork and the consensus network.
+  \item \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
+  \item \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
+  \item \textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.
+  \item \textbf{\(\rho\)\textsubscript{consensus}}: Pearson's correlation between the subnetwork and the consensus network.
 \end{tablenotes}
@@
 \begin{tablenotes}
-\mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network; \median{P}\textsubscript{SNP}: median P-value of the selected SNPs.
+\item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.
+\item \median{P}\textsubscript{SNP}: median P-value of the selected SNPs.
 \end{tablenotes}
@@
 \begin{tablenotes}
-\mean{Betweenness}: mean betweenness of the selected genes in the full network; \median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson's correlation with the consensus network.
+\item \mean{Betweenness}: mean betweenness of the selected genes in the full network.
+\item \median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson's correlation with the consensus network.
 \end{tablenotes}

### 2020-03-04 manuscript/manuscript.tex
included more of Fabienne's and Nadine's comments

@@
-\documentclass[twocolumn, 10pt]{article}
+\documentclass[twocolumn, 11pt, draft]{article}
 \usepackage{amsmath,amssymb}
@@
 
-\makeatletter
-\def\affiliation#1{\gdef\@affiliation{#1}}
-\def\abstract#1{\gdef\@abstract{#1}}
-\def\graphabst#1{\gdef\@graphabst{#1}}
-\def\keywords#1{\gdef\@keywords{#1}}
-\def\corresp#1{\gdef\@corresp{#1}}
-\def\bioauthor#1#2{\centerline{\textbf{#1}}\par #2}
-\newcommand{\MakeTitle}{
-  \newpage
-  \begin{center}
-    \Large \@title\par
-    \large \@author
-  \end{center}
-  \@affiliation\par
-  \@abstract\par
-  \vskip 1em
-}
-\makeatother
-
 \title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
-\author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$,\\
-GENESIS Study collaborators, Dominique Stoppa-Lyonnet$^{4,5,6}$,\\
+\author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
+GENESIS Study collaborators, Dominique Stoppa-Lyonnet$^{4,5,6}$, \\
 Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
@@
 
-\Large\textbf{Abstract}
-
+\begin{abstract}
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
 It does so by jointly considering the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks highly associated to this phenotype.
-We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
+In this work, we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype.
+We interrogate a familial breast cancer genome-wide association study focused on \emph{BRCA1/2} negative French women.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value = $7.8 \times 10^{-5}$ ) and more central than average.
+Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value = $7.8 \times 10^{-5}$ ) and occupy more central positions in the network than average.
 Additionally, it includes subnetworks of mechanisms related to cancer, like protein folding (\emph{HSPA1A}, \emph{HSPA1B}, and \emph{HSPA1L}) or mitochondrial ribosomes (\emph{MRPS30}, \emph{MRPS31}, \emph{MRPS18B}).
@@
 By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+\end{abstract}
 
@@
 On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the ``core'' genes directly involved in a disease's mechanism.
+The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in a disease's mechanism.
 Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
@@
 This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes.
-Additionally, as genes that contribute to a disease tend to participate in similar biological functions, guilt-by-association strategies have proved effective at identifying disease genes \cite{huang_systematic_2018}.
+Additionally, as genes that contribute to a disease tend to participate in similar biological functions, studying the neighborhood of known disease genes is effective at identifying new ones \cite{huang_systematic_2018}.
 
-Network-based, biomarker discovery methods exploit the guilt-by-association strategy to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
-In essence, each SNP has a measure of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
-Then, the problem becomes finding a functionally-related set of genes that is highly associated with the disease.
+Network-based, biomarker discovery methods exploit the aforementioned strategy to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
+In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
+Then, the problem becomes finding a functionally-related set of highly-scoring genes.
 Different solutions have been proposed to this problem, often stemming from divergent different mathematical frameworks and considerations of what the optimal solution looks like.
@@
 Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on interconnected genes with high association with the disease.
-However, they differ in their tolerance to the inclusion of lowly associated nodes, and the possible number of disconnected subnetworks in the solution.
+However, they differ in their tolerance to the inclusion of low scoring nodes, and the possible number of disconnected subnetworks in the solution.
 Lastly, other methods also consider the topology of the network, favoring solutions that are densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
 
-In this work, we analyze the effectiveness of these six methods to discover new biomarkers on GWAS data.
+In this work, we analyze the effectiveness of these six network methods for biomarker discovery on GWAS data.
+While all of them capture susceptibility mechanisms resembling that postulated by the omnigenic model, they do so in different ways, and provide a representative view of the field.
 We focus on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study of familial breast cancer conducted in the French population.
@@
 In order to map SNPs to genes we used their overlap on the genome: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
-To compute the gene association we used the 10\% of SNPs linked to the gene with lowest P-values.
+To compute the gene association score we used the 10\% of SNPs linked to the gene with lowest P-values.
 We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
@@
 This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired FDR.
-Thus formulated, the problem is NP-complete.
+Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitely long computational time.
 To solve it efficiently it is re-casted as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S):
@@
 \begin{description}
-\item[{SigMod}] SigMod aims at identifying the most densely connected gene subnetwork that is most strongly associated to the phenotype \cite{liu_sigmod:_2017}.
+\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
 It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the adjacency matrix rather than the Laplacian matrix (Section \ref{methods:notation}), to quantify solutions containing many edges.
@@
 Specifically, for the range of \(\lambda\) = \(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
-Such a large change implies that the network is strongly interconnected.
+Such a large change implies that the network is densely interconnected.
 This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other.
@@
 We focused on the GI network, as it is the network that fits better the scope of this article.
-However, at different stages of this work we also used GS and GM.
+However, at different stages of this work we also used GS and GM for comparison.
 For the GM network, we used the mapping described in Section \ref{methods:node_score}.
 For the GI network, we used the PPI as described in Section \ref{methods:gene_network}.
-For all three networks the node score used is the association of the individual SNPs with the phenotype; specifically, we used the 1~d.f.~\(\chi\)\textsuperscript{2}.
+For all three networks the node score used is an association score of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
 
 \subsection{Evaluation of methods}
+
+We evaluated multiple properties of the different method through a 5-fold subsampling setting.
+We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
+When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
+We used the 5 repetitions to estimate the average and the deviation of the different measures.
+
+\subsubsection{Properties ot the solution}
+\label{methods:comparison}
+
+We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input) of the different network methods.
+The stability was measured using the Pearson's correlation between different runs as suggested \citet{nogueira_measuring_2016}.
+
 \subsubsection{Classification accuracy of selected biomarkers}
-\label{methods:comparison}
-A desirable solution is one that is sparse, while offering a good predictive power on unseen samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression trained exclusively on those SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for LD to reduce the number of SNPs included in the model (size of the active set), while improving the generalization of the classifier.
-The value of the regularization parameter, which controls both the magnitude and the sparsity of the coefficients, was set by cross-validation.
-To that end, we used the different network-methods on a random subset of 80\% of the samples.
-On this same subset we trained our classifier exclusively on the SNPs selected by a particular method.
-When the method retrieved a list of genes (all of them except SConES), all the SNPs mapped to any of those genes were used.
-Then we evaluated performance of the classifier on the remaining 20\% of the dataset.
-We repeated this procedure 5 times to estimate the average and the deviation of the different performance measures.
-The performance measures we used were: size of the solution, size of the active set, specificity, and sensitivity.
-The size of the active set provides an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
+\label{methods:classifier}
 
-Additionally, for each of the methods, we evaluated their stability and their runtime.
-The stability of an algorithm is its sensitivity to small changes of the input, and is measured using the Pearson's correlation between different runs as suggested in \cite{nogueira_measuring_2016}.
-To obtain a baseline, we also performed the procedure using all the SNPs.
+A desirable solution offers good predictive power on unseen \emph{test} samples.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier trained exclusively on those SNPs to predict the outcome (case/control).
+The L1 penalty helps to account for LD by reducing the number of SNPs included in the model (size of the active set), while improving the generalization of the classifier.
+This penalty was set by cross-validation.
+We applied each network method to each \emph{train} set, and trained the classifier on the \emph{train} set as well using only on the selected SNPs.
+When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
+Then we evaluated the sensitivity and the specificity on the \emph{test set}.
+To obtain a baseline, we also trained the classifier on all the SNPs.
+We also evaluated the size of the active set, an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
 
-\subsubsection{Biological relevance of the genes}
+\subsubsection{Comparison to state-of-the-art}
 \label{methods:bcac}
@@
 
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015}.
-BCAC has conducted one of the largest efforts in GWAS, involving over 120\,000 women of European ancestry.
-As opposed to GENESIS, samples were not selected based on family history, and hence the study is enriched in sporadic breast cancers.
-Another difference is that BCAC is a relatively heterogeneous study on a pan-European sample, while GENESIS is a homogeneous dataset focused on the French population.
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 228\,951 women of European ancestry from the general population \cite{Michailidou2017}.
+Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic, while GENESIS is a homogeneous dataset focusing on the French high-risk population attending the family cancer clinics.
 Despite these differences, there should be shared genetic architecture.
 On top of that, that overlap should become larger at the gene level.
-For that purpose, we computed the gene association as in Section \ref{methods:node_score}.
-The iCOGS array was used for genotyping in BCAC \cite{sakoda_turning_2013}, the same array as for GENESIS \cite{sinilnikova_genesis:_2016}.
-Although imputed data are available, we used exclusively the SNPs available on GENESIS after quality control to make the results comparable.
+For that purpose, we searched for associated genes as in Section \ref{methods:node_score}.
+After imputation, it consisted of 13\,250\,642 SNPs.
+We used all of them to measure gene-level association, but only the SNPs available on GENESIS after quality control for the SNP-level comparisons.
 
@@
 
-\subsection{A conventional GWAS confirms that FGFR2 is strongly associated with familial breast cancer}
+\subsection{A conventional GWAS confirms that FGFR2 is associated with familial breast cancer}
 \label{results:conventional}
@@
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
-Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other cohorts negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
+Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1/2} carriers studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
 Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Supplementary Figure \ref{sfig:snp_gene_manhattan}B).
 These results show an overlap in the genetic architecture of the disease between the studied French population sample and other populations, especially at the gene level.
-In addition, there are other SNPs whose P-values, although higher than the conventional threshold of significance, show a strong association with familial breast cancer.
-The most prominent of such regions are 3p24 and 8q24, both of which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression (Section \ref{methods:comparison}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+In addition, there are compelling regions having low, albeit not genome-wide significant, P-values.
+The most prominent are 3p24 and 8q24, both of which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
+We applied an L1-penalized logistic regression (Section \ref{methods:classifier}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
 Unconstrained by considerations about statistical significance, it selected 100 SNPs, from all the aforementioned regions and others (Supplementary Figure \ref{sfig:snp_gene_manhattan}C).
-Yet, it is unclear why those SNPs were selected, as emphasized by the low association of some of them, which further complicates the biological interpretation.
-In this regard, and in opposition to what would be predicted by the omnigenic model, the genes to which this SNPs map to (Section \ref{required}) are not interconnected in the HINT protein-protein interaction network (Section \ref{required}).
+Yet, it is unclear why those SNPs were selected, as emphasized by the high P-value of some of them, which further complicates the biological interpretation.
+In this regard, and in opposition to what would be predicted by the omnigenic model, the genes to which this SNPs map to (Section \ref{methods:node_score}) are not interconnected in the PPIN (Section \ref{methods:gene_network}).
 This motivates exploring network methods which, by trading statistical significance for biological relevance, search susceptibility subnetworks.
@@
 SConES GI failed to recover genes in the PPIN, but it recovered one genomic region mapped to RNA gene \emph{RNU6-420P}.
-All solution subnetworks except LEAN's are, on average, more strongly associated to breast cancer than the whole PPIN (median P-values $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{fig:solution_overview}A).
+All solution subnetworks except LEAN's have, on average, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions: heinz strongly favored highly associated genes, while dmGWAS is less conservative (median gene P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select lowly associated genes.
+However, there are nuances between solutions: heinz strongly favored high scoring genes, while dmGWAS is less conservative (median gene P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
 
 The solution subnetworks present other desirable properties.
-First, four of the methods succeeded at recovering genes involved in the disease (Supplementary Figure \ref{fig:solution_overview}D), as their subnetworks were enriched in breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
-We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{michailidou_genome-wide_2015} (Supplementary Figure \ref{fig:solution_overview}D).
+First, four of the methods succeeded at recovering known breast cancer susceptibility genes (Supplementary Figure \ref{fig:solution_overview}D), as their subnetworks were enriched in breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
+We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (Supplementary Figure \ref{fig:solution_overview}D).
 Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC.
@@
 Topologically, the consensus network can be divided into a large connected component composed of 49 genes, and multiple smaller subnetworks.
-Among the latter, 14 of the 93 genes are in subnetworks of size 1 (isolated) or 2, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves (TODO TEST).
+Among the latter, 14 of the 93 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves (TODO TEST).
 Some of them are known breast cancer susceptibility gene like \emph{FGFR2} (Section \ref{results:conventional}) and \emph{SLC4A7} (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{-5}, \citet{ahmed_newly_2009}).
@@
 Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value = 0.57).
-Centrality, however, is weakly anti-correlated with association to the disease (Pearson correlation coefficient = -0.26, Supplementary Figure \ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
+The centrality in the PPIN of the genes in the consensus, however, is weakly anti-correlated with the P-value of association to the disease (Pearson correlation coefficient = -0.26, Supplementary Figure \ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
 In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
@@
 Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times of a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+Each method was run 5 times of a random subset of the samples, and tested on the remaining samples (Sections \ref{methods:comparison} and \ref{methods:classifier}).
 \textbf{(A)} Number of SNPs selected by each method and number of SNPs on the active set used by the Lasso classifier.
@@
 
-We compared the six methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
-Specifically, we measured five properties (Figure \ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression on the selected SNPs; stability; and computational runtime.
+We compared the six methods in a 5-fold subsampling setting (Sections \ref{methods:comparison} and \ref{methods:classifier}).
+Specifically, we measured five properties (Figure \ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
 
-Both solution size and active set of SNPs selected by Lasso varies greatly between the different methods (Figure \ref{fig:benchmark}A).
+Both solution size and the subset of it selected by the classifier (\emph{active set}) varies greatly between the different methods (Figure \ref{fig:benchmark}A).
 Heinz produced the smallest solutions, with an average of 182 selected SNPs.
@@
 Interestingly, heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%), although it is  high for all the methods (> 86\%).
-This suggests methods are selecting informative SNPs on average.
-
-To determine whether the selected SNPs could be used for patient classification we computed the sensitivity and the specificity of the classifier on the testing data (Figure \ref{fig:benchmark}B).
-All classifiers' sensitivities were in the 0.42 -- 0.51 range; the specificities, between 0.54 and 0.62.
-On average, SigMod had the highest sensitivity (0.51); dmGWAS, the highest specificity (0.52).
-Both heinz and SigMod had on average better sensitivity than the classifier trained on all the SNPs, but none had superior specificity.
-However, the differences are negligible, well within the 95\% confidence interval.
+To determine whether the selected SNPs could be used for patient classification we computed the performance of the classifier on the \emph{test dataset} (Figure \ref{fig:benchmark}B).
+The different classifiers displayed similar sensitivities and specificities, all in the 0.52 -- 0.56 range.
 
@@
 \subsection{The consensus overcomes the problems of the individual solutions}
-\label{sec:orgc8f05c9}
+\label{results:drawbacks}
 
@@
 Hence, heinz's solution subnetwork consists only of 4 genes, which does not provide much insight of the biology of cancer.
-Importantly, it ignores genes that are strongly associated to cancer in this dataset like \emph{FGFR2}.
+Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
 
@@
 We also discuss the limitations of such analyses, related to the lack of known interactions around some genes.
+
 A crucial step for the gene based methods is the computation of the gene score.
@@
 However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms might have more statistical power.
+Another important decision is how to handle LD in a GWAS.
+VEGAS2v2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2v2's computation time.
+With regards to SConES, less SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
+However, as mentioned in Section \ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
 
@@
 The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure \ref{sfig:biotypes_excluded}).
-The importance of these genes, like \emph{CASC16}, is highlighted in Section \ref{results:conventional}.
-Among the excluded protein-coding genes we find genes like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, both linked to breast cancer susceptibility \cite{ahmed_newly_2009}.
+Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section \ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
+Even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
 However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value = 2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
@@
 It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
+This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
 Lastly, all the methods rely heavily on how SNPs are mapped to genes.
@@
 In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns of some authors about biases introduced by adding interactions coming from targeted studies in the literature \cite{cai_broker_2010,das_hint:_2012} where a ``rich getting richer'' phenomenon is observed: popular genes have a higher proportion of their interactions described.
-On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network.
-This also supports using the largest amount of interactions.
-To clarify their impact on this study, we compared the impact of using only physical interactions from high-throughput experiment versus interactions from both high-throughput and the literature (Section \ref{methods:gene_network}).
-We conclude that for most of the methods a larger network did not greatly impact the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure \ref{sfig:lc_ht_comparison}).
+This responds to concerns of some authors about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
+Their presense might bias the results of the presented methods, as if changes the topology of the networks by adding nodes closer on average to other nodes than average.
+On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
+When we compared the impact of using a larger network containing interactions from both high-throughput experiment and the literature (Section \ref{methods:gene_network}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure \ref{sfig:lc_ht_comparison}).
+This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the selected solutions.
 
@@
 
+\section*{Author contributions}
+
+H.C-G.
+conducted all the final analyses and the writing of the manuscript.
+C.L.
+conducted analyses using LEAN, Sigmod, and dmGWAS, and provided help and feedback.
+F.L.
+and N.A.
+provided feedback on the analyses and the manuscript.
+C-A.A.
+supervised the project.
+
 \section*{Funding and acknowledgments}
 
-This project was supported by funding from Agence Nationale de la Recherche (ANR-18-CE45-0021-01).
-H.C-G.
-was funded by the European Union’s Horizon 2020 research and innovation program under the Marie Skłodowska-Curie [666003].
-Financial support for GENESIS resource and genotyping was provided by the Ligue Nationale contre le Cancer (grants PRE05/DSL, PRE07/DSL, PRE11/NA), the French National Institute of Cancer (INCa grant No b2008-029/LL-LC) and the comprehensive cancer center SiRIC, (Site de Recherche Intégrée sur le Cancer: Grant INCa-DGOS-4654).
+This project was supported by funding from Agence Nationale de la Recherche (ANR-18-CE45-0021-01) and from the European Union’s Horizon 2020 research and innovation program (Marie Skłodowska-Curie [666003]).
+Financial support for GENESIS resource and genotyping was provided by the Ligue Nationale contre le Cancer (grants PRE05/DSL, PRE07/DSL, PRE11/NA), the French National Institute of Cancer (INCa grant b2008-029/LL-LC) and the comprehensive cancer center SiRIC (Site de Recherche Intégrée sur le Cancer: INCa-DGOS-4654).
 
-GENESIS (GENE SISters) is a French national study promoted by UNICANCER (Sinilnikova et al.
-BMC Cancer 2016).
-We wish to thank the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie : Séverine Eon-Marchais, M.
+We wish to thank the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: O.
+Kulkarni, S.
+Eon-Marchais, M.
 Marcou, D.
@@
 Layet; Centre Oscar Lambret, Lille: P.
-Vennin†, C.
+Vennin, C.
 Adenis; Hôpital Jeanne de Flandre, Lille: S.
@@
 Nguyen; Polyclinique Courlancy, Reims: L.
-Demange†, C.
+Demange, C.
 Penet; Centre Eugène Marquis, Rennes: C.
@@
 Noguès, L.
-Demange†, E.
+Demange, E.
 Mouret-Fourme; CHU, Saint-Etienne: F.
@@
 Jonveaux; Gustave Roussy, Villejuif: A.
-Chompret†, O.
+Chompret, O.
 Caron).
 
-\bibliographystyle{plainnat}
+\bibliographystyle{abbrvnat}
 \bibliography{bibliography}
@@
 \textbf{(C)} SNP association as in panel (A).
-The SNPs in black are selected by a L1-penalized logistic regression (Section \ref{methods:comparison}, $\lambda = 0.03$).}
+The SNPs in black are selected by a L1-penalized logistic regression (Section \ref{methods:classifier}, $\lambda = 0.03$).}
 \end{figure}

### 2020-03-06 manuscript/manuscript.tex
updated manuscript

@@
-\documentclass[twocolumn, 11pt, draft]{article}
+\documentclass[twocolumn, 11pt]{article}
 \usepackage{amsmath,amssymb}
-\usepackage[format=hang,labelfont=bf,textfont=small,singlelinecheck=false,justification=raggedright,margin={20pt,20pt},figurename=Fig.]{caption}
+\usepackage[format=hang,labelfont=bf,textfont=small,singlelinecheck=false,justification=justified,margin={20pt,20pt},figurename=Fig.]{caption}
 \usepackage[utf8]{inputenc}
@@
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value = $7.8 \times 10^{-5}$ ) and occupy more central positions in the network than average.
+By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$ ) and occupy more central positions in the network than average.
 Additionally, it includes subnetworks of mechanisms related to cancer, like protein folding (\emph{HSPA1A}, \emph{HSPA1B}, and \emph{HSPA1L}) or mitochondrial ribosomes (\emph{MRPS30}, \emph{MRPS31}, \emph{MRPS18B}).
 We also observed a general dysregulation in the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
 \end{abstract}
@@
 GWAS have successfully identified thousands of variants underlying many common diseases \cite{buniello_nhgri-ebi_2019}.
-However, the experimental setting also presents intrinsic challenges.
+However, this experimental setting also presents intrinsic challenges.
 Some of them stem from the high-dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped.
@@
 On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in a disease's mechanism.
+The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
 Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
@@
 
-Network-based, biomarker discovery methods exploit the aforementioned strategy to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
+Network-based, biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
 In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
 Then, the problem becomes finding a functionally-related set of highly-scoring genes.
-Different solutions have been proposed to this problem, often stemming from divergent different mathematical frameworks and considerations of what the optimal solution looks like.
+Different solutions have been proposed to this problem, often stemming from divergent mathematical frameworks and considerations of what the optimal solution looks like.
 Some methods strongly constrain the problem to certain kinds of subnetworks.
-Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on star subnetworks, i.e.
+Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on \emph{star} subnetworks, i.e.
 instances were both a gene and its direct interactors are associated with the disease.
-Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on interconnected genes with high association with the disease.
+Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnected genes with high association scores.
 However, they differ in their tolerance to the inclusion of low scoring nodes, and the possible number of disconnected subnetworks in the solution.
-Lastly, other methods also consider the topology of the network, favoring solutions that are densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
+Lastly, other methods also consider the topology of the network, favoring solutions that are not only high scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
 
@@
 While all of them capture susceptibility mechanisms resembling that postulated by the omnigenic model, they do so in different ways, and provide a representative view of the field.
-We focus on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study of familial breast cancer conducted in the French population.
+We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
 After a classical GWAS approach, we use these network-based methods to recover additional breast cancer biomarkers.
@@
 Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
-Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of the corresponding case (\textpm{} 3 years).
+Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
 We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
@@
 
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy - Weinberg equilibrium in controls (P-value \textless 0.001), and those missing on more than 10\% of the samples.
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy - Weinberg equilibrium in controls (P-value \textless~0.001), and those missing on more than 10\% of the samples.
 A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
 In addition, we removed the samples with more than 10\% missing genotypes.
-After control for relatedness, 17 additional samples were removed (6 for sample identity error, 6 false ``friend/control'' having family link with other samples, 3 ``friend/control'' having a high relatedness score).
-Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with CLIS, 1 with \emph{BRCA1/2} mutation detected in the family, 1 with molecular diagnosis not received).
+After control for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, and 3 additional controls having a high relatedness score).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with CLIS, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis).
 The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
 
-We looked for population structure that could create confounding associations.
-A PCA revealed no visual differential population structure between cases and controls (Supplementary Figure \ref{sfig:pcs}).
-Independently, we did not find evidence of genomic inflation (\(\lambda\) = 1.05) either, further confirming the absence of confounding population structure.
+We looked for population structure that could produce spurious associations.
+A PCA revealed no visual differential population structure between cases and controls (Supplementary Figure~\ref{sfig:pcs}).
+Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
 
@@
 To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
-Then, we used VEGAS2v2 to compute the gene-level association score \cite{mishra_vegas2:_2015} from the SNP P-values.
-In order to map SNPs to genes we used their overlap on the genome: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
-To compute the gene association score we used the 10\% of SNPs linked to the gene with lowest P-values.
+Then, we used VEGAS2v2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P-values of the SNPs mapped to them.
+Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P-values.
+We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
 We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
 Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
-Out of the SNPs 197\,083 remaining after quality control, 164\,037 were mapped to at least one of these genes.
+Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
 
-We use such mapping to compare the outputs of methods that produce SNP- to those that produce gene-lists, and vice versa.
-In the former case, we consider any gene that can be mapped to any of the selected SNPs as selected as well.
-In the latter, we consider all the SNPs that can be mapped to that gene as selected by the method.
+We use such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
+For the former, we consider any gene that can be mapped to any of the selected SNPs as selected as well.
+For the latter, we consider all the SNPs that can be mapped to that gene as selected by the method.
 
@@
 Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
-When referring to a subnetwork S, $V_{S}$ is the set of nodes in S and $E_{S}$ is the set of edges in S.
+When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
 A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
 
-On top of a weight, nodes have other properties provided by the topology of the graph.
+On top of a weight, nodes have other properties, provided by the topology of the graph.
 In this article we focus on two: degree centrality, and betweenness centrality.
@@
 
+\subsubsection{Networks used}
+\label{methods:networks}
+
+\paragraph{Gene-gene network}
+Out of the six methods tested, five use a gene-gene interaction network (Section~\ref{methods:methods}).
+Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), for practical reasons we focused on a protein-protein interaction network (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it.
+We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
+Unless specified otherwise, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
+Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols.
+The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
+
+Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
+In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
+
+\paragraph{SNP networks}
+SConES \cite{azencott_efficient_2013} is the only of the studied methods designed to handle SNP networks.
+As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs.
+\citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
+We focused on the GI network, as it fits the scope of this article better, using the PPIN described above.
+However, at different stages of this work we also used GS and GM for comparison.
+For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
+In all three the node scores are the association score of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
 \subsubsection{Methods used}
@@
     \begin{tabular}{l|llllllll}
-      Method & Field & Nodes & Exhaustive & Solution & Components & Input & Scoring & Reference\\
+      Method & Field & Nodes & Exhaustive & Solution & Comp. & Input & Scoring & Ref.\\
       \hline
@@
       LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
-      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
+      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
       SigMod & GWAS & Genes & Yes & Module & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
@@
     \begin{tablenotes}
-      \item \textbf{\emph{Field}}: field in which the algorithm was developed.\\
-      \item \textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network.\\
-      \item \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.\\
-      \item \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.\\
-      \item \textbf{\emph{Components}}: number of connected subnetworks in the solution.\\
-      \item \textbf{\emph{Input}}: genotype data or GWAS summary statistics.\\
-      \item \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.\\
-      \item \textbf{\emph{Reference}}: original publication featuring the algorithm.
+      \footnotesize{
+        \item \textbf{\emph{Field}}: field in which the algorithm was developed.
+\textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network.
+\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
+\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.
+\textbf{\emph{Comp.}}: number of connected subnetworks in the solution.
+\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
+\textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.
+\textbf{\emph{Ref.}}: original publication featuring the algorithm.
+      }
     \end{tablenotes}
@@
 
-Beyond the assumption that genes that contribute to the same function will be nearby in the protein-protein interaction network (PPIN), they might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-That is not the only choice to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, what is the best network to use, etc.
+Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
+How to account for that is not the only choice a developer of network methods needs to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
 In consequence, multiple solutions have been proposed.
@@
 We selected methods that were open source, had an implementation available, and an accessible documentation.
-Their main differences are summarized in Table \ref{tab:method_comparison}.
+Their main differences are summarized in Table~\ref{tab:method_comparison}.
 
 \begin{description}
-\item[{dmGWAS}] dmGWAS searches the subgraph with the highest local density in low P-values \cite{jia_dmgwas:_2011}.
+\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P-values \cite{jia_dmgwas:_2011}.
 To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
@@
 \begin{equation*} 
-Z_m = \frac{\sum_{i \in S} z_i}{\sqrt{k}}
+Z_m = \frac{1}{\sqrt{k}} \sum_{i \in S} z_i
 \end{equation*}
 
-where \emph{k} is the number of genes in the subgraph; z\textsubscript{i} is the score of gene $i$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes i.e. nodes at distance \(\le\) \emph{d}. We set d = 2.
-\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than Z\textsubscript{m} \texttimes{} (1 + r). In our experiments, we set r = 0.1.
-\item Repeat 2-4 until no increment Z\textsubscript{m} \texttimes{} (1 + r) is possible.
+where \emph{k} is the number of genes in the subgraph; z\textsubscript{i} is the Z score of gene $i$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+\item Identify neighboring nodes of $S$ i.e. nodes at distance \(\le\) d. We set d~=~2.
+\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r). In our experiments, we set r~=~0.1.
+\item Repeat 2-4 until no further enlargement is possible.
 \end{enumerate}
@@
 
-where Z\textsubscript{m}(\(\pi\)) represent a vector containing 100\,000 random subsets of the same number of genes.
+where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
 
@@
 
-were p(v) = w(v) - w' is the \emph{profit} of adding a node, c(e) = w' is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$.
+were p(v)~=~w(v) - w' is the \emph{profit} of adding a node, c(e)~=~w' is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$.
 All three are positive quantities.
-heinz implements the algorithm from \citet{ljubic_algorithmic_2006}, which in practice is often fast and optimal, neither is guaranteed.
+Heinz implements the algorithm from \citet{ljubic_algorithmic_2006}, which although in practice is often fast and optimal, neither is guaranteed.
 We used BioNet's implementation of heinz, available on Bioconductor \cite{beisser_bionet:_2010,heinz}.
@@
 The former is captured by an insulated heat diffusion process: at the beginning, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
-This process continues until equilibrium is reached, and results in a similarity matrix F.
-F is used to compute the similarity matrix E that accounts also for similarities in node scores as
+This process continues until equilibrium is reached, and results in a similarity matrix $F$.
+$F$ is used to compute the similarity matrix $E$ that accounts also for similarities in node scores as
 
@@
 In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
-To obtain densely connected subnetworks, HotNet2 prunes E, only preserving edges such that w(E) \textgreater{} \(\delta\).
+To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
 Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
@@
 Note that the solution to the above problem can consist of several connected subnetworks which are disconnected from each other.
-In this case, the selected hyperparameters were \(\eta\) = 3.51, \(\lambda\) = 210.29 for SConES GS; \(\eta\) = 3.51, \(\lambda\) = 97.61 for SConES GM; and \(\eta\) = 3.51, \(\lambda\) = 45.31 for SConES GI.
+In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
 \end{description}
@@
 \item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the adjacency matrix rather than the Laplacian matrix (Section \ref{methods:notation}), to quantify solutions containing many edges.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the adjacency matrix rather than the Laplacian matrix (Section~\ref{methods:notation}), to quantify solutions containing many edges.
 
@@
 Third, it returns a single connected subnetwork, which it achieves by exploring a grid of hyperparameters and processing their respective solutions.
-Specifically, for the range of \(\lambda\) = \(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
+Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
 Such a large change implies that the network is densely interconnected.
@@
 
-\subsubsection{Gene-gene network}
-\label{methods:gene_network}
-Out of the six methods tested, five use a gene-gene interaction network (Section \ref{methods:methods}).
-Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), for practical reasons we focused on a PPIN, as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it.
-We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless specified otherwise, we used only interactions coming from high-throughput experiments to avoid biasing the topology of the network by well-studied genes with more known interactions on average.
-Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols.
-The scoring function for the nodes changed from method to method (Section \ref{methods:methods}).
+\subsection{Pathway enrichment analysis}
+\label{methods:pathway_enrichment}
 
-Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
-In that case, out of the 179\,332 interactions in HINT, we mapped 173\,797 to a pair of gene symbols.
-
-\subsubsection{SNP networks}
-\label{methods:snp_network}
-SConES \cite{azencott_efficient_2013} is the only of the studied methods designed to handle SNP networks.
-As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs.
-The authors suggested three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists (``GI network'').
-We focused on the GI network, as it is the network that fits better the scope of this article.
-However, at different stages of this work we also used GS and GM for comparison.
-For the GM network, we used the mapping described in Section \ref{methods:node_score}.
-For the GI network, we used the PPI as described in Section \ref{methods:gene_network}.
-For all three networks the node score used is an association score of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
+We searched for pathways enriched in the gene subnetworks produced by the aforementioned methods.
+We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
+The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
+Pathways with an adjusted P-value \textless 0.05 were regarded significant.
 
@@
 On top of that, that overlap should become larger at the gene level.
-For that purpose, we searched for associated genes as in Section \ref{methods:node_score}.
+For that purpose, we searched for associated genes as in Section~\ref{methods:node_score}.
 After imputation, it consisted of 13\,250\,642 SNPs.
@@
 
-We conducted association analyses in the GENESIS dataset at both the SNP and the gene levels (Section \ref{methods:node_score}).
-Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure \ref{sfig:snp_gene_manhattan}A).
+We conducted association analyses in the GENESIS dataset at both the SNP and the gene levels (Section~\ref{methods:node_score}).
+Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
 Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1/2} carriers studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
-Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Supplementary Figure \ref{sfig:snp_gene_manhattan}B).
+Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
 These results show an overlap in the genetic architecture of the disease between the studied French population sample and other populations, especially at the gene level.
-In addition, there are compelling regions having low, albeit not genome-wide significant, P-values.
-The most prominent are 3p24 and 8q24, both of which have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression (Section \ref{methods:classifier}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
-Unconstrained by considerations about statistical significance, it selected 100 SNPs, from all the aforementioned regions and others (Supplementary Figure \ref{sfig:snp_gene_manhattan}C).
+
+Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P-values.
+Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
+Exploring algorithms unconstrained by considerations about statistical significance we applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+It selected 100 SNPs, from all the aforementioned regions and others (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P-value of some of them, which further complicates the biological interpretation.
-In this regard, and in opposition to what would be predicted by the omnigenic model, the genes to which this SNPs map to (Section \ref{methods:node_score}) are not interconnected in the PPIN (Section \ref{methods:gene_network}).
-This motivates exploring network methods which, by trading statistical significance for biological relevance, search susceptibility subnetworks.
-In fact, such methods provided comparable classification accuracies to L1-penalized logistic regression (Figure \ref{fig:benchmark}D), while providing more interpretable solutions.
+Moreover, and in opposition to what the omnigenic model predicts, the genes to which this SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
+This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
+In fact, such methods provided comparable classification accuracies to L1-penalized logistic regression (Figure~\ref{fig:benchmark}D), while providing more interpretable solutions.
 
@@
 \begin{tablenotes}
+  \footnotesize{
     \item \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
-  \item \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
-  \item \textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.
-  \item \textbf{\(\rho\)\textsubscript{consensus}}: Pearson's correlation between the subnetwork and the consensus network.
+\textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
+\textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.
+\textbf{\(\rho\)\textsubscript{consensus}}: Pearson's correlation between the subnetwork and the consensus network.
+  }
 \end{tablenotes}
@@
   \centering
-  \includegraphics[width=.9\linewidth]{./figures/figure_1.png}
-  \caption{\label{fig:solution_overview} TODO.}
+  \includegraphics[height=.75\textheight]{./figures/figure_1.png}
+  \caption{\label{fig:solution_overview} Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
+As LEAN did not produce any significant gene (adjusted P-value < 0.05), it was excluded.
+Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
+\textbf{(A)} Manhattan plots of SNPs/genes; in black, the method's solution.
+The Bonferroni threshold is indicated by a red line (2.54 \texttimes{} 10\textsuperscript{-7} for SNPs, 1.53 \texttimes{} 10\textsuperscript{-6} for genes).
+\textbf{(B)} Solution networks produced by the different methods.
+\textbf{(C)} Pearson correlation between the genes selected by each of the methods.
+\textbf{(D)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
+For reference, we added a gray line with a slope of 1.}
 \end{figure}
 
-We applied six network methods to the GENESIS dataset (Section \ref{methods:methods}).
-As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained five solutions (Supplementary Figure \ref{fig:solution_overview}, Supplementary Files 1 and 2): one for each of the remaining four gene-based methods (Section \ref{methods:gene_network}), and one for SConES GI (Section \ref{methods:snp_network}).
-The solutions are very heterogeneous (Table \ref{tab:gene_solutions} and Supplementary table \ref{tab:snp_solutions}), making hard to draw joint conclusions.
-HotNet2 produced the largest solution subnetwork with 440 genes.
-SConES GI failed to recover genes in the PPIN, but it recovered one genomic region mapped to RNA gene \emph{RNU6-420P}.
-All solution subnetworks except LEAN's have, on average, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values (Supplementary Figure \ref{fig:solution_overview}A).
+We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
+As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}, Supplementary~Files~1~and~2): one for each of the remaining four gene-based methods, one for SConES GI, which works at the SNP level, and the consensus.
+
+The solutions are very heterogeneous, making it hard to draw joint conclusions, although there are some common threads (Table~\ref{tab:gene_solutions}).
+All obtained solution subnetworks have, on average, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions: heinz strongly favored high scoring genes, while dmGWAS is less conservative (median gene P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+However, there are nuances between solutions: heinz strongly favored high scoring genes i.e.
+lower P-values, while dmGWAS is less conservative (median gene P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+The solution subnetworks present other desirable properties.
+First, five of them were enriched in known breast cancer susceptibility genes (Figure~\ref{fig:solution_overview}D; consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
+Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
+This agrees with the notion that disease genes are more central than other, non-essential genes \cite{pinero_uncovering_2016}.
+We observe that this conclusion holds in this disease, as known breast cancer susceptibility genes have higher betweenness centrality than others (one-tailed Wilcoxon rank-sum test P-value~=~2.64 \texttimes{} 10\textsuperscript{-5}, Supplementary Figure~\ref{sfig:consensus_stats}C).
+Interestingly, SConES' selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than unrelated SNPs.
+However, very central nodes are also more likely to be connecting a random pair of nodes, making then more likely to be selected by the examined methods.
+Hence, further work is needed characterize the impact of centrality on these methods' results.
 
-The solution subnetworks present other desirable properties.
-First, four of the methods succeeded at recovering known breast cancer susceptibility genes (Supplementary Figure \ref{fig:solution_overview}D), as their subnetworks were enriched in breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
-We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (Supplementary Figure \ref{fig:solution_overview}D).
-Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC.
-This confirms the capability of network methods to find the same signal as in more powered studies by leveraging on prior knowledge.
-Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in three solutions (dmGWAS, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
-This agrees with the notion that disease genes are more central than other, non-essential genes \cite{pinero_uncovering_2016}.
-We observe that this conclusion holds in this disease, as known breast cancer susceptibility genes have higher betweenness centrality than others (one-tailed Wilcoxon rank-sum test P-value = 2.64 \texttimes{} 10\textsuperscript{-5}, Supplementary Figure \ref{sfig:consensus_stats}C).
-Interestingly, SConES' selected SNPs are also more central than the average SNP (Supplementary table \ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than unrelated SNPs.
-However, very central nodes are also more likely to be connecting a random pair of nodes, making then more likely to be selected by the examined methods.
-Hence, further work is needed draw conclusions.
-
-As the solutions were quite different from each other it is hard to draw joint conclusions.
+Drawing further conclusions is hard due to the differences between solutions.
+The first of them is their sizes: HotNet2 produced the largest solution (440 genes) while SConES GI failed to recover protein coding genes.
+Their topologies also differ in terms of median centrality and number of connected components (Figure~\ref{fig:solution_overview}B).
+The selected genes are also quite different (Figure~\ref{fig:solution_overview}C).
 The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
@@
 SigMod, despite being related to SConES, produces a vastly different, large solution.
-On top of recovering three breast cancer susceptibility genes, a keratin-based region of its subnetwork affects the cytoskeleton (\emph{structural constituent of cytoskeleton}, GO enrichment's adjusted P-value = 9.10 \texttimes{} 10\textsuperscript{-4}), a potentially novel susceptibility mechanism for cancer susceptibility.
-Interestingly, dmGWAS solution is also related to cytoskeleton (\emph{tubulin binding}, GO enrichment's adjusted P-value = 0.031).
-But, additionally, it includes a submodule of proteins related to \emph{unfolded protein binding} (GO enrichment's adjusted P-value = 0.045), which has been previously related to cancer susceptibility \cite{calderwood_heat_2016}.
+A keratin-based region of its subnetwork affects the cytoskeleton (\emph{structural constituent of cytoskeleton}, GO enrichment's adjusted P-value~=~9.10 \texttimes{} 10\textsuperscript{-4}), a potentially novel susceptibility mechanism for cancer susceptibility.
+Interestingly, dmGWAS solution is also related to cytoskeleton (\emph{tubulin binding}, GO enrichment's adjusted P-value~=~0.031).
+But, additionally, it includes a submodule of proteins related to \emph{unfolded protein binding} (GO enrichment's adjusted P-value~=~0.045), which has been previously related to cancer susceptibility \cite{calderwood_heat_2016}.
 Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
@@
   \caption{\label{fig:consensus}
-    Consensus subnetwork on GENESIS (Section \ref{methods:methods}).
+    Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
 Each node is represented by a pie chart, which shows the methods that selected it.
@@
 The heterogeneity in the solutions suggested, and the shared encouraging properties suggested that each method was capturing different susceptibility mechanisms.
-Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure \ref{sfig:consensus_stats}A).
-Encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure \ref{sfig:consensus_stats}B).
-To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section \ref{methods:methods}).
-This subnetwork (Figure \ref{fig:consensus}) contains 93 genes and shares the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes (Fisher's exact test P-value = 7.8 \texttimes{} 10\textsuperscript{-5}), higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value = 4.29 \texttimes{} 10\textsuperscript{-18}) and TODO pathways.
+Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure~\ref{sfig:consensus_stats}A).
+Encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure~\ref{sfig:consensus_stats}B).
+To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
+This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and shares the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes (Fisher's exact test P-value~=~7.8 \texttimes{} 10\textsuperscript{-5}), higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value~=~4.29 \texttimes{} 10\textsuperscript{-18}) and TODO pathways.
 Topologically, the consensus network can be divided into a large connected component composed of 49 genes, and multiple smaller subnetworks.
-Among the latter, 14 of the 93 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves (TODO TEST).
-Some of them are known breast cancer susceptibility gene like \emph{FGFR2} (Section \ref{results:conventional}) and \emph{SLC4A7} (VEGAS P-value = 2.70 \texttimes{} 10\textsuperscript{-5}, \citet{ahmed_newly_2009}).
+Among the latter, 14 of the 93 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by two methods (TODO TEST).
 
-TODO Yet, there are other novel functions: Globally, a GO enrichment shows the involvement of two cellular processes: unfolded protein binding, and structural constituent of cytoskeleton (adjusted P-values of 0.001, 0.001, respectively), which were already observed in different solutions (Section \ref{results:separate_networks}).
-Remarkably, many of the selected genes are related to mitochondrial translation.
-For instance, MRPS30 (VEGAS P-value = 0.001), encodes a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
-Albeit disconnected from MRPS30, the consensus network includes a 2-node subnetwork composed of two mitochondrial ribosomal protein (MRPS31 - VEGAS P-value = 7.67 \texttimes{} 10\textsuperscript{-3} - and MRPS18B - VEGAS P-value = 7.92 \texttimes{} 10\textsuperscript{-3}), which suggests an involvement of mitochondrial ribosomes in carcinogenesis \cite{required}.
+We conducted a pathway enrichment analysis of the genes in the consensus network (Section \ref{methods:pathway_enrichment}).
+After tackling the overlap between the significant pathways, we found two significant pathways: \emph{Mitochondrial translation} and \emph{Attenuation phase}.
+The former is supported by genes like \emph{MRPS30} (VEGAS P-value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
+Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
+With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
+The genes encoding these proteins are all near each other at 6p21.
+In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
+\emph{HSPA1A} was the most strongly associated one (VEGAS P-value~=~8.37 \texttimes{} 10\textsuperscript{-4}).
 
 We also examined the topological properties of the nodes.
-The genes in the consensus network have higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value = 4.29 \texttimes{} 10\textsuperscript{-18}).
-Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value = 0.57).
-The centrality in the PPIN of the genes in the consensus, however, is weakly anti-correlated with the P-value of association to the disease (Pearson correlation coefficient = -0.26, Supplementary Figure \ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
+The genes in the consensus network have higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value~=~4.29 \texttimes{} 10\textsuperscript{-18}).
+Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value~=~0.57).
+The centrality in the PPIN of the genes in the consensus, however, is weakly anti-correlated with the P-value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
 In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is \emph{COPS5} (Supplementary Figure \ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS (VEGAS P-value = 0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value = 0.006).
-
-% The consensus subnetwork is not completely connected: out of the 93 genes, the largest connected subnetwork includes only 49.
-A GO enrichment analysis showed that this component is related to three major cellular processes: unfolded protein binding, structural constituent of cytoskeleton, and poly(U) RNA binding (adjusted P-values of 0.01, 0.04, and 0.04, respectively).
-We found support in the literature of the involvement of each of these functions in the development of cancer, as discussed next.
-The consensus network also contains a protein directly involved in caspase-mediated apoptosis, \emph{CASP8} (VEGAS P-value = 1.95 \texttimes{} 10\textsuperscript{-4}).
-This is related to the enriched activity, \emph{unfolded protein binding}, which inhibits caspase-dependent apoptosis, raising the chances of developing cancer \cite{calderwood_heat_2016}.
-It involves three Hsp70 chaperones of the consensus subnetwork: HSPA1A, HSPA1B, and HSPA1L.
-The genes encoding these proteins are all near each other at 6p21.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
-\emph{HSPA1A} was the most strongly associated one (VEGAS P-value = 8.37 \texttimes{} 10\textsuperscript{-4}).
+For instance, the most central node in the consensus network is \emph{COPS5} (Supplementary Figure~\ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+Despite its lack of association in GENESIS (VEGAS P-value~=~0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value~=~0.006).
 
@@
 \centering
-\includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
+\includegraphics[width=.75\linewidth]{./figures/figure_4.pdf}
 \caption{\label{fig:benchmark}
@@
 We compared the six methods in a 5-fold subsampling setting (Sections \ref{methods:comparison} and \ref{methods:classifier}).
-Specifically, we measured five properties (Figure \ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
+Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
 
-Both solution size and the subset of it selected by the classifier (\emph{active set}) varies greatly between the different methods (Figure \ref{fig:benchmark}A).
+Both solution size and the subset of it selected by the classifier (\emph{active set}) varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
 Heinz produced the smallest solutions, with an average of 182 selected SNPs.
@@
 Interestingly, heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%), although it is  high for all the methods (> 86\%).
-To determine whether the selected SNPs could be used for patient classification we computed the performance of the classifier on the \emph{test dataset} (Figure \ref{fig:benchmark}B).
+To determine whether the selected SNPs could be used for patient classification we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
 The different classifiers displayed similar sensitivities and specificities, all in the 0.52 -- 0.56 range.
 
-Another desirable quality of an algorithm is stability of the solution with regards to different inputs (Section \ref{methods:comparison}).
-Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure \ref{fig:benchmark}C).
+Another desirable quality of an algorithm is stability of the solution with regards to different inputs (Section~\ref{methods:comparison}).
+Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure~\ref{fig:benchmark}C).
 Conversely, the other methods displayed similarly low stabilities.
 
-In terms of computational runtime, the fastest method was heinz (Figure \ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds, and HotNet2 was slowest (3 days and 14 hours on average).
+In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds, and HotNet2 was slowest (3 days and 14 hours on average).
 Including the time required to compute the gene scores slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
@@
 
-In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different (Supplementary Figure \ref{sfig:pearson_methods}A).
+In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different (Supplementary Figure~\ref{sfig:pearson_methods}A).
 That is due to the particularities of the methods which directly or indirectly provide information about the dataset.
@@
 
-In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P-value for the subnetwork (Table \ref{tab:gene_solutions}).
+In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P-value for the subnetwork (Table~\ref{tab:gene_solutions}).
 Due to this parsimonious and highly associated solution, it was the best method to select a set of good biomarkers for classification.
-(Figure \ref{fig:benchmark}B).
+(Figure~\ref{fig:benchmark}B).
 Its conservativeness stems from its preprocessing step, which models the gene P-values as a mixture model of a beta and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Figure \ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after applying the BUM model (Section \ref{methods:methods}).
+Due to the limited signal at the gene level in this dataset (Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after applying the BUM model (Section~\ref{methods:methods}).
 Hence, heinz's solution subnetwork consists only of 4 genes, which does not provide much insight of the biology of cancer.
@@
 This framework considers all nodes at distance 2 of the examined, and accepts weakly associated genes if they are linked to another, strongly associated one.
-This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure \ref{fig:issues}A).
-SigMod displays the same tendency, as the most central genes are the least associated to the disease (Figure \ref{fig:issues}B).
+This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure~\ref{fig:issues}A).
+SigMod displays the same tendency, as the most central genes are the least associated to the disease (Figure~\ref{fig:issues}B).
 This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
@@
 In fact, due to linkage disequilibrium, such genes are favored by SConES, as selecting SNPs in an LD block which overlaps with a gene favors selecting the rest of the gene.
-This might explain why the GS and GM networks, heavily affected by linkage disequilibrium, produce similar results (Supplementary Figure \ref{sfig:pearson_methods}B).
+This might explain why the GS and GM networks, heavily affected by linkage disequilibrium, produce similar results (Supplementary Figure~\ref{sfig:pearson_methods}B).
 On the other hand, SConES penalizes selecting SNPs and not their neighbors.
 This makes it conservative regarding SNPs with many interactions, for instance those mapped to hubs in the PPIN.
-For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure \ref{fig:issues}C).
-In fact SConES GS and SConES GM select regions related to breast cancer, like 16q12 (\emph{TOX3}, Section \ref{results:conventional}), 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), and 10q26 (\emph{FGFR2}, Section \ref{results:conventional}).
+For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure~\ref{fig:issues}C).
+In fact SConES GS and SConES GM select regions related to breast cancer, like 16q12 (\emph{TOX3}, Section~\ref{results:conventional}), 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), and 10q26 (\emph{FGFR2}, Section~\ref{results:conventional}).
 On top of that only SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
@@
 Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
-lower \(\lambda\), Section \ref{methods:methods}), it is unclear what is the best strategy to find them.
+lower \(\lambda\), Section~\ref{methods:methods}), it is unclear what is the best strategy to find them.
 In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hurting the accuracy with which the GS network captures the genome structure.
@@
 
-We compared the results of different network methods to the European cohort of BCAC, the largest GWAS to date in breast cancer (Section \ref{methods:bcac}).
+We compared the results of different network methods to the European cohort of BCAC, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
 This comparison is pertinent, since despite caveats we expect a significant shared genetic factors with GENESIS, and since it is a much larger study (XX times more samples, 13 million SNPs through imputation).
-We conducted a study at both the SNP and the gene levels, equivalent to the one shown in Section \ref{results:conventional} for GENESIS.
+We conducted a study at both the SNP and the gene levels, equivalent to the one shown in Section~\ref{results:conventional} for GENESIS.
 Despite the low overlap using a conventional approach (TODO), the solutions provided by the different network approaches overlap significantly (Fisher's exact test P-value $\leq X \times 10^{-X}$).
 This illustrates the ability of network guided GWAS to boost discovery.
+
+We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (Figure~\ref{fig:solution_overview}D).
+Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC.
+This confirms the capability of network methods to find the same signal as in more powered studies by leveraging on prior knowledge.
 
@@
 With regards to SConES, less SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
-However, as mentioned in Section \ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
+However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
 
@@
 Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
-The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure \ref{sfig:biotypes_excluded}).
-Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section \ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
+The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure~\ref{sfig:biotypes_excluded}).
+Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
 Even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
-However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value = 2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
+However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
 As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
@@
 Lastly, all the methods rely heavily on how SNPs are mapped to genes.
-In Section \ref{results:conventional} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+In Section~\ref{results:conventional} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
 
@@
 On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
-When we compared the impact of using a larger network containing interactions from both high-throughput experiment and the literature (Section \ref{methods:gene_network}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure \ref{sfig:lc_ht_comparison}).
+When we compared the impact of using a larger network containing interactions from both high-throughput experiment and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
 This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the selected solutions.
@@
 \begin{tablenotes}
-\item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.
-\item \median{P}\textsubscript{SNP}: median P-value of the selected SNPs.
+  \footnotesize{
+    \item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.\median{P}\textsubscript{SNP}: median P-value of the selected SNPs.
+  }
 \end{tablenotes}
@@
 \begin{tablenotes}
+  \footnotesize{
     \item \mean{Betweenness}: mean betweenness of the selected genes in the full network.
-\item \median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson's correlation with the consensus network.
+\median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson's correlation with the consensus network.
+  }
 \end{tablenotes}
@@
   \centering
-  \includegraphics[width=.9\linewidth]{./figures/sfigure_2.png}
+  \includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
   \caption{\label{sfig:snp_gene_manhattan} Association in GENESIS.
@@
 \textbf{(C)} SNP association as in panel (A).
-The SNPs in black are selected by a L1-penalized logistic regression (Section \ref{methods:classifier}, $\lambda = 0.03$).}
+The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).}
 \end{figure}
@@
 \caption{\label{sfig:consensus_names}
-Consensus subnetwork on GENESIS (Section \ref{methods:methods}).
+Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
 \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
 The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).
-This panel is identical to Figure \ref{fig:consensus}.
+This panel is identical to Figure~\ref{fig:consensus}.
 \textbf{(B)} Manhattan plot showing the genes included in the subnetwork.
@@
 \textbf{(D,E)} Proportion of the Bonferroni significant genes (in GENESIS and BCAC, respectively) included in the consensus network.
-\textbf{F} Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene, as in Figure \ref{fig:solution_overview}D}.
+\textbf{F} Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene, as in Figure~\ref{fig:solution_overview}D}.
 \end{figure}
@@
 \centering
-\includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{./figures/sfigure_5.pdf}
+\includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
 \caption{\label{sfig:consensus_stats}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
+\includegraphics[width=.85\linewidth]{./figures/sfigure_7.pdf}
 \caption{\label{sfig:lc_ht_comparison}

### 2020-03-18 manuscript/manuscript.tex
manuscript updated

@@
 In this work, we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype.
-We interrogate a familial breast cancer genome-wide association study focused on \emph{BRCA1/2} negative French women.
+We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
@@
 Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$ ) and occupy more central positions in the network than average.
-Additionally, it includes subnetworks of mechanisms related to cancer, like protein folding (\emph{HSPA1A}, \emph{HSPA1B}, and \emph{HSPA1L}) or mitochondrial ribosomes (\emph{MRPS30}, \emph{MRPS31}, \emph{MRPS18B}).
-We also observed a general dysregulation in the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
+Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
+Finally, we find a significantly large overlap between the genes in the consensus network and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
+In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
 \end{abstract}
@@
 Some of them stem from the high-dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped.
-This limits the statistical power of the experiment, as only variants with large and moderate effects can be detected.
+This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}.
 And it is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
-Additionally, to avoid false positives, a conservative multiple test correction is applied, typically Bonferroni.
-However, Bonferroni is known to be overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
+Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the aforementioned Bonferroni.
+However, Bonferroni is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
@@
 In this regard many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
-As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and between them are the edges that link them.
-These functional relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction.
+As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them.
+These relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction.
 Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}.
-In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011} \cite{pinero_uncovering_2016}.
+In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}.
 This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes.
@@
 Some methods strongly constrain the problem to certain kinds of subnetworks.
-Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on \emph{star} subnetworks, i.e.
+Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e.
 instances were both a gene and its direct interactors are associated with the disease.
@@
 However, they differ in their tolerance to the inclusion of low scoring nodes, and the possible number of disconnected subnetworks in the solution.
-Lastly, other methods also consider the topology of the network, favoring solutions that are not only high scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
+Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
 
@@
 In addition, we removed the samples with more than 10\% missing genotypes.
-After control for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, and 3 additional controls having a high relatedness score).
-Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with CLIS, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis).
+After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, and 3 additional controls having a high relatedness score).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis).
 The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
@@
 
-We use such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
-For the former, we consider any gene that can be mapped to any of the selected SNPs as selected as well.
-For the latter, we consider all the SNPs that can be mapped to that gene as selected by the method.
+We used such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
+For the former, we considered any gene that can be mapped to any of the selected SNPs as selected as well.
+For the latter, we considered all the SNPs that can be mapped to that gene as selected by the method.
 
@@
 
-\subsubsection{Networks used}
+\subsubsection{Networks}
 \label{methods:networks}
 
-\paragraph{Gene-gene network}
+\paragraph{Gene network}
 Out of the six methods tested, five use a gene-gene interaction network (Section~\ref{methods:methods}).
-Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), for practical reasons we focused on a protein-protein interaction network (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it.
+Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), we used protein-protein interaction networks (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 \paragraph{SNP networks}
-SConES \cite{azencott_efficient_2013} is the only of the studied methods designed to handle SNP networks.
+SConES \cite{azencott_efficient_2013} is the only studied methods designed to handle SNP networks.
 As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs.
 \citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
-We focused on the GI network, as it fits the scope of this article better, using the PPIN described above.
+We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
 However, at different stages of this work we also used GS and GM for comparison.
 For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
-In all three the node scores are the association score of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
-\subsubsection{Methods used}
+In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
+
+\subsubsection{Studied methods}
 \label{methods:methods}
@@
 Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-How to account for that is not the only choice a developer of network methods needs to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
-In consequence, multiple solutions have been proposed.
-In this article, we examine six of them: five that explore the PPIN, and one which explores SNP networks.
+But accounting for that is not the only choice a developer of a network method needs to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+Unsurprisingly, multiple solutions have been proposed.
+We examined six of them: five that explore the PPIN, and one which explores SNP networks.
 We selected methods that were open source, had an implementation available, and an accessible documentation.
@@
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subgraph on the network \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive value when there is.
+The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive when there is.
 This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired FDR.
@@
 All three are positive quantities.
-Heinz implements the algorithm from \citet{ljubic_algorithmic_2006}, which although in practice is often fast and optimal, neither is guaranteed.
-We used BioNet's implementation of heinz, available on Bioconductor \cite{beisser_bionet:_2010,heinz}.
+Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, although in practice is often fast and optimal, neither is guaranteed.
+We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
 
@@
 
-\item[{LEAN}] LEAN searches disregulated ``star'' gene subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
+\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
 By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
@@
 The minimum of these \emph{k} P-values is the score of the subnetwork.
-This score is transformed into a P-value through an empirical distribution obtained via a subsampling scheme, where sets of \emph{m} genes are selected randomly, and their score computed.
+This score is transformed into a P-value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
 Lastly, P-values are corrected for multiple testing through a Benjamini-Hochberg correction.
@@
 where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
-Given two hyperparameters, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
+Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
+As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
+In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
 We used the version on SConES implemented in the R package martini \cite{martini}.
-As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
-Note that the solution to the above problem can consist of several connected subnetworks which are disconnected from each other.
-In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
 \end{description}
@@
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
-Pathways with an adjusted P-value \textless 0.05 were regarded significant.
+Pathways with an adjusted P-value \textless~0.05 were deemed significant.
 
 \subsection{Evaluation of methods}
+\label{methods:comparison}
 
@@
 \subsubsection{Properties ot the solution}
-\label{methods:comparison}
+\label{methods:algorithm_comparison}
 
 We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input) of the different network methods.
-The stability was measured using the Pearson's correlation between different runs as suggested \citet{nogueira_measuring_2016}.
+The stability was measured using the Pearson's correlation between different runs as suggested by \citet{nogueira_measuring_2016}.
 
-\subsubsection{Classification accuracy of selected biomarkers}
+\subsubsection{Classification accuracy of selected SNPs}
 \label{methods:classifier}
@@
 This penalty was set by cross-validation.
-We applied each network method to each \emph{train} set, and trained the classifier on the \emph{train} set as well using only on the selected SNPs.
+We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
 When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
 Then we evaluated the sensitivity and the specificity on the \emph{test set}.
+The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
 To obtain a baseline, we also trained the classifier on all the SNPs.
-We also evaluated the size of the active set, an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
 
@@
 \label{methods:bcac}
+
 An alternative way to evaluate the results is comparing our results to an external dataset.
@@
 
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 228\,951 women of European ancestry from the general population \cite{Michailidou2017}.
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs (after imputation) genotyped on 228\,951 women of European ancestry from the general population \cite{Michailidou2017}.
 Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic, while GENESIS is a homogeneous dataset focusing on the French high-risk population attending the family cancer clinics.
-Despite these differences, there should be shared genetic architecture.
-On top of that, that overlap should become larger at the gene level.
+Despite these differences, there should be shared genetic architecture, specially at the gene level.
 For that purpose, we searched for associated genes as in Section~\ref{methods:node_score}.
-After imputation, it consisted of 13\,250\,642 SNPs.
-We used all of them to measure gene-level association, but only the SNPs available on GENESIS after quality control for the SNP-level comparisons.
+We used all the SNPs to measure gene association, but only those available after quality control on GENESIS for the SNP-level comparisons.
 
@@
 
-This work required developing computational pipelines for several GWAS analyses, such physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
-For each of those processes, a streamlined, project-agnostic pipeline with a clear interface was created.
-They are compiled in the following GitHub repository: \url{https://github.com/hclimente/gwas-tools}.
-The code that applies these pipelines to the GENESIS project, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
-Although the GENESIS dataset is not publicly available, the published code should work on any other GWAS dataset.
-All the produced gene subnetworks were deposited on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
+We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
+For each of those processes, we created a pipeline with a clear interface that should woth on any GWAS dataset.
+They are compiled in \url{https://github.com/hclimente/gwas-tools}.
+Although the GENESIS data is not public, the code that applies these pipelines to it, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
 
@@
 
-We conducted association analyses in the GENESIS dataset at both the SNP and the gene levels (Section~\ref{methods:node_score}).
+We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
 Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
@@
 Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1/2} carriers studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
-Only the peak in chromosome 10 replicated in the gene-level analysis, with \emph{FGFR2} just above the threshold of significance (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
-These results show an overlap in the genetic architecture of the disease between the studied French population sample and other populations, especially at the gene level.
+Only \emph{FGFR2} was significant at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
+These results show an overlap in the genetic architecture of the disease between the studied French population sample and other populations.
 
@@
 Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-Exploring algorithms unconstrained by considerations about statistical significance we applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
 It selected 100 SNPs, from all the aforementioned regions and others (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
@@
 The Bonferroni threshold is indicated by a red line (2.54 \texttimes{} 10\textsuperscript{-7} for SNPs, 1.53 \texttimes{} 10\textsuperscript{-6} for genes).
-\textbf{(B)} Solution networks produced by the different methods.
-\textbf{(C)} Pearson correlation between the genes selected by each of the methods.
-\textbf{(D)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
-For reference, we added a gray line with a slope of 1.}
+\textbf{(B)} Pearson correlation between the genes selected by each of the methods.
+\textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
+For reference, we added a gray line with a slope of 1.
+\textbf{(D)} Solution networks produced by the different methods.}
 \end{figure}
@@
 We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}, Supplementary~Files~1~and~2): one for each of the remaining four gene-based methods, one for SConES GI, which works at the SNP level, and the consensus.
+As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+TODO Smoothen the transition?
+The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
+By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or not.
+In fact, SConES GI retrieves 4 subnetworks in intergenic regions, and 1 overlapping an RNA gene (\emph{RNU6-420P}).
+SigMod, despite being related to SConES, produces a vastly different, large solution.
+A pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) links different parts of the network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (adjusted P-values~<~0.03).
+Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, adjusted P-value~=~0.01).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (adjusted P-values~<~0.49), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
+The second largest subnetwork (13 nodes), contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
+As SigMod, the genes are involved in mitochondrial translation (adjusted P-value~=~1.87 \texttimes{} 10\textsuperscript{-4}), and also in glycogen metabolism and transcription of nuclear receptors (adjusted P-value~<~0.04).
 
-The solutions are very heterogeneous, making it hard to draw joint conclusions, although there are some common threads (Table~\ref{tab:gene_solutions}).
+As it can be seen, the solutions are very heterogeneous, making it hard to draw joint conclusions.
+In fact, the specific genes by each solution are also different (Figure~\ref{fig:solution_overview}B).
+Another prominent difference is their sizes: HotNet2 produced the largest solution (440 genes) while SConES GI failed to recover any protein coding gene.
+Also their topologies differ, as measured by the median centrality and the number of connected components (Figure~\ref{fig:solution_overview}D).
+
+Yet, some common themes exist (Table~\ref{tab:gene_solutions}).
 All obtained solution subnetworks have, on average, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions: heinz strongly favored high scoring genes i.e.
-lower P-values, while dmGWAS is less conservative (median gene P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
-The solution subnetworks present other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (Figure~\ref{fig:solution_overview}D; consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
-Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
-This agrees with the notion that disease genes are more central than other, non-essential genes \cite{pinero_uncovering_2016}.
-We observe that this conclusion holds in this disease, as known breast cancer susceptibility genes have higher betweenness centrality than others (one-tailed Wilcoxon rank-sum test P-value~=~2.64 \texttimes{} 10\textsuperscript{-5}, Supplementary Figure~\ref{sfig:consensus_stats}C).
-Interestingly, SConES' selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than unrelated SNPs.
-However, very central nodes are also more likely to be connecting a random pair of nodes, making then more likely to be selected by the examined methods.
+However, there are nuances between solutions: heinz strongly genes with lower P-values, while dmGWAS is less conservative (median P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+Additionally, the solution subnetworks present other desirable properties.
+First, five of them were enriched in known breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
+Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
+This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P-value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
+Interestingly, SConES' selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by the examined methods.
 Hence, further work is needed characterize the impact of centrality on these methods' results.
-
-Drawing further conclusions is hard due to the differences between solutions.
-The first of them is their sizes: HotNet2 produced the largest solution (440 genes) while SConES GI failed to recover protein coding genes.
-Their topologies also differ in terms of median centrality and number of connected components (Figure~\ref{fig:solution_overview}B).
-The selected genes are also quite different (Figure~\ref{fig:solution_overview}C).
-The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
-By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or not.
-In fact, SConES GI, which adds to GM the interactions between genes, retrieves 4 subnetworks in intergenic regions, and 1 overlapping an RNA gene (\emph{RNU6-420P}).
-SigMod, despite being related to SConES, produces a vastly different, large solution.
-A keratin-based region of its subnetwork affects the cytoskeleton (\emph{structural constituent of cytoskeleton}, GO enrichment's adjusted P-value~=~9.10 \texttimes{} 10\textsuperscript{-4}), a potentially novel susceptibility mechanism for cancer susceptibility.
-Interestingly, dmGWAS solution is also related to cytoskeleton (\emph{tubulin binding}, GO enrichment's adjusted P-value~=~0.031).
-But, additionally, it includes a submodule of proteins related to \emph{unfolded protein binding} (GO enrichment's adjusted P-value~=~0.045), which has been previously related to cancer susceptibility \cite{calderwood_heat_2016}.
-Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
-The second largest subnetwork (13 nodes), contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
 
@@
 Each node is represented by a pie chart, which shows the methods that selected it.
-The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).}
+We labeled (and enlarged) the genes having a VEGAS2v2 P-value < 0.001, the two most central genes (\emph{COPS5} and \emph{OFD1}) and those that are known breast cancer susceptibility genes.
+The latter ones are also colored in pink.}
 \end{figure}
 
-The heterogeneity in the solutions suggested, and the shared encouraging properties suggested that each method was capturing different susceptibility mechanisms.
-Due to the limited overlap between methods, only 20 genes were common to more than two of them (Supplementary Figure~\ref{sfig:consensus_stats}A).
-Encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure~\ref{sfig:consensus_stats}B).
+The shared properties despite their heterogeneity suggested that each method was capturing different aspects of cancer susceptibility.
+Indeed, only 20 genes were common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure~\ref{sfig:consensus_stats}B).
 To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
-This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and shares the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes (Fisher's exact test P-value~=~7.8 \texttimes{} 10\textsuperscript{-5}), higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value~=~4.29 \texttimes{} 10\textsuperscript{-18}) and TODO pathways.
-Topologically, the consensus network can be divided into a large connected component composed of 49 genes, and multiple smaller subnetworks.
-Among the latter, 14 of the 93 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by two methods (TODO TEST).
+This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and shares the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
 
-We conducted a pathway enrichment analysis of the genes in the consensus network (Section \ref{methods:pathway_enrichment}).
-After tackling the overlap between the significant pathways, we found two significant pathways: \emph{Mitochondrial translation} and \emph{Attenuation phase}.
+A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions'.
+After tackling the overlap between the significant pathways, we found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
 The former is supported by genes like \emph{MRPS30} (VEGAS P-value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
@@
 
-We also examined the topological properties of the nodes.
-The genes in the consensus network have higher betweenness centrality than the rest of the genes (Wilcoxon rank-sum test P-value~=~4.29 \texttimes{} 10\textsuperscript{-18}).
-Interestingly, within genes in the consensus network, cancer genes are as central as non-cancer genes (Wilcoxon rank-sum test P-value~=~0.57).
-The centrality in the PPIN of the genes in the consensus, however, is weakly anti-correlated with the P-value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D), which suggests that some highly central genes were selected because they were on the shortest path between two highly associated genes.
+Topologically the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
+Among the latter, 14 of the 93 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by two methods.
+The opposite would be the case of genes with highly centrality in the PPIN, which is weakly anti-correlated with the P-value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
+This suggests that they were selected because they were on the shortest path between two highly associated genes.
 In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is \emph{COPS5} (Supplementary Figure~\ref{sfig:consensus_names}), a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS (VEGAS P-value~=~0.22), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS P-value~=~0.006).
+For instance, the most central node in the consensus network is \emph{COPS5}, a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+Despite its lack of association in GENESIS or BCAC (VEGAS2v2 P-value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS2v2 P-value~=~0.006).
 
-\subsection{Methods are comparably stable, and produce similarly good predictors}
+\subsection{Methods are comparably unstable, and produce similarly predictors}
 
@@
 \centering
-\includegraphics[width=.75\linewidth]{./figures/figure_4.pdf}
+\includegraphics[width=.72\linewidth]{./figures/figure_4.pdf}
 \caption{\label{fig:benchmark}
 Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times of a random subset of the samples, and tested on the remaining samples (Sections \ref{methods:comparison} and \ref{methods:classifier}).
+Each method was run 5 times of a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
 \textbf{(A)} Number of SNPs selected by each method and number of SNPs on the active set used by the Lasso classifier.
@@
 
-We compared the six methods in a 5-fold subsampling setting (Sections \ref{methods:comparison} and \ref{methods:classifier}).
+We compared the six methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
 Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
-
 Both solution size and the subset of it selected by the classifier (\emph{active set}) varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
@@
 To determine whether the selected SNPs could be used for patient classification we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
-The different classifiers displayed similar sensitivities and specificities, all in the 0.52 -- 0.56 range.
+The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
+Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method whose stated goal is to minimize prediction error.
 
-Another desirable quality of an algorithm is stability of the solution with regards to different inputs (Section~\ref{methods:comparison}).
+Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
 Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure~\ref{fig:benchmark}C).
@@
 
-In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds, and HotNet2 was slowest (3 days and 14 hours on average).
+In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds.
+HotNet2 was the slowest (3 days and 14 hours on average).
 Including the time required to compute the gene scores slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
-Considering that, it took 5 days on average for HotNet2 to produce results.
+Considering that, it took 5 days on average for HotNet2 to produce a result.
 
@@
 
-In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different (Supplementary Figure~\ref{sfig:pearson_methods}A).
+In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different.
 That is due to the particularities of the methods which directly or indirectly provide information about the dataset.
@@
 In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P-value for the subnetwork (Table~\ref{tab:gene_solutions}).
-Due to this parsimonious and highly associated solution, it was the best method to select a set of good biomarkers for classification.
-(Figure~\ref{fig:benchmark}B).
+Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Figure~\ref{fig:benchmark}C).
 Its conservativeness stems from its preprocessing step, which models the gene P-values as a mixture model of a beta and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after applying the BUM model (Section~\ref{methods:methods}).
-Hence, heinz's solution subnetwork consists only of 4 genes, which does not provide much insight of the biology of cancer.
+Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after that transformation.
+This small solution does not provide much insight of the biology of cancer.
 Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
@@
 In fact, due to linkage disequilibrium, such genes are favored by SConES, as selecting SNPs in an LD block which overlaps with a gene favors selecting the rest of the gene.
-This might explain why the GS and GM networks, heavily affected by linkage disequilibrium, produce similar results (Supplementary Figure~\ref{sfig:pearson_methods}B).
+This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Supplementary Figure~\ref{sfig:pearson_methods}).
 On the other hand, SConES penalizes selecting SNPs and not their neighbors.
@@
 
-We compared the results of different network methods to the European cohort of BCAC, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
-This comparison is pertinent, since despite caveats we expect a significant shared genetic factors with GENESIS, and since it is a much larger study (XX times more samples, 13 million SNPs through imputation).
-We conducted a study at both the SNP and the gene levels, equivalent to the one shown in Section~\ref{results:conventional} for GENESIS.
-Despite the low overlap using a conventional approach (TODO), the solutions provided by the different network approaches overlap significantly (Fisher's exact test P-value $\leq X \times 10^{-X}$).
-This illustrates the ability of network guided GWAS to boost discovery.
-
-We also compared the outcome of the network methods to the association tests conducted on the population of European ancestry from the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (Figure~\ref{fig:solution_overview}D).
-Encouragingly, every solution subnetwork is enriched in genes or SNPs that are Bonferroni-significant in BCAC.
-This confirms the capability of network methods to find the same signal as in more powered studies by leveraging on prior knowledge.
+We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
+This comparison is pertinent as, despite not targeting the same population(s), we expect a shared genetic architecture at the gene level \cite{required} TODO CAN'T FIND A CITATION FOR THIS, at which most examined network metods operate.
+This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
+We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
+The solutions provided by the different network approaches overlap significantly (Fisher's exact test P-value \textless~0.019).
+The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) (Figure~\ref{fig:solution_overview}C).
+However, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
 
@@
 Based on divergent considerations of what the desired set of biomarkers is, several methods for network-guided biomarker discovery have been proposed.
-We reviewed the performance of six of them on GWAS.
-Despite their differences, most of them produced a relevant subset of biomarkers, recovering known familial breast cancer genes.
-We also discuss the limitations of such analyses, related to the lack of known interactions around some genes.
-
-A crucial step for the gene based methods is the computation of the gene score.
-In this work we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms might have more statistical power.
-Another important decision is how to handle LD in a GWAS.
-VEGAS2v2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2v2's computation time.
-With regards to SConES, less SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
-However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
-
-The network methods we studied differ in what the optimal solution subnetwork looks like, which acted as a double-edged sword.
+We reviewed six of them.
+Most of them produced a relevant subset of biomarkers, although the solutions were notably different.
 In one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
@@
 While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which risk oversimplifying their richness.
-HotNet2 balances both approaches at the expense of producing a constellation of many, highly associated, small subnetworks.
-Despite their differences, all the solutions produced comparable discrimination capabilities on a linear classifier trained on them.
-On the negative side, the methods were also remarkably unstable, and hence likely to produce very different solutions in face of slightly different inputs.
+HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
+Despite these differences, all the solutions produced comparably poor discrimination capabilities on a linear classifier trained on them, and were remarkably unstable, and hence likely to produce change in face of slightly different inputs.
 
@@
 We use a straightforward aggregation to generate it, including any node that was recovered by at least two methods.
-The resulting network is a synthesis of the altered mechanism: it is smaller than the largest solutions (SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
+The resulting network is a synthesis of the altered mechanism: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
 The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
@@
 Lastly, all the methods rely heavily on how SNPs are mapped to genes.
-In Section~\ref{results:conventional} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
 
@@
 This responds to concerns of some authors about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
-Their presense might bias the results of the presented methods, as if changes the topology of the networks by adding nodes closer on average to other nodes than average.
+Their presense might bias the results of the presented methods, as they change the topology of the network by adding nodes closer to other nodes than average.
 On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
@@
 This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the selected solutions.
+
+A crucial step for the gene based methods is the computation of the gene score.
+In this work we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms might have more statistical power.
+Another important decision is how to handle LD in a GWAS.
+VEGAS2v2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2v2's computation time.
+With regards to SConES, less SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
+However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
 
@@
 
-\section*{Supplementary materials}
+\section{Supplementary materials}
 
@@
 \textbf{(D,E)} Proportion of the Bonferroni significant genes (in GENESIS and BCAC, respectively) included in the consensus network.
-\textbf{F} Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene, as in Figure~\ref{fig:solution_overview}D}.
+\textbf{F} Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene.}
 \end{figure}

### 2020-03-23 manuscript/manuscript.tex
went over text up to 3.5

@@
 \newcommand{\median}[1]{$\hat{\mbox{#1}}$}
+
+\usepackage{lmodern} % Chloe: better font rendering
+
+% Chloe's commands
+\usepackage[dvipsnames]{xcolor}
+\usepackage[normalem]{ulem} 
+\newcommand{\cazcom}[2]{{\uline{#1}}\unskip\space\textcolor{MidnightBlue}{[CA: #2]}}
+\newcommand{\caz}[2]{{\sout{#1}}\unskip\space\textcolor{MidnightBlue}{#2}}
+%%% To get the original text before my comments:
+% \newcommand{\cazcom}[2]{#1}
+% \newcommand{\caz}[2]{#1}
+%%% To get the corrected text and remove my comments:
+% \newcommand{\cazcom}[2]{#1}
+% \newcommand{\caz}[2]{#2}
 
@@
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
-It does so by jointly considering the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study the utility of six network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype.
+\caz{It does so by jointly considering}{Indeed, it allows to jointly consider} the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
+In this work, we study the utility of six \caz{network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype}{methods that identify subnetworks with higher association scores to a phenotype to discover new biomarkers for breast cancer susceptibility}.
 We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-By trading statistical astringency for biological meaningfulness, most network methods get more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$ ) and occupy more central positions in the network than average.
+By trading statistical \caz{astringency}{stringency} for biological meaningfulness, most network methods \caz{get}{give} more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$) and occupying more central positions in the network than average.
 Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
@@
 In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
+\cazcom{}{Add a sentence to the effect that the computational tools are available on github for ease of application to other data sets?}
 \end{abstract}
@@
 Then, a per-SNP statistical association test is conducted (e.g.
-logistic regression).
+\cazcom{\caz{}{based on }logistic regression}{You're doing a chi2 allelic test later on, so maybe use this here rather than talking about logistic regression?}).
 Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
@@
 However, this experimental setting also presents intrinsic challenges.
-Some of them stem from the high-dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped.
+Some of them stem from the high dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped.
 This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}.
-And it is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
-Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the aforementioned Bonferroni.
-However, Bonferroni is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
+\caz{And it}{This} is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
+Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the \caz{aforementioned}{previously mentioned} Bonferroni\caz{}{correction}.
+However, Bonferroni \caz{}{correction} is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
@@
 
-In this regard many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
+\caz{In this regard}{For this reason,} many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
 As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them.
@@
 
-Network-based, biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
+Network-based biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
 In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
@@
 instances were both a gene and its direct interactors are associated with the disease.
-Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnected genes with high association scores.
-However, they differ in their tolerance to the inclusion of low scoring nodes, and the possible number of disconnected subnetworks in the solution.
-Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
+Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnecting genes with high association scores.
+However, they differ in their tolerance to the inclusion of low-scoring nodes, and the possible number of \cazcom{\caz{disconnected subnetworks}{connected components}}{Are you worried people don't understand ``connected components''?
+If so, we should define it and use it.} in the solution.
+Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
 
@@
 
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy - Weinberg equilibrium in controls (P-value \textless~0.001), and those missing on more than 10\% of the samples.
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P-value \textless~0.001), and those missing on more than 10\% of the samples.
 A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
@@
 
-\subsubsection{Mathematical notation}
+\subsubsection{Mathematical notations}
 \label{methods:notation}
@@
 
-On top of a weight, nodes have other properties, provided by the topology of the graph.
-In this article we focus on two: degree centrality, and betweenness centrality.
+\caz{On top of}{In addition to} a weight, nodes have other properties, provided by the topology of the graph.
+In this article we focus on two\caz{}{of those}: degree centrality, and betweenness centrality.
 The degree centrality, or degree, is the number of edges that a node has.
@@
 \paragraph{Gene network}
-Out of the six methods tested, five use a gene-gene interaction network (Section~\ref{methods:methods}).
-Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), we used protein-protein interaction networks (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to it.
+Out of the six methods tested, \cazcom{five use a gene-gene interaction network}{SConES GI does too...} (Section~\ref{methods:methods}).
+Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), we used protein-protein interaction networks (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless specified otherwise, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
+Unless \caz{specified otherwise}{otherwise specified}, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
 Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols.
@@
 In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
+
+\cazcom{}{How many genes are involved in those interactions?
+In the ``SNP and gene association'' section, you describe 164\,037 SNPs mapping to 32\,767 gene symbols.
+How many of these genes and SNPs are involved in the aforementioned 142\,541 (resp.
+173\,797) interactions?
+And what happens to the genes to which you have mapped SNPs but that are not involved in any HINT association: are they discarded, or do they remain as isolated nodes?}
 
@@
 
+\cazcom{}{What are the sizes of the resulting GS, GM and GI networks?
+164\,037 SNPs/nodes (or fewer if some aren't used on the gene-gene networks and you've removed them entirely from the analysis)?
+How many edges in the different networks?
+If you're worried about space, the sizes of the gene-gene and SNP-SNP networks from the previous paragraph and this one could be given in a table in the supplementaries.}
+
 \subsubsection{Studied methods}
 \label{methods:methods}
+\cazcom{}{Maybe we could change ``studied'' in this section (title, text, tables...) to something that makes it sound a bit less like a school assignment, and also less like we decided to study those but could have studied others?
+Something like ``Network-based methods'' would have a bit more agency to it?
+Sorry if I'm nitpicking.}
 
@@
 \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.
-\textbf{\emph{Comp.}}: number of connected subnetworks in the solution.
+\textbf{\emph{Comp.}}: number of connected \caz{subnetworks}{components} in the solution.
 \textbf{\emph{Input}}: genotype data or GWAS summary statistics.
@@
 Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-But accounting for that is not the only choice a developer of a network method needs to make: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+But \caz{accounting for that is not the only choice a developer of a network method needs to make}{this is not the only aspect to model when developing a network method}: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
 Unsurprisingly, multiple solutions have been proposed.
@@
 \begin{enumerate}
-\item Select a seed node.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph $S$ as
+\item Select a seed node \caz{}{$i$ and form $S_i=\{i\}$}.
+\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph \caz{$S$}{$S_i$} as
 
 \begin{equation*} 
-Z_m = \frac{1}{\sqrt{k}} \sum_{i \in S} z_i
+Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j
 \end{equation*}
 
-where \emph{k} is the number of genes in the subgraph; z\textsubscript{i} is the Z score of gene $i$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{i}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of $S$ i.e. nodes at distance \(\le\) d. We set d~=~2.
-\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r). In our experiments, we set r~=~0.1.
+where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+\item Identify neighboring nodes of \caz{$S$}{$S_i$},  i.e. nodes at distance \(\le\) d. \caz{We set d~=~2.}{}
+\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r). \caz{In our experiments, we set r~=~0.1.}{}
 \item Repeat 2-4 until no further enlargement is possible.
+\caz{}{\item Add $S_i$ to the list of subnetworks to return.}
 \end{enumerate}
+\cazcom{}{Missing here: how to select seed nodes. Is every node a seed node? Every node that has not yet been included in a subgraph?}
 
-Lastly, the module's Z-score is normalized as
+Lastly, \caz{the module's}{a subnetwork's} Z-score is normalized as
 
@@
 We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
+\caz{}{We used the suggested hyperparameters $d = 2$ and $r = 0.1$} We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
 \end{description}
@@
 The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive when there is.
-This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired FDR.
-Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitely long computational time.
-To solve it efficiently it is re-casted as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S):
+This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired \caz{FDR}{false discovery rate (FDR)}.
+Thus formulated, the problem is NP-complete, and hence solving it would require a \caz{prohibitely}{prohibitively} long computational time.
+To solve it efficiently it is re-cast\caz{ed}{} as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
 
@@
 
-were p(v)~=~w(v) - w' is the \emph{profit} of adding a node, c(e)~=~w' is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$.
+were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, \caz{$c(e)~=~w'$}{$c(e)=-w'$} is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ \caz{}{is the smallest node weight of $G$}.
 All three are positive quantities.
-Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, although in practice is often fast and optimal, neither is guaranteed.
+Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, \caz{although} in practice is often fast and optimal, \caz{}{although} neither is guaranteed.
 We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
@@
 To that end, it considers both the local topology of the network and the scores of the nodes.
-The former is captured by an insulated heat diffusion process: at the beginning, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
-This process continues until equilibrium is reached, and results in a similarity matrix $F$.
-$F$ is used to compute the similarity matrix $E$ that accounts also for similarities in node scores as
-
+The former is captured by an insulated heat diffusion process: at \caz{the beginning}{initialization}, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
+This process continues until equilibrium is reached, and results in a \caz{similarity}{diffusion} matrix $F$.
+$F$ is used to compute the similarity matrix $E$ that \caz{accounts also for similarities in node scores as}{models exchanged heat as}
 \begin{equation*} 
@@
 \end{equation*}
-
 where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-We scored the nodes as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
+\caz{}{For any two nodes/genes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.}  To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
+Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
+We \caz{scored the nodes}{assigned initial node scores} as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
 In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
-To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
-Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
 HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
-Both parameters are set automatically by the algorithm, and are robust \cite{leiserson_pan-cancer_2015}.
+Both parameters are set automatically by the algorithm, \caz{and are robust}{which is robust to their values} \cite{leiserson_pan-cancer_2015}.
 HotNet2 is implemented in Python \cite{hotnet2}.
@@
 
+% \begin{equation}
+% \label{eq:scones}
+% \underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
+% \end{equation}
+
+\cazcom{}{Your equation for SConES was wrong.
+Use either of the following (I'd use the one that does not involve the Laplacian, then you do not need to introduce the Laplacian in the Mathematical notations.
+One way to look at it is to start from the bottom one, where the connectivity term increases with each edge that connects a selected node to a non-selected node, and as you're maximizing the whole thing, you need a minus in top of it.
+Then
+\[
+\sum_{u \in V_S} \sum_{v \not\in V_S} W_{uv} = \sum_{u \in V_S} \left(\sum_{v \in V} W_{uv} - \sum_{v \in V_S} W_{uv} \right) = \sum_{u \in V_S} \left(D_{uu} - \sum_{v \in V_S} W_{uv} \right) 
+= \sum_{u \in V_S} \sum_{v \in V_S} \left(D_{uv} - W_{uv} \right) \]
+The first equality is true because the set of all nodes is the union of the set of nodes of $S$ and the set of nodes not in $S$; the second by definition of the degree; and the third because the degree matrix is diagonal and $D_{uv}=0$ for $u \neq v$.
+Finally $L_{uv} = D_{uv} - W_{uv}$.
+}
+
 \begin{equation}
 \label{eq:scones}
-\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
+\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} {\color{red}-} \underbrace{\lambda \sum_{v \in V_S} \sum_{u {\color{red} \in} V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
+\end{equation}
+
+or
+
+\begin{equation}
+\label{eq:scones}
+\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} {\color{red}-} \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {\color{red}W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
 \end{equation}
@@
 where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
-Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
+\caz{}{The connectivity term penalizes disconnected solutions, with many edges between a node that is selected and a node that isn't.} Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
 As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
@@
 \item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but using the adjacency matrix rather than the Laplacian matrix (Section~\ref{methods:notation}), to quantify solutions containing many edges.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but \caz{using the adjacency matrix rather than the Laplacian matrix (Section~\ref{methods:notation}), to quantify solutions containing many edges}{the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones}.
 
@@
 First it is designed for gene-gene networks.
-Second, by replacing the adjacency by the Laplacian matrix, it favors subnetworks containing many edges.
+Second, \caz{by replacing the adjacency by the Laplacian matrix,}{} it favors subnetworks containing many edges.
 SConES, instead, penalizes connections between the selected and unselected nodes.
@@
 
-We searched for pathways enriched in the gene subnetworks produced by the aforementioned methods.
+We searched for pathways enriched in the gene subnetworks produced by the \caz{aforementioned}{above} methods.
 We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
@@
 When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
-We used the 5 repetitions to estimate the average and the deviation of the different measures.
+We used the 5 repetitions to estimate the average and the \caz{}{standard} deviation of the different measures.
 
@@
 
-We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input) of the different network methods.
-The stability was measured using the Pearson's correlation between different runs as suggested by \citet{nogueira_measuring_2016}.
+We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input\caz{}{, here, using different subsets of the data}) of the different network methods.
+The stability was measured using \caz{the}{} Pearson correlation between different runs as suggested by \citet{nogueira_measuring_2016}.
 
@@
 Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic, while GENESIS is a homogeneous dataset focusing on the French high-risk population attending the family cancer clinics.
-Despite these differences, there should be shared genetic architecture, specially at the gene level.
+Despite these differences, \caz{there should be}{we expect a} shared genetic architecture, especially at the gene level.
 For that purpose, we searched for associated genes as in Section~\ref{methods:node_score}.
@@
 We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
-For each of those processes, we created a pipeline with a clear interface that should woth on any GWAS dataset.
+For each of those processes, we created a pipeline with a clear interface that should work on any GWAS dataset.
 They are compiled in \url{https://github.com/hclimente/gwas-tools}.
-Although the GENESIS data is not public, the code that applies these pipelines to it, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+Although the GENESIS data is not public, the code \caz{that applies these pipelines to it}{to apply the pipelines to this data}, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
 We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
@@
 Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}), a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
-It selected 100 SNPs, from all the aforementioned regions and others (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
+We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}), \cazcom{a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy}{describe it in the methods rather than here?
+How was the regularization parameter chosen?}.
+It selected 100 SNPs, both from all aforementioned regions and from other regions (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P-value of some of them, which further complicates the biological interpretation.
 Moreover, and in opposition to what the omnigenic model predicts, the genes to which this SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
-This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
-In fact, such methods provided comparable classification accuracies to L1-penalized logistic regression (Figure~\ref{fig:benchmark}D), while providing more interpretable solutions.
+\caz{}{In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.} This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
+In fact, such methods provided comparable classification \caz{accuracies}{performance} to L1-penalized logistic regression \cazcom{(Figure~\ref{fig:benchmark}D)}{Is it B?
+I don't seel L1-penalized logistic regression on this figure; also with these types of performance I'm not sure it's worth making that point.
+Maybe frame it as ``do no worse'' than ``have comparable performance''...}, while providing more interpretable solutions.
 
@@
 \caption{\label{tab:gene_solutions}
-Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
+Summary statistics on the \caz{results}{outputs/solutions} of multiple network methods \caz{on the gene-gene interaction network}{}.
+The first row contains the summary statistics on the whole network.}
 \centering
@@
 \textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.
-\textbf{\(\rho\)\textsubscript{consensus}}: Pearson's correlation between the subnetwork and the consensus network.
+\textbf{\(\rho\)\textsubscript{consensus}}: Pearson correlation between the subnetwork and the consensus network.
   }
@@
 The Bonferroni threshold is indicated by a red line (2.54 \texttimes{} 10\textsuperscript{-7} for SNPs, 1.53 \texttimes{} 10\textsuperscript{-6} for genes).
-\textbf{(B)} Pearson correlation between the genes selected by each of the methods.
-\textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
+\textbf{(B)} \cazcom{\caz{Pearson correlation}{Overlap} between the genes selected by each of the methods\caz{}{, measured by Pearson correlation between indicator vectors}.}{You only explain how to interpret Pearson correlation in Figure 3, although it should maybe come in the Methods when you introduce stability} \textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
 For reference, we added a gray line with a slope of 1.
@@
 As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
-TODO Smoothen the transition?
-The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
+\cazcom{TODO Smoothen the transition?}{These solutions differ in many aspects.
++ go aspect by aspect (size; non-coding regions; topological properties; enriched pathways?} The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
 By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or not.
@@
 Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
-The second largest subnetwork (13 nodes), contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
-As SigMod, the genes are involved in mitochondrial translation (adjusted P-value~=~1.87 \texttimes{} 10\textsuperscript{-4}), and also in glycogen metabolism and transcription of nuclear receptors (adjusted P-value~<~0.04).
+The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
+As with SigMod, the genes are involved in mitochondrial translation (adjusted P-value~=~1.87 \texttimes{} 10\textsuperscript{-4}), and also in glycogen metabolism and transcription of nuclear receptors (adjusted P-value~<~0.04).
 
 As it can be seen, the solutions are very heterogeneous, making it hard to draw joint conclusions.
-In fact, the specific genes by each solution are also different (Figure~\ref{fig:solution_overview}B).
-Another prominent difference is their sizes: HotNet2 produced the largest solution (440 genes) while SConES GI failed to recover any protein coding gene.
+\caz{In fact, the specific genes by each solution are also different}{The overlap between the genes featured in each solution is quite small} (Figure~\ref{fig:solution_overview}B).
+Another prominent difference is their sizes: \caz{HotNet2 produced the largest solution (440 genes)}{the largest solution, produced by HotNet2, contains 440 genes,} while SConES GI failed to recover any protein coding gene.
 Also their topologies differ, as measured by the median centrality and the number of connected components (Figure~\ref{fig:solution_overview}D).
 
-Yet, some common themes exist (Table~\ref{tab:gene_solutions}).
-All obtained solution subnetworks have, on average, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
+Yet, some common themes exist \cazcom{(Table~\ref{tab:gene_solutions})}{Why this reference here?}.
+All obtained solution subnetworks have, \cazcom{on average}{1) you used a median; 2) I'm still wondering if it wouldn't make more sense to report mean or median -log10(p)?
+The conclusion will still hold}, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions: heinz strongly genes with lower P-values, while dmGWAS is less conservative (median P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+However, there are nuances between solutions: heinz strongly \caz{}{favors} genes with lower P-values, while dmGWAS is less conservative (median P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
 Additionally, the solution subnetworks present other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
+First, five of them were enriched in known breast cancer susceptibility genes (\cazcom{dmGWAS, heinz, HotNet2, and SigMod}{that's 4, not 5??}, Fisher's exact test one-sided P-value < 0.03).
 Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
 This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P-value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES' selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
-However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by the examined methods.
-Hence, further work is needed characterize the impact of centrality on these methods' results.
+Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by \caz{the examined}{any network} methods.
+\cazcom{Hence, further work is needed characterize the impact of centrality on these methods' \caz{results}{outputs}.}{Maybe more of a Discussion than a Results sentence?}
 
@@
 
-The shared properties despite their heterogeneity suggested that each method was capturing different aspects of cancer susceptibility.
-Indeed, only 20 genes were common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association was (Supplementary Figure~\ref{sfig:consensus_stats}B).
+\caz{The shared properties despite their heterogeneity suggested that each method was capturing}{Despite the heterogeneity of the solutions, their shared properties suggest that each method captures} different aspects of cancer susceptibility.
+Indeed, only 20 genes \caz{were}{are} common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods select\caz{ed}{} a gene, the higher its association \caz{was}{to the phenotype} (Supplementary Figure~\ref{sfig:consensus_stats}B).
 To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
@@
 
-A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions'.
-After tackling the overlap between the significant pathways, we found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
+A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions.
+\cazcom{After tackling the overlap between the significant pathways}{I'm not sure what you mean here?}, we found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
 The former is supported by genes like \emph{MRPS30} (VEGAS P-value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
 Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
-With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
-The genes encoding these proteins are all near each other at 6p21.
+With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: \cazcom{HSPA1A, HSPA1B, and HSPA1L}{Italics for gene/protein names?
+Or is it only for gene names?
+But HSPA1A is in italics below.}.
+The genes encoding these proteins are all near each other at \cazcom{6p21}{Isn't that the HLA complex?
+I'm always a bit wary of this region because of the strong LD, disantangling signals is hard...}.
 In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
@@
 
-\subsection{Methods are comparably unstable, and produce similarly predictors}
+\subsection{Methods are comparably unstable, and produce similar predictors}
+\cazcom{}{1) ``Network methods'' rather than ``Methods'' in title?
+2) I'm not sure we want to emphasize predictor similarity when they're all so terrible 3) Shouldn't this section appear before the previous one?
+First talk about the individual solutions, then present the consensus?.}
 
@@
 Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times of a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+Each method was run 5 times \caz{of}{on} a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
 As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
-\textbf{(A)} Number of SNPs selected by each method and number of SNPs on the active set used by the Lasso classifier.
+\textbf{(A)} Number of SNPs selected by each method \cazcom{and number of SNPs in the active set \caz{used}{found} by the \caz{Lasso}{L1-penalized logistic regression} classifier}{Not necessary if L1-penalized logreg is introduced in the Methods}.
 Points are the average over the 5 runs; lines represent the standard error of the mean.
@@
 Points are the average over the 5 runs; lines represent the standard error of the mean.
-\textbf{(C)} Pairwise Pearson's correlations of the solutions used by different methods.
-A Pearson's correlation of 1 means the two solutions are the same.
-A Pearson's correlation of 0 means that there is no SNP in common between the two solutions.
+\textbf{(C)} Pairwise Pearson correlations of the solutions used by different methods.
+A Pearson correlation of 1 means the two solutions are the same.
+A Pearson correlation of 0 means that there is no SNP in common between the two solutions.
 \textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP).
-For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) which took VEGAS2v2 on average to compute the gene scores from SNP summary statistics.}
+For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) \caz{which took VEGAS2v22}{that VEGAS2v2 took} on average to compute the gene scores from SNP summary statistics.}
 \end{figure}
 
-We compared the six methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+We compared the six \caz{}{network} methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
 Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
@@
 The largest solutions come from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
-Interestingly, heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%), although it is  high for all the methods (> 86\%).
-To determine whether the selected SNPs could be used for patient classification we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
+Interestingly, \cazcom{heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%)}{Isn't that because it's only selecting 4 genes?}, although \caz{it}{this proportion} is high for all the methods (> 86\%).
+To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
 The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
-Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method whose stated goal is to minimize prediction error.
+Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method \caz{whose stated goal is to minimize}{aiming only at minimizing} prediction error.
+\cazcom{}{Do we want to comment this suggest that in any case there isn't that much signal in the data set, or are we shooting ourselves in the foot doing that?}
 
@@
 
-In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which leverages on its ability to find efficiently the solution in a few seconds.
+In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which \caz{leverages on its ability to find efficiently the solution}{finds the solution efficiently} in a few seconds.
 HotNet2 was the slowest (3 days and 14 hours on average).
-Including the time required to compute the gene scores slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
+Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
 Considering that, it took 5 days on average for HotNet2 to produce a result.
@@
 \caption{\label{fig:issues}
-Drawbacks confronted when using network guided methods.
+Drawbacks \caz{confronted}{encountered?} when using network guided methods.
 \textbf{(A)} dmGWAS solution subnetwork.
@@
 
-In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different.
+\cazcom{In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different.
 That is due to the particularities of the methods which directly or indirectly provide information about the dataset.
-For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
+For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.}{Slightly redundant with the ``A case study: the consensus network'' section, and also maybe move the sentence about LEAN to that section?}
+
+\cazcom{}{Also the first few paragraphs of this section are somewhat redundant with the ``A case study: the consensus network'' section?}
 
@@
 We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
-This comparison is pertinent as, despite not targeting the same population(s), we expect a shared genetic architecture at the gene level \cite{required} TODO CAN'T FIND A CITATION FOR THIS, at which most examined network metods operate.
+This comparison is pertinent as, despite not targeting the same population(s), we expect a shared genetic architecture at the gene level % \cite{required}
+TODO CAN'T FIND A CITATION FOR THIS, at which most examined network metods operate.
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
@@
 
+\cazcom{}{TODO Clean out references a bit (remove URLs, shorten long authors lists with et al., etc.)}
+
 \bibliographystyle{abbrvnat}
@@
     \item \mean{Betweenness}: mean betweenness of the selected genes in the full network.
-\median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson's correlation with the consensus network.
+\median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson correlation with the consensus network.
   }
@@
 \caption{\label{sfig:pearson_methods}
-Pearson's correlation between the different solution subnetworks.
+Pearson correlation between the different solution subnetworks.
 \textbf{(A)} Correlation between selected SNPs.
@@
 \textbf{(B)} Shifts in sensitivity and specificity.
-\textbf{(C)} Shift in Pearson's correlation between benchmarks.
+\textbf{(C)} Shift in Pearson correlation between benchmarks.
 \textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) which took VEGAS2v2 on average to compute the gene scores from SNP summary statistics.
+For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) \caz{which took VEGAS2v2}{that VEGAS2v2 took} on average to compute the gene scores from SNP summary statistics.
 In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}

### 2020-03-24 manuscript/manuscript.tex
accepted some comments, and improved label visibility in figure_4

@@
 \usepackage[numbers]{natbib}
+\usepackage{lmodern} % Chloe: better font rendering 
 \newcommand{\mean}[1]{$\overline{\mbox{#1}}$}
 \newcommand{\median}[1]{$\hat{\mbox{#1}}$}
-
-\usepackage{lmodern} % Chloe: better font rendering
 
@@
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
-\caz{It does so by jointly considering}{Indeed, it allows to jointly consider} the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study the utility of six \caz{network methods to discover new biomarkers for breast cancer susceptibility by searching subnetworks with higher association scores to this phenotype}{methods that identify subnetworks with higher association scores to a phenotype to discover new biomarkers for breast cancer susceptibility}.
+Indeed, it allows to jointly consider the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
+In this work, we study the utility of six methods that identify subnetworks with higher association scores to a phenotype to discover new biomarkers for breast cancer susceptibility.
 We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-By trading statistical \caz{astringency}{stringency} for biological meaningfulness, most network methods \caz{get}{give} more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+By trading statistical stringency for biological meaningfulness, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
 Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$) and occupying more central positions in the network than average.
@@
 This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}.
-\caz{And it}{This} is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
-Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the \caz{aforementioned}{previously mentioned} Bonferroni\caz{}{correction}.
-However, Bonferroni \caz{}{correction} is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
+This is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
+Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the previously mentioned Bonferroni correction.
+However, Bonferroni correction is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
@@
 
-\caz{In this regard}{For this reason,} many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
+For this reason, many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
 As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them.
@@
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless \caz{specified otherwise}{otherwise specified}, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
+Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
 Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols.
@@
 Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-But \caz{accounting for that is not the only choice a developer of a network method needs to make}{this is not the only aspect to model when developing a network method}: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
 Unsurprisingly, multiple solutions have been proposed.
@@
 where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of \caz{$S$}{$S_i$},  i.e. nodes at distance \(\le\) d. \caz{We set d~=~2.}{}
-\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r). \caz{In our experiments, we set r~=~0.1.}{}
+\item Identify neighboring nodes of \caz{$S$}{$S_i$},  i.e. nodes at distance \(\le\) d.
+\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
 \item Repeat 2-4 until no further enlargement is possible.
@@
 
-Lastly, \caz{the module's}{a subnetwork's} Z-score is normalized as
+Lastly, a subnetwork's Z-score is normalized as
 
@@
 We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-\caz{}{We used the suggested hyperparameters $d = 2$ and $r = 0.1$} We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
+We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
+We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
 \end{description}
@@
 The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive when there is.
-This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired \caz{FDR}{false discovery rate (FDR)}.
-Thus formulated, the problem is NP-complete, and hence solving it would require a \caz{prohibitely}{prohibitively} long computational time.
-To solve it efficiently it is re-cast\caz{ed}{} as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
+This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
+Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
+To solve it efficiently it is re-cast as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
 
@@
 
-were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, \caz{$c(e)~=~w'$}{$c(e)=-w'$} is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ \caz{}{is the smallest node weight of $G$}.
+were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, \caz{$c(e)~=~w'$}{$c(e)=-w'$} is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
 All three are positive quantities.
-Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, \caz{although} in practice is often fast and optimal, \caz{}{although} neither is guaranteed.
+Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
 We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
@@
 To that end, it considers both the local topology of the network and the scores of the nodes.
-The former is captured by an insulated heat diffusion process: at \caz{the beginning}{initialization}, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
-This process continues until equilibrium is reached, and results in a \caz{similarity}{diffusion} matrix $F$.
-$F$ is used to compute the similarity matrix $E$ that \caz{accounts also for similarities in node scores as}{models exchanged heat as}
+The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
+This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
+$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
 \begin{equation*} 
@@
 where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-\caz{}{For any two nodes/genes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.}  To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
+For any two nodes/genes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
+To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
 Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
-We \caz{scored the nodes}{assigned initial node scores} as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
+We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
 In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
 HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
-Both parameters are set automatically by the algorithm, \caz{and are robust}{which is robust to their values} \cite{leiserson_pan-cancer_2015}.
+Both parameters are set automatically by the algorithm, which is robust to their values \cite{leiserson_pan-cancer_2015}.
 HotNet2 is implemented in Python \cite{hotnet2}.
@@
 where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
-\caz{}{The connectivity term penalizes disconnected solutions, with many edges between a node that is selected and a node that isn't.} Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
+The connectivity term penalizes disconnected solutions, with many edges between a node that is selected and a node that is not.
+Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
 As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
@@
 \item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but \caz{using the adjacency matrix rather than the Laplacian matrix (Section~\ref{methods:notation}), to quantify solutions containing many edges}{the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones}.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
 
@@
 First it is designed for gene-gene networks.
-Second, \caz{by replacing the adjacency by the Laplacian matrix,}{} it favors subnetworks containing many edges.
+Second, it favors subnetworks containing many edges.
 SConES, instead, penalizes connections between the selected and unselected nodes.
@@
 
-We searched for pathways enriched in the gene subnetworks produced by the \caz{aforementioned}{above} methods.
+We searched for pathways enriched in the gene subnetworks produced by the above methods.
 We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
@@
 When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
-We used the 5 repetitions to estimate the average and the \caz{}{standard} deviation of the different measures.
+We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
 
@@
 
-We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input\caz{}{, here, using different subsets of the data}) of the different network methods.
-The stability was measured using \caz{the}{} Pearson correlation between different runs as suggested by \citet{nogueira_measuring_2016}.
+We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different subsets of the data) of the different network methods.
+The stability was measured using Pearson correlation between different runs as suggested by \citet{nogueira_measuring_2016}.
 
@@
 Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic, while GENESIS is a homogeneous dataset focusing on the French high-risk population attending the family cancer clinics.
-Despite these differences, \caz{there should be}{we expect a} shared genetic architecture, especially at the gene level.
+Despite these differences, we expect a shared genetic architecture, especially at the gene level.
 For that purpose, we searched for associated genes as in Section~\ref{methods:node_score}.
@@
 They are compiled in \url{https://github.com/hclimente/gwas-tools}.
-Although the GENESIS data is not public, the code \caz{that applies these pipelines to it}{to apply the pipelines to this data}, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
 We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
@@
 \caz{}{In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.} This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
-In fact, such methods provided comparable classification \caz{accuracies}{performance} to L1-penalized logistic regression \cazcom{(Figure~\ref{fig:benchmark}D)}{Is it B?
+In fact, such methods provided comparable classification performance to L1-penalized logistic regression \cazcom{(Figure~\ref{fig:benchmark}B)}{Is it B?
 I don't seel L1-penalized logistic regression on this figure; also with these types of performance I'm not sure it's worth making that point.
@@
 The Bonferroni threshold is indicated by a red line (2.54 \texttimes{} 10\textsuperscript{-7} for SNPs, 1.53 \texttimes{} 10\textsuperscript{-6} for genes).
-\textbf{(B)} \cazcom{\caz{Pearson correlation}{Overlap} between the genes selected by each of the methods\caz{}{, measured by Pearson correlation between indicator vectors}.}{You only explain how to interpret Pearson correlation in Figure 3, although it should maybe come in the Methods when you introduce stability} \textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
+\textbf{(B)} \cazcom{Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.}{You only explain how to interpret Pearson correlation in Figure 3, although it should maybe come in the Methods when you introduce stability} \textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
 For reference, we added a gray line with a slope of 1.
@@
 This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions: heinz strongly \caz{}{favors} genes with lower P-values, while dmGWAS is less conservative (median P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+However, there are nuances between solutions: heinz strongly favors genes with lower P-values, while dmGWAS is less conservative (median P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
 Additionally, the solution subnetworks present other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (\cazcom{dmGWAS, heinz, HotNet2, and SigMod}{that's 4, not 5??}, Fisher's exact test one-sided P-value < 0.03).
+First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
 Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
@@
 Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
-However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by \caz{the examined}{any network} methods.
-\cazcom{Hence, further work is needed characterize the impact of centrality on these methods' \caz{results}{outputs}.}{Maybe more of a Discussion than a Results sentence?}
+However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by any network methods.
+\cazcom{Hence, further work is needed characterize the impact of centrality on these methods' outputs.}{Maybe more of a Discussion than a Results sentence?}
 
@@
 
-\caz{The shared properties despite their heterogeneity suggested that each method was capturing}{Despite the heterogeneity of the solutions, their shared properties suggest that each method captures} different aspects of cancer susceptibility.
-Indeed, only 20 genes \caz{were}{are} common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods select\caz{ed}{} a gene, the higher its association \caz{was}{to the phenotype} (Supplementary Figure~\ref{sfig:consensus_stats}B).
+Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
+Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods select a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
 To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
@@
 Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times \caz{of}{on} a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+Each method was run 5 times on a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
 As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
-\textbf{(A)} Number of SNPs selected by each method \cazcom{and number of SNPs in the active set \caz{used}{found} by the \caz{Lasso}{L1-penalized logistic regression} classifier}{Not necessary if L1-penalized logreg is introduced in the Methods}.
+\textbf{(A)} Number of SNPs selected by each method and number of SNPs in the active set found by the classifier.
 Points are the average over the 5 runs; lines represent the standard error of the mean.
@@
 \textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP).
-For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) \caz{which took VEGAS2v22}{that VEGAS2v2 took} on average to compute the gene scores from SNP summary statistics.}
+For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2v2 took on average to compute the gene scores from SNP summary statistics.}
 \end{figure}
@@
 The largest solutions come from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
-Interestingly, \cazcom{heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%)}{Isn't that because it's only selecting 4 genes?}, although \caz{it}{this proportion} is high for all the methods (> 86\%).
+Interestingly, \cazcom{heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%)}{Isn't that because it's only selecting 4 genes?}, although this proportion is high for all the methods (> 86\%).
 To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
 The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
-Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method \caz{whose stated goal is to minimize}{aiming only at minimizing} prediction error.
+Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
 \cazcom{}{Do we want to comment this suggest that in any case there isn't that much signal in the data set, or are we shooting ourselves in the foot doing that?}
@@
 
-In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which \caz{leverages on its ability to find efficiently the solution}{finds the solution efficiently} in a few seconds.
+In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which finds the solution efficiently in a few seconds.
 HotNet2 was the slowest (3 days and 14 hours on average).
@@
 \caption{\label{fig:issues}
-Drawbacks \caz{confronted}{encountered?} when using network guided methods.
+Drawbacks encountered when using network guided methods.
 \textbf{(A)} dmGWAS solution subnetwork.

### 2020-03-26 manuscript/manuscript.tex
comments on discussion

@@
 \caption{\label{tab:gene_solutions}
-Summary statistics on the \caz{results}{outputs/solutions} of multiple network methods \caz{on the gene-gene interaction network}{}.
+Summary statistics on the \caz{results}{outputs|solutions} of multiple network methods \caz{on the gene-gene interaction network}{}.
 The first row contains the summary statistics on the whole network.}
@@
 
-\subsection{The consensus overcomes the problems of the individual solutions}
+\subsection{The consensus overcomes the limitations of the individual solutions}
 \label{results:drawbacks}
+\cazcom{}{I replaced ``problems'' with ``limitations'' in the section title.}
 
@@
 lower \(\lambda\), Section~\ref{methods:methods}), it is unclear what is the best strategy to find them.
-In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hurting the accuracy with which the GS network captures the genome structure.
+In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- \caz{hurting}{limiting|hindering} the accuracy with which the GS network captures the genome structure.
 
@@
 We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
-This comparison is pertinent as, despite not targeting the same population(s), we expect a shared genetic architecture at the gene level % \cite{required}
-TODO CAN'T FIND A CITATION FOR THIS, at which most examined network metods operate.
+This comparison is pertinent as, \caz{despite not targeting}{although they do not target} the same population(s), we expect a shared genetic architecture at the gene level % \cite{required}
+\cazcom{TODO CAN'T FIND A CITATION FOR THIS}{Maybe we don't need to justify it?
+Or maybe Nadine/Fabienne can provide something?}, at which most examined network metods operate.
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
 We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
-The solutions provided by the different network approaches overlap significantly (Fisher's exact test P-value \textless~0.019).
+The solutions provided by the different network approaches overlap significantly \caz{}{with BCAC findings} (Fisher's exact test P-value \textless~0.019).
 The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) (Figure~\ref{fig:solution_overview}C).
@@
 
-In this article we evaluate the viability of a systems biology take on genetic studies by examining a GWAS dataset on familial breast cancer.
+In this article we evaluate the viability of a systems biology take on genetic \caz{}{association} studies by examining a GWAS dataset on familial breast cancer.
 Such an approach addresses two of the largest issues with GWAS: interpretability and an overly conservative statistical framework that hinders discovery.
-This is achieved by considering the biological context of each of the genes and SNPs.
-Based on divergent considerations of what the desired set of biomarkers is, several methods for network-guided biomarker discovery have been proposed.
+This is achieved by \caz{considering}{using interaction networks to consider} the biological context of each of the genes and SNPs.
+\cazcom{}{I'd like to say here that this is in line with the recently proposed omnigenic model, itself supported by a large body of work on using networks to understand disease.
+I would also like to say that although some folks use the omnigenic model to claim that GWAS don't work, such network methods are a way of bridging classical GWAS and the omnigenic model.}  Based on \caz{divergent}{various|diverse} considerations of what the desired set of biomarkers is, several methods for network-guided biomarker discovery have been proposed \caz{}{in the literature|in recent years}.
 We reviewed six of them.
-Most of them produced a relevant subset of biomarkers, although the solutions were notably different.
-In one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
+
+Most of these methods produced a relevant subset of biomarkers, although the solutions were notably different.
+\caz{In}{At} one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
 On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which risk oversimplifying their richness.
+While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, \cazcom{which risk oversimplifying their richness}{Are you arguing that these methods give richer insights but that we're not equipped to analyze them, resulting in poorer insights?
+I find this a bit convoluted...}.
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
-Despite these differences, all the solutions produced comparably poor discrimination capabilities on a linear classifier trained on them, and were remarkably unstable, and hence likely to produce change in face of slightly different inputs.
-
-To overcome the problems posed by the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
-We use a straightforward aggregation to generate it, including any node that was recovered by at least two methods.
-The resulting network is a synthesis of the altered mechanism: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
+Despite these differences, all solutions \caz{produced comparably poor discrimination capabilities on a linear classifier trained on them}{have equally poor discriminative power|are equally poor at discriminating cases from controls}\caz{, and were remarkably unstable, and hence likely to produce change in face of slightly different inputs.}{.
+In addition, all methods are remarkably unstable, yielding different solutions for slightly different inputs.}
+\cazcom{}{I think we need to discuss this.
+Why aren't we able to classify cases from controls (multiple non-genetic factors, disease heterogeneity...)?
+Why is our approach still useful if we aren't, in fact, discovering biomarkers?
+(it points at relevant pathways, etc) Why are the results unstable?
+(correlations in the data i.e.
+everything is connected i.e.
+the omnigenic model?) Why can we still do something with the results if they are unstable?
+(not clear to me at this stage, if we don't have stability at the pathway/subnetwork level or for the consensus network)}
+
+To overcome the \caz{problems posed by}{limitations of} the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
+We use a straightforward aggregation \caz{to generate it}{}, including any node that was recovered by at least two methods.
+The resulting network is a synthesis of the altered mechanism\caz{}{s}: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
 The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
@@
 Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
-Even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
+In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
 However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
@@
 It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
-This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
-Lastly, all the methods rely heavily on how SNPs are mapped to genes.
+\cazcom{}{Say a word about their ability to recover SNPs that have not been mapped to genes, i.e.
+mostly intergenic genes?} This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
+Lastly, all the methods rely heavily on \cazcom{how SNPs are mapped to genes}{maybe discuss the promising possibility of creating more complex mapping, relying on eQTLs and contact maps?}.
 In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
@@
 In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns of some authors about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
+This responds to concerns \caz{of some authors}{} about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
 Their presense might bias the results of the presented methods, as they change the topology of the network by adding nodes closer to other nodes than average.
@@
 In this work we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms might have more statistical power.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and \cazcom{other algorithms}{cite a few among SKAT, MAGMA, MAGENTA, PEGASUS etc.} might have more statistical power.
 Another important decision is how to handle LD in a GWAS.

### 2020-04-14 manuscript/manuscript.tex
updated figures and addressed most of chloé's comments

--- 2020-03-24 manuscript/manuscript.tex
@@
 In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
-\cazcom{}{Add a sentence to the effect that the computational tools are available on github for ease of application to other data sets?}
+Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
 \end{abstract}
@@
 Then, a per-SNP statistical association test is conducted (e.g.
-\cazcom{\caz{}{based on }logistic regression}{You're doing a chi2 allelic test later on, so maybe use this here rather than talking about logistic regression?}).
+based on logistic regression).
 Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
@@
 Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnecting genes with high association scores.
-However, they differ in their tolerance to the inclusion of low-scoring nodes, and the possible number of \cazcom{\caz{disconnected subnetworks}{connected components}}{Are you worried people don't understand ``connected components''?
-If so, we should define it and use it.} in the solution.
+However, they differ in their tolerance to the inclusion of low-scoring nodes, and the topology of the solution.
 Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
@@
 
-\caz{On top of}{In addition to} a weight, nodes have other properties, provided by the topology of the graph.
-In this article we focus on two\caz{}{of those}: degree centrality, and betweenness centrality.
+In addition to a weight, nodes have other properties, provided by the topology of the graph.
+In this article we focus on two of those: degree centrality, and betweenness centrality.
 The degree centrality, or degree, is the number of edges that a node has.
@@
 The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
-Lastly, the \emph{Laplacian matrix} $L_G$ is defined as $L_G = D_G - W_G$.
 
@@
 \paragraph{Gene network}
-Out of the six methods tested, \cazcom{five use a gene-gene interaction network}{SConES GI does too...} (Section~\ref{methods:methods}).
-Although their respective statistical frameworks are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.), we used protein-protein interaction networks (PPIN), as they are interpretable, well characterized, and most of the methods were designed to scale appropriately to them.
+The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
+Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to scale appropriately to them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
 Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
-Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols.
+Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
+12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
 The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
@@
 In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
-
-\cazcom{}{How many genes are involved in those interactions?
-In the ``SNP and gene association'' section, you describe 164\,037 SNPs mapping to 32\,767 gene symbols.
-How many of these genes and SNPs are involved in the aforementioned 142\,541 (resp.
-173\,797) interactions?
-And what happens to the genes to which you have mapped SNPs but that are not involved in any HINT association: are they discarded, or do they remain as isolated nodes?}
+Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
 
 \paragraph{SNP networks}
-SConES \cite{azencott_efficient_2013} is the only studied methods designed to handle SNP networks.
+SConES \cite{azencott_efficient_2013} is the only network method designed to handle SNP networks.
 As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs.
@@
 In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
+The properties of these three subnetworks is available in Supplementary Table \ref{tab:snp_solutions}.
 
-\cazcom{}{What are the sizes of the resulting GS, GM and GI networks?
-164\,037 SNPs/nodes (or fewer if some aren't used on the gene-gene networks and you've removed them entirely from the analysis)?
-How many edges in the different networks?
-If you're worried about space, the sizes of the gene-gene and SNP-SNP networks from the previous paragraph and this one could be given in a table in the supplementaries.}
-
-\subsubsection{Studied methods}
+\subsubsection{Network methods}
 \label{methods:methods}
-\cazcom{}{Maybe we could change ``studied'' in this section (title, text, tables...) to something that makes it sound a bit less like a school assignment, and also less like we decided to study those but could have studied others?
-Something like ``Network-based methods'' would have a bit more agency to it?
-Sorry if I'm nitpicking.}
 
 \begin{table}[htbp]
-  \caption{\label{tab:method_comparison} Summary of the differences between the studied algorithms.}
+  \caption{Summary of the differences between the network methods.}
+  \label{tab:method_comparison}
   \centering
@@
       SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
-      SigMod & GWAS & Genes & Yes & Module & 1 & Summary & -log\textsubscript{10}(P) & \cite{liu_sigmod:_2017}\\
+      SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \cite{liu_sigmod:_2017}\\
     \end{tabular}
@@
 \textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.
-\textbf{\emph{Comp.}}: number of connected \caz{subnetworks}{components} in the solution.
+\textbf{\emph{Comp.}}: number of connected components in the solution.
 \textbf{\emph{Input}}: genotype data or GWAS summary statistics.
 \textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.
-\textbf{\emph{Ref.}}: original publication featuring the algorithm.
+In the case of SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
+        cumulative distribution function of the standard Normal distribution. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
       }
@@
 \begin{enumerate}
-\item Select a seed node \caz{}{$i$ and form $S_i=\{i\}$}.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph \caz{$S$}{$S_i$} as
+\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
+\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph $S_i$ as
 
@@
 where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of \caz{$S$}{$S_i$},  i.e. nodes at distance \(\le\) d.
+\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
 \item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
 \item Repeat 2-4 until no further enlargement is possible.
-\caz{}{\item Add $S_i$ to the list of subnetworks to return.}
-\end{enumerate}
-\cazcom{}{Missing here: how to select seed nodes. Is every node a seed node? Every node that has not yet been included in a subgraph?}
-
-Lastly, a subnetwork's Z-score is normalized as
+\item Add $S_i$ to the list of subnetworks to return. Its Z-score is normalized as
 
@@
 where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
+\end{enumerate}
 
+DmGWAS carries out this process on every gene in the network.
 We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
@@
 
-were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, \caz{$c(e)~=~w'$}{$c(e)=-w'$} is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
+were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
 All three are positive quantities.
@@
 
-% \begin{equation}
-% \label{eq:scones}
-% \underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
-% \end{equation}
-
-\cazcom{}{Your equation for SConES was wrong.
-Use either of the following (I'd use the one that does not involve the Laplacian, then you do not need to introduce the Laplacian in the Mathematical notations.
-One way to look at it is to start from the bottom one, where the connectivity term increases with each edge that connects a selected node to a non-selected node, and as you're maximizing the whole thing, you need a minus in top of it.
-Then
-\[
-\sum_{u \in V_S} \sum_{v \not\in V_S} W_{uv} = \sum_{u \in V_S} \left(\sum_{v \in V} W_{uv} - \sum_{v \in V_S} W_{uv} \right) = \sum_{u \in V_S} \left(D_{uu} - \sum_{v \in V_S} W_{uv} \right) 
-= \sum_{u \in V_S} \sum_{v \in V_S} \left(D_{uv} - W_{uv} \right) \]
-The first equality is true because the set of all nodes is the union of the set of nodes of $S$ and the set of nodes not in $S$; the second by definition of the degree; and the third because the degree matrix is diagonal and $D_{uv}=0$ for $u \neq v$.
-Finally $L_{uv} = D_{uv} - W_{uv}$.
-}
-
 \begin{equation}
 \label{eq:scones}
-\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} {\color{red}-} \underbrace{\lambda \sum_{v \in V_S} \sum_{u {\color{red} \in} V_S} L_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
-\end{equation}
-
-or
-
-\begin{equation}
-\label{eq:scones}
-\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} {\color{red}-} \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {\color{red}W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
+\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
 \end{equation}
@@
 SConES, instead, penalizes connections between the selected and unselected nodes.
-Third, it returns a single connected subnetwork, which it achieves by exploring a grid of hyperparameters and processing their respective solutions.
+Third, it explores the grid of hyperparameters differently, and processes their respective solutions.
 Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
@@
 Pathways with an adjusted P-value \textless~0.05 were deemed significant.
+As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
 
@@
 We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different subsets of the data) of the different network methods.
-The stability was measured using Pearson correlation between different runs as suggested by \citet{nogueira_measuring_2016}.
+The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by \citet{nogueira_measuring_2016}.
 
@@
 A desirable solution offers good predictive power on unseen \emph{test} samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier trained exclusively on those SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for LD by reducing the number of SNPs included in the model (size of the active set), while improving the generalization of the classifier.
-This penalty was set by cross-validation.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
+The L1 penalty helps to account for LD by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
+This penalty was set by cross-validation, choosing the value that minimized misclassification error.
 We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
@@
 Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}), \cazcom{a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy}{describe it in the methods rather than here?
-How was the regularization parameter chosen?}.
+We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
 It selected 100 SNPs, both from all aforementioned regions and from other regions (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
@@
 Moreover, and in opposition to what the omnigenic model predicts, the genes to which this SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
-\caz{}{In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.} This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
-In fact, such methods provided comparable classification performance to L1-penalized logistic regression \cazcom{(Figure~\ref{fig:benchmark}B)}{Is it B?
-I don't seel L1-penalized logistic regression on this figure; also with these types of performance I'm not sure it's worth making that point.
-Maybe frame it as ``do no worse'' than ``have comparable performance''...}, while providing more interpretable solutions.
+In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
+This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
+In fact, such methods provided comparably (poor) classification performance to L1-penalized logistic regression (Figure~\ref{fig:benchmark}B), while providing more interpretable solutions.
 
@@
   \begin{threeparttable}
-\caption{\label{tab:gene_solutions}
-Summary statistics on the \caz{results}{outputs/solutions} of multiple network methods \caz{on the gene-gene interaction network}{}.
-The first row contains the summary statistics on the whole network.}
+\caption{Summary statistics on the solutions of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
+\label{tab:gene_solutions}
 \centering
@@
   \includegraphics[height=.75\textheight]{./figures/figure_1.png}
-  \caption{\label{fig:solution_overview} Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
+  \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
 As LEAN did not produce any significant gene (adjusted P-value < 0.05), it was excluded.
@@
 The Bonferroni threshold is indicated by a red line (2.54 \texttimes{} 10\textsuperscript{-7} for SNPs, 1.53 \texttimes{} 10\textsuperscript{-6} for genes).
-\textbf{(B)} \cazcom{Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.}{You only explain how to interpret Pearson correlation in Figure 3, although it should maybe come in the Methods when you introduce stability} \textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in GENESIS (left) and BCAC (middle), and breast cancer susceptibility genes (right, only at the gene-level).
+\textbf{(B)} Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
+\textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
 For reference, we added a gray line with a slope of 1.
 \textbf{(D)} Solution networks produced by the different methods.}
+  \label{fig:solution_overview}
 \end{figure}
@@
 As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
-\cazcom{TODO Smoothen the transition?}{These solutions differ in many aspects.
-+ go aspect by aspect (size; non-coding regions; topological properties; enriched pathways?} The 4-gene solution selected by heinz includes the breast cancer susceptibility gene \emph{TOX3}, in region 16q12.
-By dealing with SNP networks, SConES studies the association of non-coding regions, as well as SNPs in any gene, coding or not.
-In fact, SConES GI retrieves 4 subnetworks in intergenic regions, and 1 overlapping an RNA gene (\emph{RNU6-420P}).
-SigMod, despite being related to SConES, produces a vastly different, large solution.
-A pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) links different parts of the network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (adjusted P-values~<~0.03).
+These solutions differ in many aspects, making it hard to draw joint conclusions.
+For starters, the overlap between the genes featured in each solution is quite small (Figure~\ref{fig:solution_overview}B).
+Another prominent difference is their sizes: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
+While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
+Their topologies also differ, as measured by the median centrality (Table~\ref{tab:gene_solutions}) and the number of connected components (Figure~\ref{fig:solution_overview}D).
+Only two methods have more than one connected component: SConES, as described above, and HotNet2.
+HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
+The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
+Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
+It linked different parts of SigMod's soluton network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (adjusted P-values~<~0.03).
 Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, adjusted P-value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (adjusted P-values~<~0.49), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-Lastly, HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
-The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
-As with SigMod, the genes are involved in mitochondrial translation (adjusted P-value~=~1.87 \texttimes{} 10\textsuperscript{-4}), and also in glycogen metabolism and transcription of nuclear receptors (adjusted P-value~<~0.04).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (adjusted P-values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (adjusted P-value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (adjusted P-value~<~0.04).
 
-As it can be seen, the solutions are very heterogeneous, making it hard to draw joint conclusions.
-\caz{In fact, the specific genes by each solution are also different}{The overlap between the genes featured in each solution is quite small} (Figure~\ref{fig:solution_overview}B).
-Another prominent difference is their sizes: \caz{HotNet2 produced the largest solution (440 genes)}{the largest solution, produced by HotNet2, contains 440 genes,} while SConES GI failed to recover any protein coding gene.
-Also their topologies differ, as measured by the median centrality and the number of connected components (Figure~\ref{fig:solution_overview}D).
-
-Yet, some common themes exist \cazcom{(Table~\ref{tab:gene_solutions})}{Why this reference here?}.
-All obtained solution subnetworks have, \cazcom{on average}{1) you used a median; 2) I'm still wondering if it wouldn't make more sense to report mean or median -log10(p)?
-The conclusion will still hold}, lower association P-values than the whole PPIN (median P-value $\ll 0.46$), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
+Despite their differences, some common themes exist as well.
+All obtained solution subnetworks have lower association P-values than the whole PPIN (median P-value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
@@
 Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
-However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by any network methods.
-\cazcom{Hence, further work is needed characterize the impact of centrality on these methods' outputs.}{Maybe more of a Discussion than a Results sentence?}
 
@@
   \includegraphics[width=.5\linewidth]{./figures/figure_3.pdf}
-  \caption{\label{fig:consensus}
-    Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
+  \caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
 Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the genes having a VEGAS2v2 P-value < 0.001, the two most central genes (\emph{COPS5} and \emph{OFD1}) and those that are known breast cancer susceptibility genes.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
 The latter ones are also colored in pink.}
+  \label{fig:consensus}
 \end{figure}
@@
 A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions.
-\cazcom{After tackling the overlap between the significant pathways}{I'm not sure what you mean here?}, we found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
+We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
 The former is supported by genes like \emph{MRPS30} (VEGAS P-value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
 Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
-With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: \cazcom{HSPA1A, HSPA1B, and HSPA1L}{Italics for gene/protein names?
-Or is it only for gene names?
-But HSPA1A is in italics below.}.
-The genes encoding these proteins are all near each other at \cazcom{6p21}{Isn't that the HLA complex?
-I'm always a bit wary of this region because of the strong LD, disantangling signals is hard...}.
+With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
+The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
 In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
@@
 In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is \emph{COPS5}, a gene related to multiple hallmarks of cancer and which is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+For instance, the most central node in the consensus network is \emph{COPS5}, whose protein is a component of the COP9 signalosome, which regulates multiple signalling pathways.
+\emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
 Despite its lack of association in GENESIS or BCAC (VEGAS2v2 P-value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS2v2 P-value~=~0.006).
 
-\subsection{Methods are comparably unstable, and produce similar predictors}
-\cazcom{}{1) ``Network methods'' rather than ``Methods'' in title?
-2) I'm not sure we want to emphasize predictor similarity when they're all so terrible 3) Shouldn't this section appear before the previous one?
-First talk about the individual solutions, then present the consensus?.}
+\subsection{Network methods boost biomarker discovery}
+
+We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
+This comparison is pertinent as, although they do not target the same population(s), we expect a shared genetic architecture at the gene level, at which most examined network metods operate.
+This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
+We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
+The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P-value \textless~0.019).
+The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) (Figure~\ref{fig:solution_overview}C).
+However, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+
+\subsection{Network methods share limitations}
 
@@
 \includegraphics[width=.72\linewidth]{./figures/figure_4.pdf}
-\caption{\label{fig:benchmark}
-Comparison of network-based GWAS methods on GENESIS.
+\caption{Comparison of network-based GWAS methods on GENESIS.
 Each method was run 5 times on a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
@@
 For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2v2 took on average to compute the gene scores from SNP summary statistics.}
+\label{fig:benchmark}
 \end{figure}
 
-We compared the six \caz{}{network} methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
 Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
-Both solution size and the subset of it selected by the classifier (\emph{active set}) varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
+The solution size varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
 Heinz produced the smallest solutions, with an average of 182 selected SNPs.
 The largest solutions come from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
-Interestingly, \cazcom{heinz has the highest proportion of the selected SNPs that go into the active set (99.9\%)}{Isn't that because it's only selecting 4 genes?}, although this proportion is high for all the methods (> 86\%).
 To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
@@
 Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-\cazcom{}{Do we want to comment this suggest that in any case there isn't that much signal in the data set, or are we shooting ourselves in the foot doing that?}
 
@@
 
-\subsection{The consensus overcomes the problems of the individual solutions}
+\subsection{The consensus overcomes the limitations of the individual solutions}
 \label{results:drawbacks}
@@
 \includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
-\caption{\label{fig:issues}
-Drawbacks encountered when using network guided methods.
+\caption{Drawbacks encountered when using network guided methods.
 \textbf{(A)} dmGWAS solution subnetwork.
@@
 \textbf{(C)} Genomic regions where either SConES GS, GM or GI select SNPs.}
+\label{fig:issues}
 \end{figure}
 
-\cazcom{In practice, and despite their similarities and their involvement in cancer mechanisms, the solutions are remarkably different.
-That is due to the particularities of the methods which directly or indirectly provide information about the dataset.
-For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.}{Slightly redundant with the ``A case study: the consensus network'' section, and also maybe move the sentence about LEAN to that section?}
-
-\cazcom{}{Also the first few paragraphs of this section are somewhat redundant with the ``A case study: the consensus network'' section?}
+As shown above, and despite their similarities, the solutions remain remarkably different.
+This is due to the particularities of the methods, which directly or indirectly provide information about the dataset.
+For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
 
@@
 Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after that transformation.
-This small solution does not provide much insight of the biology of cancer.
+Yet, this small solution does not provide much insight into the biology of cancer.
 Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
@@
 lower \(\lambda\), Section~\ref{methods:methods}), it is unclear what is the best strategy to find them.
-In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hurting the accuracy with which the GS network captures the genome structure.
+In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
 
-\subsection{Network methods boost biomarker discovery}
+\subsection{Adjusting for instability leads to similar conclusions}
 
-We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
-This comparison is pertinent as, despite not targeting the same population(s), we expect a shared genetic architecture at the gene level % \cite{required}
-TODO CAN'T FIND A CITATION FOR THIS, at which most examined network metods operate.
-This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
-We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
-The solutions provided by the different network approaches overlap significantly (Fisher's exact test P-value \textless~0.019).
-The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) (Figure~\ref{fig:solution_overview}C).
-However, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+Most of the network methods, even the consensus, were highly unstable, raising questions about the veracity of the results.
+We built a consensus network using the 30 solutions from the 6 different methods on 5 different splits containing 80\% of the data (Section \ref{methods:comparison}).
+Such network would capture the subnetworks more often found altered, and hence should be more resistant to noise.
+We used only genes selected in at least 7 of the settings (percentile 99\%).
+The resulting stablility-based consensus was composed of 68 genes.
+This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P-value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P-values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively), the selected genes are more central than average (P-value = 1.1 \texttimes{} 10\textsuperscript{-14}), and 19 of them are isolated.
+
+However, the shared global properties of the network, the lack of stability results in different genes being selected.
+In this case, the most central gene of the network is \emph{CUL3}, which is absent from the consensus network.
+Indeed, \emph{CUL3} has a low association score in both GENESIS and BCAC (P-values of 0.04 and 0.26, respectively).
+\emph{CUL3} is a component of Cullin-RING ubiquitin ligases.
+It impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
 
@@
 
-In this article we evaluate the viability of a systems biology take on genetic studies by examining a GWAS dataset on familial breast cancer.
-Such an approach addresses two of the largest issues with GWAS: interpretability and an overly conservative statistical framework that hinders discovery.
-This is achieved by considering the biological context of each of the genes and SNPs.
-Based on divergent considerations of what the desired set of biomarkers is, several methods for network-guided biomarker discovery have been proposed.
-We reviewed six of them.
-Most of them produced a relevant subset of biomarkers, although the solutions were notably different.
-In one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
+In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been put into questions \cite{boyle_expanded_2017}.
+The omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
+The practical consequence is that larger and larger GWAS will lead to the discovery an uninformative wide-spread pleiotropy.
+Other than such problems, discovery in GWAS is hindered by a conservative statistical framework.
+Network methods tackle both issues by using both the association score and an interaction networks to consider the biological context of each of the genes and SNPs.
+Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
+In this article we evaluate the relevance of six network methods by examining a GWAS dataset on familial breast cancer.
+
+Most of these methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
+On median, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
+However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by these network methods.
+Hence, further work is needed characterize the impact of centrality on these methods' outputs.
+Despite their similarities, the solutions were notably different.
+At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
 On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which risk oversimplifying their richness.
+While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
-Despite these differences, all the solutions produced comparably poor discrimination capabilities on a linear classifier trained on them, and were remarkably unstable, and hence likely to produce change in face of slightly different inputs.
+Despite these differences, all solutions are equally bad at discriminating cases from controls.
+In addition, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
+This might partly be caused by the instability of the P-values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
+Hence, heinz's conservative transformation of P-values, which favors only the most extreme ones, leads to improved stability.
+Another source of instability might be the redundancy inherent to biological networks.
+We built a stable consensus, which accounts for that instability.
+Encouragingly, this network shared the same biological and topological properties as the individual solutions, despite the few genes in common due to instability.
+Overall, this supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.
+\cazcom{}{I think we need to discuss this.
+Why aren't we able to classify cases from controls (multiple non-genetic factors, disease heterogeneity...)?
+Why is our approach still useful if we aren't, in fact, discovering biomarkers?
+(it points at relevant pathways, etc) }
 
-To overcome the problems posed by the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
-We use a straightforward aggregation to generate it, including any node that was recovered by at least two methods.
-The resulting network is a synthesis of the altered mechanism: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
+To overcome the limitations of the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
+We use a straightforward aggregation, including any node that was recovered by at least two methods.
+The resulting network is a synthesis of the altered mechanisms: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
 The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
@@
 It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
+Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
 This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
-Lastly, all the methods rely heavily on how SNPs are mapped to genes.
+Lastly, all the methods rely heavily on.
 In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+In fact, the presented case is paradigmatic, since the genes are in the most gene-dense region of the genome \cite{Xie2003}.
+Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the network, but might be more resilient to them when the overlapping genes are unrelated.
+More targeted mappings of SNPs to genes (e.g.
+eQTLs, SNPs associated to the expression of a gene), altogether with a stringent LD pruning, might address such problems.
 
@@
 In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns of some authors about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
+This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
 Their presense might bias the results of the presented methods, as they change the topology of the network by adding nodes closer to other nodes than average.
@@
 In this work we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms might have more statistical power.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
 Another important decision is how to handle LD in a GWAS.
@@
 H.C-G.
-conducted all the final analyses and the writing of the manuscript.
+conducted all the final analyses and the wrote the manuscript.
 C.L.
-conducted analyses using LEAN, Sigmod, and dmGWAS, and provided help and feedback.
+conducted analyses using LEAN, Sigmod, and dmGWAS, and provided technical and scientific help and feedback.
 F.L.
@@
 
-\cazcom{}{TODO Clean out references a bit (remove URLs, shorten long authors lists with et al., etc.)}
-
 \bibliographystyle{abbrvnat}
@@
 \begin{threeparttable}
-\caption{\label{tab:snp_solutions}
-Summary statistics on the results of SConES on the three SNP-SNP interaction networks. The first row within each block contains the summary statistics on the whole network.}
+\caption{Summary statistics on the results of SConES on the three SNP-SNP interaction networks. The first row within each block contains the summary statistics on the whole network.}
+\label{tab:snp_solutions}
 \centering
@@
 \begin{threeparttable}
-  \caption{\label{tab:scones_gene_solutions}
-Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
+  \caption{Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
+\label{tab:scones_gene_solutions}
 \centering
@@
 \includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
-\caption{\label{sfig:pcs} GENESIS shows no differential population structure between cases and controls.
+\caption{GENESIS shows no differential population structure between cases and controls.
 \textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
 Cases are colored in green, controls in orange.}
+\label{sfig:pcs}
 \end{figure}
@@
   \includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
-  \caption{\label{sfig:snp_gene_manhattan} Association in GENESIS.
+  \caption{Association in GENESIS.
 The red line represents the Bonferroni threshold.
@@
 The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).}
+  \label{sfig:snp_gene_manhattan}
 \end{figure}
@@
 \includegraphics[width=.9\linewidth]{./figures/sfigure_3.pdf}
-\caption{\label{sfig:pearson_methods}
-Pearson correlation between the different solution subnetworks.
+\caption{Pearson correlation between the different solution subnetworks.
 \textbf{(A)} Correlation between selected SNPs.
@@
 In general, the solutions display a very low overlap.}
+\label{sfig:pearson_methods}
 \end{figure}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_4.pdf}
-\caption{\label{sfig:consensus_names}
-Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
+\includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
+\caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
 \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-The labeled genes have a VEGAS2v2 P-value < 0.001 and/or are known breast cancer susceptibility genes (colored in pink).
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+The latter ones are also colored in pink.
 This panel is identical to Figure~\ref{fig:consensus}.
-\textbf{(B)} Manhattan plot showing the genes included in the subnetwork.
-\textbf{(C)} Every gene name is indicated.
-\textbf{(D,E)} Proportion of the Bonferroni significant genes (in GENESIS and BCAC, respectively) included in the consensus network.
-\textbf{F} Proportion of the selected genes by each of the methods on the GENESIS data that is a known breast cancer susceptibility gene.}
+\textbf{(B)} Same network, but every gene name is indicated.}
+\label{sfig:consensus_names}
 \end{figure}
@@
 \includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
-\caption{\label{sfig:consensus_stats}
-Genes on the consensus network.
+\caption{Genes on the consensus network.
 Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
@@
 The blue line represents a fitted generalized linear model.}
+\label{sfig:consensus_stats}
 \end{figure}
@@
 \includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
-\caption{\label{sfig:biotypes_excluded}
-Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
+\caption{Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
+\label{sfig:biotypes_excluded}
 \end{figure}
@@
 \includegraphics[width=.85\linewidth]{./figures/sfigure_7.pdf}
-\caption{\label{sfig:lc_ht_comparison}
-Comparison of benchmark on high-throughput interactions to benchmark on both high-throughput and literature curated interactions.
-Grey lines represent no change between the benchmarks (1 for ratios, 0 for differences).
+\caption{Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
+Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
 \textbf{(A)} Ratios of the selected features between both benchmarks and of the active set.
@@
 \textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) \caz{which took VEGAS2v2}{that VEGAS2v2 took} on average to compute the gene scores from SNP summary statistics.
+For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2v2 took on average to compute the gene scores from SNP summary statistics.
 In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}
+\label{sfig:lc_ht_comparison}
 \end{figure}
@@
   \centering
-  \includegraphics[width=.9\linewidth]{./figures/sfigure_8.png}
-  \caption{\label{sfig:scones_gsm} TODO.}
+  \includegraphics[width=\linewidth]{./figures/sfigure_8.png}
+  \caption{Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.
+\textbf{(A)} Manhattan plots of SNPs; in black,the method’s solution.
+The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
+\textbf{(B)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
+For reference, we added a gray line with a slope of 1.
+\textbf{(C)} Solution networks.}
+  \label{sfig:scones_gsm}
 \end{figure}

### 2020-04-20 manuscript/manuscript.tex
feedback on stability network

@@
 Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$) and occupying more central positions in the network than average.
-Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
+\cazcom{Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer}{Update to talk about the stable consensus and \textit{CUL3} instead?
+If we find it more reliable?}.
 Finally, we find a significantly large overlap between the genes in the consensus network and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
@@
 
-Most of the network methods, even the consensus, were highly unstable, raising questions about the veracity of the results.
-We built a consensus network using the 30 solutions from the 6 different methods on 5 different splits containing 80\% of the data (Section \ref{methods:comparison}).
-Such network would capture the subnetworks more often found altered, and hence should be more resistant to noise.
-We used only genes selected in at least 7 of the settings (percentile 99\%).
-The resulting stablility-based consensus was composed of 68 genes.
-This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P-value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P-values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively), the selected genes are more central than average (P-value = 1.1 \texttimes{} 10\textsuperscript{-14}), and 19 of them are isolated.
+Most of the network methods, even the consensus, were highly unstable, raising questions about the \caz{veracity}{reliability} of the results.
+We built a \caz{}{new, stable} consensus network using the 30 solutions \caz{from}{obtain by running} the 6 different methods on 5 different splits \caz{}{each} containing 80\% of the data (Section \ref{methods:comparison}).
+Such a network \caz{would}{is expected to} capture the subnetworks more often found altered, and hence should be more resistant to noise.
+We used only genes selected in at least 7 of the settings (\cazcom{percentile 99\%}{I'm reluctant to talk about percentiles here, as the genes aren't ranked with each a different value -- I'd rather say ``corresponding to selecting 1\% of the genes'' or something like that}).
+The resulting stablility-based consensus was composed of 68 genes \cazcom{}{Figure???
+Also add  a figure with all gene names in the supplementaries, like Supp Fig 4?}.
+This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P-value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P-values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P-value = 1.1 \texttimes{} 10\textsuperscript{-14}); and \cazcom{19 of them are isolated}{what property of the previous consensus network is this?}.
 
-However, the shared global properties of the network, the lack of stability results in different genes being selected.
-In this case, the most central gene of the network is \emph{CUL3}, which is absent from the consensus network.
+However, \caz{the shared global properties of the network}{although this new network exhibits similar global properties as the previous one}, the lack of stability results in different genes being selected.
+In this case, the most central gene of the network is \emph{CUL3}, which is absent from the \caz{}{previous} consensus network.
 Indeed, \emph{CUL3} has a low association score in both GENESIS and BCAC (P-values of 0.04 and 0.26, respectively).
@@
 
-In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been put into questions \cite{boyle_expanded_2017}.
+In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
 The omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
-The practical consequence is that larger and larger GWAS will lead to the discovery an uninformative wide-spread pleiotropy.
-Other than such problems, discovery in GWAS is hindered by a conservative statistical framework.
-Network methods tackle both issues by using both the association score and an interaction networks to consider the biological context of each of the genes and SNPs.
+The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
+\caz{Other than such problems}{In addition}, discovery in GWAS is hindered by a conservative statistical framework.
+Network methods tackle \caz{both}{these two} issues by using both the association score and an interaction network to \caz{consider}{take into consideration} the biological context of each of the genes and SNPs.
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
@@
 Most of these methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
-On median, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
+\caz{On median}{In general} \cazcom{}{I don't think ``On median'' exists, unfortunately}, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
 However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by these network methods.
-Hence, further work is needed characterize the impact of centrality on these methods' outputs.
+Hence, further work is needed to characterize the impact of centrality on these methods' outputs.
 Despite their similarities, the solutions were notably different.
@@
 Another source of instability might be the redundancy inherent to biological networks.
-We built a stable consensus, which accounts for that instability.
+\cazcom{We built a stable consensus, which accounts for that instability.
 Encouragingly, this network shared the same biological and topological properties as the individual solutions, despite the few genes in common due to instability.
-Overall, this supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.
+Overall, this supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.}{This comes too early, the next paragraph discusses why and how we built the consensus network and this bit should only arrive afterwards}
 \cazcom{}{I think we need to discuss this.
@@
 This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
-Their presense might bias the results of the presented methods, as they change the topology of the network by adding nodes closer to other nodes than average.
+Their presence might bias the results of the presented methods, as they change the topology of the network by adding nodes closer to other nodes than average.
 On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.

### 2020-04-22 manuscript/manuscript.tex
addressed chloé's comments

--- 2020-04-14 manuscript/manuscript.tex
@@
 By trading statistical stringency for biological meaningfulness, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-Interestingly, a combination of solution subnetworks provided a concise subnetwork of 93 genes, enriched in known breast cancer susceptibility genes (\emph{BABAM1}, \emph{BLM}, \emph{CASP8}, \emph{FGFR2}, and \emph{TOX3}, Fisher's exact test P-value~=~$7.8 \times 10^{-5}$) and occupying more central positions in the network than average.
-Additionally, it shows a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-Finally, we find a significantly large overlap between the genes in the consensus network and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
+Importantly, we find a significantly large overlap between the genes in the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
+For instance, we show a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
+Yet, network methods are notably unstable, producing different results when the input data changes.
+Hence, we produce a stable consensus subnetwork, formed by the most consistently selected genes.
+The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P-value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
+The network seems organized around \emph{CUL3}, a ubiquitin ligase related that regulates the protein levels of several genes involved in cancer progression.
 In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
@@
 We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
-The latter ones are also colored in pink.}
+The latter ones are also colored in pink.
+All gene names are indicated in Supplementary Figure \ref{sfig:consensus-names}.}
   \label{fig:consensus}
@@
 
-\subsection{Adjusting for instability leads to similar conclusions}
-
-Most of the network methods, even the consensus, were highly unstable, raising questions about the veracity of the results.
-We built a consensus network using the 30 solutions from the 6 different methods on 5 different splits containing 80\% of the data (Section \ref{methods:comparison}).
-Such network would capture the subnetworks more often found altered, and hence should be more resistant to noise.
-We used only genes selected in at least 7 of the settings (percentile 99\%).
-The resulting stablility-based consensus was composed of 68 genes.
-This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P-value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P-values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively), the selected genes are more central than average (P-value = 1.1 \texttimes{} 10\textsuperscript{-14}), and 19 of them are isolated.
-
-However, the shared global properties of the network, the lack of stability results in different genes being selected.
-In this case, the most central gene of the network is \emph{CUL3}, which is absent from the consensus network.
+\subsection{Adjusting for instability preserves global network properties}
+\label{results:stable-consensus}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.5\linewidth]{./figures/figure_5.pdf}
+  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
+Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+The latter ones are also colored in pink.
+All gene names are indicated in Supplementary Figure \ref{sfig:stable-consensus-names}.}
+  \label{fig:stable-consensus}
+\end{figure}
+
+Most of the network methods, even the consensus, were highly unstable, raising questions about the reliability of the results.
+We built a new, stable consensus network using the 30 solutions obtained by running the 6 different methods on 5 different splits each containing 80\% of the data (Section \ref{methods:comparison}).
+Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
+We used only genes selected in at least 7 of the settings, corresponding to selecting 1\% of the genes.
+The resulting stablility-based consensus was composed of 68 genes (Figure \ref{fig:stable-consensus}).
+This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P-value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P-values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P-value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
+
+However, although this new network exhibits similar global properties as the previous one, the lack of stability results in different genes being selected.
+In this case, the most central gene of the network is \emph{CUL3}, which is absent from the previous consensus network.
 Indeed, \emph{CUL3} has a low association score in both GENESIS and BCAC (P-values of 0.04 and 0.26, respectively).
@@
 
-In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been put into questions \cite{boyle_expanded_2017}.
+In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
 The omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
-The practical consequence is that larger and larger GWAS will lead to the discovery an uninformative wide-spread pleiotropy.
-Other than such problems, discovery in GWAS is hindered by a conservative statistical framework.
-Network methods tackle both issues by using both the association score and an interaction networks to consider the biological context of each of the genes and SNPs.
+The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
+In addition, discovery in GWAS is hindered by a conservative statistical framework.
+Network methods tackle these two issues by using both the association score and an interaction network to take into consideration the biological context of each of the genes and SNPs.
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
@@
 Most of these methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
-On median, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
+In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
 However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by these network methods.
-Hence, further work is needed characterize the impact of centrality on these methods' outputs.
+Hence, further work is needed to characterize the impact of centrality on these methods' outputs.
 Despite their similarities, the solutions were notably different.
@@
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
-Despite these differences, all solutions are equally bad at discriminating cases from controls.
-In addition, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
+To overcome the limitations of the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
+We use a straightforward aggregation, including any node that was recovered by at least two methods.
+The resulting network is a synthesis of the altered mechanisms: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
+The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
+However, thanks to its smaller size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.
+
+Despite these differences, all solutions share two drawbacks.
+First, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
 This might partly be caused by the instability of the P-values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
@@
 Overall, this supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.
-\cazcom{}{I think we need to discuss this.
-Why aren't we able to classify cases from controls (multiple non-genetic factors, disease heterogeneity...)?
-Why is our approach still useful if we aren't, in fact, discovering biomarkers?
-(it points at relevant pathways, etc) }
-
-To overcome the limitations of the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
-We use a straightforward aggregation, including any node that was recovered by at least two methods.
-The resulting network is a synthesis of the altered mechanisms: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
-The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
-However, thanks to its smaller size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.
+Second, all solutions are equally bad at discriminating cases from controls.
+Yet, the classification accuracy of network methods is similar to that of a classifier, which suggests that cases and controls are difficult to separate in this dataset.
+This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
 
@@
 \textbf{(B)} Same network, but every gene name is indicated.}
-\label{sfig:consensus_names}
+\label{sfig:consensus-names}
 \end{figure}
@@
   \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
+  \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf}
+  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
+\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+The latter ones are also colored in pink.
+This panel is identical to Figure~\ref{fig:stable-consensus}.
+\textbf{(B)} Same network, but every gene name is indicated.}
+  \label{sfig:stable-consensus-names}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
 \caption{Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
@@
 \centering
-\includegraphics[width=.85\linewidth]{./figures/sfigure_7.pdf}
+\includegraphics[width=.85\linewidth]{./figures/sfigure_8.pdf}
 \caption{Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
@@
   \centering
-  \includegraphics[width=\linewidth]{./figures/sfigure_8.png}
+  \includegraphics[width=\linewidth]{./figures/sfigure_9.png}
   \caption{Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.

### 2020-04-22 manuscript/manuscript.tex
where to comment on predictive power

@@
 To obtain a baseline, we also trained the classifier on all the SNPs.
+\caz{}{We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
+Indeed, [TODO : can you add info about heritability of non-BRCA familial breast cancer, explain that only part of this heritability is going to be genetic, and conclude that we do not expect to look at a woman's SNPs and tell whether she has cancer or not?].
+However, classification accuracy may be one additional informative criterion on which to evaluate solutions}.
 
@@
 Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
+\cazcom{}{TODO : Add a sentence to the effect of ``Although these classification performance may seem low, remember we are not really hoping to separate cases from controls well from SNP data''?}
 

### 2020-04-26 manuscript/manuscript.tex
final proofread of the article

@@
 \usepackage{lmodern} % Chloe: better font rendering
+\usepackage[shortcuts]{extdash} 
 \newcommand{\mean}[1]{$\overline{\mbox{#1}}$}
@@
 Indeed, it allows to jointly consider the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study the utility of six methods that identify subnetworks with higher association scores to a phenotype to discover new biomarkers for breast cancer susceptibility.
-We interrogate a familial breast cancer genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women.
+In this work, we study six network methods which identify subnetworks with high association scores to a phenotype.
+Specifically, we examine their utility to discover new biomarkers for breast cancer susceptibility by interrogating a genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women with a family history.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
 By trading statistical stringency for biological meaningfulness, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+For instance, we show a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
 Importantly, we find a significantly large overlap between the genes in the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
-For instance, we show a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-Yet, network methods are notably unstable, producing different results when the input data changes.
-Hence, we produce a stable consensus subnetwork, formed by the most consistently selected genes.
-The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P-value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
+Yet, network methods are notably unstable, producing different results when the input data changes slightly.
+To account for that, we produce a stable consensus subnetwork, formed by the most consistently selected genes.
+The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
 The network seems organized around \emph{CUL3}, a ubiquitin ligase related that regulates the protein levels of several genes involved in cancer progression.
@@
 based on logistic regression).
-Those SNPs with a P-value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
+Those SNPs with a P\=/value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
 Once the risk SNPs have been discovered, they can be used for risk assessment, and to deepen our understanding of the disease.
@@
 We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
-After a classical GWAS approach, we use these network-based methods to recover additional breast cancer biomarkers.
+After a classical GWAS approach, we use these network methods to recover additional breast cancer biomarkers.
 Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus network of predisposition to familial breast cancer.
@@
 
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P-value \textless~0.001), and those missing on more than 10\% of the samples.
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those missing on more than 10\% of the samples.
 A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
 In addition, we removed the samples with more than 10\% missing genotypes.
-After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, and 3 additional controls having a high relatedness score).
+After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
 Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis).
@@
 To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
-Then, we used VEGAS2v2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P-values of the SNPs mapped to them.
-Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P-values.
+Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
+Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P\=/values.
 We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
@@
 
-In addition, we use several matrices that describe different properties of a graph.
+In addition, we use two matrices that describe different properties of a graph.
 The described matrices are square, and have as many rows and columns as nodes are in the network.
@@
 In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
-The properties of these three subnetworks is available in Supplementary Table \ref{tab:snp_solutions}.
+The properties of these three subnetworks are available in Supplementary Table \ref{tab:snp_solutions}.
 
@@
         \item \textbf{\emph{Field}}: field in which the algorithm was developed.
-\textbf{\emph{Nodes}}: the type of network, either gene (protein-protein interaction network usually) or a SNP network.
+\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
 \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
-\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than being dense in high scores and connected.
+\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than containing high scoring, connected nodes.
 \textbf{\emph{Comp.}}: number of connected components in the solution.
 \textbf{\emph{Input}}: genotype data or GWAS summary statistics.
-\textbf{\emph{Scoring}}: how SNP/gene P-values are transformed into node scores.
+\textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
 In the case of SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
@@
 
-Genes that contribute to the same function are nearby in the PPIN, might be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
+Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
 But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
@@
 \begin{description}
-\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P-values \cite{jia_dmgwas:_2011}.
+\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
 To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
@@
 \item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for the current subgraph $S_i$ as
+\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
 
 \begin{equation*} 
-Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j
+Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
 \end{equation*}
 
-where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P-value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
 \item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
@@
 \begin{equation*}
-Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)}
+Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
 \end{equation*}
@@
 We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
-We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% modules into the solution subnetwork.
+We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
 \end{description}
@@
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subgraph on the network \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P-value into a score that is negative under no association with the phenotype, and positive when there is.
-This transformation is achieved by modelling the distribution of P-values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
+The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
+This transformation is achieved by modelling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
 Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
@@
 where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-For any two nodes/genes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
+For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
 To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
 Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
-We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P-value) to those likely to be.
-In this dataset, the threshold separating both was a P-value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
+We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
 HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
@@
 By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
-For a particular subnetwork of size \emph{m}, the P-values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
-Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P-values lower or equal to p\textsubscript{k} under the null hypothesis.
-The minimum of these \emph{k} P-values is the score of the subnetwork.
-This score is transformed into a P-value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
-Lastly, P-values are corrected for multiple testing through a Benjamini-Hochberg correction.
+For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
+Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
+The minimum of these \emph{k} P\=/values is the score of the subnetwork.
+This score is transformed into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
+Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
 We used the implementation of LEAN from the LEANR R package \cite{leanr}.
@@
 where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
-The connectivity term penalizes disconnected solutions, with many edges between a node that is selected and a node that is not.
+The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
 Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
@@
 \begin{description}
-\item[{Consensus}] In addition, we built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
+\item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
 \end{description}
@@
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
-Pathways with an adjusted P-value \textless~0.05 were deemed significant.
+Pathways with an adjusted P\=/value \textless~0.05 were deemed significant.
 As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
@@
 
-We evaluated multiple properties of the different method through a 5-fold subsampling setting.
+We evaluated multiple properties of the different methods (described below) through a 5-fold subsampling setting.
 We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
@@
 
-We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different subsets of the data) of the different network methods.
+We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
 The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by \citet{nogueira_measuring_2016}.
@@
 We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for LD by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
+The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
 This penalty was set by cross-validation, choosing the value that minimized misclassification error.
@@
 To obtain a baseline, we also trained the classifier on all the SNPs.
-\caz{}{We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
+We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
 Indeed, [TODO : can you add info about heritability of non-BRCA familial breast cancer, explain that only part of this heritability is going to be genetic, and conclude that we do not expect to look at a woman's SNPs and tell whether she has cancer or not?].
-However, classification accuracy may be one additional informative criterion on which to evaluate solutions}.
+However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
 
@@
 Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic, while GENESIS is a homogeneous dataset focusing on the French high-risk population attending the family cancer clinics.
-Despite these differences, we expect a shared genetic architecture, especially at the gene level.
-For that purpose, we searched for associated genes as in Section~\ref{methods:node_score}.
+Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
+For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
 We used all the SNPs to measure gene association, but only those available after quality control on GENESIS for the SNP-level comparisons.
@@
 We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
-Two genomic regions have a P-value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
+Two genomic regions have a P\=/value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
@@
 
-Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P-values.
+Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
 Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
@@
 It selected 100 SNPs, both from all aforementioned regions and from other regions (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
-Yet, it is unclear why those SNPs were selected, as emphasized by the high P-value of some of them, which further complicates the biological interpretation.
-Moreover, and in opposition to what the omnigenic model predicts, the genes to which this SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
+Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
+Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which this SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
 In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
@@
 \hline
-Consensus & 55 & 117 & 74\,062 & 0.0051 & 1\\
 dmGWAS & 194 & 450 & 49\,115 & 0.19 & 0.41\\
-heinz & 4 & 3 & 113\,633 & 0.0012 & 0.21\\
+heinz & 4 & 3 & 113\,633 & 0.001 & 0.21\\
 HotNet2 & 440 & 374 & 7\,739 & 0.048 & 0.31\\
@@
 SConES GI & 0 (1) & 0 & - & - & -\\
-SigMod & 142 & 249 & 92\,603 & 0.0083 & 0.73\\
+SigMod & 142 & 249 & 92\,603 & 0.008 & 0.73\\
+\hline
+Consensus & 93 & 186 & 50\,737 & 0.006 & 1\\
+Stable consensus & 68 & 49 & 94\,854 & 0.005 & 0.54 \\
 \end{tabular}
@@
 \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
-\textbf{\median{P}\textsubscript{gene}}: median P-value of the selected genes.
+\textbf{\median{P}\textsubscript{gene}}: median P\=/value of the selected genes.
 \textbf{\(\rho\)\textsubscript{consensus}}: Pearson correlation between the subnetwork and the consensus network.
@@
   \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-As LEAN did not produce any significant gene (adjusted P-value < 0.05), it was excluded.
+As LEAN did not produce any significant gene (adjusted P\=/value~<~0.05), it was excluded.
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
 \textbf{(A)} Manhattan plots of SNPs/genes; in black, the method's solution.
-The Bonferroni threshold is indicated by a red line (2.54 \texttimes{} 10\textsuperscript{-7} for SNPs, 1.53 \texttimes{} 10\textsuperscript{-6} for genes).
+The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).
 \textbf{(B)} Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
@@
 We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (adjusted P-value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+As none of the networks examined by LEAN was significant (adjusted P\=/value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
 These solutions differ in many aspects, making it hard to draw joint conclusions.
 For starters, the overlap between the genes featured in each solution is quite small (Figure~\ref{fig:solution_overview}B).
-Another prominent difference is their sizes: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
+Another prominent difference is their size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
 While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
@@
 Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
-It linked different parts of SigMod's soluton network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (adjusted P-values~<~0.03).
-Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, adjusted P-value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (adjusted P-values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (adjusted P-value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (adjusted P-value~<~0.04).
-
-Despite their differences, some common themes exist as well.
-All obtained solution subnetworks have lower association P-values than the whole PPIN (median P-value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P-values as well (Figure~\ref{fig:solution_overview}A).
+It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (adjusted P\=/values~<~0.03).
+Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, adjusted P\=/value~=~0.01).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (adjusted P\=/value~<~0.04).
+
+Despite their differences, there are additional common themes.
+All obtained solution subnetworks have lower association P\=/values than the whole PPIN (median P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Figure~\ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions: heinz strongly favors genes with lower P-values, while dmGWAS is less conservative (median P-values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
-Additionally, the solution subnetworks present other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P-value < 0.03).
-Second, the genes in four solution subnetworks display on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P-value < 1.4 \texttimes{} 10\textsuperscript{-21}).
-This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P-value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
+However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+Additionally, the solution subnetworks presented other desirable properties.
+First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value < 0.03).
+Second, the genes in four solution subnetworks displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value < 1.4 \texttimes{} 10\textsuperscript{-21}).
+This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
 Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
@@
 Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
-Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods select a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
+Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
 To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
-This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and shares the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
+This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
 
@@
 We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
-The former is supported by genes like \emph{MRPS30} (VEGAS P-value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
+The former is supported by genes like \emph{MRPS30} (VEGAS2 P\=/value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
 Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
@@
 The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their association.
-\emph{HSPA1A} was the most strongly associated one (VEGAS P-value~=~8.37 \texttimes{} 10\textsuperscript{-4}).
+In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their effects.
+\emph{HSPA1A} was the most strongly associated gene (VEGAS2 P\=/value~=~8.37~\texttimes{}~10\textsuperscript{-4}).
 
 Topologically the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
-Among the latter, 14 of the 93 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by two methods.
-The opposite would be the case of genes with highly centrality in the PPIN, which is weakly anti-correlated with the P-value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
+Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
+The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
 This suggests that they were selected because they were on the shortest path between two highly associated genes.
-In view of this, we hypothesize that highly central genes might contribute to the heritability through consistent alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is \emph{COPS5}, whose protein is a component of the COP9 signalosome, which regulates multiple signalling pathways.
+In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
+For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signalling pathways.
 \emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS or BCAC (VEGAS2v2 P-value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P-values (median VEGAS2v2 P-value~=~0.006).
+Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
 
@@
 We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
-This comparison is pertinent as, although they do not target the same population(s), we expect a shared genetic architecture at the gene level, at which most examined network metods operate.
+Although performed on different populations, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
 We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
-The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P-value \textless~0.019).
-The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) (Figure~\ref{fig:solution_overview}C).
-However, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
+The gene-based network methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (Figure~\ref{fig:solution_overview}C).
+Interestingly, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
 
@@
 \textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP).
-For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2v2 took on average to compute the gene scores from SNP summary statistics.}
+For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.}
 \label{fig:benchmark}
@@
 Heinz produced the smallest solutions, with an average of 182 selected SNPs.
-The largest solutions come from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
+The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
 To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
@@
 
-In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which finds the solution efficiently in a few seconds.
+In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which returned a solution in a few seconds.
 HotNet2 was the slowest (3 days and 14 hours on average).
@@
 \caption{Drawbacks encountered when using network guided methods.
-\textbf{(A)} dmGWAS solution subnetwork.
-Genes with a P-value < 0.1 are highlighted in red.
-\textbf{(B)} Centrality degree and -log\textsubscript{10} of the VEGAS2v2 P-value for the nodes in SigMod solution subnetwork.
-\textbf{(C)} Genomic regions where either SConES GS, GM or GI select SNPs.}
+\textbf{(A)}~DmGWAS solution subnetwork.
+Genes with a P\=/value < 0.1 are highlighted in red.
+\textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value for the nodes in SigMod solution subnetwork.
+\textbf{(C)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
+SNPs that were not selected in the studied network, but were selected in another one, are displayed in background color.
+}
 \label{fig:issues}
@@
 
-As shown above, and despite their similarities, the solutions remain remarkably different.
-This is due to the particularities of the methods, which directly or indirectly provide information about the dataset.
-For instance, the fact that LEAN did not provide any biomarkers implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
-
-In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P-value for the subnetwork (Table~\ref{tab:gene_solutions}).
+As shown above, and despite their similarities, the network methods produced remarkably different solutions.
+This is due to the particularities of each methods, and directly or indirectly provide information about the dataset.
+For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
+
+In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value for the subnetwork (Table~\ref{tab:gene_solutions}).
 Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Figure~\ref{fig:benchmark}C).
-Its conservativeness stems from its preprocessing step, which models the gene P-values as a mixture model of a beta and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of them are retain a positive score after that transformation.
-Yet, this small solution does not provide much insight into the biology of cancer.
+Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
+Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
+Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
 Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
 
-On the other end of the spectrum, we have large solutions provided by dmGWAS, HotNet2, and SigMod.
+On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
 dmGWAS' subnetwork is the least associated subnetwork on average.
 This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
-This framework considers all nodes at distance 2 of the examined, and accepts weakly associated genes if they are linked to another, strongly associated one.
+It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
 This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure~\ref{fig:issues}A).
-SigMod displays the same tendency, as the most central genes are the least associated to the disease (Figure~\ref{fig:issues}B).
+SigMod displayed the same tendency, as the most central genes are the least associated to the disease (Figure~\ref{fig:issues}B).
 This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
@@
 
-By virtue of using a SNP subnetwork, SConES analyzes each SNP in their context.
-It therefore selects SNPs in genes none of whose interactors are associated to the disease, as well as SNPs in non-coding regions or in non-interacting genes.
-In fact, due to linkage disequilibrium, such genes are favored by SConES, as selecting SNPs in an LD block which overlaps with a gene favors selecting the rest of the gene.
+By virtue of using a SNP subnetwork, SConES analyzes each SNP in their functional context.
+It therefore can select SNPs in genes without any associated interactor, as well as SNPs in non-coding regions or in non-interacting genes.
+In fact, due to linkage disequilibrium, SConES favors such genes, as selecting SNPs in an LD-block which overlaps with a gene favors selecting the rest of the gene.
 This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Supplementary Figure~\ref{sfig:pearson_methods}).
 On the other hand, SConES penalizes selecting SNPs and not their neighbors.
-This makes it conservative regarding SNPs with many interactions, for instance those mapped to hubs in the PPIN.
+This makes it conservative regarding SNPs with many interactions, like those mapped to hub genes in the PPIN.
 For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure~\ref{fig:issues}C).
-In fact SConES GS and SConES GM select regions related to breast cancer, like 16q12 (\emph{TOX3}, Section~\ref{results:conventional}), 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), and 10q26 (\emph{FGFR2}, Section~\ref{results:conventional}).
-On top of that only SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
-We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a module.
+In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
+On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
+We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a gene and a majority of its neighbors.
 Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
-lower \(\lambda\), Section~\ref{methods:methods}), it is unclear what is the best strategy to find them.
-In addition, due to the iCOGS SNP array design, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
+a lower \(\lambda\) in Equation~\ref{eq:scones}), it is unclear what the best strategy to find them is.
+In addition, due to the design of the iCOGS array, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
 
@@
 
-Most of the network methods, even the consensus, were highly unstable, raising questions about the reliability of the results.
-We built a new, stable consensus network using the 30 solutions obtained by running the 6 different methods on 5 different splits each containing 80\% of the data (Section \ref{methods:comparison}).
+Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
+We built a new, stable consensus network using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:comparison}).
 Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
-We used only genes selected in at least 7 of the settings, corresponding to selecting 1\% of the genes.
+We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once.
 The resulting stablility-based consensus was composed of 68 genes (Figure \ref{fig:stable-consensus}).
-This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P-value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P-values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P-value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
+This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
 
 However, although this new network exhibits similar global properties as the previous one, the lack of stability results in different genes being selected.
-In this case, the most central gene of the network is \emph{CUL3}, which is absent from the previous consensus network.
-Indeed, \emph{CUL3} has a low association score in both GENESIS and BCAC (P-values of 0.04 and 0.26, respectively).
-\emph{CUL3} is a component of Cullin-RING ubiquitin ligases.
-It impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
+In this case, the most central gene is \emph{CUL3}, which is absent from the previous consensus network and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
+This gene is a component of Cullin-RING ubiquitin ligases.
+Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
 
@@
 In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
-The omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
+On the one hand, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
 The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
-In addition, discovery in GWAS is hindered by a conservative statistical framework.
+On the other hand, discovery in GWAS is hindered by a conservative statistical framework.
 Network methods tackle these two issues by using both the association score and an interaction network to take into consideration the biological context of each of the genes and SNPs.
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
-In this article we evaluate the relevance of six network methods by examining a GWAS dataset on familial breast cancer.
-
-Most of these methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
+In this article we evaluated the relevance of six of them by examining a GWAS dataset on familial breast cancer.
+
+Most of the network methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
 In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
@@
 Hence, further work is needed to characterize the impact of centrality on these methods' outputs.
-Despite their similarities, the solutions were notably different.
+Despite these similarities, the solutions were notably different.
 At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
@@
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
-To overcome the limitations of the individual methods while exploiting their strengths, we propose combining them into a consensus subnetwork.
-We use a straightforward aggregation, including any node that was recovered by at least two methods.
-The resulting network is a synthesis of the altered mechanisms: it is smaller than the largest solutions (HotNet2, SigMod and dmGWAS), which makes it more manageable, and includes the majority of the strongly associated smaller solutions (SConES and heinz).
-The consensus subnetwork captures mechanisms and genes known to be related to cancer, recovering known breast cancer susceptibility genes as well as genome regions associated to breast cancer susceptibility.
-However, thanks to its smaller size and its network structure, it provides compelling hypotheses of non-canonical mechanisms involved in carcinogenesis, like mitochondrial translation and chaperone activity.
-
-Despite these differences, all solutions share two drawbacks.
-First, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
-This might partly be caused by the instability of the P-values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
-Hence, heinz's conservative transformation of P-values, which favors only the most extreme ones, leads to improved stability.
-Another source of instability might be the redundancy inherent to biological networks.
-We built a stable consensus, which accounts for that instability.
-Encouragingly, this network shared the same biological and topological properties as the individual solutions, despite the few genes in common due to instability.
-Overall, this supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.
-Second, all solutions are equally bad at discriminating cases from controls.
+Additionally, all solutions share two drawbacks.
+First, all solutions are equally bad at discriminating cases from controls.
 Yet, the classification accuracy of network methods is similar to that of a classifier, which suggests that cases and controls are difficult to separate in this dataset.
 This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
+Second, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
+This might partly be caused by the instability of the P\=/values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
+Hence, heinz's conservative transformation of P\=/values, which favors only the most extreme ones, leads to improved stability.
+Another source of instability might be the redundancy inherent to biological networks.
+
+To overcome the limitations of the individual methods while exploiting their strengths, we proposed combining them into a consensus subnetwork.
+We use a straightforward strategy of including any node that was recovered by multiple methods.
+We proposed two networks: a consensus network that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus network, that tackled the instability of the methods.
+They both synthesized the altered mechanisms: they both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer.
+Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but who are related to cancer at the expression level and whose neighborhood has consistent high association scores.
+Crucially, the consensus was as unstable as the tested methods, while the stable consensus shared these properties while accounting for instability.
+This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.
 
@@
 Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
-In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P-value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
-However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P-value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P-values of 0.43 and 0.47).
+In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
+However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
 As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
@@
 In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}.
-Their presence might bias the results of the presented methods, as they change the topology of the network by adding nodes closer to other nodes than average.
+This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
 On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
-When we compared the impact of using a larger network containing interactions from both high-throughput experiment and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
-This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the selected solutions.
+When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
+This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the solutions.
 
 A crucial step for the gene based methods is the computation of the gene score.
-In this work we used VEGAS2v2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P-value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
+In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
 Another important decision is how to handle LD in a GWAS.
-VEGAS2v2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2v2's computation time.
+VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
 With regards to SConES, less SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
@@
 
-In order to produce the consensus network, we had to face the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
+In order to produce the consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
 To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
@@
   \footnotesize{
-    \item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.\median{P}\textsubscript{SNP}: median P-value of the selected SNPs.
+    \item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.\median{P}\textsubscript{SNP}: median P\=/value of the selected SNPs.
   }
@@
     \item \mean{Betweenness}: mean betweenness of the selected genes in the full network.
-\median{P}\textsubscript{gene}: median P-value of the selected genes; $\rho_{consensus}$: Pearson correlation with the consensus network.
+\median{P}\textsubscript{gene}: median P\=/value of the selected genes; $\rho_{consensus}$: Pearson correlation with the consensus network.
   }
@@
 The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
-\textbf{(B)} Gene association, measured by P-value of VEGAS2v2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P-values.
+\textbf{(B)} Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values.
 The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
@@
 \textbf{(A)} Number of methods selecting every gene in the subnetwork.
-\textbf{(B)} VEGAS P-values of association of the genes, with regards to the number of methods that selected them.
+\textbf{(B)} VEGAS2 P\=/values of association of the genes, with regards to the number of methods that selected them.
 \textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network.
 To improve visualization, we removed outliers.
-\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS P-value of the genes in the consensus network.
+\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus network.
 The blue line represents a fitted generalized linear model.}
@@
 \textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2v2 took on average to compute the gene scores from SNP summary statistics.
+For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
 In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}

### 2020-04-26 manuscript/manuscript.tex
addessed TODOs regarding classification performance

@@
 We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
-Indeed, [TODO : can you add info about heritability of non-BRCA familial breast cancer, explain that only part of this heritability is going to be genetic, and conclude that we do not expect to look at a woman's SNPs and tell whether she has cancer or not?].
+Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
 However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
@@
 Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-\cazcom{}{TODO : Add a sentence to the effect of ``Although these classification performance may seem low, remember we are not really hoping to separate cases from controls well from SNP data''?}
+It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
 

### 2020-05-01 manuscript/manuscript.tex
discuss centrality + address last of fabiennes comments

@@
-\documentclass[twocolumn, 11pt]{article}
+\documentclass[draft,twocolumn, 11pt]{article}
 \usepackage{amsmath,amssymb}
@@
 \author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
-GENESIS Study collaborators, Dominique Stoppa-Lyonnet$^{4,5,6}$, \\
+GENESIS Study collaborators$^{1,2,3}$, Dominique Stoppa-Lyonnet$^{4,5,6}$, \\
 Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
@@
 In this work, we study six network methods which identify subnetworks with high association scores to a phenotype.
-Specifically, we examine their utility to discover new biomarkers for breast cancer susceptibility by interrogating a genome-wide association study (GWAS) focused on \emph{BRCA1/2} negative French women with a family history.
+Specifically, we examine their utility to discover new biomarkers for breast cancer susceptibility by interrogating a genome-wide association study (GWAS) focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
 We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
@@
 The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
-The network seems organized around \emph{CUL3}, a ubiquitin ligase related that regulates the protein levels of several genes involved in cancer progression.
+The network seems organized around \emph{CUL3}, encoding an ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
 In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
@@
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
-On top of that, recent large-sampled studies suggest that most of the genome contributes to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
+On top of that, recent large-sampled studies suggest that numerous loci spread all along the genome contribute to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
 The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
@@
 This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes.
-Additionally, as genes that contribute to a disease tend to participate in similar biological functions, studying the neighborhood of known disease genes is effective at identifying new ones \cite{huang_systematic_2018}.
+Therefore, studying the neighborhood of disease-associated genes is effective at identifying new ones that are involved in the same biological functions \cite{huang_systematic_2018}.
 
@@
 The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
-It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as functional candidate variants in selected genes and pathways.
+It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
 
@@
 
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those missing on more than 10\% of the samples.
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
 A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
@@
 After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
-Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1/2} mutation detected in the family, 1 with unknown molecular diagnosis).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
 The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
@@
 We looked for population structure that could produce spurious associations.
-A PCA revealed no visual differential population structure between cases and controls (Supplementary Figure~\ref{sfig:pcs}).
+A principal component analysis revealed no visual differential population structure between cases and controls (Supplementary Figure~\ref{sfig:pcs}).
 Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
@@
 The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to scale appropriately to them.
+Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to run efficiently on networks of their size.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 SConES \cite{azencott_efficient_2013} is the only network method designed to handle SNP networks.
-As in gene networks, two SNPs are linked in a SNP network when there is evidence of shared functionality between two SNPs.
+As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
 \citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
@@
 \textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
-In the case of SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
+In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
         cumulative distribution function of the standard Normal distribution. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
@@
 \begin{description}
-\item[{heinz}] The goal of heinz is to identify the highest-scored connected subgraph on the network \cite{dittrich_identifying_2008}.
+\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
 The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
@@
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
-Pathways with an adjusted P\=/value \textless~0.05 were deemed significant.
+We adjusted the P-values for multiple testing as in \citet{Benjamini1995} (BH).Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
 As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
@@
 
-We evaluated multiple properties of the different methods (described below) through a 5-fold subsampling setting.
+We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
 We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
@@
 
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs (after imputation) genotyped on 228\,951 women of European ancestry from the general population \cite{Michailidou2017}.
-Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic, while GENESIS is a homogeneous dataset focusing on the French high-risk population attending the family cancer clinics.
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
+Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focus on the French high-risk population attending the family cancer clinics.
 Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
 For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
-We used all the SNPs to measure gene association, but only those available after quality control on GENESIS for the SNP-level comparisons.
+We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
 
@@
 
-\subsection{A conventional GWAS confirms that FGFR2 is associated with familial breast cancer}
+\subsection{Conventional SNP- and gene-based analyses confirm that \emph{FGFR2} locus is associated with familial breast cancer}
 \label{results:conventional}
@@
 We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
-Two genomic regions have a P\=/value lower than the Bonferroni threshold in chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
+Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
-Variants in both \emph{FGFR2} and \emph{TOX3} were related to breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1/2} carriers studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for \emph{BRCA1/2} \cite{rinella_genetic_2013}.
-Only \emph{FGFR2} was significant at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
-These results show an overlap in the genetic architecture of the disease between the studied French population sample and other populations.
+Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
+In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
 
@@
 We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
-It selected 100 SNPs, both from all aforementioned regions and from other regions (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
+It selected 100 SNPs, both from all aforementioned regions and new ones (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
-Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which this SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
+Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
 In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
@@
   \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-As LEAN did not produce any significant gene (adjusted P\=/value~<~0.05), it was excluded.
+As LEAN did not produce any significant gene (BH adjusted P\=/value~<~0.05), it was excluded.
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
@@
 We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (adjusted P\=/value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+As none of the networks examined by LEAN was significant (BH adjusted P\=/value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
 These solutions differ in many aspects, making it hard to draw joint conclusions.
@@
 Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
-It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (adjusted P\=/values~<~0.03).
-Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, adjusted P\=/value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (adjusted P\=/value~<~0.04).
+It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
+Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
 
@@
 Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
 The latter ones are also colored in pink.
@@
 
-We compared the results of different network methods to the European cohort of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date in breast cancer (Section~\ref{methods:bcac}).
-Although performed on different populations, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
+We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
+Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
@@
 \caption{Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times on a random subset of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
 As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
@@
 
-\subsection{The consensus overcomes the limitations of the individual solutions}
+\subsection{Network topology matters, and might lead to ambiguous results}
 \label{results:drawbacks}
@@
 Genes with a P\=/value < 0.1 are highlighted in red.
-\textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value for the nodes in SigMod solution subnetwork.
+\textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
+We highlighted the genes selected by each method, and the ones selected by more than one (``Consensus'').
+We labeled the three most central genes that were picked by any method.
 \textbf{(C)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
@@
 This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure~\ref{fig:issues}A).
-SigMod displayed the same tendency, as the most central genes are the least associated to the disease (Figure~\ref{fig:issues}B).
 This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
@@
 Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
+
+Another peculiarity of network methods is their relationship to degree centrality.
+On the one hand we observed that highly central genes often had no association to disease (Figure~\ref{fig:issues}B).
+On the other, network methods favor central genes, as they often connect high scoring nodes.
+This was specially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
+As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
+With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration.
+Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it is negatively correlated with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.
 
@@
 Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
 The latter ones are also colored in pink.
@@
 In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
-However, very central nodes are also more likely to be connecting a given random pair of nodes, making them more likely to be selected by these network methods.
-Hence, further work is needed to characterize the impact of centrality on these methods' outputs.
+However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by these network methods.
+Across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
+Yet, further work is needed to characterize the impact of centrality on network methods' outputs.
 Despite these similarities, the solutions were notably different.
@@
 Additionally, all solutions share two drawbacks.
-First, all solutions are equally bad at discriminating cases from controls.
-Yet, the classification accuracy of network methods is similar to that of a classifier, which suggests that cases and controls are difficult to separate in this dataset.
+First, they are all equally bad at discriminating cases from controls.
+Yet, the classification accuracy of network methods is similar to that of a machine learning classifier, which suggests that cases and controls are difficult to separate in this dataset.
 This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
@@
 We use a straightforward strategy of including any node that was recovered by multiple methods.
-We proposed two networks: a consensus network that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus network, that tackled the instability of the methods.
+We proposed two networks: a consensus network that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus network, that addressed the instability of the methods.
 They both synthesized the altered mechanisms: they both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer.
@@
 Crucially, the consensus was as unstable as the tested methods, while the stable consensus shared these properties while accounting for instability.
-This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence it validates the conclusions obtained on the individual solutions and the consensus.
+This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence validates the conclusions obtained on the individual solutions and the consensus.
 
 The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
-In consequence, they show their shortcomings in front of understudied genes, especially those not in the network.
+In consequence, they show their shortcomings with respect to understudied genes, especially those not in the network.
 Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
@@
 This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
-Lastly, all the methods rely heavily on.
+Lastly, all the methods are heavily affected by how SNPs are mapped to genes.
 In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
@@
 Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the network, but might be more resilient to them when the overlapping genes are unrelated.
-More targeted mappings of SNPs to genes (e.g.
+Making use of more targeted mappings of SNPs to genes (e.g.
 eQTLs, SNPs associated to the expression of a gene), altogether with a stringent LD pruning, might address such problems.
@@
 VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
-With regards to SConES, less SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
+With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
 However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
@@
 supervised the project.
+N.A and D.S-L lead the GENESIS study.
 
@@
 
-We wish to thank the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: O.
-Kulkarni, S.
+We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S.
 Eon-Marchais, M.
@@
 \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
 The latter ones are also colored in pink.
@@
 \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significant in the BCAC dataset.
+We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
 The latter ones are also colored in pink.

### 2020-08-20 manuscript/manuscript.tex
compiled final version
updated manuscript's affiliations and contributions

@@
-\documentclass[draft,twocolumn, 11pt]{article}
+\documentclass[twocolumn, 11pt]{article}
 \usepackage{amsmath,amssymb}
@@
 \title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
-\author{Héctor Climente-González$^{1,2,3}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
-GENESIS Study collaborators$^{1,2,3}$, Dominique Stoppa-Lyonnet$^{4,5,6}$, \\
+\author{Héctor Climente-González$^{1,2,3,4,\spadesuit}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
+GENESIS Study collaborators$^{1,2,3}$, Dominique Stoppa-Lyonnet$^{5,6,7}$, \\
 Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
-\date{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
+\date{\small{
+  \raggedright{
+  $^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
   $^{2}$INSERM, U900, F-75005 Paris, France;\\
   $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
-  $^{4}$Service de Génétique, Institut Curie, F-75005 Paris, France;\\
-  $^{5}$INSERM, U830, F-75005 Paris, France;\\
-  $^{6}$Université Paris Descartes.\\
+  $^{4}$RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, 103-0027, Japan; \\
+  $^{5}$Service de Génétique, Institut Curie, F-75005 Paris, France;\\
+  $^{6}$INSERM, U830, F-75005 Paris, France;\\
+  $^{7}$Université Paris Descartes.\\
+  }
+  \raggedright{$^{\spadesuit}$Corresponding author: \href{hector.climente@riken.jp}{hector.climente@riken.jp}\\}
+}
 }
@@
 
-H.C-G.
-conducted all the final analyses and the wrote the manuscript.
-C.L.
-conducted analyses using LEAN, Sigmod, and dmGWAS, and provided technical and scientific help and feedback.
-F.L.
-and N.A.
-provided feedback on the analyses and the manuscript.
-C-A.A.
-supervised the project.
-N.A and D.S-L lead the GENESIS study.
+\begin{description}
+  \item[Conceptualization] Héctor Climente-González, Christine Lonjou, Chloé-Agathe Azencott.
+  \item[Data curation] Christine Lonjou, GENESIS Study collaborators.
+  \item[Formal Analysis] Héctor Climente-González, Christine Lonjou.
+  \item[Funding acquisition] Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott.
+  \item[Investigation] Héctor Climente-González, Christine Lonjou.
+  \item[Methodology] Héctor Climente-González, Christine Lonjou, Chloé-Agathe Azencott.
+  \item[Project administration] Chloé-Agathe Azencott.
+  \item[Resources] GENESIS Study collaborators, Dominique Stoppa-Lyonnet, Nadine Andrieu.
+  \item[Software] Héctor Climente-González, Christine Lonjou.
+  \item[Supervision] Christine Lonjou, Fabienne Lesueur, Nadine Andrieu, Chloé-Agathe Azencott.
+  \item[Validation] Christine Lonjou, Fabienne Lesueur.
+  \item[Visualization] Héctor Climente-González.
+  \item[Writing – original draft] Héctor Climente-González.
+  \item[Writing – review \& editing] Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, Nadine Andrieu, Chloé-Agathe Azencott.
+\end{description}
 

### 2020-08-20 manuscript/plos/manuscript_plos.tex
commit before moving to plos template

--- 2020-08-20 manuscript/manuscript.tex
@@
 
-\title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
+\title{Comparing and combining network-guided biomarker discovery to discover genetic susceptibility mechanisms to breast cancer on the GENESIS study}
 \author{Héctor Climente-González$^{1,2,3,4,\spadesuit}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
-GENESIS Study collaborators$^{1,2,3}$, Dominique Stoppa-Lyonnet$^{5,6,7}$, \\
-Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
+Dominique Stoppa-Lyonnet$^{5,6,7,\clubsuit}$, Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
 \date{\small{
@@
   $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
-  $^{4}$RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, 103-0027, Japan; \\
+  $^{4}$RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, Japan; \\
   $^{5}$Service de Génétique, Institut Curie, F-75005 Paris, France;\\
@@
   $^{7}$Université Paris Descartes.\\
+  $^{\clubsuit}$For the GENESIS study group\\
+  $^{\spadesuit}$Corresponding author: \href{hector.climente@riken.jp}{hector.climente@riken.jp}\\
   }
-  \raggedright{$^{\spadesuit}$Corresponding author: \href{hector.climente@riken.jp}{hector.climente@riken.jp}\\}
   }
@@
 In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
-Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
+Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.\\
+
+\raggedright{\textbf{Author summary}}
+
+In genome-wide association studies (GWAS), thousands of genomes are scanned to identify variants associated with a complex trait.
+Over the last 15 years, GWAS have advanced our understanding of the genetics of complex diseases, and in particular of hereditary cancers.
+Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
+An elegant explanation has been proposed with the omnigenic model: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated, so that many other genes are able to alter the function of the core genes.
+These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
+In this study, we characterize six such network methods on GENESIS, a GWAS dataset for familial breast cancer in the French population.
+Combining these approaches allows us to identify potentially novel breast cancer susceptibility genes, and provides a mechanistic explanation for their role in the development of the disease.
+Our pipeline can easily be applied to other diseases.
 \end{abstract}
@@
 
-\section{Methods}
+\section{Results}
+
+\subsection{Conventional SNP- and gene-based analyses confirm that \emph{FGFR2} locus is associated with familial breast cancer}
+\label{results:conventional}
+
+We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
+Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
+The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
+Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
+In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
+
+Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
+Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
+We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
+It selected 100 SNPs, both from all aforementioned regions and new ones (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
+Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
+Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
+In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
+This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
+In fact, such methods provided comparably (poor) classification performance to L1-penalized logistic regression (Figure~\ref{fig:benchmark}B), while providing more interpretable solutions.
+
+\subsection{Network methods successfully identify genes associated with breast cancer}
+\label{results:separate_networks}
+\begin{table}[htbp]
+  \centering
+  \begin{threeparttable}
+\caption{Summary statistics on the solutions of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
+\label{tab:gene_solutions}
+\centering
+\begin{tabular}{lrrrrr}
+Network & \# genes & \# edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & \(\rho\)\textsubscript{consensus}\\
+\hline
+HINT HT & 13\,619 & 142\,541 & 16\,706 & 0.46 & 0.066\\
+\hline
+dmGWAS & 194 & 450 & 49\,115 & 0.19 & 0.41\\
+heinz & 4 & 3 & 113\,633 & 0.001 & 0.21\\
+HotNet2 & 440 & 374 & 7\,739 & 0.048 & 0.31\\
+LEAN & 0 & 0 & - & - & -\\
+SConES GI & 0 (1) & 0 & - & - & -\\
+SigMod & 142 & 249 & 92\,603 & 0.008 & 0.73\\
+\hline
+Consensus & 93 & 186 & 50\,737 & 0.006 & 1\\
+Stable consensus & 68 & 49 & 94\,854 & 0.005 & 0.54 \\
+\end{tabular}
+\begin{tablenotes}
+  \footnotesize{
+    \item \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
+\textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
+\textbf{\median{P}\textsubscript{gene}}: median P\=/value of the selected genes.
+\textbf{\(\rho\)\textsubscript{consensus}}: Pearson correlation between the subnetwork and the consensus network.
+  }
+\end{tablenotes}
+\end{threeparttable}
+\end{table}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[height=.75\textheight]{./figures/figure_1.png}
+  \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
+As LEAN did not produce any significant gene (BH adjusted P\=/value~<~0.05), it was excluded.
+Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
+\textbf{(A)} Manhattan plots of SNPs/genes; in black, the method's solution.
+The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).
+\textbf{(B)} Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
+\textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
+For reference, we added a gray line with a slope of 1.
+\textbf{(D)} Solution networks produced by the different methods.}
+  \label{fig:solution_overview}
+\end{figure}
+
+We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
+As none of the networks examined by LEAN was significant (BH adjusted P\=/value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+These solutions differ in many aspects, making it hard to draw joint conclusions.
+For starters, the overlap between the genes featured in each solution is quite small (Figure~\ref{fig:solution_overview}B).
+Another prominent difference is their size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
+While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
+Their topologies also differ, as measured by the median centrality (Table~\ref{tab:gene_solutions}) and the number of connected components (Figure~\ref{fig:solution_overview}D).
+Only two methods have more than one connected component: SConES, as described above, and HotNet2.
+HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
+The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
+Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
+It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
+Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
+
+Despite their differences, there are additional common themes.
+All obtained solution subnetworks have lower association P\=/values than the whole PPIN (median P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Figure~\ref{fig:solution_overview}A).
+This exemplifies the trade-off between statistical significance and biological relevance.
+However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+Additionally, the solution subnetworks presented other desirable properties.
+First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value < 0.03).
+Second, the genes in four solution subnetworks displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value < 1.4 \texttimes{} 10\textsuperscript{-21}).
+This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
+Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+
+\subsection{A case study: the consensus network}
+\label{results:consensus}
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.5\linewidth]{./figures/figure_3.pdf}
+  \caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
+Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+All gene names are indicated in Supplementary Figure \ref{sfig:consensus-names}.}
+  \label{fig:consensus}
+\end{figure}
+
+Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
+Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
+To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
+This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
+
+A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions.
+We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
+The former is supported by genes like \emph{MRPS30} (VEGAS2 P\=/value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
+Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
+With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
+The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
+In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their effects.
+\emph{HSPA1A} was the most strongly associated gene (VEGAS2 P\=/value~=~8.37~\texttimes{}~10\textsuperscript{-4}).
+
+Topologically the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
+Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
+The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
+This suggests that they were selected because they were on the shortest path between two highly associated genes.
+In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
+For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signalling pathways.
+\emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
+
+\subsection{Network methods boost biomarker discovery}
+
+We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
+Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
+This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
+We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
+The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
+The gene-based network methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (Figure~\ref{fig:solution_overview}C).
+Interestingly, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+
+\subsection{Network methods share limitations}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.72\linewidth]{./figures/figure_4.pdf}
+\caption{Comparison of network-based GWAS methods on GENESIS.
+Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
+\textbf{(A)} Number of SNPs selected by each method and number of SNPs in the active set found by the classifier.
+Points are the average over the 5 runs; lines represent the standard error of the mean.
+A grey diagonal line with slope 1 is added for comparison.
+For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs.
+\textbf{(B)} Sensitivity and specificity on test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
+In addition, the performance of the classifier trained on all SNPs is displayed.
+Points are the average over the 5 runs; lines represent the standard error of the mean.
+\textbf{(C)} Pairwise Pearson correlations of the solutions used by different methods.
+A Pearson correlation of 1 means the two solutions are the same.
+A Pearson correlation of 0 means that there is no SNP in common between the two solutions.
+\textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP).
+For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.}
+\label{fig:benchmark}
+\end{figure}
+
+We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
+The solution size varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
+Heinz produced the smallest solutions, with an average of 182 selected SNPs.
+The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
+To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
+The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
+Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
+It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
+
+Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
+Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure~\ref{fig:benchmark}C).
+Conversely, the other methods displayed similarly low stabilities.
+
+In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which returned a solution in a few seconds.
+HotNet2 was the slowest (3 days and 14 hours on average).
+Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
+Considering that, it took 5 days on average for HotNet2 to produce a result.
+
+\subsection{Network topology matters, and might lead to ambiguous results}
+\label{results:drawbacks}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
+\caption{Drawbacks encountered when using network guided methods.
+\textbf{(A)}~DmGWAS solution subnetwork.
+Genes with a P\=/value < 0.1 are highlighted in red.
+\textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
+We highlighted the genes selected by each method, and the ones selected by more than one (``Consensus'').
+We labeled the three most central genes that were picked by any method.
+\textbf{(C)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
+SNPs that were not selected in the studied network, but were selected in another one, are displayed in background color.
+}
+\label{fig:issues}
+\end{figure}
+
+As shown above, and despite their similarities, the network methods produced remarkably different solutions.
+This is due to the particularities of each methods, and directly or indirectly provide information about the dataset.
+For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
+
+In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value for the subnetwork (Table~\ref{tab:gene_solutions}).
+Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Figure~\ref{fig:benchmark}C).
+Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
+Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
+Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
+Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
+
+On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
+dmGWAS' subnetwork is the least associated subnetwork on average.
+This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
+It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
+This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure~\ref{fig:issues}A).
+This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
+In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
+Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
+
+Another peculiarity of network methods is their relationship to degree centrality.
+On the one hand we observed that highly central genes often had no association to disease (Figure~\ref{fig:issues}B).
+On the other, network methods favor central genes, as they often connect high scoring nodes.
+This was specially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
+As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
+With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration.
+Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it is negatively correlated with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.
+
+By virtue of using a SNP subnetwork, SConES analyzes each SNP in their functional context.
+It therefore can select SNPs in genes without any associated interactor, as well as SNPs in non-coding regions or in non-interacting genes.
+In fact, due to linkage disequilibrium, SConES favors such genes, as selecting SNPs in an LD-block which overlaps with a gene favors selecting the rest of the gene.
+This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Supplementary Figure~\ref{sfig:pearson_methods}).
+On the other hand, SConES penalizes selecting SNPs and not their neighbors.
+This makes it conservative regarding SNPs with many interactions, like those mapped to hub genes in the PPIN.
+For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure~\ref{fig:issues}C).
+In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
+On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
+We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a gene and a majority of its neighbors.
+Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
+a lower \(\lambda\) in Equation~\ref{eq:scones}), it is unclear what the best strategy to find them is.
+In addition, due to the design of the iCOGS array, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
+
+\subsection{Adjusting for instability preserves global network properties}
+\label{results:stable-consensus}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.5\linewidth]{./figures/figure_5.pdf}
+  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
+Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+All gene names are indicated in Supplementary Figure \ref{sfig:stable-consensus-names}.}
+  \label{fig:stable-consensus}
+\end{figure}
+
+Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
+We built a new, stable consensus network using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:comparison}).
+Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
+We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once.
+The resulting stablility-based consensus was composed of 68 genes (Figure \ref{fig:stable-consensus}).
+This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
+
+However, although this new network exhibits similar global properties as the previous one, the lack of stability results in different genes being selected.
+In this case, the most central gene is \emph{CUL3}, which is absent from the previous consensus network and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
+This gene is a component of Cullin-RING ubiquitin ligases.
+Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
+
+\section{Discussion}
+
+In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
+On the one hand, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
+The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
+On the other hand, discovery in GWAS is hindered by a conservative statistical framework.
+Network methods tackle these two issues by using both the association score and an interaction network to take into consideration the biological context of each of the genes and SNPs.
+Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
+In this article we evaluated the relevance of six of them by examining a GWAS dataset on familial breast cancer.
+
+Most of the network methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
+In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
+However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by these network methods.
+Across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
+Yet, further work is needed to characterize the impact of centrality on network methods' outputs.
+Despite these similarities, the solutions were notably different.
+At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
+On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
+While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
+HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
+Additionally, all solutions share two drawbacks.
+First, they are all equally bad at discriminating cases from controls.
+Yet, the classification accuracy of network methods is similar to that of a machine learning classifier, which suggests that cases and controls are difficult to separate in this dataset.
+This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
+Second, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
+This might partly be caused by the instability of the P\=/values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
+Hence, heinz's conservative transformation of P\=/values, which favors only the most extreme ones, leads to improved stability.
+Another source of instability might be the redundancy inherent to biological networks.
+
+To overcome the limitations of the individual methods while exploiting their strengths, we proposed combining them into a consensus subnetwork.
+We use a straightforward strategy of including any node that was recovered by multiple methods.
+We proposed two networks: a consensus network that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus network, that addressed the instability of the methods.
+They both synthesized the altered mechanisms: they both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer.
+Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but who are related to cancer at the expression level and whose neighborhood has consistent high association scores.
+Crucially, the consensus was as unstable as the tested methods, while the stable consensus shared these properties while accounting for instability.
+This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence validates the conclusions obtained on the individual solutions and the consensus.
+
+The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
+In consequence, they show their shortcomings with respect to understudied genes, especially those not in the network.
+Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
+The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure~\ref{sfig:biotypes_excluded}).
+Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
+In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
+However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
+As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
+As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
+It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
+Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
+This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
+Lastly, all the methods are heavily affected by how SNPs are mapped to genes.
+In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+In fact, the presented case is paradigmatic, since the genes are in the most gene-dense region of the genome \cite{Xie2003}.
+Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the network, but might be more resilient to them when the overlapping genes are unrelated.
+Making use of more targeted mappings of SNPs to genes (e.g.
+eQTLs, SNPs associated to the expression of a gene), altogether with a stringent LD pruning, might address such problems.
+
+As not all databases compile the same interactions, the choice of the PPIN determines the final output.
+In this work we used exclusively interactions from HINT from high-throughput experiments.
+This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
+On the other hand, Huang et al.
+\cite{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
+When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
+This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the solutions.
+
+A crucial step for the gene based methods is the computation of the gene score.
+In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
+Another important decision is how to handle LD in a GWAS.
+VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
+With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
+However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
+
+In order to produce the consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
+To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
+They are available on GitHub: \url{https://github.com/hclimente/gwas-tools}.
+Importantly, those methods that had a permissive license were compiled into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
+
+\section{Materials and methods}
 \subsection{GENESIS}
@@
 As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
-\citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
+Azencott et al.
+\cite{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
 We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
@@
 All three are positive quantities.
-Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
+Heinz implements the algorithm from Ljubić et al.
+\cite{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
 We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
@@
 Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
-We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+We assigned initial node scores as in Nakka et al.
+\cite{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
 In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
@@
 Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
-As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
+As in Azencott et al.
+\cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
 In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
@@
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
-We adjusted the P-values for multiple testing as in \citet{Benjamini1995} (BH).Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
+We adjusted the P-values for multiple testing as in Benjamini and Hochberg \citet{Benjamini1995} (BH).
+Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
 As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
@@
 We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
-The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by \citet{nogueira_measuring_2016}.
+The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by Nogueira and Brown \cite{nogueira_measuring_2016}.
 
@@
 We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
-
-\section{Results}
-
-\subsection{Conventional SNP- and gene-based analyses confirm that \emph{FGFR2} locus is associated with familial breast cancer}
-\label{results:conventional}
-
-We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
-Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
-The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
-Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
-In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
-
-Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
-Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
-It selected 100 SNPs, both from all aforementioned regions and new ones (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
-Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
-Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
-In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
-This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
-In fact, such methods provided comparably (poor) classification performance to L1-penalized logistic regression (Figure~\ref{fig:benchmark}B), while providing more interpretable solutions.
-
-\subsection{Network methods successfully identify genes associated with breast cancer}
-\label{results:separate_networks}
-\begin{table}[htbp]
-  \centering
-  \begin{threeparttable}
-\caption{Summary statistics on the solutions of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
-\label{tab:gene_solutions}
-\centering
-\begin{tabular}{lrrrrr}
-Network & \# genes & \# edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & \(\rho\)\textsubscript{consensus}\\
-\hline
-HINT HT & 13\,619 & 142\,541 & 16\,706 & 0.46 & 0.066\\
-\hline
-dmGWAS & 194 & 450 & 49\,115 & 0.19 & 0.41\\
-heinz & 4 & 3 & 113\,633 & 0.001 & 0.21\\
-HotNet2 & 440 & 374 & 7\,739 & 0.048 & 0.31\\
-LEAN & 0 & 0 & - & - & -\\
-SConES GI & 0 (1) & 0 & - & - & -\\
-SigMod & 142 & 249 & 92\,603 & 0.008 & 0.73\\
-\hline
-Consensus & 93 & 186 & 50\,737 & 0.006 & 1\\
-Stable consensus & 68 & 49 & 94\,854 & 0.005 & 0.54 \\
-\end{tabular}
-\begin{tablenotes}
-  \footnotesize{
-    \item \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
-\textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
-\textbf{\median{P}\textsubscript{gene}}: median P\=/value of the selected genes.
-\textbf{\(\rho\)\textsubscript{consensus}}: Pearson correlation between the subnetwork and the consensus network.
-  }
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[height=.75\textheight]{./figures/figure_1.png}
-  \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-As LEAN did not produce any significant gene (BH adjusted P\=/value~<~0.05), it was excluded.
-Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
-\textbf{(A)} Manhattan plots of SNPs/genes; in black, the method's solution.
-The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{(B)} Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
-\textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(D)} Solution networks produced by the different methods.}
-  \label{fig:solution_overview}
-\end{figure}
-
-We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (BH adjusted P\=/value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
-These solutions differ in many aspects, making it hard to draw joint conclusions.
-For starters, the overlap between the genes featured in each solution is quite small (Figure~\ref{fig:solution_overview}B).
-Another prominent difference is their size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
-While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
-Their topologies also differ, as measured by the median centrality (Table~\ref{tab:gene_solutions}) and the number of connected components (Figure~\ref{fig:solution_overview}D).
-Only two methods have more than one connected component: SConES, as described above, and HotNet2.
-HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
-The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
-Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
-It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
-Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
-
-Despite their differences, there are additional common themes.
-All obtained solution subnetworks have lower association P\=/values than the whole PPIN (median P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Figure~\ref{fig:solution_overview}A).
-This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
-Additionally, the solution subnetworks presented other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value < 0.03).
-Second, the genes in four solution subnetworks displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value < 1.4 \texttimes{} 10\textsuperscript{-21}).
-This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
-
-\subsection{A case study: the consensus network}
-\label{results:consensus}
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.5\linewidth]{./figures/figure_3.pdf}
-  \caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
-Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-All gene names are indicated in Supplementary Figure \ref{sfig:consensus-names}.}
-  \label{fig:consensus}
-\end{figure}
-
-Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
-Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
-To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
-This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
-
-A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions.
-We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
-The former is supported by genes like \emph{MRPS30} (VEGAS2 P\=/value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
-Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
-With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
-The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their effects.
-\emph{HSPA1A} was the most strongly associated gene (VEGAS2 P\=/value~=~8.37~\texttimes{}~10\textsuperscript{-4}).
-
-Topologically the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
-Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
-The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
-This suggests that they were selected because they were on the shortest path between two highly associated genes.
-In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signalling pathways.
-\emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
-
-\subsection{Network methods boost biomarker discovery}
-
-We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
-Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
-This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
-We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
-The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
-The gene-based network methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (Figure~\ref{fig:solution_overview}C).
-Interestingly, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
-
-\subsection{Network methods share limitations}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.72\linewidth]{./figures/figure_4.pdf}
-\caption{Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
-As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
-\textbf{(A)} Number of SNPs selected by each method and number of SNPs in the active set found by the classifier.
-Points are the average over the 5 runs; lines represent the standard error of the mean.
-A grey diagonal line with slope 1 is added for comparison.
-For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs.
-\textbf{(B)} Sensitivity and specificity on test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
-In addition, the performance of the classifier trained on all SNPs is displayed.
-Points are the average over the 5 runs; lines represent the standard error of the mean.
-\textbf{(C)} Pairwise Pearson correlations of the solutions used by different methods.
-A Pearson correlation of 1 means the two solutions are the same.
-A Pearson correlation of 0 means that there is no SNP in common between the two solutions.
-\textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP).
-For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.}
-\label{fig:benchmark}
-\end{figure}
-
-We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
-Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
-The solution size varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
-Heinz produced the smallest solutions, with an average of 182 selected SNPs.
-The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
-To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
-The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
-Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
-
-Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
-Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure~\ref{fig:benchmark}C).
-Conversely, the other methods displayed similarly low stabilities.
-
-In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which returned a solution in a few seconds.
-HotNet2 was the slowest (3 days and 14 hours on average).
-Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
-Considering that, it took 5 days on average for HotNet2 to produce a result.
-
-\subsection{Network topology matters, and might lead to ambiguous results}
-\label{results:drawbacks}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
-\caption{Drawbacks encountered when using network guided methods.
-\textbf{(A)}~DmGWAS solution subnetwork.
-Genes with a P\=/value < 0.1 are highlighted in red.
-\textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
-We highlighted the genes selected by each method, and the ones selected by more than one (``Consensus'').
-We labeled the three most central genes that were picked by any method.
-\textbf{(C)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
-SNPs that were not selected in the studied network, but were selected in another one, are displayed in background color.
-}
-\label{fig:issues}
-\end{figure}
-
-As shown above, and despite their similarities, the network methods produced remarkably different solutions.
-This is due to the particularities of each methods, and directly or indirectly provide information about the dataset.
-For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
-
-In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value for the subnetwork (Table~\ref{tab:gene_solutions}).
-Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Figure~\ref{fig:benchmark}C).
-Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
-Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
-Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
-
-On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
-dmGWAS' subnetwork is the least associated subnetwork on average.
-This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
-It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
-This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure~\ref{fig:issues}A).
-This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
-In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
-Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
-
-Another peculiarity of network methods is their relationship to degree centrality.
-On the one hand we observed that highly central genes often had no association to disease (Figure~\ref{fig:issues}B).
-On the other, network methods favor central genes, as they often connect high scoring nodes.
-This was specially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
-As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
-With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration.
-Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it is negatively correlated with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.
-
-By virtue of using a SNP subnetwork, SConES analyzes each SNP in their functional context.
-It therefore can select SNPs in genes without any associated interactor, as well as SNPs in non-coding regions or in non-interacting genes.
-In fact, due to linkage disequilibrium, SConES favors such genes, as selecting SNPs in an LD-block which overlaps with a gene favors selecting the rest of the gene.
-This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Supplementary Figure~\ref{sfig:pearson_methods}).
-On the other hand, SConES penalizes selecting SNPs and not their neighbors.
-This makes it conservative regarding SNPs with many interactions, like those mapped to hub genes in the PPIN.
-For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure~\ref{fig:issues}C).
-In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
-On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
-We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a gene and a majority of its neighbors.
-Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
-a lower \(\lambda\) in Equation~\ref{eq:scones}), it is unclear what the best strategy to find them is.
-In addition, due to the design of the iCOGS array, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
-
-\subsection{Adjusting for instability preserves global network properties}
-\label{results:stable-consensus}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.5\linewidth]{./figures/figure_5.pdf}
-  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
-Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-All gene names are indicated in Supplementary Figure \ref{sfig:stable-consensus-names}.}
-  \label{fig:stable-consensus}
-\end{figure}
-
-Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
-We built a new, stable consensus network using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:comparison}).
-Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
-We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once.
-The resulting stablility-based consensus was composed of 68 genes (Figure \ref{fig:stable-consensus}).
-This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
-
-However, although this new network exhibits similar global properties as the previous one, the lack of stability results in different genes being selected.
-In this case, the most central gene is \emph{CUL3}, which is absent from the previous consensus network and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
-This gene is a component of Cullin-RING ubiquitin ligases.
-Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
-
-\section{Discussion}
-
-In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
-On the one hand, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
-The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
-On the other hand, discovery in GWAS is hindered by a conservative statistical framework.
-Network methods tackle these two issues by using both the association score and an interaction network to take into consideration the biological context of each of the genes and SNPs.
-Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
-In this article we evaluated the relevance of six of them by examining a GWAS dataset on familial breast cancer.
-
-Most of the network methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
-In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
-However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by these network methods.
-Across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
-Yet, further work is needed to characterize the impact of centrality on network methods' outputs.
-Despite these similarities, the solutions were notably different.
-At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
-On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
-HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
-Additionally, all solutions share two drawbacks.
-First, they are all equally bad at discriminating cases from controls.
-Yet, the classification accuracy of network methods is similar to that of a machine learning classifier, which suggests that cases and controls are difficult to separate in this dataset.
-This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
-Second, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
-This might partly be caused by the instability of the P\=/values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
-Hence, heinz's conservative transformation of P\=/values, which favors only the most extreme ones, leads to improved stability.
-Another source of instability might be the redundancy inherent to biological networks.
-
-To overcome the limitations of the individual methods while exploiting their strengths, we proposed combining them into a consensus subnetwork.
-We use a straightforward strategy of including any node that was recovered by multiple methods.
-We proposed two networks: a consensus network that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus network, that addressed the instability of the methods.
-They both synthesized the altered mechanisms: they both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer.
-Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but who are related to cancer at the expression level and whose neighborhood has consistent high association scores.
-Crucially, the consensus was as unstable as the tested methods, while the stable consensus shared these properties while accounting for instability.
-This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence validates the conclusions obtained on the individual solutions and the consensus.
-
-The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
-In consequence, they show their shortcomings with respect to understudied genes, especially those not in the network.
-Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
-The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure~\ref{sfig:biotypes_excluded}).
-Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
-In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
-However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
-As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
-As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
-It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
-Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
-This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
-Lastly, all the methods are heavily affected by how SNPs are mapped to genes.
-In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
-In fact, the presented case is paradigmatic, since the genes are in the most gene-dense region of the genome \cite{Xie2003}.
-Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the network, but might be more resilient to them when the overlapping genes are unrelated.
-Making use of more targeted mappings of SNPs to genes (e.g.
-eQTLs, SNPs associated to the expression of a gene), altogether with a stringent LD pruning, might address such problems.
-
-As not all databases compile the same interactions, the choice of the PPIN determines the final output.
-In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
-On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
-When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
-This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the solutions.
-
-A crucial step for the gene based methods is the computation of the gene score.
-In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
-Another important decision is how to handle LD in a GWAS.
-VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
-With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
-However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
-
-In order to produce the consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
-To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
-They are available on GitHub: \url{https://github.com/hclimente/gwas-tools}.
-Importantly, those methods that had a permissive license were compiled into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
 
@@
 
-\section*{Funding and acknowledgments}
-
-This project was supported by funding from Agence Nationale de la Recherche (ANR-18-CE45-0021-01) and from the European Union’s Horizon 2020 research and innovation program (Marie Skłodowska-Curie [666003]).
-Financial support for GENESIS resource and genotyping was provided by the Ligue Nationale contre le Cancer (grants PRE05/DSL, PRE07/DSL, PRE11/NA), the French National Institute of Cancer (INCa grant b2008-029/LL-LC) and the comprehensive cancer center SiRIC (Site de Recherche Intégrée sur le Cancer: INCa-DGOS-4654).
-
-We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S.
-Eon-Marchais, M.
-Marcou, D.
-Le Gal, L.
-Toulemonde, J.
-Beauvallet, N.
-Mebirouk, E.
-Cavaciuti), the biological resource centre (S.
-Mazoyer, F.
-Damiola, L.
-Barjhoux, C.
-Verny-Pierre, V.
-Sornin) and all the GENESIS collaborating cancer clinics clinics (Clinique Sainte Catherine, Avignon: H.
-Dreyfus; Hôpital Saint Jacques, Besançon: M-A.
-Collonge-Rame; Institut Bergonié, Bordeaux: M.Longy, A.
-Floquet, E.
-Barouk-Simonet; CHU, Brest: S.
-Audebert; Centre François Baclesse, Caen: P.
-Berthet; Hôpital Dieu, Chambéry: S.
-Fert-Ferrer; Centre Jean Perrin, Clermont-Ferrand: Y-J.
-Bignon; Hôpital Pasteur, Colmar: J-M.
-Limacher; Hôpital d’Enfants CHU – Centre Georges François Leclerc, Dijon: L.
-Faivre-Olivier; CHU, Fort de France: O.
-Bera; CHU Albert Michallon, Grenoble: D.
-Leroux; Hôpital Flaubert, Le Havre: V.
-Layet; Centre Oscar Lambret, Lille: P.
-Vennin, C.
-Adenis; Hôpital Jeanne de Flandre, Lille: S.
-Lejeune-Dumoulin, S.
-Manouvier-Hanu; CHRU Dupuytren, Limoges: L.
-Venat-Bouvet; Centre Léon Bérard, Lyon: C.
-Lasset, V.
-Bonadona; Hôpital Edouard Herriot, Lyon: S.
-Giraud; Institut Paoli-Calmettes, Marseille: F.
-Eisinger, L.
-Huiart; Centre Val d’Aurelle – Paul Lamarque, Montpellier: I.
-Coupier; CHU Arnaud de Villeneuve, Montpellier: I.
-Coupier, P.
-Pujol; Centre René Gauducheau, Nantes: C.
-Delnatte; Centre Catherine de Sienne, Nantes: A.
-Lortholary; Centre Antoine Lacassagne, Nice: M.
-Frénay, V.
-Mari; Hôpital Caremeau, Nîmes: J.
-Chiesa; Réseau Oncogénétique Poitou Charente, Niort: P.
-Gesta; Institut Curie, Paris: D.
-Stoppa-Lyonnet, M.
-Gauthier-Villars, B.
-Buecher, A.
-de Pauw, C.
-Abadie, M.
-Belotti; Hôpital Saint-Louis, Paris: O.
-Cohen-Haguenauer; Centre Viggo-Petersen, Paris: F.
-Cornélis; Hôpital Tenon, Paris: A.
-Fajac; GH Pitié Salpétrière et Hôpital Beaujon, Paris: C.
-Colas, F.
-Soubrier, P.
-Hammel, A.
-Fajac; Institut Jean Godinot, Reims: C.
-Penet, T.
-D.
-Nguyen; Polyclinique Courlancy, Reims: L.
-Demange, C.
-Penet; Centre Eugène Marquis, Rennes: C.
-Dugast; Centre Henri Becquerel, Rouen: A.
-Chevrier, T.
-Frebourg, J.
-Tinat, I.
-Tennevet, A.
-Rossi; Hôpital René Huguenin/Institut Curie, Saint Cloud: C.
-Noguès, L.
-Demange, E.
-Mouret-Fourme; CHU, Saint-Etienne: F.
-Prieur; Centre Paul Strauss, Strasbourg: J-P.
-Fricker, H.
-Schuster; Hôpital Civil, Strasbourg: O.
-Caron, C.
-Maugard; Institut Claudius Regaud, Toulouse: L.
-Gladieff, V.
-Feillel; Hôpital Bretonneau, Tours: I.
-Mortemousque; Centre Alexis Vautrin, Vandoeuvre-les-Nancy: E.
-Luporsi; Hôpital de Bravois, Vandoeuvre-les-Nancy: P.
-Jonveaux; Gustave Roussy, Villejuif: A.
-Chompret, O.
-Caron).
-
-\bibliographystyle{abbrvnat}
+\section*{Acknowledgments}
+
+We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie), the biological resource centre and all the GENESIS collaborating cancer clinics clinics.
+
+\bibliographystyle{vancouver}
 \bibliography{bibliography}
@@
   \centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
-\caption{GENESIS shows no differential population structure between cases and controls.
-\textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
-Cases are colored in green, controls in orange.}
-\label{sfig:pcs}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
   \includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
+\caption{GENESIS shows no differential population structure between cases and controls.
+\textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
+Cases are colored in green, controls in orange.}
+\label{sfig:pcs}
+\end{figure}
+
 \end{document}

### 2020-08-20 manuscript/plos/plos_latex_template.tex
commit before moving to plos template

--- 2020-08-20 manuscript/manuscript.tex
@@
-\documentclass[twocolumn, 11pt]{article}
+% Template for PLoS
+% Version 3.5 March 2018
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your equations and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline equations, please be sure to include all portions of an equation in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display equations when possible in order to fit size of the column. 
+%
+% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-\usepackage[format=hang,labelfont=bf,textfont=small,singlelinecheck=false,justification=justified,margin={20pt,20pt},figurename=Fig.]{caption}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{grffile}
-\usepackage{longtable}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{capt-of}
-\usepackage{hyperref}
-\usepackage{threeparttable}
-\usepackage[numbers]{natbib}
-\usepackage{lmodern} % Chloe: better font rendering
-\usepackage[shortcuts]{extdash} 
-\newcommand{\mean}[1]{$\overline{\mbox{#1}}$}
-\newcommand{\median}[1]{$\hat{\mbox{#1}}$}
-
-% Chloe's commands
-\usepackage[dvipsnames]{xcolor}
-\usepackage[normalem]{ulem} 
-\newcommand{\cazcom}[2]{{\uline{#1}}\unskip\space\textcolor{MidnightBlue}{[CA: #2]}}
-\newcommand{\caz}[2]{{\sout{#1}}\unskip\space\textcolor{MidnightBlue}{#2}}
-%%% To get the original text before my comments:
-% \newcommand{\cazcom}[2]{#1}
-% \newcommand{\caz}[2]{#1}
-%%% To get the corrected text and remove my comments:
-% \newcommand{\cazcom}[2]{#1}
-% \newcommand{\caz}[2]{#2}
-
-\title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
-\author{Héctor Climente-González$^{1,2,3,4,\spadesuit}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
-GENESIS Study collaborators$^{1,2,3}$, Dominique Stoppa-Lyonnet$^{5,6,7}$, \\
-Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
-\date{\small{
-  \raggedright{
-  $^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
-  $^{2}$INSERM, U900, F-75005 Paris, France;\\
-  $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
-  $^{4}$RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, 103-0027, Japan; \\
-  $^{5}$Service de Génétique, Institut Curie, F-75005 Paris, France;\\
-  $^{6}$INSERM, U830, F-75005 Paris, France;\\
-  $^{7}$Université Paris Descartes.\\
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
 }
-  \raggedright{$^{\spadesuit}$Corresponding author: \href{hector.climente@riken.jp}{hector.climente@riken.jp}\\}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
+
+%% END MACROS SECTION
+
+\begin{document}
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Title of submission to PLOS journals} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
-}
-\hypersetup{
-  pdfauthor={Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, GENESIS Study collaborators, Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott},
-  pdftitle={Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer},
-  pdfkeywords={},
-  pdfsubject={},
-  pdflang={English}}
-\begin{document}
-
-\onecolumn
-\maketitle
-
-\begin{abstract}
-Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
-Indeed, it allows to jointly consider the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study six network methods which identify subnetworks with high association scores to a phenotype.
-Specifically, we examine their utility to discover new biomarkers for breast cancer susceptibility by interrogating a genome-wide association study (GWAS) focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
-We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-By trading statistical stringency for biological meaningfulness, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-For instance, we show a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-Importantly, we find a significantly large overlap between the genes in the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
-Yet, network methods are notably unstable, producing different results when the input data changes slightly.
-To account for that, we produce a stable consensus subnetwork, formed by the most consistently selected genes.
-The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
-The network seems organized around \emph{CUL3}, encoding an ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
-In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
-Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
-\end{abstract}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Name1 Surname\textsuperscript{1,2\Yinyang},
+Name2 Surname\textsuperscript{2\Yinyang},
+Name3 Surname\textsuperscript{2,3\textcurrency},
+Name4 Surname\textsuperscript{2},
+Name5 Surname\textsuperscript{2\ddag},
+Name6 Surname\textsuperscript{2\ddag},
+Name7 Surname\textsuperscript{1,2,3*},
+with the Lorem Ipsum Consortium\textsuperscript{\textpilcrow}
+\\
+\bigskip
+\textbf{1} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\textbf{3} Affiliation Dept/Program/Center, Institution Name, City, State, Country
+\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+\Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+\ddag These authors also contributed equally to this work.
+
+% Current address notes
+\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address 
+% \textcurrency c Insert third current address
+
+% Deceased author note
+\dag Deceased
+
+% Group/Consortium Author Note
+\textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* correspondingauthor@institute.edu
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author summary}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+\linenumbers
 
@@
 
+% Use "Eq" instead of "Equation" for equation citations.
+\section*{Introduction}
+Lorem ipsum dolor sit~\cite{bib1} amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam Eq~(\ref{eq:schemeP}) sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.~\cite{bib2} Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+\begin{eqnarray}
+\label{eq:schemeP}
+	\mathrm{P_Y} = \underbrace{H(Y_n) - H(Y_n|\mathbf{V}^{Y}_{n})}_{S_Y} + \underbrace{H(Y_n|\mathbf{V}^{Y}_{n})- H(Y_n|\mathbf{V}^{X,Y}_{n})}_{T_{X\rightarrow Y}},
+\end{eqnarray}
+
+\section*{Materials and methods}
+\subsection*{Etiam eget sapien nibh}
+
+% For figure citations, please use "Fig" instead of "Figure".
+Nulla mi mi, Fig~\ref{fig1} venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, \nameref{S1_Video} vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+% Place figure captions after the first paragraph in which they are cited.
+\begin{figure}[!h]
+\caption{{\bf Bold the figure title.}
+Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
+\label{fig1}
+\end{figure}
+
+% Results and Discussion can be combined.
+\section*{Results}
+Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+% Place tables after the first paragraph in which they are cited.
+\begin{table}[!ht]
+\begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+\centering
+\caption{
+{\bf Table caption Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.}}
+\begin{tabular}{|l+l|l|l|l|l|l|l|}
+\hline
+\multicolumn{4}{|l|}{\bf Heading1} & \multicolumn{4}{|l|}{\bf Heading2}\\ \thickhline
+$cell1 row1$ & cell2 row 1 & cell3 row 1 & cell4 row 1 & cell5 row 1 & cell6 row 1 & cell7 row 1 & cell8 row 1\\ \hline
+$cell1 row2$ & cell2 row 2 & cell3 row 2 & cell4 row 2 & cell5 row 2 & cell6 row 2 & cell7 row 2 & cell8 row 2\\ \hline
+$cell1 row3$ & cell2 row 3 & cell3 row 3 & cell4 row 3 & cell5 row 3 & cell6 row 3 & cell7 row 3 & cell8 row 3\\ \hline
+\end{tabular}
+\begin{flushleft} Table notes Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+\end{flushleft}
+\label{table1}
+\end{adjustwidth}
+\end{table}
+
+%PLOS does not support heading levels beyond the 3rd (no 4th level headings).
+\subsection*{\lorem\ and \ipsum\ nunc blandit a tortor}
+\subsubsection*{3rd level heading} 
+Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
+Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
+
+\begin{enumerate}
+	\item{react}
+	\item{diffuse free particles}
+	\item{increment time by dt and go to 1}
+\end{enumerate}
+
+\subsection*{Sed ac quam id nisi malesuada congue}
+
+Nulla mi mi, venenatis sed ipsum varius, volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero.
+
+\begin{itemize}
+	\item First bulleted item.
+	\item Second bulleted item.
+	\item Third bulleted item.
+\end{itemize}
+
+\section*{Discussion}
+Nulla mi mi, venenatis sed ipsum varius, Table~\ref{table1} volutpat euismod diam.
+Proin rutrum vel massa non gravida.
+Quisque tempor sem et dignissim rutrum.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Morbi at justo vitae nulla elementum commodo eu id massa.
+In vitae diam ac augue semper tincidunt eu ut eros.
+Fusce fringilla erat porttitor lectus cursus, vel sagittis arcu lobortis.
+Aliquam in enim semper, aliquam massa id, cursus neque.
+Praesent faucibus semper libero~\cite{bib3}.
+
+\section*{Conclusion}
+
+CO\textsubscript{2} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+Quisque augue sem, tincidunt sit amet feugiat eget, ullamcorper sed velit.
+
+Sed non aliquet felis.
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Mauris commodo justo ac dui pretium imperdiet.
+Sed suscipit iaculis mi at feugiat.
+Ut neque ipsum, luctus id lacus ut, laoreet scelerisque urna.
+Phasellus venenatis, tortor nec vestibulum mattis, massa tortor interdum felis, nec pellentesque metus tortor nec nisl.
+Ut ornare mauris tellus, vel dapibus arcu suscipit sed.
+Nam condimentum sem eget mollis euismod.
+Nullam dui urna, gravida venenatis dui et, tincidunt sodales ex.
+Nunc est dui, sodales sed mauris nec, auctor sagittis leo.
+Aliquam tincidunt, ex in facilisis elementum, libero lectus luctus est, non vulputate nisl augue at dolor.
+For more information, see \nameref{S1_Appendix}.
+
+\section*{Supporting information}
+
+% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
+\paragraph*{S1 Fig.}
+\label{S1_Fig}
+{\bf Bold the title sentence.} Add descriptive text after the title of the item (optional).
+
+\paragraph*{S2 Fig.}
+\label{S2_Fig}
+{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\paragraph*{S1 File.}
+\label{S1_File}
+{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\paragraph*{S1 Video.}
+\label{S1_Video}
+{\bf Lorem ipsum.}  Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\paragraph*{S1 Appendix.}
+\label{S1_Appendix}
+{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\paragraph*{S1 Table.}
+\label{S1_Table}
+{\bf Lorem ipsum.} Maecenas convallis mauris sit amet sem ultrices gravida.
+Etiam eget sapien nibh.
+Sed ac ipsum eget enim egestas ullamcorper nec euismod ligula.
+Curabitur fringilla pulvinar lectus consectetur pellentesque.
+
+\section*{Acknowledgments}
+Cras egestas velit mauris, eu mollis turpis pellentesque sit amet.
+Interdum et malesuada fames ac ante ipsum primis in faucibus.
+Nam id pretium nisi.
+Sed ac quam id nisi malesuada congue.
+Sed interdum aliquet augue, at pellentesque quam rhoncus vitae.
+
+\nolinenumbers
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% or
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here. See http://journals.plos.org/plosone/s/latex for 
+% step-by-step instructions.
+% 
+\begin{thebibliography}{10}
+
+\bibitem{bib1}
+Conant GC, Wolfe KH.
+\newblock {{T}urning a hobby into a job: how duplicated genes find new
+  functions}.
+\newblock Nat Rev Genet. 2008 Dec;9(12):938--950.
+
+\bibitem{bib2}
+Ohno S.
+\newblock Evolution by gene duplication.
+\newblock London: George Alien \& Unwin Ltd. Berlin, Heidelberg and New York:
+  Springer-Verlag.; 1970.
+
+\bibitem{bib3}
+Magwire MM, Bayer F, Webster CL, Cao C, Jiggins FM.
+\newblock {{S}uccessive increases in the resistance of {D}rosophila to viral
+  infection through a transposon insertion followed by a {D}uplication}.
+\newblock PLoS Genet. 2011 Oct;7(10):e1002337.
+
+\end{thebibliography}
+
 \end{document}

### 2020-08-20 manuscript/plos/original.tex
prepared review environment

--- 2020-08-20 manuscript/manuscript.tex
@@
-\documentclass[twocolumn, 11pt]{article}
+% Template for PLoS
+% Version 3.5 March 2018
+%
+% % % % % % % % % % % % % % % % % % % % % %
+%
+% -- IMPORTANT NOTE
+%
+% This template contains comments intended 
+% to minimize problems and delays during our production 
+% process. Please follow the template instructions
+% whenever possible.
+%
+% % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Once your paper is accepted for publication, 
+% PLEASE REMOVE ALL TRACKED CHANGES in this file 
+% and leave only the final text of your manuscript. 
+% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
+% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
+%
+%
+% There are no restrictions on package use within the LaTeX files except that 
+% no packages listed in the template may be deleted.
+%
+% Please do not include colors or graphics in the text.
+%
+% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
+%
+% % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- FIGURES AND TABLES
+%
+% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
+%
+% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
+% - Figures should be uploaded separately from your manuscript file. 
+% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
+% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
+% For figure citations, please use "Fig" instead of "Figure".
+% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
+%
+% Tables should be cell-based and may not contain:
+% - spacing/line breaks within cells to alter layout or alignment
+% - do not nest tabular environments (no tabular environments within tabular environments)
+% - no graphics or colored text (cell background color/shading OK)
+% See http://journals.plos.org/plosone/s/tables for table guidelines.
+%
+% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+%
+% -- eqnarrayS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
+%
+% IMPORTANT
+% Below are a few tips to help format your eqnarrays and other special characters according to our specifications.
+For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
+%
+% For inline eqnarrays, please be sure to include all portions of an eqnarray in the math environment.
+For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
+%
+% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
+%
+% Please add line breaks to long display eqnarrays when possible in order to fit size of the column. 
+%
+% For inline eqnarrays, please do not include punctuation (commas, etc) within the math environment unless this is part of the eqnarray.
+%
+% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
+%
+% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
+%
+% % % % % % % % % % % % % % % % % % % % % % % % 
+%
+% Please contact latex@plos.org with any questions.
+%
+% % % % % % % % % % % % % % % % % % % % % % % %
+
+\documentclass[10pt,letterpaper]{article}
+\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
+
+% amsmath and amssymb packages, useful for mathematical formulas and symbols
 \usepackage{amsmath,amssymb}
-\usepackage[format=hang,labelfont=bf,textfont=small,singlelinecheck=false,justification=justified,margin={20pt,20pt},figurename=Fig.]{caption}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{grffile}
-\usepackage{longtable}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{capt-of}
-\usepackage{hyperref}
-\usepackage{threeparttable}
-\usepackage[numbers]{natbib}
+
+% Use adjustwidth environment to exceed column width (see example table in text)
+\usepackage{changepage}
+
+% Use Unicode characters when possible
+\usepackage[utf8x]{inputenc}
+
+% textcomp package and marvosym package for additional characters
+\usepackage{textcomp,marvosym}
+
+% cite package, to clean up citations in the main text. Do not remove.
+\usepackage{cite}
+
+% Use nameref to cite supporting information files (see Supporting Information section for more info)
+\usepackage{nameref,hyperref}
+
+% line numbers
+\usepackage[right]{lineno}
+
+% ligatures disabled
+\usepackage{microtype}
+\DisableLigatures[f]{encoding = *, family = * }
+
+% color can be used to apply background shading to table cells only
+\usepackage[table]{xcolor}
+
+% array package and thick rules for tables
+\usepackage{array}
+
+% author's packages
+\usepackage[shortcuts]{extdash}
 \usepackage{lmodern} % Chloe: better font rendering
-\usepackage[shortcuts]{extdash} 
+
+% create "+" rule type for thick vertical lines
+\newcolumntype{+}{!{\vrule width 2pt}}
+
+% create \thickcline for thick horizontal lines of variable length
+\newlength\savedwidth
+\newcommand\thickcline[1]{%
+  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+  \cline{#1}%
+  \noalign{\vskip\arrayrulewidth}%
+  \noalign{\global\arrayrulewidth\savedwidth}%
+}
+
+% \thickhline command for thick horizontal lines that span the table
+\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
+\hline
+\noalign{\global\arrayrulewidth\savedwidth}}
+
+% Remove comment for double spacing
+%\usepackage{setspace} 
+%\doublespacing
+
+% Text layout
+\raggedright
+\setlength{\parindent}{0.5cm}
+\textwidth 5.25in 
+\textheight 8.75in
+
+% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
+% Captions will be left justified
+\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
+\renewcommand{\figurename}{Fig}
+
+% Use the PLoS provided BiBTeX style
+\bibliographystyle{plos2015}
+
+% Remove brackets from numbering in List of References
+\makeatletter
+\renewcommand{\@biblabel}[1]{\quad#1.}
+\makeatother
+
+% Header and Footer with logo
+\usepackage{lastpage,fancyhdr,graphicx}
+\usepackage{epstopdf}
+%\pagestyle{myheadings}
+\pagestyle{fancy}
+\fancyhf{}
+%\setlength{\headheight}{27.023pt}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+\rfoot{\thepage/\pageref{LastPage}}
+\renewcommand{\headrulewidth}{0pt}
+\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+\fancyheadoffset[L]{2.25in}
+\fancyfootoffset[L]{2.25in}
+\lfoot{\today}
+
+%% Include all macros below
+
+\newcommand{\lorem}{{\bf LOREM}}
+\newcommand{\ipsum}{{\bf IPSUM}}
 \newcommand{\mean}[1]{$\overline{\mbox{#1}}$}
@@
 
-% Chloe's commands
-\usepackage[dvipsnames]{xcolor}
-\usepackage[normalem]{ulem} 
-\newcommand{\cazcom}[2]{{\uline{#1}}\unskip\space\textcolor{MidnightBlue}{[CA: #2]}}
-\newcommand{\caz}[2]{{\sout{#1}}\unskip\space\textcolor{MidnightBlue}{#2}}
-%%% To get the original text before my comments:
-% \newcommand{\cazcom}[2]{#1}
-% \newcommand{\caz}[2]{#1}
-%%% To get the corrected text and remove my comments:
-% \newcommand{\cazcom}[2]{#1}
-% \newcommand{\caz}[2]{#2}
-
-\title{Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer}
-\author{Héctor Climente-González$^{1,2,3,4,\spadesuit}$, Christine Lonjou$^{1,2,3}$, Fabienne Lesueur$^{1,2,3}$, \\
-GENESIS Study collaborators$^{1,2,3}$, Dominique Stoppa-Lyonnet$^{5,6,7}$, \\
-Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
-\date{\small{
-  \raggedright{
-  $^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
-  $^{2}$INSERM, U900, F-75005 Paris, France;\\
-  $^{3}$MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
-  $^{4}$RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, 103-0027, Japan; \\
-  $^{5}$Service de Génétique, Institut Curie, F-75005 Paris, France;\\
-  $^{6}$INSERM, U830, F-75005 Paris, France;\\
-  $^{7}$Université Paris Descartes.\\
+%% END MACROS SECTION
+
+\begin{document}
+\vspace*{0.2in}
+
+% Title must be 250 characters or less.
+\begin{flushleft}
+{\Large
+\textbf\newline{Comparing and combining network-guided biomarker discovery to discover genetic susceptibility mechanisms to breast cancer on the GENESIS study} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
-  \raggedright{$^{\spadesuit}$Corresponding author: \href{hector.climente@riken.jp}{hector.climente@riken.jp}\\}
-}
-}
-\hypersetup{
-  pdfauthor={Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, GENESIS Study collaborators, Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott},
-  pdftitle={Combining network-guided GWAS to discover susceptibility mechanisms for breast cancer},
-  pdfkeywords={},
-  pdfsubject={},
-  pdflang={English}}
-\begin{document}
-
-\onecolumn
-\maketitle
-
-\begin{abstract}
+\newline
+% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
+\\
+Héctor Climente-González\textsuperscript{1,2,3,4*}, 
+Christine Lonjou\textsuperscript{1,2,3}, 
+Fabienne Lesueur\textsuperscript{1,2,3},
+Dominique Stoppa-Lyonnet\textsuperscript{5,6,7\textcurrency}, 
+Nadine Andrieu\textsuperscript{1,2,3}, 
+Chloé-Agathe Azencott\textsuperscript{3,1,2},
+with the GENESIS study group\textsuperscript{\textpilcrow}
+\\
+\bigskip
+\textbf{1} Institut Curie, PSL Research University, F-75005 Paris, France;\\
+\textbf{2} INSERM, U900, F-75005 Paris, France;\\
+\textbf{3} MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
+\textbf{4} RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, Japan; \\
+\textbf{5} Service de Génétique, Institut Curie, F-75005 Paris, France;\\
+\textbf{6} INSERM, U830, F-75005 Paris, France;\\
+\textbf{7} Université Paris Descartes.\\
+\bigskip
+
+% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
+% 
+% Remove or comment out the author notes below if they aren't used.
+%
+% Primary Equal Contribution Note
+% \Yinyang These authors contributed equally to this work.
+
+% Additional Equal Contribution Note
+% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
+% \ddag These authors also contributed equally to this work.
+
+% Current address notes
+\textcurrency For the GENESIS study group % change symbol to "\textcurrency a" if more than one current address note
+% \textcurrency b Insert second current address 
+% \textcurrency c Insert third current address
+
+% Deceased author note
+% \dag Deceased
+
+% Group/Consortium Author Note
+\textpilcrow Membership list can be found in the Acknowledgments section.
+
+% Use the asterisk to denote corresponding authorship and provide email address in note below.
+* hector.climente@riken.jp
+
+\end{flushleft}
+% Please keep the abstract below 300 words
+\section*{Abstract}
 Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
@@
 Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
-\end{abstract}
-
+
+% Please keep the Author Summary between 150 and 200 words
+% Use first person. PLOS ONE authors please skip this step. 
+% Author Summary not valid for PLOS ONE submissions.   
+\section*{Author summary}
+In genome-wide association studies (GWAS), thousands of genomes are scanned to identify variants associated with a complex trait.
+Over the last 15 years, GWAS have advanced our understanding of the genetics of complex diseases, and in particular of hereditary cancers.
+Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
+An elegant explanation has been proposed with the omnigenic model: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated, so that many other genes are able to alter the function of the core genes.
+These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
+In this study, we characterize six such network methods on GENESIS, a GWAS dataset for familial breast cancer in the French population.
+Combining these approaches allows us to identify potentially novel breast cancer susceptibility genes, and provides a mechanistic explanation for their role in the development of the disease.
+Our pipeline can easily be applied to other diseases.
+
+\linenumbers
+
+% Use "Eq" instead of "eqnarray" for eqnarray citations.
+% For figure citations, please use "Fig" instead of "Figure".
+% Place figure captions after the first paragraph in which they are cited.
 \section{Introduction}
@@
 
-\section{Methods}
-\subsection{GENESIS}
-
-The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
-Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
-Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
-We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
-The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
-It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
-
-\subsection{Preprocessing and quality control}
-
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
-A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
-In addition, we removed the samples with more than 10\% missing genotypes.
-After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
-Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
-The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
-
-We looked for population structure that could produce spurious associations.
-A principal component analysis revealed no visual differential population structure between cases and controls (Supplementary Figure~\ref{sfig:pcs}).
-Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
-
-\subsection{High-score subnetwork search algorithms}
-\subsubsection{SNP and gene association}
-\label{methods:node_score}
-To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
-Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
-Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P\=/values.
-We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
-We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
-Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
-Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
-
-We used such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
-For the former, we considered any gene that can be mapped to any of the selected SNPs as selected as well.
-For the latter, we considered all the SNPs that can be mapped to that gene as selected by the method.
-
-\subsubsection{Mathematical notations}
-\label{methods:notation}
-In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
-$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
-Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
-When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
-A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
-
-In addition to a weight, nodes have other properties, provided by the topology of the graph.
-In this article we focus on two of those: degree centrality, and betweenness centrality.
-The degree centrality, or degree, is the number of edges that a node has.
-The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.
-
-In addition, we use two matrices that describe different properties of a graph.
-The described matrices are square, and have as many rows and columns as nodes are in the network.
-The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
-The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; the diagonal is zero.
-The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
-
-\subsubsection{Networks}
-\label{methods:networks}
-
-\paragraph{Gene network}
-The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to run efficiently on networks of their size.
-We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
-Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
-12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
-The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
-
-Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
-In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
-Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
-
-\paragraph{SNP networks}
-SConES \cite{azencott_efficient_2013} is the only network method designed to handle SNP networks.
-As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
-\citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
-We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
-However, at different stages of this work we also used GS and GM for comparison.
-For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
-In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
-The properties of these three subnetworks are available in Supplementary Table \ref{tab:snp_solutions}.
-
-\subsubsection{Network methods}
-\label{methods:methods}
-
-\begin{table}[htbp]
-  \caption{Summary of the differences between the network methods.}
-  \label{tab:method_comparison}
-  \centering
-  \begin{threeparttable}
-    \begin{tabular}{l|llllllll}
-      Method & Field & Nodes & Exhaustive & Solution & Comp. & Input & Scoring & Ref.\\
-      \hline
-      dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
-      heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
-      HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
-      LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
-      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
-      SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \cite{liu_sigmod:_2017}\\
-    \end{tabular}
-    \begin{tablenotes}
-      \footnotesize{
-        \item \textbf{\emph{Field}}: field in which the algorithm was developed.
-\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
-\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
-\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than containing high scoring, connected nodes.
-\textbf{\emph{Comp.}}: number of connected components in the solution.
-\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
-\textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
-In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
-        cumulative distribution function of the standard Normal distribution. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
-      }
-    \end{tablenotes}
-  \end{threeparttable}
-\end{table}
-
-Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
-Unsurprisingly, multiple solutions have been proposed.
-We examined six of them: five that explore the PPIN, and one which explores SNP networks.
-We selected methods that were open source, had an implementation available, and an accessible documentation.
-Their main differences are summarized in Table~\ref{tab:method_comparison}.
-
-\begin{description}
-\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
-To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
-
-\begin{enumerate}
-\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
-
-\begin{equation*} 
-Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
-\end{equation*}
-
-where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
-\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
-\item Repeat 2-4 until no further enlargement is possible.
-\item Add $S_i$ to the list of subnetworks to return. Its Z-score is normalized as
-
-\begin{equation*}
-Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
-\end{equation*}
-
-where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
-\end{enumerate}
-
-DmGWAS carries out this process on every gene in the network.
-We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
-We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
-\end{description}
-
-\begin{description}
-\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
-This transformation is achieved by modelling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
-Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
-To solve it efficiently it is re-cast as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
-
-\begin{equation*}
-p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
-\end{equation*}
-
-were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
-All three are positive quantities.
-Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
-We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
-
-\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}.
-To that end, it considers both the local topology of the network and the scores of the nodes.
-The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
-This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
-$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
-\begin{equation*} 
-E = F \operatorname{diag}(w(V)), 
-\end{equation*}
-where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
-To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
-Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
-We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
-In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
-HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
-Both parameters are set automatically by the algorithm, which is robust to their values \cite{leiserson_pan-cancer_2015}.
-HotNet2 is implemented in Python \cite{hotnet2}.
-
-\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
-By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
-For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
-Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
-The minimum of these \emph{k} P\=/values is the score of the subnetwork.
-This score is transformed into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
-Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
-We used the implementation of LEAN from the LEANR R package \cite{leanr}.
-\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem
-
-\begin{equation}
-\label{eq:scones}
-\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
-\end{equation}
-
-where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
-The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
-Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
-As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
-In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
-We used the version on SConES implemented in the R package martini \cite{martini}.
-\end{description}
-
-\begin{description}
-\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
-
-\begin{equation*}
-\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
-\end{equation*}
-
-As SConES, this optimization problem can also be solved by a graph min-cut approach.
-
-SigMod presents three important differences with SConES.
-First it is designed for gene-gene networks.
-Second, it favors subnetworks containing many edges.
-SConES, instead, penalizes connections between the selected and unselected nodes.
-Third, it explores the grid of hyperparameters differently, and processes their respective solutions.
-Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
-Such a large change implies that the network is densely interconnected.
-This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other.
-A score is assigned to each candidate solution by summing their node scores and normalizing by size.
-The candidate solution with the highest standardized score is the chosen solution.
-SigMod is implemented in an R package \cite{sigmod}.
-\end{description}
-
-\begin{description}
-\item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
-\end{description}
-
-\subsection{Pathway enrichment analysis}
-\label{methods:pathway_enrichment}
-
-We searched for pathways enriched in the gene subnetworks produced by the above methods.
-We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
-The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
-We adjusted the P-values for multiple testing as in \citet{Benjamini1995} (BH).Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
-As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
-
-\subsection{Evaluation of methods}
-\label{methods:comparison}
-
-We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
-We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
-When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
-We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
-
-\subsubsection{Properties ot the solution}
-\label{methods:algorithm_comparison}
-
-We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
-The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by \citet{nogueira_measuring_2016}.
-
-\subsubsection{Classification accuracy of selected SNPs}
-\label{methods:classifier}
-
-A desirable solution offers good predictive power on unseen \emph{test} samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
-We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
-This penalty was set by cross-validation, choosing the value that minimized misclassification error.
-We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
-When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
-Then we evaluated the sensitivity and the specificity on the \emph{test set}.
-The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
-To obtain a baseline, we also trained the classifier on all the SNPs.
-We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
-Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
-However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
-
-\subsubsection{Comparison to state-of-the-art}
-\label{methods:bcac}
-
-An alternative way to evaluate the results is comparing our results to an external dataset.
-For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}.
-Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
-
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
-Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focus on the French high-risk population attending the family cancer clinics.
-Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
-For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
-We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
-
-\subsection{Code availability}
-
-We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
-For each of those processes, we created a pipeline with a clear interface that should work on any GWAS dataset.
-They are compiled in \url{https://github.com/hclimente/gwas-tools}.
-Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
-We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
-
+% Results and Discussion can be combined.
 \section{Results}
@@
 We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
-Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
+Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
 Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
-In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
+In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (\nameref{sfig:snp_gene_manhattan}B).
 
@@
 We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
-It selected 100 SNPs, both from all aforementioned regions and new ones (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
+It selected 100 SNPs, both from all aforementioned regions and new ones (\nameref{sfig:snp_gene_manhattan}C).
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
@@
 This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
-In fact, such methods provided comparably (poor) classification performance to L1-penalized logistic regression (Figure~\ref{fig:benchmark}B), while providing more interpretable solutions.
+In fact, such methods provided comparably (poor) classification performance to L1-penalized logistic regression (Fig~\ref{fig:benchmark}B), while providing more interpretable solutions.
 
@@
 \label{results:separate_networks}
-\begin{table}[htbp]
+
+We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
+As none of the networks examined by LEAN was significant (BH adjusted P\=/value < 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+These solutions differ in many aspects, making it hard to draw joint conclusions.
+For starters, the overlap between the genes featured in each solution is quite small (Fig~\ref{fig:solution_overview}B).
+Another prominent difference is their size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
+While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
+Their topologies also differ, as measured by the median centrality (Table~\ref{tab:gene_solutions}) and the number of connected components (Fig~\ref{fig:solution_overview}D).
+Only two methods have more than one connected component: SConES, as described above, and HotNet2.
+HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
+The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
+Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
+It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
+Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
+
+\begin{table}[!ht]
+  \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
   \centering
-  \begin{threeparttable}
-\caption{Summary statistics on the solutions of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
-\label{tab:gene_solutions}
-\centering
+  \caption{
+  {\bf Summary statistics on the solutions of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}}
   \begin{tabular}{lrrrrr}
-Network & \# genes & \# edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & \(\rho\)\textsubscript{consensus}\\
-\hline
+  {\bf Network } & {\bf \# genes } & {\bf \# edges } & {\bf \mean{Betweenness} } & {\bf \median{P}\textsubscript{gene} } & {\bf \(\rho\)\textsubscript{consensus} }\\
+  \thickhline
   HINT HT & 13\,619 & 142\,541 & 16\,706 & 0.46 & 0.066\\
@@
   \end{tabular}
-\begin{tablenotes}
-  \footnotesize{
-    \item \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
+  \begin{flushleft} \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
 \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
@@
 \textbf{\(\rho\)\textsubscript{consensus}}: Pearson correlation between the subnetwork and the consensus network.
-  }
-\end{tablenotes}
-\end{threeparttable}
+  \end{flushleft}
+  \label{tab:gene_solutions}
+  \end{adjustwidth}
 \end{table}
 
-\begin{figure}[htbp]
+\begin{figure}[!ht]
   \centering
-  \includegraphics[height=.75\textheight]{./figures/figure_1.png}
-  \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-As LEAN did not produce any significant gene (BH adjusted P\=/value~<~0.05), it was excluded.
+  \includegraphics[height=\textwidth]{./figures/figure_1.png}
+  \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
+} As LEAN did not produce any significant gene (BH adjusted P\=/value~<~0.05), it was excluded.
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
@@
 
-We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (BH adjusted P\=/value < 0.05), we obtained six solutions (Figure~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
-These solutions differ in many aspects, making it hard to draw joint conclusions.
-For starters, the overlap between the genes featured in each solution is quite small (Figure~\ref{fig:solution_overview}B).
-Another prominent difference is their size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
-While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
-Their topologies also differ, as measured by the median centrality (Table~\ref{tab:gene_solutions}) and the number of connected components (Figure~\ref{fig:solution_overview}D).
-Only two methods have more than one connected component: SConES, as described above, and HotNet2.
-HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
-The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
-Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
-It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
-Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
-
 Despite their differences, there are additional common themes.
-All obtained solution subnetworks have lower association P\=/values than the whole PPIN (median P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Figure~\ref{fig:solution_overview}A).
+All obtained solution subnetworks have lower association P\=/values than the whole PPIN (median P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Fig~\ref{fig:solution_overview}A).
 This exemplifies the trade-off between statistical significance and biological relevance.
@@
 This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
 
@@
 \label{results:consensus}
-\begin{figure}[htbp]
+
+Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
+Indeed, only 20 genes are common to more than two solutions (\nameref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (\nameref{sfig:consensus_stats}B).
+To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
+This subnetwork (Fig~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
+
+\begin{figure}[!ht]
   \centering
-  \includegraphics[width=.5\linewidth]{./figures/figure_3.pdf}
-  \caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
+  \includegraphics[width=.7\linewidth]{./figures/figure_3.pdf}
+  \caption{ {\bf Consensus subnetwork on GENESIS (Section~\ref{methods:methods}) }.
 Each node is represented by a pie chart, which shows the methods that selected it.
@@
 The latter ones are also colored in pink.
-All gene names are indicated in Supplementary Figure \ref{sfig:consensus-names}.}
+All gene names are indicated in \nameref{sfig:consensus-names}.}
   \label{fig:consensus}
 \end{figure}
-
-Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
-Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
-To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
-This subnetwork (Figure~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
 
@@
 Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
-The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
+The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}D).
 This suggests that they were selected because they were on the shortest path between two highly associated genes.
@@
 The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
-The gene-based network methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (Figure~\ref{fig:solution_overview}C).
+The gene-based network methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (Fig~\ref{fig:solution_overview}C).
 Interestingly, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
@@
 
-\begin{figure}[htbp]
+We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+Specifically, we measured five properties (Fig~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
+The solution size varies greatly between the different methods (Fig~\ref{fig:benchmark}A).
+Heinz produced the smallest solutions, with an average of 182 selected SNPs.
+The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
+To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Fig~\ref{fig:benchmark}B).
+The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
+Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
+It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics[width=.72\linewidth]{./figures/figure_4.pdf}
-\caption{Comparison of network-based GWAS methods on GENESIS.
-Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+  \includegraphics[width=\linewidth]{./figures/figure_4.pdf}
+  \caption{{ \bf Comparison of network-based GWAS methods on GENESIS.
+} Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
 As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
@@
 
-We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
-Specifically, we measured five properties (Figure~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
-The solution size varies greatly between the different methods (Figure~\ref{fig:benchmark}A).
-Heinz produced the smallest solutions, with an average of 182 selected SNPs.
-The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
-To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Figure~\ref{fig:benchmark}B).
-The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
-Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
-
 Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
-Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Figure~\ref{fig:benchmark}C).
+Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Fig~\ref{fig:benchmark}C).
 Conversely, the other methods displayed similarly low stabilities.
 
-In terms of computational runtime, the fastest method was heinz (Figure~\ref{fig:benchmark}D), which returned a solution in a few seconds.
+In terms of computational runtime, the fastest method was heinz (Fig~\ref{fig:benchmark}D), which returned a solution in a few seconds.
 HotNet2 was the slowest (3 days and 14 hours on average).
@@
 
-\begin{figure}[htbp]
+As shown above, and despite their similarities, the network methods produced remarkably different solutions.
+This is due to the particularities of each methods, and directly or indirectly provide information about the dataset.
+For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
+
+In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value for the subnetwork (Table~\ref{tab:gene_solutions}).
+Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Fig~\ref{fig:benchmark}C).
+Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
+Due to the limited signal at the gene level in this dataset (\nameref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
+Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
+Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
+
+On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
+dmGWAS' subnetwork is the least associated subnetwork on average.
+This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
+It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
+This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
+This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
+In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
+Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
+
+\begin{figure}[!ht]
   \centering
-\includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
-\caption{Drawbacks encountered when using network guided methods.
-\textbf{(A)}~DmGWAS solution subnetwork.
+  \includegraphics[width=\linewidth]{./figures/figure_2.pdf}
+  \caption{{\bf Drawbacks encountered when using network guided methods.} \textbf{(A)}~DmGWAS solution subnetwork.
 Genes with a P\=/value < 0.1 are highlighted in red.
@@
 
-As shown above, and despite their similarities, the network methods produced remarkably different solutions.
-This is due to the particularities of each methods, and directly or indirectly provide information about the dataset.
-For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
-
-In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value for the subnetwork (Table~\ref{tab:gene_solutions}).
-Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Figure~\ref{fig:benchmark}C).
-Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
-Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
-Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
-
-On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
-dmGWAS' subnetwork is the least associated subnetwork on average.
-This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
-It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
-This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Figure~\ref{fig:issues}A).
-This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
-In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
-Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
-
 Another peculiarity of network methods is their relationship to degree centrality.
-On the one hand we observed that highly central genes often had no association to disease (Figure~\ref{fig:issues}B).
+On the one hand we observed that highly central genes often had no association to disease (Fig~\ref{fig:issues}B).
 On the other, network methods favor central genes, as they often connect high scoring nodes.
@@
 In fact, due to linkage disequilibrium, SConES favors such genes, as selecting SNPs in an LD-block which overlaps with a gene favors selecting the rest of the gene.
-This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Supplementary Figure~\ref{sfig:pearson_methods}).
+This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (\nameref{sfig:pearson_methods}).
 On the other hand, SConES penalizes selecting SNPs and not their neighbors.
 This makes it conservative regarding SNPs with many interactions, like those mapped to hub genes in the PPIN.
-For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure~\ref{fig:issues}C).
+For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Fig~\ref{fig:issues}C).
 In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
@@
 
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.5\linewidth]{./figures/figure_5.pdf}
-  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
-Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-All gene names are indicated in Supplementary Figure \ref{sfig:stable-consensus-names}.}
-  \label{fig:stable-consensus}
-\end{figure}
-
 Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
@@
 We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once.
-The resulting stablility-based consensus was composed of 68 genes (Figure \ref{fig:stable-consensus}).
+The resulting stablility-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}).
 This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=.7\linewidth]{./figures/figure_5.pdf}
+  \caption{{\bf Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+All gene names are indicated in \nameref{sfig:stable-consensus-names}.}
+  \label{fig:stable-consensus}
+  \end{figure}
 
@@
 Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
-The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure~\ref{sfig:biotypes_excluded}).
+The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (\nameref{sfig:biotypes_excluded}).
 Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
@@
 This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
-On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
-When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
+On the other hand, Huang et al.
+\cite{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
+When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
 This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the solutions.
@@
 Importantly, those methods that had a permissive license were compiled into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
+
+\section{Materials and methods}
+
+\subsection{GENESIS}
+
+The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
+Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
+Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
+We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
+The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
+It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
+
+\subsection{Preprocessing and quality control}
+
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
+A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
+In addition, we removed the samples with more than 10\% missing genotypes.
+After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
+The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
+
+We looked for population structure that could produce spurious associations.
+A principal component analysis revealed no visual differential population structure between cases and controls (\nameref{sfig:pcs}).
+Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
+
+\subsection{High-score subnetwork search algorithms}
+\subsubsection{SNP and gene association}
+\label{methods:node_score}
+To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
+Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
+Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P\=/values.
+We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
+We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
+Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
+Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
+
+We used such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
+For the former, we considered any gene that can be mapped to any of the selected SNPs as selected as well.
+For the latter, we considered all the SNPs that can be mapped to that gene as selected by the method.
+
+\subsubsection{Mathematical notations}
+\label{methods:notation}
+In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
+$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
+Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
+When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
+A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
+
+In addition to a weight, nodes have other properties, provided by the topology of the graph.
+In this article we focus on two of those: degree centrality, and betweenness centrality.
+The degree centrality, or degree, is the number of edges that a node has.
+The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.
+
+In addition, we use two matrices that describe different properties of a graph.
+The described matrices are square, and have as many rows and columns as nodes are in the network.
+The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
+The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; the diagonal is zero.
+The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
+
+\subsubsection{Networks}
+\label{methods:networks}
+
+\paragraph{Gene network}
+The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
+Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to run efficiently on networks of their size.
+We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
+Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
+Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
+12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
+The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
+
+Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
+In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
+Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
+
+\paragraph{SNP networks}
+SConES \cite{azencott_efficient_2013} is the only network method designed to handle SNP networks.
+As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
+Azencott et al.
+\cite{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
+We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
+However, at different stages of this work we also used GS and GM for comparison.
+For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
+In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
+The properties of these three subnetworks are available in \nameref{stab:snp_solutions}.
+
+\subsubsection{Network methods}
+\label{methods:methods}
+
+Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
+But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+Unsurprisingly, multiple solutions have been proposed.
+We examined six of them: five that explore the PPIN, and one which explores SNP networks.
+We selected methods that were open source, had an implementation available, and an accessible documentation.
+Their main differences are summarized in Table~\ref{tab:method_comparison}.
+
+\begin{table}[!ht]
+  \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+  \centering
+  \caption{
+  {\bf Summary of the differences between the network methods.}}
+  \begin{tabular}{l|llllllll}
+  {\bf Method} & {\bf Field } & {\bf Nodes} & {\bf Exhaustive} & {\bf Solution} & {\bf Comp.} & {\bf Input} & {\bf Scoring} & {\bf Ref.}\\ \thickhline
+  dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
+  heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
+  HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
+  LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
+  SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
+  SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \cite{liu_sigmod:_2017}\\
+  \end{tabular}
+  \begin{flushleft} \textbf{\emph{Field}}: field in which the algorithm was developed.
+\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
+\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
+\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than containing high scoring, connected nodes.
+\textbf{\emph{Comp.}}: number of connected components in the solution.
+\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
+\textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
+In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the cumulative distribution function of the standard Normal distribution.
+\textbf{\emph{Ref.}}: original publication featuring the algorithm.
+  \end{flushleft}
+  \label{tab:method_comparison}
+  \end{adjustwidth}
+\end{table}
+
+\begin{description}
+\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
+To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
+
+\begin{enumerate}
+\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
+\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
+
+\begin{eqnarray} 
+Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
+\end{eqnarray}
+
+where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
+\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
+\item Repeat 2-4 until no further enlargement is possible.
+\item Add $S_i$ to the list of subnetworks to return. Its Z-score is normalized as
+
+\begin{eqnarray}
+Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
+\end{eqnarray}
+
+where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
+\end{enumerate}
+
+DmGWAS carries out this process on every gene in the network.
+We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
+We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
+We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
+\end{description}
+
+\begin{description}
+\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
+The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
+This transformation is achieved by modelling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
+Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
+To solve it efficiently it is re-cast as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
+
+\begin{eqnarray}
+p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
+\end{eqnarray}
+
+were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
+All three are positive quantities.
+Heinz implements the algorithm from Ljubić et al.
+\cite{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
+We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
+
+\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}.
+To that end, it considers both the local topology of the network and the scores of the nodes.
+The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
+This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
+$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
+\begin{eqnarray} 
+E = F \operatorname{diag}(w(V)), 
+\end{eqnarray}
+where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
+For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
+To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
+Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
+We assigned initial node scores as in Nakka et al.
+\cite{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
+HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
+Both parameters are set automatically by the algorithm, which is robust to their values \cite{leiserson_pan-cancer_2015}.
+HotNet2 is implemented in Python \cite{hotnet2}.
+
+\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
+By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
+For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
+Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
+The minimum of these \emph{k} P\=/values is the score of the subnetwork.
+This score is transformed into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
+Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
+We used the implementation of LEAN from the LEANR R package \cite{leanr}.
+\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem
+
+\begin{eqnarray}
+\label{eq:scones}
+\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
+\end{eqnarray}
+
+where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
+The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
+Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
+As in Azencott et al.
+\cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
+In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
+We used the version on SConES implemented in the R package martini \cite{martini}.
+\end{description}
+
+\begin{description}
+\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
+
+\begin{eqnarray}
+\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
+\end{eqnarray}
+
+As SConES, this optimization problem can also be solved by a graph min-cut approach.
+
+SigMod presents three important differences with SConES.
+First it is designed for gene-gene networks.
+Second, it favors subnetworks containing many edges.
+SConES, instead, penalizes connections between the selected and unselected nodes.
+Third, it explores the grid of hyperparameters differently, and processes their respective solutions.
+Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
+Such a large change implies that the network is densely interconnected.
+This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other.
+A score is assigned to each candidate solution by summing their node scores and normalizing by size.
+The candidate solution with the highest standardized score is the chosen solution.
+SigMod is implemented in an R package \cite{sigmod}.
+\end{description}
+
+\begin{description}
+\item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
+\end{description}
+
+\subsection{Pathway enrichment analysis}
+\label{methods:pathway_enrichment}
+
+We searched for pathways enriched in the gene subnetworks produced by the above methods.
+We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
+The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
+We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH).
+Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
+As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
+
+\subsection{Evaluation of methods}
+\label{methods:comparison}
+
+We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
+We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
+When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
+We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
+
+\subsubsection{Properties ot the solution}
+\label{methods:algorithm_comparison}
+
+We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
+The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by Nogueira and Brown \cite{nogueira_measuring_2016}.
+
+\subsubsection{Classification accuracy of selected SNPs}
+\label{methods:classifier}
+
+A desirable solution offers good predictive power on unseen \emph{test} samples.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
+The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
+This penalty was set by cross-validation, choosing the value that minimized misclassification error.
+We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
+When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
+Then we evaluated the sensitivity and the specificity on the \emph{test set}.
+The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
+To obtain a baseline, we also trained the classifier on all the SNPs.
+We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
+Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
+However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
+
+\subsubsection{Comparison to state-of-the-art}
+\label{methods:bcac}
+
+An alternative way to evaluate the results is comparing our results to an external dataset.
+For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}.
+Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
+
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
+Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focus on the French high-risk population attending the family cancer clinics.
+Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
+For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
+We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
+
+\subsection{Code availability}
+
+We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
+For each of those processes, we created a pipeline with a clear interface that should work on any GWAS dataset.
+They are compiled in \url{https://github.com/hclimente/gwas-tools}.
+Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
+
+\section{Supporting information}
+
+% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
+\paragraph*{S1 Table.}
+\label{stab:snp_solutions}
+{\bf Summary statistics on the results of SConES on the three SNP-SNP interaction networks.} The first row within each block contains the summary statistics on the whole network.
+
+\paragraph*{S2 Table.}
+\label{stab:scones_gene_solutions}
+{\bf Summary statistics on the results of multiple network methods on the gene-gene interaction network.} The first row contains the summary statistics on the whole network.
+
+\paragraph*{S1 Fig.}
+\label{sfig:pcs}
+{\bf GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
+Cases are colored in green, controls in orange.
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
+
+\paragraph*{S2 Fig.}
+\label{sfig:snp_gene_manhattan}
+{\bf Association in GENESIS.
+The red line represents the Bonferroni threshold.} \textbf{(A)} SNP association, measured from the outcome of a 1 d.f.
+$\chi^2$ allelic test.
+Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
+The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
+\textbf{(B)} Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values.
+The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
+\textbf{(C)} SNP association as in panel (A).
+The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).
+%\includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
+
+\paragraph*{S3 Fig.}
+\label{sfig:pearson_methods}
+{\bf Pearson correlation between the different solution subnetworks.
+} \textbf{(A)} Correlation between selected SNPs.
+\textbf{(B)} Correlation between selected genes.
+In general, the solutions display a very low overlap.
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_3.pdf}
+
+\paragraph*{S4 Fig.}
+\label{sfig:consensus-names}
+{\bf Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
+} \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+This panel is identical to Fig~\ref{fig:consensus}.
+\textbf{(B)} Same network, but every gene name is indicated.
+% \includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
+
+\paragraph*{S5 Fig.}
+\label{sfig:consensus_stats}
+{\bf Genes on the consensus network.
+Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
+} \textbf{(A)} Number of methods selecting every gene in the subnetwork.
+\textbf{(B)} VEGAS2 P\=/values of association of the genes, with regards to the number of methods that selected them.
+\textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network.
+To improve visualization, we removed outliers.
+\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus network.
+The blue line represents a fitted generalized linear model.
+% \includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
+
+\paragraph*{S6 Fig.}
+\label{sfig:stable-consensus-names}
+{\bf Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
+} \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+This panel is identical to Fig~\ref{fig:stable-consensus}.
+\textbf{(B)} Same network, but every gene name is indicated.
+% \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf}
+
+\paragraph*{S7 Fig.}
+\label{sfig:biotypes_excluded}
+{\bf Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
+
+\paragraph*{S8 Fig.}
+\label{sfig:lc_ht_comparison}
+{\bf Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
+} Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
+\textbf{(A)} Ratios of the selected features between both benchmarks and of the active set.
+\textbf{(B)} Shifts in sensitivity and specificity.
+\textbf{(C)} Shift in Pearson correlation between benchmarks.
+\textbf{(D)} Ratio between the runtimes of the benchmarks.
+For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
+In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.
+% \includegraphics[width=.85\linewidth]{./figures/sfigure_8.pdf}
+
+\paragraph*{S9 Fig.}
+\label{sfig:scones_gsm}
+{\bf Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)} Manhattan plots of SNPs; in black,the method’s solution.
+The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
+\textbf{(B)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
+For reference, we added a gray line with a slope of 1.
+\textbf{(C)} Solution networks.
+% \includegraphics[width=\linewidth]{./figures/sfigure_9.png}
+
+\section*{Acknowledgments}
+We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie), the biological resource centre and all the GENESIS collaborating cancer clinics clinics.
 
@@
 
-\section*{Funding and acknowledgments}
-
-This project was supported by funding from Agence Nationale de la Recherche (ANR-18-CE45-0021-01) and from the European Union’s Horizon 2020 research and innovation program (Marie Skłodowska-Curie [666003]).
-Financial support for GENESIS resource and genotyping was provided by the Ligue Nationale contre le Cancer (grants PRE05/DSL, PRE07/DSL, PRE11/NA), the French National Institute of Cancer (INCa grant b2008-029/LL-LC) and the comprehensive cancer center SiRIC (Site de Recherche Intégrée sur le Cancer: INCa-DGOS-4654).
-
-We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S.
-Eon-Marchais, M.
-Marcou, D.
-Le Gal, L.
-Toulemonde, J.
-Beauvallet, N.
-Mebirouk, E.
-Cavaciuti), the biological resource centre (S.
-Mazoyer, F.
-Damiola, L.
-Barjhoux, C.
-Verny-Pierre, V.
-Sornin) and all the GENESIS collaborating cancer clinics clinics (Clinique Sainte Catherine, Avignon: H.
-Dreyfus; Hôpital Saint Jacques, Besançon: M-A.
-Collonge-Rame; Institut Bergonié, Bordeaux: M.Longy, A.
-Floquet, E.
-Barouk-Simonet; CHU, Brest: S.
-Audebert; Centre François Baclesse, Caen: P.
-Berthet; Hôpital Dieu, Chambéry: S.
-Fert-Ferrer; Centre Jean Perrin, Clermont-Ferrand: Y-J.
-Bignon; Hôpital Pasteur, Colmar: J-M.
-Limacher; Hôpital d’Enfants CHU – Centre Georges François Leclerc, Dijon: L.
-Faivre-Olivier; CHU, Fort de France: O.
-Bera; CHU Albert Michallon, Grenoble: D.
-Leroux; Hôpital Flaubert, Le Havre: V.
-Layet; Centre Oscar Lambret, Lille: P.
-Vennin, C.
-Adenis; Hôpital Jeanne de Flandre, Lille: S.
-Lejeune-Dumoulin, S.
-Manouvier-Hanu; CHRU Dupuytren, Limoges: L.
-Venat-Bouvet; Centre Léon Bérard, Lyon: C.
-Lasset, V.
-Bonadona; Hôpital Edouard Herriot, Lyon: S.
-Giraud; Institut Paoli-Calmettes, Marseille: F.
-Eisinger, L.
-Huiart; Centre Val d’Aurelle – Paul Lamarque, Montpellier: I.
-Coupier; CHU Arnaud de Villeneuve, Montpellier: I.
-Coupier, P.
-Pujol; Centre René Gauducheau, Nantes: C.
-Delnatte; Centre Catherine de Sienne, Nantes: A.
-Lortholary; Centre Antoine Lacassagne, Nice: M.
-Frénay, V.
-Mari; Hôpital Caremeau, Nîmes: J.
-Chiesa; Réseau Oncogénétique Poitou Charente, Niort: P.
-Gesta; Institut Curie, Paris: D.
-Stoppa-Lyonnet, M.
-Gauthier-Villars, B.
-Buecher, A.
-de Pauw, C.
-Abadie, M.
-Belotti; Hôpital Saint-Louis, Paris: O.
-Cohen-Haguenauer; Centre Viggo-Petersen, Paris: F.
-Cornélis; Hôpital Tenon, Paris: A.
-Fajac; GH Pitié Salpétrière et Hôpital Beaujon, Paris: C.
-Colas, F.
-Soubrier, P.
-Hammel, A.
-Fajac; Institut Jean Godinot, Reims: C.
-Penet, T.
-D.
-Nguyen; Polyclinique Courlancy, Reims: L.
-Demange, C.
-Penet; Centre Eugène Marquis, Rennes: C.
-Dugast; Centre Henri Becquerel, Rouen: A.
-Chevrier, T.
-Frebourg, J.
-Tinat, I.
-Tennevet, A.
-Rossi; Hôpital René Huguenin/Institut Curie, Saint Cloud: C.
-Noguès, L.
-Demange, E.
-Mouret-Fourme; CHU, Saint-Etienne: F.
-Prieur; Centre Paul Strauss, Strasbourg: J-P.
-Fricker, H.
-Schuster; Hôpital Civil, Strasbourg: O.
-Caron, C.
-Maugard; Institut Claudius Regaud, Toulouse: L.
-Gladieff, V.
-Feillel; Hôpital Bretonneau, Tours: I.
-Mortemousque; Centre Alexis Vautrin, Vandoeuvre-les-Nancy: E.
-Luporsi; Hôpital de Bravois, Vandoeuvre-les-Nancy: P.
-Jonveaux; Gustave Roussy, Villejuif: A.
-Chompret, O.
-Caron).
-
-\bibliographystyle{abbrvnat}
-\bibliography{bibliography}
-
-\clearpage
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\setcounter{table}{0}
-
-\section{Supplementary materials}
-
-\begin{table}[htbp]
-\begin{threeparttable}
-\caption{Summary statistics on the results of SConES on the three SNP-SNP interaction networks. The first row within each block contains the summary statistics on the whole network.}
-\label{tab:snp_solutions}
-\centering
-\begin{tabular}{lrrllr}
-Network & SNPs & Edges & Subnetworks & \mean{Betweenness} & \median{P}\textsubscript{SNP}\\
-\hline
-GS & 197\,083 & 197\,060 & - & 2.03 \texttimes{} 10\textsuperscript{7} & 0.49\\
-SConES GS & 1\,590 & 1\,585 & 5 & 2.52 \texttimes{} 10\textsuperscript{7} & 0.023\\
-\hline
-GM & 197\,083 & 6\,442\,446 & - & 3.99 \texttimes{} 10\textsuperscript{6} & 0.49\\
-SConES GM & 1\,692 & 177\,611 & 5 & 4.40 \texttimes{} 10\textsuperscript{6} & 0.055\\
-\hline
-GI & 197\,083 & 28\,733\,720 & - & 1.46 \texttimes{} 10\textsuperscript{6} & 0.49\\
-SConES GI & 408 & 539 & 5 & 9.33 \texttimes{} 10\textsuperscript{6} & 0.076\\
-\end{tabular}
-\begin{tablenotes}
-  \footnotesize{
-    \item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.\median{P}\textsubscript{SNP}: median P\=/value of the selected SNPs.
-  }
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[htbp]
-\begin{threeparttable}
-  \caption{Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
-\label{tab:scones_gene_solutions}
-\centering
-\begin{tabular}{lrrrlr}
-Network & Genes & Edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & $\rho_{consensus}$\\
-\hline
-SConES GS & 5 & 0 & 9\,805 & 2.7 \texttimes{} 10\textsuperscript{-5} & 0.19\\
-SConES GM & 28 & 2 & 4\,267 & 0.067 & 0.12\\
-\end{tabular}
-\begin{tablenotes}
-  \footnotesize{
-    \item \mean{Betweenness}: mean betweenness of the selected genes in the full network.
-\median{P}\textsubscript{gene}: median P\=/value of the selected genes; $\rho_{consensus}$: Pearson correlation with the consensus network.
-  }
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
-\caption{GENESIS shows no differential population structure between cases and controls.
-\textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
-Cases are colored in green, controls in orange.}
-\label{sfig:pcs}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
-  \caption{Association in GENESIS.
-The red line represents the Bonferroni threshold.
-\textbf{(A)} SNP association, measured from the outcome of a 1 d.f.
-$\chi^2$ allelic test.
-Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
-The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
-\textbf{(B)} Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values.
-The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
-\textbf{(C)} SNP association as in panel (A).
-The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).}
-  \label{sfig:snp_gene_manhattan}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_3.pdf}
-\caption{Pearson correlation between the different solution subnetworks.
-\textbf{(A)} Correlation between selected SNPs.
-\textbf{(B)} Correlation between selected genes.
-In general, the solutions display a very low overlap.}
-\label{sfig:pearson_methods}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
-\caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
-\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-This panel is identical to Figure~\ref{fig:consensus}.
-\textbf{(B)} Same network, but every gene name is indicated.}
-\label{sfig:consensus-names}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
-\caption{Genes on the consensus network.
-Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
-\textbf{(A)} Number of methods selecting every gene in the subnetwork.
-\textbf{(B)} VEGAS2 P\=/values of association of the genes, with regards to the number of methods that selected them.
-\textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network.
-To improve visualization, we removed outliers.
-\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus network.
-The blue line represents a fitted generalized linear model.}
-\label{sfig:consensus_stats}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf}
-  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
-\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-This panel is identical to Figure~\ref{fig:stable-consensus}.
-\textbf{(B)} Same network, but every gene name is indicated.}
-  \label{sfig:stable-consensus-names}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
-\caption{Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
-\label{sfig:biotypes_excluded}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.85\linewidth]{./figures/sfigure_8.pdf}
-\caption{Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
-Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
-\textbf{(A)} Ratios of the selected features between both benchmarks and of the active set.
-\textbf{(B)} Shifts in sensitivity and specificity.
-\textbf{(C)} Shift in Pearson correlation between benchmarks.
-\textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
-In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}
-\label{sfig:lc_ht_comparison}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\linewidth]{./figures/sfigure_9.png}
-  \caption{Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.
-\textbf{(A)} Manhattan plots of SNPs; in black,the method’s solution.
-The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
-\textbf{(B)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(C)} Solution networks.}
-  \label{sfig:scones_gsm}
-\end{figure}
+\nolinenumbers
+
+\bibliography{bibliography.bib}
+
+% Either type in your references using
+% \begin{thebibliography}{}
+% \bibitem{}
+% Text
+% \end{thebibliography}
+%
+% or
+%
+% Compile your BiBTeX database using our plos2015.bst
+% style file and paste the contents of your .bbl file
+% here. See http://journals.plos.org/plosone/s/latex for 
+% step-by-step instructions.
+%
 

### 2020-08-21 manuscript/plos/review_1.tex
added computational resources, explicit how Pearson correlation is computed, and first grammarly pass

--- 2020-08-20 manuscript/plos/original.tex
@@
 Specifically, we examine their utility to discover new biomarkers for breast cancer susceptibility by interrogating a genome-wide association study (GWAS) focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
-We perform an in-depth benchmarking of the methods with regards to size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
+We perform an in-depth benchmarking of the methods with regards to the size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
 By trading statistical stringency for biological meaningfulness, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
@@
 The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
-The network seems organized around \emph{CUL3}, encoding an ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
-In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely lack of statistical power and of interpretable solutions.
+The network seems organized around \emph{CUL3}, encoding a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
+In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely a lack of statistical power and interpretable solutions.
 Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
@@
 Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
-An elegant explanation has been proposed with the omnigenic model: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated, so that many other genes are able to alter the function of the core genes.
+An elegant explanation has been proposed with the omnigenic model: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated so that many other genes can alter the function of the core genes.
 These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
@@
 Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the previously mentioned Bonferroni correction.
-However, Bonferroni correction is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
+However, Bonferroni correction is overly conservative when the statistical tests are correlated, as it is the case in GWAS \cite{wang_statistical_2018}.
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
@@
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
-Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the PPIN (Section~\ref{methods:networks}).
+Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the protein-protein interaction network (PPIN, Section~\ref{methods:networks}).
 In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
@@
 We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (BH adjusted P\=/value < 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+As none of the networks examined by LEAN was significant (Benjamini-Hochberg [BH] correction adjusted P\=/value < 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
 These solutions differ in many aspects, making it hard to draw joint conclusions.
@@
 It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
-Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
+Interestingly, the dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
 But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
@@
 This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
 
@@
 Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
-With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
+With regards to the attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
 The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making hard to disentangle their effects.
+In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making it hard to disentangle their effects.
 \emph{HSPA1A} was the most strongly associated gene (VEGAS2 P\=/value~=~8.37~\texttimes{}~10\textsuperscript{-4}).
@@
 In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signalling pathways.
+For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signaling pathways.
 \emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
@@
 We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
-Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
+Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
@@
 
-We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:comparison}).
+We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:benchmark}).
 Specifically, we measured five properties (Fig~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
@@
   \caption{{ \bf Comparison of network-based GWAS methods on GENESIS.
-} Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:comparison}).
+} Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:benchmark}).
 As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
@@
 For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs.
-\textbf{(B)} Sensitivity and specificity on test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
-In addition, the performance of the classifier trained on all SNPs is displayed.
+\textbf{(B)} Sensitivity and specificity on the test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
+The performance of the classifier trained on all SNPs is also displayed.
 Points are the average over the 5 runs; lines represent the standard error of the mean.
@@
 Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
-Both heinz and LEAN displayed a high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Fig~\ref{fig:benchmark}C).
+Both heinz and LEAN displayed high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Fig~\ref{fig:benchmark}C).
 Conversely, the other methods displayed similarly low stabilities.
@@
 As shown above, and despite their similarities, the network methods produced remarkably different solutions.
-This is due to the particularities of each methods, and directly or indirectly provide information about the dataset.
+This is due to the particularities of each method, and directly or indirectly provide information about the dataset.
 For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
@@
 This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
-It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
+It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated gene if it was linked to another, strongly associated one.
 This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
@@
 Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
-We built a new, stable consensus network using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:comparison}).
+We built a new, stable consensus network using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
 Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
 We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once.
-The resulting stablility-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}).
-This network shares most of the proterties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
+The resulting stability-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}).
+This network shares most of the properties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
 
@@
 On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While this deepens our understanding of the disease and provide biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
+While this deepens our understanding of the disease and provides biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
@@
 The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
-In consequence, they show their shortcomings with respect to understudied genes, especially those not in the network.
+In consequence, they show their shortcomings on understudied genes, especially those not in the network.
 Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
@@
 When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
-This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the solutions.
-
-A crucial step for the gene based methods is the computation of the gene score.
+This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best-known genes are artificially pushed into the solutions.
+
+A crucial step for the gene-based methods is the computation of the gene score.
 In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
@@
 
-In order to produce the consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
+To produce the two consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
 To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
@@
 We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
-The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
+The genotyping was performed using the iCOGS array, a custom Illumina array designed to study the genetic susceptibility to hormone-related cancers \cite{sakoda_turning_2013}.
 It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
@@
 
-In addition to a weight, nodes have other properties, provided by the topology of the graph.
-In this article we focus on two of those: degree centrality, and betweenness centrality.
+In addition to weights, nodes have other properties, provided by the topology of the graph.
+We focus on two of those: degree centrality, and betweenness centrality.
 The degree centrality, or degree, is the number of edges that a node has.
@@
 
-In addition, we use two matrices that describe different properties of a graph.
+We also use two matrices that describe two different properties of a graph.
 The described matrices are square, and have as many rows and columns as nodes are in the network.
@@
 The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to run efficiently on networks of their size.
+Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well-characterized, and they were designed to run efficiently on networks of their size.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
-However, at different stages of this work we also used GS and GM for comparison.
+However, at different stages we also used GS and GM for comparison.
 For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
@@
 We examined six of them: five that explore the PPIN, and one which explores SNP networks.
-We selected methods that were open source, had an implementation available, and an accessible documentation.
+We selected methods that were open-source, and had an implementation available and accessible documentation.
 Their main differences are summarized in Table~\ref{tab:method_comparison}.
@@
 The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
-This transformation is achieved by modelling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
+This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
 Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
@@
 To that end, it considers both the local topology of the network and the scores of the nodes.
-The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
+The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors while retaining part of its own (hence, \emph{insulated}).
 This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
@@
 We assigned initial node scores as in Nakka et al.
-\cite{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+\cite{nakka_gene_2016}, assigning a score of 0 for the genes with a low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
 In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
@@
 \item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
-By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
+By imposing this restriction, LEAN can exhaustively test all such subnetworks (one per node).
 For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
@@
 We searched for pathways enriched in the gene subnetworks produced by the above methods.
-We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
+We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
 We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH).
-Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
-As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
-
-\subsection{Evaluation of methods}
-\label{methods:comparison}
+Pathways with a BH adjusted P\=/value \textless~0.05 were deemed significant.
+As the significant pathways are often overlapping and redundant, the results were manually curated afterward.
+
+\subsection{Benchmark of methods}
+\label{methods:benchmark}
 
@@
 
-\subsubsection{Properties ot the solution}
+\subsubsection{Properties of the solution}
 \label{methods:algorithm_comparison}
 
-We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
-The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by Nogueira and Brown \cite{nogueira_measuring_2016}.
+We compared the runtime, the number of selected features (genes or SNPs), and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
+Nogueira and Brown \cite{nogueira_measuring_2016} proposed quantifying the stability of a method using the Pearson correlation between the genes selected on different subsamples.
+This correlation was calculated between vectors with the length of the total number of features, containing a 0 at position $i$ if feature $i$ was not selected, and 1 if it was.
 
@@
 A desirable solution offers good predictive power on unseen \emph{test} samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provides good classification accuracy.
 We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
+The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}) while improving the generalization of the classifier.
 This penalty was set by cross-validation, choosing the value that minimized misclassification error.
@@
 Then we evaluated the sensitivity and the specificity on the \emph{test set}.
-The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
+The active set gave an estimate of a plausible, more sparse solution with comparable predictive power to the original solution.
 To obtain a baseline, we also trained the classifier on all the SNPs.
@@
 
-An alternative way to evaluate the results is comparing our results to an external dataset.
-For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}.
+An alternative way to evaluate the results is by comparing our results to an external dataset.
+For that purpose, we used the 153 genes associated to familial breast cancer on DisGeNET \cite{pinero_disgenet:_2017}.
 Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
@@
 Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
-Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focus on the French high-risk population attending the family cancer clinics.
+Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focuses on the French high-risk population attending the family cancer clinics.
 Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
@@
 We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
+
+\subsection{Computational resources}
+
+We ran all the computations on a Slurm cluster, running Ubuntu 16.04.2 on the nodes.
+The CPU models on the nodes were Intel Xeon CPU E5-2450 v2 at 2.50GHz and Intel Xeon E5-2440 at 2.40GHz.
+The nodes running heinz and HotNet2 had 20GB of memory; the ones running dmGWAS, LEAN, SConES, and SigMod, 60GB.
+For the benchmark (Section \ref{methods:benchmark}), we ran each of the methods on the same Ubuntu 16.04.2 node, with a CPU Intel Xeon E5-2450 v2 at 2.50GHz, and 60GB of memory.
 
@@
 This panel is identical to Fig~\ref{fig:consensus}.
-\textbf{(B)} Same network, but every gene name is indicated.
+\textbf{(B)} The same network, but every gene name is indicated.
 % \includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
@@
 This panel is identical to Fig~\ref{fig:stable-consensus}.
-\textbf{(B)} Same network, but every gene name is indicated.
+\textbf{(B)} The same network, but every gene name is indicated.
 % \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf}
@@
 \label{sfig:lc_ht_comparison}
-{\bf Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
+{\bf Comparison of the benchmark on high-throughput (HT) interactions to the benchmark on both high-throughput and literature curated interactions (HT+LC).
 } Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
@@
 \label{sfig:scones_gsm}
-{\bf Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)} Manhattan plots of SNPs; in black,the method’s solution.
+{\bf Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)} Manhattan plots of SNPs; in black, the method’s solution.
 The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
@@
 \section*{Acknowledgments}
-We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie), the biological resource centre and all the GENESIS collaborating cancer clinics clinics.
+We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie), the biological resource center and all the GENESIS collaborating cancer clinics.
 

### 2020-08-24 manuscript/genome_research/manuscript_genome_research.tex
added genome research submission

--- 2020-05-01 manuscript/manuscript.tex
@@
 \usepackage{threeparttable}
-\usepackage[numbers]{natbib}
+\usepackage[round]{natbib}
 \usepackage{lmodern} % Chloe: better font rendering
@@
 Nadine Andrieu$^{1,2,3}$, Chloé-Agathe Azencott$^{3,1,2}$}
-\date{$^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
+\date{\small{
+  \raggedright{
+  $^{1}$Institut Curie, PSL Research University, F-75005 Paris, France;\\
   $^{2}$INSERM, U900, F-75005 Paris, France;\\
@@
   $^{6}$Université Paris Descartes.\\
-}
+}}}
+
 \hypersetup{
@@
 
-In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \cite{bush_chapter_2012}.
+In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \citep{bush_chapter_2012}.
 To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering the disease of interest, and the controls from the general population.
@@
 
-GWAS have successfully identified thousands of variants underlying many common diseases \cite{buniello_nhgri-ebi_2019}.
+GWAS have successfully identified thousands of variants underlying many common diseases \citep{buniello_nhgri-ebi_2019}.
 However, this experimental setting also presents intrinsic challenges.
 Some of them stem from the high dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped.
-This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}.
-This is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
+This limits the statistical power of the experiment, as only variants with larger effects can be detected \citep{visscher_10_2017}.
+This is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \citep{visscher_10_2017}.
 Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the previously mentioned Bonferroni correction.
-However, Bonferroni correction is overly conservative when the statistical tests are correlated, as is the case in GWAS \cite{wang_statistical_2018}.
+However, Bonferroni correction is overly conservative when the statistical tests are correlated, as is the case in GWAS \citep{wang_statistical_2018}.
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
-On top of that, recent large-sampled studies suggest that numerous loci spread all along the genome contribute to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
+On top of that, recent large-sampled studies suggest that numerous loci spread all along the genome contribute to a degree to any complex trait, in accordance with the infinitesimal model \citep{barton_infinitesimal_2017}.
+The recently proposed omnigenic model \citep{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
 Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
 
-For this reason, many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
+For this reason, many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \citep{furlong_human_2013}.
 As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them.
 These relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction.
-Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}.
-In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}.
+Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \citep{barabasi_network_2011}.
+In fact, an examination of biological networks shows that disease genes have differential properties \citep{barabasi_network_2011,pinero_uncovering_2016}.
 This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes.
-Therefore, studying the neighborhood of disease-associated genes is effective at identifying new ones that are involved in the same biological functions \cite{huang_systematic_2018}.
-
-Network-based biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
+Therefore, studying the neighborhood of disease-associated genes is effective at identifying new ones that are involved in the same biological functions \citep{huang_systematic_2018}.
+
+Network-based biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \citep{azencott_network-guided_2016}.
 In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
@@
 Some methods strongly constrain the problem to certain kinds of subnetworks.
-Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e.
+Such is the extreme case of LEAN \citep{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e.
 instances were both a gene and its direct interactors are associated with the disease.
-Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnecting genes with high association scores.
+Other algorithms, like dmGWAS \citep{jia_dmgwas:_2011} and heinz \citep{dittrich_identifying_2008}, focus on larger subnetworks interconnecting genes with high association scores.
 However, they differ in their tolerance to the inclusion of low-scoring nodes, and the topology of the solution.
-Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
+Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring, but also densely interconnected; such is the case of HotNet2 \citep{leiserson_pan-cancer_2015}, SConES \citep{azencott_efficient_2013}, and SigMod \citep{liu_sigmod:_2017}.
 
@@
 While all of them capture susceptibility mechanisms resembling that postulated by the omnigenic model, they do so in different ways, and provide a representative view of the field.
-We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
+We worked on the GENESIS dataset \citep{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
 After a classical GWAS approach, we use these network methods to recover additional breast cancer biomarkers.
@@
 
-\section{Methods}
-\subsection{GENESIS}
-
-The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
-Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
-Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
-We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
-The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \cite{sakoda_turning_2013}.
-It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
-
-\subsection{Preprocessing and quality control}
-
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
-A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
-In addition, we removed the samples with more than 10\% missing genotypes.
-After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
-Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
-The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
-
-We looked for population structure that could produce spurious associations.
-A principal component analysis revealed no visual differential population structure between cases and controls (Supplementary Figure~\ref{sfig:pcs}).
-Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
-
-\subsection{High-score subnetwork search algorithms}
-\subsubsection{SNP and gene association}
-\label{methods:node_score}
-To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
-Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
-Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P\=/values.
-We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
-We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
-Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
-Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
-
-We used such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
-For the former, we considered any gene that can be mapped to any of the selected SNPs as selected as well.
-For the latter, we considered all the SNPs that can be mapped to that gene as selected by the method.
-
-\subsubsection{Mathematical notations}
-\label{methods:notation}
-In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
-$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
-Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
-When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
-A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
-
-In addition to a weight, nodes have other properties, provided by the topology of the graph.
-In this article we focus on two of those: degree centrality, and betweenness centrality.
-The degree centrality, or degree, is the number of edges that a node has.
-The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.
-
-In addition, we use two matrices that describe different properties of a graph.
-The described matrices are square, and have as many rows and columns as nodes are in the network.
-The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
-The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; the diagonal is zero.
-The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
-
-\subsubsection{Networks}
-\label{methods:networks}
-
-\paragraph{Gene network}
-The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to run efficiently on networks of their size.
-We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
-Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
-12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
-The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
-
-Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
-In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
-Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
-
-\paragraph{SNP networks}
-SConES \cite{azencott_efficient_2013} is the only network method designed to handle SNP networks.
-As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
-\citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
-We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
-However, at different stages of this work we also used GS and GM for comparison.
-For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
-In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
-The properties of these three subnetworks are available in Supplementary Table \ref{tab:snp_solutions}.
-
-\subsubsection{Network methods}
-\label{methods:methods}
-
-\begin{table}[htbp]
-  \caption{Summary of the differences between the network methods.}
-  \label{tab:method_comparison}
-  \centering
-  \begin{threeparttable}
-    \begin{tabular}{l|llllllll}
-      Method & Field & Nodes & Exhaustive & Solution & Comp. & Input & Scoring & Ref.\\
-      \hline
-      dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
-      heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
-      HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
-      LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
-      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
-      SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \cite{liu_sigmod:_2017}\\
-    \end{tabular}
-    \begin{tablenotes}
-      \footnotesize{
-        \item \textbf{\emph{Field}}: field in which the algorithm was developed.
-\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
-\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
-\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than containing high scoring, connected nodes.
-\textbf{\emph{Comp.}}: number of connected components in the solution.
-\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
-\textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
-In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
-        cumulative distribution function of the standard Normal distribution. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
-      }
-    \end{tablenotes}
-  \end{threeparttable}
-\end{table}
-
-Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
-Unsurprisingly, multiple solutions have been proposed.
-We examined six of them: five that explore the PPIN, and one which explores SNP networks.
-We selected methods that were open source, had an implementation available, and an accessible documentation.
-Their main differences are summarized in Table~\ref{tab:method_comparison}.
-
-\begin{description}
-\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
-To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
-
-\begin{enumerate}
-\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
-
-\begin{equation*} 
-Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
-\end{equation*}
-
-where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
-\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
-\item Repeat 2-4 until no further enlargement is possible.
-\item Add $S_i$ to the list of subnetworks to return. Its Z-score is normalized as
-
-\begin{equation*}
-Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
-\end{equation*}
-
-where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
-\end{enumerate}
-
-DmGWAS carries out this process on every gene in the network.
-We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
-We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
-\end{description}
-
-\begin{description}
-\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
-This transformation is achieved by modelling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
-Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
-To solve it efficiently it is re-cast as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
-
-\begin{equation*}
-p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
-\end{equation*}
-
-were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
-All three are positive quantities.
-Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
-We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
-
-\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}.
-To that end, it considers both the local topology of the network and the scores of the nodes.
-The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
-This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
-$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
-\begin{equation*} 
-E = F \operatorname{diag}(w(V)), 
-\end{equation*}
-where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
-To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
-Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
-We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
-In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
-HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
-Both parameters are set automatically by the algorithm, which is robust to their values \cite{leiserson_pan-cancer_2015}.
-HotNet2 is implemented in Python \cite{hotnet2}.
-
-\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
-By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
-For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
-Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
-The minimum of these \emph{k} P\=/values is the score of the subnetwork.
-This score is transformed into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
-Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
-We used the implementation of LEAN from the LEANR R package \cite{leanr}.
-\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem
-
-\begin{equation}
-\label{eq:scones}
-\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
-\end{equation}
-
-where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
-The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
-Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
-As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
-In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
-We used the version on SConES implemented in the R package martini \cite{martini}.
-\end{description}
-
-\begin{description}
-\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
-
-\begin{equation*}
-\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
-\end{equation*}
-
-As SConES, this optimization problem can also be solved by a graph min-cut approach.
-
-SigMod presents three important differences with SConES.
-First it is designed for gene-gene networks.
-Second, it favors subnetworks containing many edges.
-SConES, instead, penalizes connections between the selected and unselected nodes.
-Third, it explores the grid of hyperparameters differently, and processes their respective solutions.
-Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
-Such a large change implies that the network is densely interconnected.
-This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other.
-A score is assigned to each candidate solution by summing their node scores and normalizing by size.
-The candidate solution with the highest standardized score is the chosen solution.
-SigMod is implemented in an R package \cite{sigmod}.
-\end{description}
-
-\begin{description}
-\item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
-\end{description}
-
-\subsection{Pathway enrichment analysis}
-\label{methods:pathway_enrichment}
-
-We searched for pathways enriched in the gene subnetworks produced by the above methods.
-We conducted an hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
-The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
-We adjusted the P-values for multiple testing as in \citet{Benjamini1995} (BH).Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
-As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
-
-\subsection{Evaluation of methods}
-\label{methods:comparison}
-
-We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
-We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
-When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
-We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
-
-\subsubsection{Properties ot the solution}
-\label{methods:algorithm_comparison}
-
-We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
-The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by \citet{nogueira_measuring_2016}.
-
-\subsubsection{Classification accuracy of selected SNPs}
-\label{methods:classifier}
-
-A desirable solution offers good predictive power on unseen \emph{test} samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
-We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
-This penalty was set by cross-validation, choosing the value that minimized misclassification error.
-We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
-When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
-Then we evaluated the sensitivity and the specificity on the \emph{test set}.
-The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
-To obtain a baseline, we also trained the classifier on all the SNPs.
-We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
-Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
-However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
-
-\subsubsection{Comparison to state-of-the-art}
-\label{methods:bcac}
-
-An alternative way to evaluate the results is comparing our results to an external dataset.
-For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \cite{pinero_disgenet:_2017}.
-Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
-
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
-Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focus on the French high-risk population attending the family cancer clinics.
-Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
-For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
-We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
-
-\subsection{Code availability}
-
-We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
-For each of those processes, we created a pipeline with a clear interface that should work on any GWAS dataset.
-They are compiled in \url{https://github.com/hclimente/gwas-tools}.
-Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
-We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
-
 \section{Results}
@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.7\linewidth]{../figures/sfigure_2.png}
+  \caption{Association in GENESIS.
+The red line represents the Bonferroni threshold.
+\textbf{(A)} SNP association, measured from the outcome of a 1 d.f.
+$\chi^2$ allelic test.
+Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
+The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
+\textbf{(B)} Gene association, measured by P\=/value of VEGAS2 \citep{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values.
+The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
+\textbf{(C)} SNP association as in panel (A).
+The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).}
+  \label{sfig:snp_gene_manhattan}
+\end{figure}
+
 We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
-Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (Supplementary Figure~\ref{sfig:snp_gene_manhattan}A).
+Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (Figure~\ref{sfig:snp_gene_manhattan}A).
 The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
-Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
-In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B).
+Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \citep{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \citep{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \citep{rinella_genetic_2013}.
+In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (Figure~\ref{sfig:snp_gene_manhattan}B).
 
 Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
-Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
+Both of them have been associated to breast cancer susceptibility in the past \citep{brisbin_meta-analysis_2011,search_newly_2009}.
 We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
-It selected 100 SNPs, both from all aforementioned regions and new ones (Supplementary Figure~\ref{sfig:snp_gene_manhattan}C).
+It selected 100 SNPs, both from all aforementioned regions and new ones (Figure~\ref{sfig:snp_gene_manhattan}C).
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
@@
   \centering
-  \includegraphics[height=.75\textheight]{./figures/figure_1.png}
+  \includegraphics[height=.75\textheight]{../figures/figure_1.png}
   \caption{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
@@
 Interestingly, dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \citep{nielsen_hereditary_2016}.
 As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
@@
 Second, the genes in four solution subnetworks displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value < 1.4 \texttimes{} 10\textsuperscript{-21}).
-This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+This agrees with the notion that disease genes are more central than other non-essential genes \citep{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
+Interestingly, SConES selected SNPs that are also more central than the average SNP (Table~\ref{tab:snp_solutions}), suggesting that causal SNPs are also more central than non associated SNPs.
+
+\begin{table}[htbp]
+  \begin{threeparttable}
+  \caption{Summary statistics on the results of SConES on the three SNP-SNP interaction networks. The first row within each block contains the summary statistics on the whole network.}
+  \label{tab:snp_solutions}
+  \centering
+  \begin{tabular}{lrrllr}
+  Network & SNPs & Edges & Subnetworks & \mean{Betweenness} & \median{P}\textsubscript{SNP}\\
+  \hline
+  GS & 197\,083 & 197\,060 & - & 2.03 \texttimes{} 10\textsuperscript{7} & 0.49\\
+  SConES GS & 1\,590 & 1\,585 & 5 & 2.52 \texttimes{} 10\textsuperscript{7} & 0.023\\
+  \hline
+  GM & 197\,083 & 6\,442\,446 & - & 3.99 \texttimes{} 10\textsuperscript{6} & 0.49\\
+  SConES GM & 1\,692 & 177\,611 & 5 & 4.40 \texttimes{} 10\textsuperscript{6} & 0.055\\
+  \hline
+  GI & 197\,083 & 28\,733\,720 & - & 1.46 \texttimes{} 10\textsuperscript{6} & 0.49\\
+  SConES GI & 408 & 539 & 5 & 9.33 \texttimes{} 10\textsuperscript{6} & 0.076\\
+  \end{tabular}
+  \begin{tablenotes}
+    \footnotesize{
+      \item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.\median{P}\textsubscript{SNP}: median P\=/value of the selected SNPs.
+    }
+  \end{tablenotes}
+  \end{threeparttable}
+  \end{table}
 
@@
   \centering
-  \includegraphics[width=.5\linewidth]{./figures/figure_3.pdf}
+  \includegraphics[width=.5\linewidth]{../figures/figure_3.pdf}
   \caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
@@
 The latter ones are also colored in pink.
-All gene names are indicated in Supplementary Figure \ref{sfig:consensus-names}.}
+All gene names are indicated in Figure \ref{sfig:consensus-names}.}
   \label{fig:consensus}
@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.675\linewidth]{../figures/sfigure_4.pdf}
+  \caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
+\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+This panel is identical to Figure~\ref{fig:consensus}.
+\textbf{(B)} Same network, but every gene name is indicated.}
+  \label{sfig:consensus-names}
+  \end{figure}
+
 Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
-Indeed, only 20 genes are common to more than two solutions (Supplementary Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Supplementary Figure~\ref{sfig:consensus_stats}B).
+Indeed, only 20 genes are common to more than two solutions (Figure~\ref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Figure~\ref{sfig:consensus_stats}B).
 To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.65\textwidth]{../figures/sfigure_5.pdf}
+  \caption{Genes on the consensus network.
+Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
+\textbf{(A)} Number of methods selecting every gene in the subnetwork.
+\textbf{(B)} VEGAS2 P\=/values of association of the genes, with regards to the number of methods that selected them.
+\textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network.
+To improve visualization, we removed outliers.
+\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus network.
+The blue line represents a fitted generalized linear model.}
+  \label{sfig:consensus_stats}
+  \end{figure}
+
 A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions.
 We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
-The former is supported by genes like \emph{MRPS30} (VEGAS2 P\=/value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
-Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
+The former is supported by genes like \emph{MRPS30} (VEGAS2 P\=/value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \citep{quigley_5p12_2014}.
+Interestingly, increased mitochondrial translation has been found in cancer cells \citep{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
 With regards to attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
@@
 Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
-The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, Supplementary Figure~\ref{sfig:consensus_stats}D).
+The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, Figure~\ref{sfig:consensus_stats}D).
 This suggests that they were selected because they were on the shortest path between two highly associated genes.
-In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
+In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \citep{boyle_expanded_2017}.
 For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signalling pathways.
-\emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
+\emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \citep{liu_jab1_cops5_2018}.
 Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
@@
 
-We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
+We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \citep{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
 Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network metods operate.
@@
 \centering
-\includegraphics[width=.72\linewidth]{./figures/figure_4.pdf}
+\includegraphics[width=.72\linewidth]{../figures/figure_4.pdf}
 \caption{Comparison of network-based GWAS methods on GENESIS.
@@
 \centering
-\includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
+\includegraphics[width=.8\linewidth]{../figures/figure_2.pdf}
 \caption{Drawbacks encountered when using network guided methods.
@@
 Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (Supplementary Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
+Due to the limited signal at the gene level in this dataset (Figure~\ref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
 Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
@@
 dmGWAS' subnetwork is the least associated subnetwork on average.
-This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
+This is due to the greedy framework it uses, which has a bias for larger solutions \citep{nikolayeva_network_2018}.
 It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated genes if it was linked to another, strongly associated one.
@@
 With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration.
-Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it is negatively correlated with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.
+Overexpression of \emph{FN1} has been observed in breast cancer \citep{Ioachim2002}, and it is negatively correlated with poor prognosis in other cancer types \citep{Yi2016,Sponziello2016}.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.9\linewidth]{../figures/sfigure_3.pdf}
+  \caption{Pearson correlation between the different solution subnetworks.
+\textbf{(A)} Correlation between selected SNPs.
+\textbf{(B)} Correlation between selected genes.
+In general, the solutions display a very low overlap.}
+  \label{sfig:pearson_methods}
+  \end{figure}
 
@@
 In fact, due to linkage disequilibrium, SConES favors such genes, as selecting SNPs in an LD-block which overlaps with a gene favors selecting the rest of the gene.
-This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Supplementary Figure~\ref{sfig:pearson_methods}).
+This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Figure~\ref{sfig:pearson_methods}).
 On the other hand, SConES penalizes selecting SNPs and not their neighbors.
@@
 For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Figure~\ref{fig:issues}C).
-In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
-On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
+In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \citep{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \citep{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
+On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \citep{breyer_expressed_2014}).
 We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a gene and a majority of its neighbors.
@@
   \centering
-  \includegraphics[width=.5\linewidth]{./figures/figure_5.pdf}
+  \includegraphics[width=.5\linewidth]{../figures/figure_5.pdf}
   \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
@@
 The latter ones are also colored in pink.
-All gene names are indicated in Supplementary Figure \ref{sfig:stable-consensus-names}.}
+All gene names are indicated in Figure \ref{sfig:stable-consensus-names}.}
   \label{fig:stable-consensus}
+\end{figure}
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.675\linewidth]{../figures/sfigure_6.pdf}
+  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
+\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+The latter ones are also colored in pink.
+This panel is identical to Figure~\ref{fig:stable-consensus}.
+\textbf{(B)} Same network, but every gene name is indicated.}
+  \label{sfig:stable-consensus-names}
 \end{figure}
@@
 This gene is a component of Cullin-RING ubiquitin ligases.
-Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
+Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \citep{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \citep{Loignon2009}.
 
@@
 
-In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
+In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \citep{boyle_expanded_2017}.
 On the one hand, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
@@
 Most of the network methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
-In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
+In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \citep{pinero_uncovering_2016}.
 However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by these network methods.
@@
 Second, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
-This might partly be caused by the instability of the P\=/values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
+This might partly be caused by the instability of the P\=/values themselves in low statistical power settings \citep{halseyFickleValueGenerates2015}.
 Hence, heinz's conservative transformation of P\=/values, which favors only the most extreme ones, leads to improved stability.
@@
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.9\linewidth]{../figures/sfigure_7.pdf}
+  \caption{Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
+  \label{sfig:biotypes_excluded}
+  \end{figure}
+
 The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
@@
 Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
-The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Supplementary Figure~\ref{sfig:biotypes_excluded}).
+The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (Figure~\ref{sfig:biotypes_excluded}).
 Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
-In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
+In addition, even protein-coding genes linked to breast cancer susceptibility \citep{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
 However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
 As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
-As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
+As disease genes tend to be more central \citep{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
 It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
-Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
+Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \citep{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
 This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
@@
 In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
-In fact, the presented case is paradigmatic, since the genes are in the most gene-dense region of the genome \cite{Xie2003}.
+In fact, the presented case is paradigmatic, since the genes are in the most gene-dense region of the genome \citep{Xie2003}.
 Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the network, but might be more resilient to them when the overlapping genes are unrelated.
@@
 In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
+This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \citep{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
 On the other hand, \citet{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
-When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Supplementary Figure~\ref{sfig:lc_ht_comparison}).
+When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (Figure~\ref{sfig:lc_ht_comparison}).
 This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best known genes are artificially pushed into the solutions.
 
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.85\linewidth]{../figures/sfigure_8.pdf}
+  \caption{Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
+Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
+\textbf{(A)} Ratios of the selected features between both benchmarks and of the active set.
+\textbf{(B)} Shifts in sensitivity and specificity.
+\textbf{(C)} Shift in Pearson correlation between benchmarks.
+\textbf{(D)} Ratio between the runtimes of the benchmarks.
+For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
+In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}
+  \label{sfig:lc_ht_comparison}
+  \end{figure}
+
 A crucial step for the gene based methods is the computation of the gene score.
-In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
+In this work we used VEGAS2 \citep{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \citep{nakka_gene_2016}), and other algorithms \citep{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
 Another important decision is how to handle LD in a GWAS.
@@
 In order to produce the consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
-To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
+To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \citep{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
 They are available on GitHub: \url{https://github.com/hclimente/gwas-tools}.
@@
 
-\section*{Author contributions}
-
-H.C-G.
-conducted all the final analyses and the wrote the manuscript.
-C.L.
-conducted analyses using LEAN, Sigmod, and dmGWAS, and provided technical and scientific help and feedback.
-F.L.
-and N.A.
-provided feedback on the analyses and the manuscript.
-C-A.A.
-supervised the project.
-N.A and D.S-L lead the GENESIS study.
-
-\section*{Funding and acknowledgments}
+\section{Methods}
+\subsection{GENESIS}
+
+The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \citep{sinilnikova_genesis:_2016}.
+Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
+Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
+We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
+The genotyping was performed using the iCOGS array, a custom Illumina array designed to study genetic susceptibility of hormone-related cancers \citep{sakoda_turning_2013}.
+It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
+
+\subsection{Preprocessing and quality control}
+
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
+A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
+In addition, we removed the samples with more than 10\% missing genotypes.
+After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
+The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
+
+We looked for population structure that could produce spurious associations.
+A principal component analysis revealed no visual differential population structure between cases and controls (Figure~\ref{sfig:pcs}).
+Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
+
+\begin{figure}[htbp]
+  \centering
+  \includegraphics[width=.9\linewidth]{../figures/sfigure_1.pdf}
+  \caption{GENESIS shows no differential population structure between cases and controls.
+\textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
+Cases are colored in green, controls in orange.}
+  \label{sfig:pcs}
+  \end{figure}
+
+\subsection{High-score subnetwork search algorithms}
+\subsubsection{SNP and gene association}
+\label{methods:node_score}
+To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \citep{chang_second-generation_2015}.
+Then, we used VEGAS2 \citep{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
+Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P\=/values.
+We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
+We used the 62\,193 genes described in GENCODE 31 \citep{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
+Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
+Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
+
+We used such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
+For the former, we considered any gene that can be mapped to any of the selected SNPs as selected as well.
+For the latter, we considered all the SNPs that can be mapped to that gene as selected by the method.
+
+\subsubsection{Mathematical notations}
+\label{methods:notation}
+In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
+$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
+Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
+When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
+A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
+
+In addition to a weight, nodes have other properties, provided by the topology of the graph.
+In this article we focus on two of those: degree centrality, and betweenness centrality.
+The degree centrality, or degree, is the number of edges that a node has.
+The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.
+
+In addition, we use two matrices that describe different properties of a graph.
+The described matrices are square, and have as many rows and columns as nodes are in the network.
+The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
+The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; the diagonal is zero.
+The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
+
+\subsubsection{Networks}
+\label{methods:networks}
+
+\paragraph{Gene network}
+The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
+Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well characterized, and they were designed to run efficiently on networks of their size.
+We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \citep{das_hint:_2012}.
+Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
+Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
+12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
+The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
+
+Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
+In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
+Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
+
+\paragraph{SNP networks}
+SConES \citep{azencott_efficient_2013} is the only network method designed to handle SNP networks.
+As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
+\citet{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
+We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
+However, at different stages of this work we also used GS and GM for comparison.
+For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
+In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
+The properties of these three subnetworks are available in Table \ref{tab:snp_solutions}.
+
+\subsubsection{Network methods}
+\label{methods:methods}
+
+\begin{table}[htbp]
+  \caption{Summary of the differences between the network methods.}
+  \label{tab:method_comparison}
+  \centering
+  \begin{threeparttable}
+    \begin{tabular}{l|llllllll}
+      Method & Field & Nodes & Exhaustive & Solution & Comp. & Input & Scoring & Ref.\\
+      \hline
+      dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \citep{jia_dmgwas:_2011}\\
+      heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \citep{dittrich_identifying_2008}\\
+      HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \citep{leiserson_pan-cancer_2015}\\
+      LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \citep{gwinner_network-based_2016}\\
+      SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \citep{azencott_efficient_2013}\\
+      SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \citep{liu_sigmod:_2017}\\
+    \end{tabular}
+    \begin{tablenotes}
+      \footnotesize{
+        \item \textbf{\emph{Field}}: field in which the algorithm was developed.
+\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
+\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
+\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than containing high scoring, connected nodes.
+\textbf{\emph{Comp.}}: number of connected components in the solution.
+\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
+\textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
+In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the
+        cumulative distribution function of the standard Normal distribution. \textbf{\emph{Ref.}}: original publication featuring the algorithm.
+      }
+    \end{tablenotes}
+  \end{threeparttable}
+\end{table}
+
+Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
+But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+Unsurprisingly, multiple solutions have been proposed.
+We examined six of them: five that explore the PPIN, and one which explores SNP networks.
+We selected methods that were open source, had an implementation available, and an accessible documentation.
+Their main differences are summarized in Table~\ref{tab:method_comparison}.
+
+\begin{description}
+\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \citep{jia_dmgwas:_2011}.
+To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
+
+\begin{enumerate}
+\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
+\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
+
+\begin{equation*} 
+Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
+\end{equation*}
+
+where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
+\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
+\item Repeat 2-4 until no further enlargement is possible.
+\item Add $S_i$ to the list of subnetworks to return. Its Z-score is normalized as
+
+\begin{equation*}
+Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
+\end{equation*}
+
+where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
+\end{enumerate}
+
+DmGWAS carries out this process on every gene in the network.
+We used the implementation of dmGWAS in the dmGWAS 3.0 R package \citep{dmgwas}.
+We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
+We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
+\end{description}
+
+\begin{description}
+\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \citep{dittrich_identifying_2008}.
+The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
+This transformation is achieved by modelling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
+Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
+To solve it efficiently it is re-cast as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
+
+\begin{equation*}
+p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
+\end{equation*}
+
+were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
+All three are positive quantities.
+Heinz implements the algorithm from \citet{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
+We used BioNet's implementation of heinz \citep{beisser_bionet:_2010,heinz}.
+
+\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \citep{leiserson_pan-cancer_2015}.
+To that end, it considers both the local topology of the network and the scores of the nodes.
+The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors, while retaining part of its own (hence, \emph{insulated}).
+This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
+$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
+\begin{equation*} 
+E = F \operatorname{diag}(w(V)), 
+\end{equation*}
+where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
+For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
+To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
+Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
+We assigned initial node scores as in \citet{nakka_gene_2016}, assigning a score of 0 for the genes with low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \citep{scheid_twilight;_2005}.
+HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
+Both parameters are set automatically by the algorithm, which is robust to their values \citep{leiserson_pan-cancer_2015}.
+HotNet2 is implemented in Python \citep{hotnet2}.
+
+\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \citep{gwinner_network-based_2016}.
+By imposing this restriction, LEAN is able to exhaustively test all such subnetworks (one per node).
+For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
+Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
+The minimum of these \emph{k} P\=/values is the score of the subnetwork.
+This score is transformed into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
+Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
+We used the implementation of LEAN from the LEANR R package \citep{leanr}.
+\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \citep{azencott_efficient_2013}. Specifically, it solves the problem
+
+\begin{equation}
+\label{eq:scones}
+\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
+\end{equation}
+
+where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
+The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
+Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
+As in \citet{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
+In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
+We used the version on SConES implemented in the R package martini \citep{martini}.
+\end{description}
+
+\begin{description}
+\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \citep{liu_sigmod:_2017}.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
+
+\begin{equation*}
+\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
+\end{equation*}
+
+As SConES, this optimization problem can also be solved by a graph min-cut approach.
+
+SigMod presents three important differences with SConES.
+First it is designed for gene-gene networks.
+Second, it favors subnetworks containing many edges.
+SConES, instead, penalizes connections between the selected and unselected nodes.
+Third, it explores the grid of hyperparameters differently, and processes their respective solutions.
+Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
+Such a large change implies that the network is densely interconnected.
+This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other.
+A score is assigned to each candidate solution by summing their node scores and normalizing by size.
+The candidate solution with the highest standardized score is the chosen solution.
+SigMod is implemented in an R package \citep{sigmod}.
+\end{description}
+
+\begin{description}
+\item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
+\end{description}
+
+\subsection{Pathway enrichment analysis}
+\label{methods:pathway_enrichment}
+
+We searched for pathways enriched in the gene subnetworks produced by the above methods.
+We conducted an hypergeometric test on pathways from Reactome \citep{Jassal2019} using the R package ReactomePA \citep{Yu2016}.
+The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
+We adjusted the P-values for multiple testing as in \citet{Benjamini1995} (BH).Pathways with an BH adjusted P\=/value \textless~0.05 were deemed significant.
+As the significant pathways are often overlapping and redundant, the results were manually curated afterwards.
+
+\subsection{Evaluation of methods}
+\label{methods:comparison}
+
+We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
+We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
+When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
+We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
+
+\subsubsection{Properties ot the solution}
+\label{methods:algorithm_comparison}
+
+We compared the runtime, the number of selected genes/SNPs, and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
+The stability was quantified using the Pearson correlation to measure the overlap between different runs as suggested by \citet{nogueira_measuring_2016}.
+
+\subsubsection{Classification accuracy of selected SNPs}
+\label{methods:classifier}
+
+A desirable solution offers good predictive power on unseen \emph{test} samples.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provide good classification accuracy.
+We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
+The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}), while improving the generalization of the classifier.
+This penalty was set by cross-validation, choosing the value that minimized misclassification error.
+We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
+When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
+Then we evaluated the sensitivity and the specificity on the \emph{test set}.
+The active set gave an estimate of a plausible, more sparse solution with a comparable predictive power to the original solution.
+To obtain a baseline, we also trained the classifier on all the SNPs.
+We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
+Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \citep{Metcalfe2008}.
+However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
+
+\subsubsection{Comparison to state-of-the-art}
+\label{methods:bcac}
+
+An alternative way to evaluate the results is comparing our results to an external dataset.
+For that purpose, we recovered a list of 153 genes associated to familial breast cancer from DisGeNET \citep{pinero_disgenet:_2017}.
+Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
+
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \citep{Michailidou2017}.
+Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focus on the French high-risk population attending the family cancer clinics.
+Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
+For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
+We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \citep{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
+
+\section{Data access}
+
+We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
+For each of those processes, we created a pipeline with a clear interface that should work on any GWAS dataset.
+They are compiled in \url{https://github.com/hclimente/gwas-tools}.
+Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
+
+\section{Acknowledgments}
 
@@
 
+\section{Disclosure declaration}
+
+We declare no conflicts of interest.
+
+\section{Author contributions}
+
+H.C-G.
+conducted all the final analyses and the wrote the manuscript.
+C.L.
+conducted analyses using LEAN, Sigmod, and dmGWAS, and provided technical and scientific help and feedback.
+F.L.
+and N.A.
+provided feedback on the analyses and the manuscript.
+C-A.A.
+supervised the project.
+N.A and D.S-L lead the GENESIS study.
+
 \bibliographystyle{abbrvnat}
@@
 
-\clearpage
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\setcounter{table}{0}
-
-\section{Supplementary materials}
-
-\begin{table}[htbp]
-\begin{threeparttable}
-\caption{Summary statistics on the results of SConES on the three SNP-SNP interaction networks. The first row within each block contains the summary statistics on the whole network.}
-\label{tab:snp_solutions}
-\centering
-\begin{tabular}{lrrllr}
-Network & SNPs & Edges & Subnetworks & \mean{Betweenness} & \median{P}\textsubscript{SNP}\\
-\hline
-GS & 197\,083 & 197\,060 & - & 2.03 \texttimes{} 10\textsuperscript{7} & 0.49\\
-SConES GS & 1\,590 & 1\,585 & 5 & 2.52 \texttimes{} 10\textsuperscript{7} & 0.023\\
-\hline
-GM & 197\,083 & 6\,442\,446 & - & 3.99 \texttimes{} 10\textsuperscript{6} & 0.49\\
-SConES GM & 1\,692 & 177\,611 & 5 & 4.40 \texttimes{} 10\textsuperscript{6} & 0.055\\
-\hline
-GI & 197\,083 & 28\,733\,720 & - & 1.46 \texttimes{} 10\textsuperscript{6} & 0.49\\
-SConES GI & 408 & 539 & 5 & 9.33 \texttimes{} 10\textsuperscript{6} & 0.076\\
-\end{tabular}
-\begin{tablenotes}
-  \footnotesize{
-    \item \mean{Betweenness}: mean betweenness of the selected SNPs in the corresponding full network.\median{P}\textsubscript{SNP}: median P\=/value of the selected SNPs.
-  }
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{table}[htbp]
-\begin{threeparttable}
-  \caption{Summary statistics on the results of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}
-\label{tab:scones_gene_solutions}
-\centering
-\begin{tabular}{lrrrlr}
-Network & Genes & Edges & \mean{Betweenness} & \median{P}\textsubscript{gene} & $\rho_{consensus}$\\
-\hline
-SConES GS & 5 & 0 & 9\,805 & 2.7 \texttimes{} 10\textsuperscript{-5} & 0.19\\
-SConES GM & 28 & 2 & 4\,267 & 0.067 & 0.12\\
-\end{tabular}
-\begin{tablenotes}
-  \footnotesize{
-    \item \mean{Betweenness}: mean betweenness of the selected genes in the full network.
-\median{P}\textsubscript{gene}: median P\=/value of the selected genes; $\rho_{consensus}$: Pearson correlation with the consensus network.
-  }
-\end{tablenotes}
-\end{threeparttable}
-\end{table}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
-\caption{GENESIS shows no differential population structure between cases and controls.
-\textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
-Cases are colored in green, controls in orange.}
-\label{sfig:pcs}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
-  \caption{Association in GENESIS.
-The red line represents the Bonferroni threshold.
-\textbf{(A)} SNP association, measured from the outcome of a 1 d.f.
-$\chi^2$ allelic test.
-Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
-The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
-\textbf{(B)} Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values.
-The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
-\textbf{(C)} SNP association as in panel (A).
-The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).}
-  \label{sfig:snp_gene_manhattan}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_3.pdf}
-\caption{Pearson correlation between the different solution subnetworks.
-\textbf{(A)} Correlation between selected SNPs.
-\textbf{(B)} Correlation between selected genes.
-In general, the solutions display a very low overlap.}
-\label{sfig:pearson_methods}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
-\caption{Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
-\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-This panel is identical to Figure~\ref{fig:consensus}.
-\textbf{(B)} Same network, but every gene name is indicated.}
-\label{sfig:consensus-names}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
-\caption{Genes on the consensus network.
-Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
-\textbf{(A)} Number of methods selecting every gene in the subnetwork.
-\textbf{(B)} VEGAS2 P\=/values of association of the genes, with regards to the number of methods that selected them.
-\textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network.
-To improve visualization, we removed outliers.
-\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus network.
-The blue line represents a fitted generalized linear model.}
-\label{sfig:consensus_stats}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf}
-  \caption{Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
-\textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-This panel is identical to Figure~\ref{fig:stable-consensus}.
-\textbf{(B)} Same network, but every gene name is indicated.}
-  \label{sfig:stable-consensus-names}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
-\caption{Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
-\label{sfig:biotypes_excluded}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.85\linewidth]{./figures/sfigure_8.pdf}
-\caption{Comparison of benchmark on high-throughput (HT) interactions to benchmark on both high-throughput and literature curated interactions (HT+LC).
-Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
-\textbf{(A)} Ratios of the selected features between both benchmarks and of the active set.
-\textbf{(B)} Shifts in sensitivity and specificity.
-\textbf{(C)} Shift in Pearson correlation between benchmarks.
-\textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
-In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.}
-\label{sfig:lc_ht_comparison}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=\linewidth]{./figures/sfigure_9.png}
-  \caption{Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.
-\textbf{(A)} Manhattan plots of SNPs; in black,the method’s solution.
-The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
-\textbf{(B)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(C)} Solution networks.}
-  \label{sfig:scones_gsm}
-\end{figure}
-
 \end{document}

### 2020-08-26 manuscript/plos/review_1.tex
updated figures and added supplementary tables

@@
 Importantly, we find a significantly large overlap between the genes in the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
-Yet, network methods are notably unstable, producing different results when the input data changes slightly.
+Yet, network methods are notably unstable: the average Pearson correlation across all the methods between the genes selected on different subsamples of the data is 0.45, versus 0.57 simply using Bonferroni correction.
 To account for that, we produce a stable consensus subnetwork, formed by the most consistently selected genes.
@@
 Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
-It linked different parts of SigMod's solution network to four processes: protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
-Interestingly, the dmGWAS solution is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
+It linked different parts of SigMod's solution network to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
+Interestingly, the dmGWAS solution (\nameref{stab:dmgwas_pwy}) is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
 But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
+As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
 
@@
   \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-} As LEAN did not produce any significant gene (BH adjusted P\=/value~<~0.05), it was excluded.
+} As LEAN did not produce any significant network (BH adjusted P\=/value~<~0.05), it was excluded.
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
@@
 
-A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions.
+A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions (\nameref{stab:consensus_pwy}).
 We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
@@
 This is due to the particularities of each method, and directly or indirectly provide information about the dataset.
-For instance, the fact that LEAN did not return any biomarker implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
+For instance, the fact that LEAN did not return any gene implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
 
@@
 We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
-We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
+We used the function \texttt{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
 \end{description}
@@
 We searched for pathways enriched in the gene subnetworks produced by the above methods.
-We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the R package ReactomePA \cite{Yu2016}.
+We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the ReactomePA R package \cite{Yu2016}.
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
@@
 {\bf Summary statistics on the results of multiple network methods on the gene-gene interaction network.} The first row contains the summary statistics on the whole network.
+
+\paragraph*{S3 Table.}
+\label{stab:sigmod_pwy}
+{\bf Pathway enrichment analyses of the genes in SigMod solution.}
+
+\paragraph*{S4 Table.}
+\label{stab:dmgwas_pwy}
+{\bf Pathway enrichment analyses of the genes in dmGWAS solution.}
+
+\paragraph*{S5 Table.}
+\label{stab:hotnet2_pwy}
+{\bf Pathway enrichment analyses of the genes in HotNet2 solution.}
+
+\paragraph*{S6 Table.}
+\label{stab:consensus_pwy}
+{\bf Pathway enrichment analyses of the genes in the consensus network.}
 

### 2020-08-27 results/params/heinz/run.sh
start parametrization comparison

@@ -1,0 +1,6 @@
+cut -f2,9 ../../preprocessing/scored_genes.vegas.txt | sed 's/Top-0.1-pvalue/Pvalue/' >scored_genes.top10.txt
+for i in `seq 0 0.05 1`
+do
+  heinz.nf --vegas scored_genes.top10.txt --tab2 ../../preprocessing/hint.ht_complex.hgnc.pseudo.tab2 --fdr $i -profile bigmem -resume
+  mv selected_genes.heinz.txt selected_genes.heinz.$i.txt
+done

### 2020-08-28 results/params/heinz/run.sh
added parameter analyses

@@ -1,6 +1,7 @@
-cut -f2,9 ../../preprocessing/scored_genes.vegas.txt | sed 's/Top-0.1-pvalue/Pvalue/' >scored_genes.top10.txt
+# cut -f2,9 ../../preprocessing/scored_genes.vegas.txt | sed 's/Top-0.1-pvalue/Pvalue/' >scored_genes.top10.txt
+
 for i in `seq 0 0.05 1`
 do
   heinz.nf --vegas scored_genes.top10.txt --tab2 ../../preprocessing/hint.ht_complex.hgnc.pseudo.tab2 --fdr $i -profile bigmem -resume
-  mv selected_genes.heinz.txt selected_genes.heinz.$i.txt
+  mv selected_genes.heinz.txt heinz_$i.txt
 done

### 2020-08-28 results/params/sigmod/run.sh

@@ -1,0 +1,3 @@
+# wget https://github.com/YuanlongLiu/SigMod/raw/master/SigMod_v2.zip && unzip SigMod_v2.zip
+# cut -f2,9 ../../preprocessing/scored_genes.vegas.txt | sed 's/Top-0.1-pvalue/Pvalue/' >scored_genes.top10.txt
+sigmod.nf --sigmod SigMod_v2 --vegas scored_genes.top10.txt --tab2 ../../preprocessing/hint.ht_complex.hgnc.pseudo.tab2 -profile bigmem -resume 

### 2020-10-07 manuscript/plos/review_1.tex
fixed many reviewers' issues

@@
 
+\usepackage[shortcuts]{extdash}
+
 \begin{document}
@@
 {\Large
-\textbf\newline{Comparing and combining network-guided biomarker discovery to discover genetic susceptibility mechanisms to breast cancer on the GENESIS study} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Biological networks and GWAS: comparing and combining network methods to understand the genetics of familial breast cancer susceptibility in the GENESIS study} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 % Use the asterisk to denote corresponding authorship and provide email address in note below.
-* hector.climente@riken.jp
+* hector.climente(at)riken.jp
 
@@
 \section*{Abstract}
-Systems biology provides a comprehensive approach to biomarker discovery and biological hypothesis building.
-Indeed, it allows to jointly consider the statistical association between gene variation and a phenotype, and the biological context of each gene, represented as a network.
+Systems biology provides a comprehensive approach to uncovering the genetics of complex diseases and building hypotheses.
+Indeed, it allows to jointly consider the statistical association between genetic variation and a phenotype, and the biological context of each gene, represented as a network.
 In this work, we study six network methods which identify subnetworks with high association scores to a phenotype.
-Specifically, we examine their utility to discover new biomarkers for breast cancer susceptibility by interrogating a genome-wide association study (GWAS) focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
-We perform an in-depth benchmarking of the methods with regards to the size of the solution subnetwork, their utility as biomarkers, and the stability and the runtime of the methods.
-By trading statistical stringency for biological meaningfulness, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-For instance, we show a general alteration of the neighborhood of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-Importantly, we find a significantly large overlap between the genes in the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
+Specifically, we examine their utility to discover new breast cancer susceptibility genes by interrogating a genome-wide association study (GWAS) focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
+We perform an in-depth benchmarking of the methods with regards to the size of the solution, their predictive power, and the stability and the runtime of the methods.
+Network methods include prior knowledge in the analysis: they might tolerate selecting a gene with a high P-value, as long as it connects other two, low P-value genes.
+Thanks to this, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+For instance, at least two network methods selected 14 neighbors of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
+Importantly, we find significant overlaps between the genes in five of the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
 Yet, network methods are notably unstable: the average Pearson correlation across all the methods between the genes selected on different subsamples of the data is 0.45, versus 0.57 simply using Bonferroni correction.
-To account for that, we produce a stable consensus subnetwork, formed by the most consistently selected genes.
-The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than average.
+To account for that, we produce a stable consensus solution, formed by the most consistently selected genes.
+The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than the average gene.
 The network seems organized around \emph{CUL3}, encoding a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
@@
 Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
-An elegant explanation has been proposed with the omnigenic model: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated so that many other genes can alter the function of the core genes.
+The omnigenic model offers an elegant explanation: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated so that many other genes can alter the function of the core genes.
 These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
@@
 Combining these approaches allows us to identify potentially novel breast cancer susceptibility genes, and provides a mechanistic explanation for their role in the development of the disease.
-Our pipeline can easily be applied to other diseases.
-
+We provide ready-to-use implementations of all the examined methods.
 \linenumbers
@@
 Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}.
-In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}.
-This is particularly true for cancer driver genes, which tend to be key players in connecting different, densely-connected communities of genes.
+In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}: they tend to occupy central positions in the network, interconecting different modules TODO.
 Therefore, studying the neighborhood of disease-associated genes is effective at identifying new ones that are involved in the same biological functions \cite{huang_systematic_2018}.
 
-Network-based biomarker discovery methods exploit this relatedness to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
-In essence, each SNP has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
+Network-based discovery methods exploit the differential properties described above to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
+In essence, each gene has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
 Then, the problem becomes finding a functionally-related set of highly-scoring genes.
@@
 instances were both a gene and its direct interactors are associated with the disease.
-Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, focus on larger subnetworks interconnecting genes with high association scores.
+Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, do not impose such strong constraints, and search for subnetworks interconnecting genes with high association scores.
 However, they differ in their tolerance to the inclusion of low-scoring nodes, and the topology of the solution.
@@
 
-In this work, we analyze the effectiveness of these six network methods for biomarker discovery on GWAS data.
-While all of them capture susceptibility mechanisms resembling that postulated by the omnigenic model, they do so in different ways, and provide a representative view of the field.
+In this work, we analyze the effectiveness of these six network methods on GWAS data.
+They do so by using different interpretations of the omnigenic model, providing a representative view of the field.
 We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
-After a classical GWAS approach, we use these network methods to recover additional breast cancer biomarkers.
-Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus network of predisposition to familial breast cancer.
+After a classical GWAS approach, we use these network methods to recover additional breast cancer susceptibility genes.
+Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus solution of predisposition to familial breast cancer.
 
@@
 
-We conducted association analyses in the GENESIS dataset at both SNP and gene levels (Section~\ref{methods:node_score}).
-Two genomic regions have a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
-The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16}, and it is located near the protein-coding gene \emph{TOX3}.
+We conducted association analyses in the GENESIS dataset (Section~\ref{methods:data}) at both SNP and gene levels (Section~\ref{methods:conventional}).
+Two genomic regions had a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
+The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16} the protein-coding gene \emph{TOX3}.
 Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
-In our studied population, only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (\nameref{sfig:snp_gene_manhattan}B).
-
-Closer examination reveals two regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
+In GENESIS only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (\nameref{sfig:snp_gene_manhattan}B).
+
+Closer examination reveals two other regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
 Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression on the whole dataset (Section~\ref{methods:classifier}).
-It selected 100 SNPs, both from all aforementioned regions and new ones (\nameref{sfig:snp_gene_manhattan}C).
+We applied an L1-penalized logistic regression using all the GENESIS genotypes as input, and the phenotype (cancer/healthy) as outcome (Section~\ref{methods:classifier}).
+The algorithm selected 100 SNPs, both from all aforementioned regions and new ones (\nameref{sfig:snp_gene_manhattan}C).
 Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
-Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:node_score}) are not interconnected in the protein-protein interaction network (PPIN, Section~\ref{methods:networks}).
-In addition, the classification performance of the method is very low, and L1-penalized methods select only one of several correlated variables and are prone to instability, which further complicates interpretation.
-This motivates exploring network methods, which trade statistical significance for biological relevance to find susceptibility subnetworks.
-In fact, such methods provided comparably (poor) classification performance to L1-penalized logistic regression (Fig~\ref{fig:benchmark}B), while providing more interpretable solutions.
+Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:snp2gene}) are not interconnected in the protein-protein interaction network (PPIN, Section~\ref{methods:networks}).
+Moreover, the classification performance of the model is low (sensitivity = 55\%, specificity = 55\%, Section \ref{methods:benchmark}).
+Together, these issues motivate exploring network methods, which consider not only statistical association, but also the location of each gene in a PPIN to find susceptibility genes.
 
@@
 We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (Benjamini-Hochberg [BH] correction adjusted P\=/value < 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+As none of the networks examined by LEAN was significant (Benjamini-Hochberg [BH] correction adjusted P\=/value $<$ 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
 These solutions differ in many aspects, making it hard to draw joint conclusions.
-For starters, the overlap between the genes featured in each solution is quite small (Fig~\ref{fig:solution_overview}B).
-Another prominent difference is their size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
-While SConES GI failed to recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
-Their topologies also differ, as measured by the median centrality (Table~\ref{tab:gene_solutions}) and the number of connected components (Fig~\ref{fig:solution_overview}D).
+For starters, the overlap between the genes featured in each solution is quite small (Fig~\ref{fig:solution_overview}A).
+Despite that, the methods tend to agree on the genes with the strongest signal: genes selected by more methods tended to have lower P-value of association (Fig~\ref{fig:solution_overview}B).
+Another prominent difference is the solution size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
+While SConES GI did not recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
+Their topologies also differ (Fig~\ref{fig:solution_overview}C), as measured by the median centrality  and the number of connected components (Table~\ref{tab:gene_solutions}).
 Only two methods have more than one connected component: SConES, as described above, and HotNet2.
@@
 The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
-Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences in the underlying mechanisms.
-It linked different parts of SigMod's solution network to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~<~0.03).
+Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences between the solutions of the different methods.
+It linked different parts of SigMod's solution network to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
 Interestingly, the dmGWAS solution (\nameref{stab:dmgwas_pwy}) is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~<~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~<~0.04).
+But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~$<$~0.04).
 
@@
   \caption{
-  {\bf Summary statistics on the solutions of multiple network methods on the gene-gene interaction network. The first row contains the summary statistics on the whole network.}}
-  \begin{tabular}{lrrrrr}
-  {\bf Network } & {\bf \# genes } & {\bf \# edges } & {\bf \mean{Betweenness} } & {\bf \median{P}\textsubscript{gene} } & {\bf \(\rho\)\textsubscript{consensus} }\\
+  {\bf Summary statistics on the solutions of multiple network methods on the PPIN. The first row contains the summary statistics on the whole PPIN.}}
+  \begin{tabular}{lrrrrrr}
+  {\bf Network } & {\bf \# genes } & {\bf \# edges } & {\bf \# components } & {\bf \mean{Betweenness} } & {\bf \median{P}\textsubscript{gene} } & {\bf \# genes in consensus } \\
   \thickhline
-  HINT HT & 13\,619 & 142\,541 & 16\,706 & 0.46 & 0.066\\
+  HINT HT           & 13\,619 & 142\,541  & 15  & 16\,706   & 0.46  & 93/93 \\
   \hline
-  dmGWAS & 194 & 450 & 49\,115 & 0.19 & 0.41\\
-  heinz & 4 & 3 & 113\,633 & 0.001 & 0.21\\
-  HotNet2 & 440 & 374 & 7\,739 & 0.048 & 0.31\\
-  LEAN & 0 & 0 & - & - & -\\
-  SConES GI & 0 (1) & 0 & - & - & -\\
-  SigMod & 142 & 249 & 92\,603 & 0.008 & 0.73\\
+  dmGWAS            & 194     & 450       & 1   & 49\,115   & 0.19  & 55/93 \\
+  heinz             & 4       & 3         & 1   & 113\,633  & 0.001 & 4/93  \\
+  HotNet2           & 440     & 374       & 130 & 7\,739    & 0.048 & 63/93 \\
+  LEAN              & 0       & 0         & 0   & -         & -     & 0/93  \\
+  SConES GI         & 0 (1)   & 0         & 0   & -         & -     & 0/93  \\
+  SigMod            & 142     & 249       & 11  & 92\,603   & 0.008 & 84/93 \\
   \hline
-  Consensus & 93 & 186 & 50\,737 & 0.006 & 1\\
-  Stable consensus & 68 & 49 & 94\,854 & 0.005 & 0.54 \\
+  Consensus         & 93      & 186       & 21  & 50\,737   & 0.006 & 93/93 \\
+  Stable consensus  & 68      & 49        & 32  & 94\,854   & 0.005 & 43/93 \\
   \end{tabular}
-  \begin{flushleft} \textbf{\emph{\# genes}}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
+  \begin{flushleft} \textbf{\# genes}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
+{\bf \# components }: number of connected components.
 \textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
-\textbf{\median{P}\textsubscript{gene}}: median P\=/value of the selected genes.
-\textbf{\(\rho\)\textsubscript{consensus}}: Pearson correlation between the subnetwork and the consensus network.
+\textbf{\median{P}\textsubscript{gene}}: median VEGAS2 P\=/value of the selected genes.
+\textbf{\# genes in consensus}: Number of genes in common between the method's solution, and the 93 genes in the consensus solution.
   \end{flushleft}
@@
   \centering
-  \includegraphics[height=\textwidth]{./figures/figure_1.png}
+  \includegraphics[width=\textwidth]{./figures/figure_1.png}
   \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-} As LEAN did not produce any significant network (BH adjusted P\=/value~<~0.05), it was excluded.
+} As LEAN did not produce any significant network (BH adjusted P\=/value~$<$~0.05), it was excluded.
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
-\textbf{(A)} Manhattan plots of SNPs/genes; in black, the method's solution.
+\textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
+\textbf{(B)}~VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods.
+\textbf{(C)}~Manhattan plots of SNPs/genes; in black, the method's solution.
 The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{(B)} Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
-\textbf{(C)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(D)} Solution networks produced by the different methods.}
+\textbf{(D)}~Solution networks produced by the different methods.}
   \label{fig:solution_overview}
@@
 Despite their differences, there are additional common themes.
-All obtained solution subnetworks have lower association P\=/values than the whole PPIN (median P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Fig~\ref{fig:solution_overview}A).
-This exemplifies the trade-off between statistical significance and biological relevance.
-However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
-Additionally, the solution subnetworks presented other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value < 0.03).
-Second, the genes in four solution subnetworks displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value < 1.4 \texttimes{} 10\textsuperscript{-21}).
+All obtained solutions have lower association P\=/values than the whole PPIN (median VEGAS2 P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Fig~\ref{fig:solution_overview}D).
+This exemplifies the trade-off between controlling for type I error and biological relevance.
+However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+Additionally, the solutions presented other desirable properties.
+First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value$<$0.03).
+Second, the genes in four solutions displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value$<$1.4 \texttimes{} 10\textsuperscript{-21}).
 This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (Supplementary table~\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
-
-\subsection{A case study: the consensus network}
+Interestingly, SConES selected SNPs that are also more central than the average SNP (\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
+
+\subsection{A case study: the consensus solution}
 \label{results:consensus}
 
-Despite the heterogeneity of the solutions, their shared properties suggest that each method captures different aspects of cancer susceptibility.
-Indeed, only 20 genes are common to more than two solutions (\nameref{sfig:consensus_stats}A), but encouragingly, the more methods selected a gene, the higher its association score to the phenotype (\nameref{sfig:consensus_stats}B).
-To leverage on their strengths and compensate their respective weaknesses, we built a consensus subnetwork that captures the mechanisms most shared among the solution subnetworks (Section~\ref{methods:methods}).
-This subnetwork (Fig~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
+Despite their shared properties, the differences between the solutions of the different methods suggest that each of them captures different aspects of cancer susceptibility.
+Indeed, out of the 668 genes that are selected by at least one method, only 93 are selected by at least two, 20 by three, and none by four or more.
+Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig \ref{fig:solution_overview}B), a relationship that plateaus at 2.
+Hence, to leverage on their strengths and compensate their respective weaknesses, we built a consensus solution that captures the genes shared among at least two solutions (Section~\ref{methods:methods}).
+This solution (Fig~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
 
@@
   \includegraphics[width=.7\linewidth]{./figures/figure_3.pdf}
-  \caption{ {\bf Consensus subnetwork on GENESIS (Section~\ref{methods:methods}) }.
+  \caption{ {\bf Consensus solution on GENESIS (Section~\ref{methods:methods}) }.
 Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
 The latter ones are also colored in pink.
@@
 
-A pathway enrichment analysis of the genes in the consensus network also shows similar pathways as the individual solutions (\nameref{stab:consensus_pwy}).
+A pathway enrichment analysis of the genes in the consensus solution also shows similar pathways as the individual solutions (\nameref{stab:consensus_pwy}).
 We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
@@
 Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
-The opposite would be the case of highly central genes in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}D).
-This suggests that they were selected because they were on the shortest path between two highly associated genes.
+In the former we find some genes that are highly central in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
+This suggests that these genes were selected because they were on the shortest path between another two highly associated genes.
 In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus network is COPS5, a component of the COP9 signalosome which regulates multiple signaling pathways.
+For instance, the most central node in the consensus solution is COPS5, a component of the COP9 signalosome which regulates multiple signaling pathways.
 \emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus subnetwork have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
-
-\subsection{Network methods boost biomarker discovery}
+Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus solution have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
+
+\subsection{Network methods boost discovery}
+\label{results:boost}
 
@@
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
-We computed a gene association score on BCAC, in an equivalent way to the one described in Section~\ref{methods:node_score}.
-The solutions provided by the different network approaches overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
-The gene-based network methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (Fig~\ref{fig:solution_overview}C).
-Interestingly, while SConES GI at the SNP-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+We computed a gene association score on BCAC (Section~\ref{methods:bcac}).
+The solutions provided by the different  network methods overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
+The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (\nameref{sfig:additional-benchmarks}A).
+interestingly, while scones gi at the snp-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
 
 \subsection{Network methods share limitations}
+\label{results:benchmark}
 
 We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:benchmark}).
-Specifically, we measured five properties (Fig~\ref{fig:benchmark}): size of the solution subnetwork; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
+Specifically, we measured five properties (Fig~\ref{fig:benchmark}): size of the solution; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
 The solution size varies greatly between the different methods (Fig~\ref{fig:benchmark}A).
-Heinz produced the smallest solutions, with an average of 182 selected SNPs.
+Heinz produced the smallest solutions, with an average of 182 selected SNPs (Section~\ref{methods:snp2gene}).
 The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
-To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (Fig~\ref{fig:benchmark}B).
+LEAN did not produce any solution in any of the subsamples.
+To determine whether the selected snps could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (\nameref{sfig:additional-benchmarks}B).
 The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
 Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data.
+It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data \cite{deloscamposComplexTraitPredictionEra2018}.
 
@@
 As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
-\textbf{(A)} Number of SNPs selected by each method and number of SNPs in the active set found by the classifier.
-Points are the average over the 5 runs; lines represent the standard error of the mean.
-A grey diagonal line with slope 1 is added for comparison.
+\textbf{(A)}~Number of SNPs selected by each method and number of SNPs in the active set found by the classifier(Section~\ref{methods:classifier}).
+Points are the average over the 5 runs; the error bars represent the standard error of the mean.
+A grey diagonal line with slope 1 is added for comparison, indicating the upper bound of the active set (i.e.
+the number of SNPs in the solution).
 For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs.
-\textbf{(B)} Sensitivity and specificity on the test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
-The performance of the classifier trained on all SNPs is also displayed.
-Points are the average over the 5 runs; lines represent the standard error of the mean.
-\textbf{(C)} Pairwise Pearson correlations of the solutions used by different methods.
+\textbf{(C)}~Pairwise Pearson correlations of the solutions produced by different methods.
 A Pearson correlation of 1 means the two solutions are the same.
 A Pearson correlation of 0 means that there is no SNP in common between the two solutions.
-\textbf{(D)} Runtime of the evaluated methods, by type of network used (gene or SNP).
-For gene network-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.}
+\textbf{(D)}~Runtime of the evaluated methods, by type of network used (PPIN or SNP).
+For gene-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.}
   \label{fig:benchmark}
@@
 Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
-Both heinz and LEAN displayed high stability in our benchmark, consistently selecting the same genes and no genes over the 5 subsamples, respectively (Fig~\ref{fig:benchmark}C).
-Conversely, the other methods displayed similarly low stabilities.
+Heinz was high stabile in our benchmark, while the other methods displayed similarly low stabilities (Fig~\ref{fig:benchmark}C).
 
@@
 As shown above, and despite their similarities, the network methods produced remarkably different solutions.
-This is due to the particularities of each method, and directly or indirectly provide information about the dataset.
+This is due to the fact that each of them models the optimal solution differently.
+Importantly, understanding which assumptions they make allows to understand the dataset more deeply.
 For instance, the fact that LEAN did not return any gene implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
 
-In this dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value for the subnetwork (Table~\ref{tab:gene_solutions}).
-Due to this parsimonious and highly associated solution, it was the best method to stably select a set of biomarkers (Fig~\ref{fig:benchmark}C).
+In the GENESIS dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value (Table~\ref{tab:gene_solutions}).
+By repeatedly selecting this compact solution, heinz was the most stable method (Fig~\ref{fig:benchmark}C).
 Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
@@
 On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
-dmGWAS' subnetwork is the least associated subnetwork on average.
+dmGWAS' solution is the lowest scoring solution on average.
 This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
-It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated gene if it was linked to another, strongly associated one.
+It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated gene if it was linked to another, high scoring one.
 This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
 This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
-In the same line, HotNet2's subnetwork is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
-Lastly, SigMod misses some of the most strongly associated, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
+In the same line, HotNet2's solution is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
+Lastly, SigMod misses some of the highest scoring, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
 
@@
   \includegraphics[width=\linewidth]{./figures/figure_2.pdf}
-  \caption{{\bf Drawbacks encountered when using network guided methods.} \textbf{(A)}~DmGWAS solution subnetwork.
-Genes with a P\=/value < 0.1 are highlighted in red.
+  \caption{{\bf Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
 \textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
@@
 We labeled the three most central genes that were picked by any method.
-\textbf{(C)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
+\textbf{(C)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its centrality degree.
+\textbf{(D)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
 SNPs that were not selected in the studied network, but were selected in another one, are displayed in background color.
@@
 Another peculiarity of network methods is their relationship to degree centrality.
-On the one hand we observed that highly central genes often had no association to disease (Fig~\ref{fig:issues}B).
-On the other, network methods favor central genes, as they often connect high scoring nodes.
+We studied randomly rewirings of the PPIN while preserving node centrality (Section~\ref{methods:rewiring}).
+In this setting, network methods favored central genes, as they often connect high scoring nodes (Fig~\ref{fig:issues}B).
+This is despite the fact that highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
 This was specially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
@@
 It therefore can select SNPs in genes without any associated interactor, as well as SNPs in non-coding regions or in non-interacting genes.
-In fact, due to linkage disequilibrium, SConES favors such genes, as selecting SNPs in an LD-block which overlaps with a gene favors selecting the rest of the gene.
-This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (\nameref{sfig:pearson_methods}).
-On the other hand, SConES penalizes selecting SNPs and not their neighbors.
-This makes it conservative regarding SNPs with many interactions, like those mapped to hub genes in the PPIN.
-For this reason, SConES GI did not select any protein coding gene, despite selecting similar regions as SConES GS (Fig~\ref{fig:issues}C).
-In fact SConES GS and SConES GM select regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
+We compared the solution of SConES in the GI network (using PPIN information), to the one using only positional information (GS network), or positional and gene annotations (GM network).
+Importantly, SConES produces similar results on the GS and GM networks (\nameref{sfig:pearson_methods}).
+While the solutions on those two greatly overlap with SConES GI's, they contain additional gene-coding segments (Fig~\ref{fig:issues}C).
+In fact the formers' solutions are chromosome regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
 On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
-We hypothesize that the lack of results on the PPIN network of SConES GI and LEAN are due to the same cause: the absence of joint association of a gene and a majority of its neighbors.
-Although in the case of SConES other hyperparameters could lead to a more informative solution (e.g.
-a lower \(\lambda\) in Equation~\ref{eq:scones}), it is unclear what the best strategy to find them is.
-In addition, due to the design of the iCOGS array, the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
 
@@
 Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
-We built a new, stable consensus network using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
+We built a new, stable consensus solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
 Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
@@
   \includegraphics[width=.7\linewidth]{./figures/figure_5.pdf}
-  \caption{{\bf Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+  \caption{{\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}).} Each node is represented by a pie chart, which shows the methods that selected it.
+We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
 The latter ones are also colored in pink.
@@
 
-  However, although this new network exhibits similar global properties as the previous one, the lack of stability results in different genes being selected.
-In this case, the most central gene is \emph{CUL3}, which is absent from the previous consensus network and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
+Despite these similarities, the consensus and the stable consensus solutions include different genes.
+In this case, the most central gene is \emph{CUL3}, which is absent from the previous consensus solution and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
 This gene is a component of Cullin-RING ubiquitin ligases.
@@
 Network methods tackle these two issues by using both the association score and an interaction network to take into consideration the biological context of each of the genes and SNPs.
-Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided biomarker discovery have been proposed in recent years.
-In this article we evaluated the relevance of six of them by examining a GWAS dataset on familial breast cancer.
-
-Most of the network methods produced a relevant subset of biomarkers, recapitulating known breast cancer susceptibility genes.
+Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided discovery have been proposed in recent years.
+In this article we evaluated the relevance of six of them by examining GENESIS, a GWAS dataset on familial breast cancer.
+
+DmGWAS, Heinz, HotNet2, SConES and SigMod relevant solutions, which recapitulate known breast cancer susceptibility genes.
 In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
-However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by these network methods.
-Across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
-Yet, further work is needed to characterize the impact of centrality on network methods' outputs.
-Despite these similarities, the solutions were notably different.
-At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, providing a conveniently short list of biomarkers, at the expense of not shedding much light on the etiology of the disease.
+However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by network methods (Section \ref{results:drawbacks}).
+Yet, across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
+Despite these similarities, the solutions of the methods were notably different.
+At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, at the expense of not shedding much light on the etiology of the disease.
 On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
 While this deepens our understanding of the disease and provides biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
+For instance, examining the network properties of the genes in the solution (e.g.
+centrality and neighbors).
+Those might drive the selection of genes, especially of those with low association scores.
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
@@
 First, they are all equally bad at discriminating cases from controls.
-Yet, the classification accuracy of network methods is similar to that of a machine learning classifier, which suggests that cases and controls are difficult to separate in this dataset.
+Yet, the classification accuracy of network methods is similar to that of a machine learning classifier, which suggests that cases and controls are difficult to separate in the GENESIS dataset.
 This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
@@
 Hence, heinz's conservative transformation of P\=/values, which favors only the most extreme ones, leads to improved stability.
-Another source of instability might be the redundancy inherent to biological networks.
-
-To overcome the limitations of the individual methods while exploiting their strengths, we proposed combining them into a consensus subnetwork.
+Another source of instability might be the redundancy inherent to biological networks, since they are subject to an evolutionary pressure to avoid single points of failure \cite{wagnerAlternativeRoutesMutational2007}.
+Hence, often network methods will have multiple paths connecting two high-scoring nodes.
+
+To overcome the limitations of the individual methods while exploiting their strengths, we proposed combining them into a consensus solution.
 We use a straightforward strategy of including any node that was recovered by multiple methods.
-We proposed two networks: a consensus network that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus network, that addressed the instability of the methods.
+We proposed two networks: a consensus solution that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus solution, that addressed the instability of the methods.
 They both synthesized the altered mechanisms: they both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer.
 Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but who are related to cancer at the expression level and whose neighborhood has consistent high association scores.
-Crucially, the consensus was as unstable as the tested methods, while the stable consensus shared these properties while accounting for instability.
+Importantly, while the consensus approach was as unstable as the individual network-guided methods, the stable consensus network retained the ability to provide compelling hypotheses and had better stability.
 This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence validates the conclusions obtained on the individual solutions and the consensus.
+
+Across this article, we have compared our results to significant genes and SNPs in the BCAC study \cite{Michailidou2017}.
+Network methods show modest precision, but much higher recall at recovering BCAC biomarkers (Section \ref{results:boost}).
+While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in explotatory settings.
+Nonetheless, BCAC is not an ideal counterfactual in which GENESIS had more samples.
+Yet, there are caveats which make it a non-optimal comparison.
+First, the studied populations are non-overlapping: BCAC focuses on a pan-European cohort, while GENESIS targets the French population specifically.
+Second, the study designs differ: a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focuses on the French high-risk population attending the family cancer clinics.
+Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
+Lastly GWAS are unlikely to identify all genes relevant for the disease: some might only show up in rare-variant studies; others might have too low effect sizes.
+Network methods account for this by including genes with low association scores but with relevant topological properties.
+Hence, network methods and GWAS are unlikely to capture exactly the same sets of genes.
+This might partly excuse the low precisions displayed in Section \ref{results:boost} and the low AUC displayed in Section \ref{results:benchmark}.
 
@@
 In consequence, they show their shortcomings on understudied genes, especially those not in the network.
-Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the protein-protein interaction network.
+Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the PPIN.
 The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (\nameref{sfig:biotypes_excluded}).
 Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
-In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, located near \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the network.
-However, on average protein-coding genes absent from the PPIN are less associated with this phenotype (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
-As we are using interactions from high-throughput experiments, such difference cannot be due to well-known genes having more known interactions.
+In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, overlapping with \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the PPIN.
+However, on average protein-coding genes absent from the PPIN are less associated with breast cancer susceptibility (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
+This cannot be due to well-known genes having more known interactions because we are only using interactions from high-throughput experiments.
 As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
-It is worth noting that network approaches that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
+It is worth noting that  network methods that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
 Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
 This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
-Lastly, all the methods are heavily affected by how SNPs are mapped to genes.
-In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
-In fact, the presented case is paradigmatic, since the genes are in the most gene-dense region of the genome \cite{Xie2003}.
-Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the network, but might be more resilient to them when the overlapping genes are unrelated.
-Making use of more targeted mappings of SNPs to genes (e.g.
-eQTLs, SNPs associated to the expression of a gene), altogether with a stringent LD pruning, might address such problems.
+Lastly, all the methods are heavily affected by how SNPs are mapped to genes, and other strategies (e.g.
+eQTLs, SNPs associated to the expression of a gene) might lead to different results.
 
@@
 On the other hand, Huang et al.
-\cite{huang_systematic_2018} found that the best predictor of the performance of a network for disease gene discovery is the size of the network, which supports using the largest amount of interactions.
-When we compared the impact of using a larger network containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
+\cite{huang_systematic_2018} found that network methods performed better on larger methods than in smaller networks at identifying disease genes.
+This would support using the largest networks in our experiments.
+When we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
 This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best-known genes are artificially pushed into the solutions.
@@
 In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms \cite{nakka_gene_2016,ionita-laza_sequence_2013,wang_combat:_2017} might have more statistical power.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms (e.g.
+PEGASUS \cite{nakka_gene_2016}, SKAT \cite{ionita-laza_sequence_2013} or COMBAT \cite{wang_combat:_2017}) might have more statistical power.
 Another important decision is how to handle LD in a GWAS.
@@
 
-To produce the two consensus networks, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
+How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners.
+In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+In fact, the presented case is paradigmatic, since all three genes are located in the HLA region, the most gene-dense region of the genome \cite{Xie2003}.
+Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the PPIN.
+But the opposite case is also true: when nearby genes are not functionally related (and hence disconnected in the PPIN), network methods might disregard them even if they have high association scores.
+LD also affects SConES, since it penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Equation~\ref{eq:scones}.
+Due to linkage disequilibrium, nearby SNPs' P-values are be correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
+Hence, SConES will tend to select LD-blocks formed by low P-value SNPs.
+This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Section \ref{results:drawbacks}).
+However, this same behavior raises the burden of proof required to select SNPs with many interactions, like those mapped to hub genes in the PPIN.
+For this reason, SConES GI did not select any protein coding gene.
+We hypothesize that this is caused by the absence of joint association of a gene and a majority of its neighbors.
+This is supported by the lack of results from LEAN as well.
+Yet, a different combination of parameters could lead to a more informative SConES' solution (e.g.
+a lower \(\lambda\) in Equation~\ref{eq:scones}), although it is unclear how to find it.
+In addition, due to the design of the iCOGS array (Section~\ref{methods:data}), the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
+A stringent LD pruning might address such problems.
+
+To produce the two consensus solutions, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
 To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
-They are available on GitHub: \url{https://github.com/hclimente/gwas-tools}.
+They are available on GitHub: \url{https://github.com/hclimente/gwas-tools} (Section \ref{methods:data}).
 Importantly, those methods that had a permissive license were compiled into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
@@
 
-\subsection{GENESIS}
+\subsection{GENESIS dataset, preprocessing and quality control}
+\label{methods:data}
 
@@
 Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
-We focused on the 2\,577 samples of European ancestry, of which 1\,279 are controls and 1\,298 are cases.
+We focused on the 2\,577 samples of European ancestry, of which 1\,279 were controls and 1\,298 were cases.
 The genotyping was performed using the iCOGS array, a custom Illumina array designed to study the genetic susceptibility to hormone-related cancers \cite{sakoda_turning_2013}.
 It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
-
-\subsection{Preprocessing and quality control}
 
@@
 
-\subsection{High-score subnetwork search algorithms}
-\subsubsection{SNP and gene association}
-\label{methods:node_score}
-To measure association between a genotype and the phenotype, we performed a per-SNP 1~d.f.~\(\chi\)\textsuperscript{2} allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
+\subsection{SNP- and gene-based GWAS}
+\label{methods:conventional}
+
+To measure association between a genotype and susceptibility to breast cancer, we performed a per-SNP 1~d.f.~$\chi^2$ allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
+In order to obtain significant SNPs, we performed Bonferroni correction, to keep family-wise error rate below 5\%.
+The threshold used was [$]\frac {0.05} {197083} = 2.54 \times 10^{-7}[/$].
+
 Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
@@
 Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
-
-We used such mapping to compare the outputs of methods that produce SNP-lists to those that produce gene-lists, and vice versa.
-For the former, we considered any gene that can be mapped to any of the selected SNPs as selected as well.
-For the latter, we considered all the SNPs that can be mapped to that gene as selected by the method.
+We also used Bonferroni correction to obtain significant genes; in this case, the threshold of signigicance was [$]\frac {0.05} {32767} = 1.53 \times 10^{-6}[/$].
+
+\subsection{High-score subnetwork search algorithms}
 
@@
 The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Yet, we used protein-protein interaction networks (PPIN) for all of them except SConES, as they are interpretable, well-characterized, and they were designed to run efficiently on networks of their size.
+Yet, we used protein-protein interaction networks (PPIN) for all them except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the network.
+Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the PPIN.
 Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
@@
 However, at different stages we also used GS and GM for comparison.
-For the GM network, we used the mapping described in Section~\ref{methods:node_score}.
+For the GM network, we used the mapping described in Section~\ref{methods:snp2gene}.
 In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
@@
 Their main differences are summarized in Table~\ref{tab:method_comparison}.
+We scored both SNPs and genes with the P-values (or transformations of them) computed in Section~\ref{methods:conventional}.
 
@@
 \textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
-\textbf{\emph{Solution}}: additional properties are enforced on the solution subnetwork, other than containing high scoring, connected nodes.
+\textbf{\emph{Solution}}: additional properties are enforced on the solution, other than containing high scoring, connected nodes.
 \textbf{\emph{Comp.}}: number of connected components in the solution.
@@
 \item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
-To that end it searches candidate subnetwork solutions using a greedy, ``seed and extend'', heuristic:
+To that end it searches candidate solutions using a greedy, ``seed and extend'', heuristic:
 
@@
 
-DmGWAS carries out this process on every gene in the network.
+DmGWAS carries out this process on every gene in the PPIN.
 We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
 We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
-We used the function \texttt{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
-\end{description}
-
-\begin{description}
+We used the function \texttt{simpleChoose} to select the solution, which aggregates the top 1\% subnetworks.
+
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P\=/value into a score that is negative under no association with the phenotype, and positive when there is.
+The authors propose a transformation of the genes’ P\=value into a score that is negative under weak association with the phenotype, and positive under a strong one.
 This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
@@
 \item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}.
-To that end, it considers both the local topology of the network and the scores of the nodes.
+To that end, it considers both the local topology of the PPIN and the scores of the nodes.
 The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors while retaining part of its own (hence, \emph{insulated}).
-This process continues until equilibrium is reached, and results in a diffusion matrix $F$.
+This process continues until a stationary state is reached, in which the temperature of nodes does not change anymore, and results in a diffusion matrix $F$.
 $F$ is used to compute the similarity matrix $E$ that models exchanged heat as
@@
 For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
-To obtain densely connected subnetworks, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
-Lastly, HotNet2 evaluates the statistical significance of the subnetworks by comparing their size to the size of networks obtained by permuting the node scores.
+To obtain densely connected solutions, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
+Lastly, HotNet2 evaluates the statistical significance of the solutions by comparing their size to the size of PPINs obtained by permuting the node scores.
 We assigned initial node scores as in Nakka et al.
 \cite{nakka_gene_2016}, assigning a score of 0 for the genes with a low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
-In this dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
+In the GENESIS dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
 HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
@@
 We used the version on SConES implemented in the R package martini \cite{martini}.
-\end{description}
-
-\begin{description}
-\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
+
+\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected subnetwork \cite{liu_sigmod:_2017}.
 It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
@@
 SigMod presents three important differences with SConES.
-First it is designed for gene-gene networks.
-Second, it favors subnetworks containing many edges.
+First it is designed for PPINs.
+Second, it favors solutions containing many edges.
 SConES, instead, penalizes connections between the selected and unselected nodes.
@@
 SigMod is implemented in an R package \cite{sigmod}.
+
+\item[{Consensus}] We built a consensus solution by retaining the genes that were selected by at least two of the six methods (using SConES GI for SConES).
+Any edge between the selected genes in the PPIN was included.
 \end{description}
 
+We performed all the computations in the cluster described in Section \ref{methods:computation}.
+
+\subsubsection{Parameter space}
+\label{methods:parameters}
+
+We used the network methods with the parameters recommended by their authors, or with the default parameters in their absence.
+Additionally, we explored the parameter space of the different methods to study how they alter the output.
+
 \begin{description}
-\item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).
+\item[{dmGWAS}] We essayed multiple values for $r$ (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and $d$ (1, 2, and 3). 
+\item[{heinz}] We essayed multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
+\item[{HotNet2}] We essayed different thresholds to decide which genes would receive a score of 0, and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
+\item[{LEAN}] We used the following significance cutoffs for LEAN's P-values (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, and 1).
+\item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
+\item[{SigMod}] We used essayed multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
 \end{description}
+
+\subsubsection{Comparing SNP-methods to gene-methods, and vice versa}
+\label{methods:snp2gene}
+
+In multiple steps of this article, we needed to compare the outcome of a methods that works on genes with the outcome of another methods that works on SNPs, or vice versa.
+For this purpose, we used the SNP-gene correspondance described in Section \ref{methods:conventional}.
+To convert a list of SNPs into a list of genes, we included all the genes that can be mapped to any of the SNPs.
+Conversely, to convert a list of genes into a list of SNPs, we included all the SNPs that can be mapped to any of the genes.
 
@@
 
-We searched for pathways enriched in the gene subnetworks produced by the above methods.
+We searched for pathways enriched in the gene solutions produced by the above methods.
 We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the ReactomePA R package \cite{Yu2016}.
-The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:node_score}).
+The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:conventional}).
 We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH).
 Pathways with a BH adjusted P\=/value \textless~0.05 were deemed significant.
-As the significant pathways are often overlapping and redundant, the results were manually curated afterward.
 
@@
 We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
-We applied each method to 5 random subsets of the original dataset containing 80\% of the samples (\emph{train set}).
+We applied each method to 5 random subsets of the original GENESIS dataset containing 80\% of the samples (\emph{train set}).
 When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
 We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
+Every method and repetition ran on the same computational settings (Section \ref{methods:computation}).
 
@@
 
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,642 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
-Hence, a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focuses on the French high-risk population attending the family cancer clinics.
-Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
-For that purpose, we searched associated genes as in Section~\ref{methods:node_score}.
-We provided VEGAS2 with both the summary statistics of all available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,641 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
+Through imputation, BCAC includes more SNPs than the iCOGS array used for GENESIS (Section~\ref{methods:data}).
+Yet, in all the comparisons in this paper, we focused on the subset of the GENESIS SNPs that passed quality control (Section~\ref{methods:data}).
+Hence, we used the same Bonferroni threshold as in Section~\ref{methods:conventional} to determine the significant SNPs in BCAC.
+We also computed gene-scores in the BCAC data using VEGAS2, as in Section~\ref{methods:data}.
+In this case, we did use the summary statistics of all 13\,250\,641 available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
+Since the these genotypes did not include chromosome X, we excluded that chromosome in this analysis.
+All comparisons included only the common genes between GENESIS and BCAC, so we used a less  stringent Bonferroni threshold ($1.66 \times 10^{-6}$) to call gene significance.
+
+\subsubsection{Network rewirings}
+\label{methods:rewiring}
+
+Rewiring the PPIN while preserving the number of edges of each gene allows to study the impact of the topology on the output of network methods.
+We produced 100 such rewirings by swapping edges in the PPIN.
+We scored the genes as described in section \ref{methods:methods}.
+We only applied only four methods on the rewirings: heinz, dmGWAS, LEAN and SigMod.
+We excluded HotNet2, since it takes notably longer to run than the other methods, while taking up more computational resources.
 
 \subsection{Computational resources}
+\label{methods:computation}
 
@@
 
-\subsection{Code availability}
+\subsection{Code and data availability}
+\label{methods:code}
 
@@
 Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
-We deposited all the produced gene subnetworks on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
+We deposited all the produced gene solutions on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
+
+Summary statistics for SNPs and genes are available at \url{https://github.com/hclimente/genewa}.
+We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact \href{mailto:nadine.andrieu@curie.fr}{nadine.andrieu@curie.fr}.
 
@@
 \label{stab:scones_gene_solutions}
-{\bf Summary statistics on the results of multiple network methods on the gene-gene interaction network.} The first row contains the summary statistics on the whole network.
+{\bf Summary statistics on the results of multiple network methods on the PPIN.} The first row contains the summary statistics on the whole PPIN.
 
@@
 \label{stab:consensus_pwy}
-{\bf Pathway enrichment analyses of the genes in the consensus network.}
+{\bf Pathway enrichment analyses of the genes in the consensus solution.}
 
@@
 \label{sfig:pcs}
-{\bf GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)} Eight main principal components computed on the genotypes of GENESIS.
+{\bf GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)}~Eight main principal components computed on the genotypes of GENESIS.
 Cases are colored in green, controls in orange.
@@
 {\bf Association in GENESIS.
-The red line represents the Bonferroni threshold.} \textbf{(A)} SNP association, measured from the outcome of a 1 d.f.
-$\chi^2$ allelic test.
+The red line represents the Bonferroni threshold.} \textbf{(A)}~SNP association, measured from the outcome of a 1 d.f.
+$\chi^2$ allelic test (Section~\ref{methods:conventional}).
 Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
 The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
-\textbf{(B)} Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values.
+\textbf{(B)}~Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values (Section~\ref{methods:conventional}).
 The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
-\textbf{(C)} SNP association as in panel (A).
+\textbf{(C)}~SNP association as in panel (A).
 The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).
@@
 \label{sfig:pearson_methods}
-{\bf Pearson correlation between the different solution subnetworks.
-} \textbf{(A)} Correlation between selected SNPs.
-\textbf{(B)} Correlation between selected genes.
+{\bf Pearson correlation between the different solutions.
+} \textbf{(A)}~Correlation between selected SNPs.
+\textbf{(B)}~Correlation between selected genes.
 In general, the solutions display a very low overlap.
@@
 \label{sfig:consensus-names}
-{\bf Consensus subnetwork on GENESIS (Section~\ref{methods:methods}).
-} \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+{\bf Consensus solution on GENESIS (Section~\ref{methods:methods}).
+} \textbf{(A)}~Each node is represented by a pie chart, which shows the methods that selected it.
 We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
@@
 This panel is identical to Fig~\ref{fig:consensus}.
-\textbf{(B)} The same network, but every gene name is indicated.
+\textbf{(B)}~The same network, but every gene name is indicated.
+We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
 % \includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
@@
 \label{sfig:consensus_stats}
-{\bf Genes on the consensus network.
-Breast cancer susceptibility genes are colored in pink; the rest are colored in grey.
-} \textbf{(A)} Number of methods selecting every gene in the subnetwork.
-\textbf{(B)} VEGAS2 P\=/values of association of the genes, with regards to the number of methods that selected them.
-\textbf{(C)} Comparison of betweenness centrality of the genes in the consensus network and the other genes in the PPIN and not in the consensus network.
-To improve visualization, we removed outliers.
-\textbf{(D)} Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus network.
-The blue line represents a fitted generalized linear model.
+{\bf Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus solution.} The blue line represents a fitted generalized linear model.
 % \includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
@@
 \paragraph*{S6 Fig.}
+\label{sfig:additional-benchmarks}
+{\bf Additional benchmarks of the network methods} \textbf{(A)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
+For reference, we added a gray line with a slope of 1.
+This panel is identical to Fig~\ref{fig:consensus}.
+\textbf{(B)}~Sensitivity and specificity on the test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
+The performance of the classifier trained on all SNPs is also displayed.
+Points are the average over the 5 runs; the error bars represent the standard error of the mean.
+% \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf
+
+\paragraph*{S7 Fig.}
 \label{sfig:stable-consensus-names}
-{\bf Stable consensus subnetwork on GENESIS (Section~\ref{results:stable-consensus}).
-} \textbf{(A)} Each node is represented by a pie chart, which shows the methods that selected it.
+{\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}).
+} \textbf{(A)}~Each node is represented by a pie chart, which shows the methods that selected it.
 We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
@@
 This panel is identical to Fig~\ref{fig:stable-consensus}.
-\textbf{(B)} The same network, but every gene name is indicated.
-% \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf}
-
-\paragraph*{S7 Fig.}
+\textbf{(B)}~The same network, but every gene name is indicated.
+We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
+% \includegraphics[width=.675\linewidth]{./figures/sfigure_7.pdf}
+
+\paragraph*{S8 Fig.}
 \label{sfig:biotypes_excluded}
-{\bf Biotypes of genes from the annotation that are not present in the HINT protein-protein interaction network.}
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf}
-
-\paragraph*{S8 Fig.}
+{\bf Biotypes of genes from the annotation that are not present in the HINT PPIN.}
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf}
+
+\paragraph*{S9 Fig.}
 \label{sfig:lc_ht_comparison}
@@
 } Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
-\textbf{(A)} Ratios of the selected features between both benchmarks and of the active set.
-\textbf{(B)} Shifts in sensitivity and specificity.
-\textbf{(C)} Shift in Pearson correlation between benchmarks.
-\textbf{(D)} Ratio between the runtimes of the benchmarks.
-For gene network-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
+\textbf{(A)}~Ratios of the selected features between both benchmarks and of the active set (Section~\ref{methods:classifier}).
+\textbf{(B)}~Shifts in sensitivity and specificity.
+\textbf{(C)}~Shift in Pearson correlation between benchmarks.
+\textbf{(D)}~Ratio between the runtimes of the benchmarks.
+For gene-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
 In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.
-% \includegraphics[width=.85\linewidth]{./figures/sfigure_8.pdf}
-
-\paragraph*{S9 Fig.}
+% \includegraphics[width=.85\linewidth]{./figures/sfigure_9.pdf}
+
+\paragraph*{S10 Fig.}
 \label{sfig:scones_gsm}
-{\bf Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)} Manhattan plots of SNPs; in black, the method’s solution.
+{\bf Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)}~Manhattan plots of SNPs (Section \ref{methods:conventional}); in black, the method’s solution.
 The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
-\textbf{(B)} Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
+\textbf{(B)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
 For reference, we added a gray line with a slope of 1.
-\textbf{(C)} Solution networks.
-% \includegraphics[width=\linewidth]{./figures/sfigure_9.png}
+\textbf{(C)}~Solution networks.
+% \includegraphics[width=\linewidth]{./figures/sfigure_10.png}
 
 \section*{Acknowledgments}
-We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie), the biological resource center and all the GENESIS collaborating cancer clinics.
+
+We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S.
+Eon-Marchais, M.
+Marcou, D.
+Le Gal, L.
+Toulemonde, J.
+Beauvallet, N.
+Mebirouk, E.
+Cavaciuti), the biological resource centre (S.
+Mazoyer, F.
+Damiola, L.
+Barjhoux, C.
+Verny-Pierre, V.
+Sornin).
+We wish to pay a tribute to Olga M.
+Sinilnikova, who was one of the initiators and principal investigators of GENESIS and who died prematurely on June 30, 2014.
+
+We thank all the GENESIS collaborating cancer clinics clinics (Clinique Sainte Catherine, Avignon: H.
+Dreyfus; Hôpital Saint Jacques, Besançon: M-A.
+Collonge-Rame; Institut Bergonié, Bordeaux: M.Longy, A.
+Floquet, E.
+Barouk-Simonet; CHU, Brest: S.
+Audebert; Centre François Baclesse, Caen: P.
+Berthet; Hôpital Dieu, Chambéry: S.
+Fert-Ferrer; Centre Jean Perrin, Clermont-Ferrand: Y-J.
+Bignon; Hôpital Pasteur, Colmar: J-M.
+Limacher; Hôpital d’Enfants CHU – Centre Georges François Leclerc, Dijon: L.
+Faivre-Olivier; CHU, Fort de France: O.
+Bera; CHU Albert Michallon, Grenoble: D.
+Leroux; Hôpital Flaubert, Le Havre: V.
+Layet; Centre Oscar Lambret, Lille: P.
+Vennin, C.
+Adenis; Hôpital Jeanne de Flandre, Lille: S.
+Lejeune-Dumoulin, S.
+Manouvier-Hanu; CHRU Dupuytren, Limoges: L.
+Venat-Bouvet; Centre Léon Bérard, Lyon: C.
+Lasset, V.
+Bonadona; Hôpital Edouard Herriot, Lyon: S.
+Giraud; Institut Paoli-Calmettes, Marseille: F.
+Eisinger, L.
+Huiart; Centre Val d’Aurelle – Paul Lamarque, Montpellier: I.
+Coupier; CHU Arnaud de Villeneuve, Montpellier: I.
+Coupier, P.
+Pujol; Centre René Gauducheau, Nantes: C.
+Delnatte; Centre Catherine de Sienne, Nantes: A.
+Lortholary; Centre Antoine Lacassagne, Nice: M.
+Frénay, V.
+Mari; Hôpital Caremeau, Nîmes: J.
+Chiesa; Réseau Oncogénétique Poitou Charente, Niort: P.
+Gesta; Institut Curie, Paris: D.
+Stoppa-Lyonnet, M.
+Gauthier-Villars, B.
+Buecher, A.
+de Pauw, C.
+Abadie, M.
+Belotti; Hôpital Saint-Louis, Paris: O.
+Cohen-Haguenauer; Centre Viggo-Petersen, Paris: F.
+Cornélis; Hôpital Tenon, Paris: A.
+Fajac; GH Pitié Salpétrière et Hôpital Beaujon, Paris: C.
+Colas, F.
+Soubrier, P.
+Hammel, A.
+Fajac; Institut Jean Godinot, Reims: C.
+Penet, T.
+D.
+Nguyen; Polyclinique Courlancy, Reims: L.
+Demange*, C.
+Penet; Centre Eugène Marquis, Rennes: C.
+Dugast*; Centre Henri Becquerel, Rouen: A.
+Chevrier, T.
+Frebourg, J.
+Tinat, I.
+Tennevet, A.
+Rossi; Hôpital René Huguenin/Institut Curie, Saint Cloud: C.
+Noguès, L.
+Demange*, E.
+Mouret-Fourme; CHU, Saint-Etienne: F.
+Prieur; Centre Paul Strauss, Strasbourg: J-P.
+Fricker, H.
+Schuster; Hôpital Civil, Strasbourg: O.
+Caron, C.
+Maugard; Institut Claudius Regaud, Toulouse: L.
+Gladieff, V.
+Feillel; Hôpital Bretonneau, Tours: I.
+Mortemousque; Centre Alexis Vautrin, Vandoeuvre-les-Nancy: E.
+Luporsi; Hôpital de Bravois, Vandoeuvre-les-Nancy: P.
+Jonveaux; Gustave Roussy, Villejuif: A.
+Chompret*, O.
+Caron).
+*Deceased prematurely
 

### 2020-10-07 manuscript/plos/review_1.tex
added diff file

@@
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes’ P\=value into a score that is negative under weak association with the phenotype, and positive under a strong one.
+The authors propose a transformation of the genes' P\=value into a score that is negative under weak association with the phenotype, and positive under a strong one.
 This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
@@
 Summary statistics for SNPs and genes are available at \url{https://github.com/hclimente/genewa}.
-We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact \href{mailto:nadine.andrieu@curie.fr}{nadine.andrieu@curie.fr}.
+We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact \href{mailto:nadine.andrieu@curie.fr}{nadine.andrieu(at)curie.fr}.
 

### 2020-10-12 manuscript/plos/review_1.tex
fixed typos

@@
 Indeed, it allows to jointly consider the statistical association between genetic variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study six network methods which identify subnetworks with high association scores to a phenotype.
-Specifically, we examine their utility to discover new breast cancer susceptibility genes by interrogating a genome-wide association study (GWAS) focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
-We perform an in-depth benchmarking of the methods with regards to the size of the solution, their predictive power, and the stability and the runtime of the methods.
+In this work, we study six network methods which identify subnetworks with high scores of association with a phenotype.
+More specifically, we examine their use to discover new breast cancer susceptibility genes by interrogating a genome-wide association study (GWAS) data set focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
+We perform an in-depth benchmarking of these six methods with respect to the size of the solution, its predictive power, and the stability and the runtime of the methods.
 Network methods include prior knowledge in the analysis: they might tolerate selecting a gene with a high P-value, as long as it connects other two, low P-value genes.
-Thanks to this, most network methods give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+This alows them to give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
 For instance, at least two network methods selected 14 neighbors of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
@@
 Yet, network methods are notably unstable: the average Pearson correlation across all the methods between the genes selected on different subsamples of the data is 0.45, versus 0.57 simply using Bonferroni correction.
-To account for that, we produce a stable consensus solution, formed by the most consistently selected genes.
-The stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than the average gene.
-The network seems organized around \emph{CUL3}, encoding a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
-In conclusion, this article shows the pertinence of network-based analyses to tackle known issues with GWAS, namely a lack of statistical power and interpretable solutions.
+To compensate this instability, we produce a stable consensus solution, formed by the most consistently selected genes.
+This stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than the average gene.
+The network seems organized around \emph{CUL3}, which encodes a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
+In conclusion, this article shows the pertinence of network-based analyses to address known limitations of GWAS, namely their lack of statistical power and the difficulty of their interpretation.
 Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
@@
 In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \cite{bush_chapter_2012}.
-To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering the disease of interest, and the controls from the general population.
-Then, a per-SNP statistical association test is conducted (e.g.
-based on logistic regression).
+To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering from the disease of interest, and the controls, taken from the general population.
+Then, a statistical test of association (e.g.
+based on logistic regression) is conducted between each individual SNP and the phenotype.
 Those SNPs with a P\=/value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
@@
 On top of that, recent large-sampled studies suggest that numerous loci spread all along the genome contribute to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are very functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
-Hence, a comprehensive statistical framework which includes the structure of biological data might address the aforementioned issues.
+The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are strongly functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
+Hence, a comprehensive statistical framework which includes the structure of biological data might help alleviate the aforementioned issues.
 
@@
 Network-based discovery methods exploit the differential properties described above to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
-In essence, each gene has a score of association with the disease, given by the experiment, and functionally biological relationships, given by a network built on prior knowledge.
+In essence, each gene is assigned a score of association with the disease, computed from the GWAS data, and biological relationships, given by a network built on prior knowledge.
 Then, the problem becomes finding a functionally-related set of highly-scoring genes.
-Different solutions have been proposed to this problem, often stemming from divergent mathematical frameworks and considerations of what the optimal solution looks like.
-Some methods strongly constrain the problem to certain kinds of subnetworks.
-Such is the extreme case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e.
+Multiple solutions have been proposed to this problem, often stemming from different mathematical frameworks and considerations of what the optimal solution looks like.
+For example, some methods restrict the problem to specific types of subnetworks.
+Such is the case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e.
 instances were both a gene and its direct interactors are associated with the disease.
@@
 
-In this work, we analyze the effectiveness of these six network methods on GWAS data.
+In this work, we analyze the application of these six network methods on GWAS data.
 They do so by using different interpretations of the omnigenic model, providing a representative view of the field.
@@
 As none of the networks examined by LEAN was significant (Benjamini-Hochberg [BH] correction adjusted P\=/value $<$ 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+
 These solutions differ in many aspects, making it hard to draw joint conclusions.
 For starters, the overlap between the genes featured in each solution is quite small (Fig~\ref{fig:solution_overview}A).
-Despite that, the methods tend to agree on the genes with the strongest signal: genes selected by more methods tended to have lower P-value of association (Fig~\ref{fig:solution_overview}B).
-Another prominent difference is the solution size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contained only 4 genes.
-While SConES GI did not recover any protein coding gene, by dealing with SNP networks it retrieved four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
-Their topologies also differ (Fig~\ref{fig:solution_overview}C), as measured by the median centrality  and the number of connected components (Table~\ref{tab:gene_solutions}).
-Only two methods have more than one connected component: SConES, as described above, and HotNet2.
-HotNet2 produced 135 subnetworks, 115 of which have less than five genes.
+However, the methods tend to agree on the genes with the strongest signal: genes selected by more methods tended to have lower P-value of association (Fig~\ref{fig:solution_overview}B).
+
+Another prominent difference is the solution size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contains only 4 genes.
+While SConES GI did not recover any protein coding gene, working with SNP networks rather than gene networks allowed it to retriev four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
+
+The topologies of the six solutions differ as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality  and the number of connected components (Table~\ref{tab:gene_solutions}).
+Only two methods yield more than one connected component: SConES, as described above, and HotNet2.
+HotNet2 produced 135 subnetworks, 115 of which have fewer than five genes.
 The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
+
 Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences between the solutions of the different methods.
-It linked different parts of SigMod's solution network to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
+It linked different parts of SigMod's solution to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
 Interestingly, the dmGWAS solution (\nameref{stab:dmgwas_pwy}) is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-But, additionally, it includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+However, it additionally includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
 As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~$<$~0.04).
@@
 All obtained solutions have lower association P\=/values than the whole PPIN (median VEGAS2 P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Fig~\ref{fig:solution_overview}D).
-This exemplifies the trade-off between controlling for type I error and biological relevance.
-However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+This illustrates the trade-off between controlling for type I error and biological relevance.
+However, there are nuances between solutions in this regard: heinz strongly favors genes with lower P\=/values, while dmGWAS is less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tends to select whole LD-blocks; and HotNet2 and SigMod are less likely to select low scoring genes.
+
 Additionally, the solutions presented other desirable properties.
@@
 
-Topologically the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
+Topologically, the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
 Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
-In the former we find some genes that are highly central in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
+The large connected component contains genes that are highly central in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
 This suggests that these genes were selected because they were on the shortest path between another two highly associated genes.
@@
 
-We compared the results of different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
-Although BCAC case-control studies do not necessarily target cases with a family history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
+We compared the results obtained with different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
+Although BCAC case-control studies do not necessarily target cases with a familial history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
 This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
@@
 The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (\nameref{sfig:additional-benchmarks}A).
-interestingly, while scones gi at the snp-level achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+Interestingly, while SConES GI, at the SNP-level, achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
 
@@
 We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:benchmark}).
-Specifically, we measured five properties (Fig~\ref{fig:benchmark}): size of the solution; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
+This allowed us to measure five properties (Fig~\ref{fig:benchmark}): size of the solution; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
 The solution size varies greatly between the different methods (Fig~\ref{fig:benchmark}A).
@@
 LEAN did not produce any solution in any of the subsamples.
-To determine whether the selected snps could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (\nameref{sfig:additional-benchmarks}B).
+To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (\nameref{sfig:additional-benchmarks}B).
 The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
 Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-It should be considered that, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data \cite{deloscamposComplexTraitPredictionEra2018}.
+Of course, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data \cite{deloscamposComplexTraitPredictionEra2018}.
 
@@
 
-Another desirable quality of an algorithm is the stability of the solution with regards to small changes in the input (Section~\ref{methods:algorithm_comparison}).
-Heinz was high stabile in our benchmark, while the other methods displayed similarly low stabilities (Fig~\ref{fig:benchmark}C).
+Another desirable quality of a selection algorithm is the stability of the solution with respect to small changes in the input (Section~\ref{methods:algorithm_comparison}).
+Heinz was highly stable in our benchmark, while the other methods displayed similarly low stabilities (Fig~\ref{fig:benchmark}C).
 
@@
 Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
-Considering that, it took 5 days on average for HotNet2 to produce a result.
+Including this first step, it therefore took 5 days on average for HotNet2 to produce a result.
 
@@
 As shown above, and despite their similarities, the network methods produced remarkably different solutions.
-This is due to the fact that each of them models the optimal solution differently.
-Importantly, understanding which assumptions they make allows to understand the dataset more deeply.
+This is due to the fact that each of them models the problem differently.
+Importantly, understanding which assumptions they make allows to understand the results more in depth.
 For instance, the fact that LEAN did not return any gene implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
@@
 Another peculiarity of network methods is their relationship to degree centrality.
-We studied randomly rewirings of the PPIN while preserving node centrality (Section~\ref{methods:rewiring}).
+We studied random rewirings of the PPIN while preserving node centrality (Section~\ref{methods:rewiring}).
 In this setting, network methods favored central genes, as they often connect high scoring nodes (Fig~\ref{fig:issues}B).
 This is despite the fact that highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
-This was specially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
+This was especially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
 As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
@@
 By virtue of using a SNP subnetwork, SConES analyzes each SNP in their functional context.
-It therefore can select SNPs in genes without any associated interactor, as well as SNPs in non-coding regions or in non-interacting genes.
+It can therefore select SNPs located in genes not included in the PPIN, as well as SNPs in non-coding regions or in non-interacting genes.
 We compared the solution of SConES in the GI network (using PPIN information), to the one using only positional information (GS network), or positional and gene annotations (GM network).
@@
 
-\subsection{Adjusting for instability preserves global network properties}
+\subsection{Building a stable consensus network preserves global network properties}
 \label{results:stable-consensus}
@@
 Despite these similarities, the consensus and the stable consensus solutions include different genes.
-In this case, the most central gene is \emph{CUL3}, which is absent from the previous consensus solution and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
+In the stable consensus network, the most central gene is \emph{CUL3}, which is absent from the previous consensus solution and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
 This gene is a component of Cullin-RING ubiquitin ligases.
@@
 In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
-On the one hand, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
+First, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
 The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
-On the other hand, discovery in GWAS is hindered by a conservative statistical framework.
-Network methods tackle these two issues by using both the association score and an interaction network to take into consideration the biological context of each of the genes and SNPs.
+Second, discovery in GWAS is hindered by a conservative statistical framework.
+Network methods elegantly address these two issues by using both association scores and an interaction network to take into consideration the biological context of each of the genes and SNPs.
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided discovery have been proposed in recent years.
-In this article we evaluated the relevance of six of them by examining GENESIS, a GWAS dataset on familial breast cancer.
-
-DmGWAS, Heinz, HotNet2, SConES and SigMod relevant solutions, which recapitulate known breast cancer susceptibility genes.
+In this article we evaluated the relevance of six of these methods by applying them to the study of GENESIS, a GWAS dataset on familial breast cancer.
+
+DmGWAS, Heinz, HotNet2, SConES and SigMod all yield interesting solutions, which include (but are not limited to) known breast cancer susceptibility genes.
 In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
@@
 Yet, across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
-Despite these similarities, the solutions of the methods were notably different.
+Despite these similarities, the solutions obtained were notably different.
 At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, at the expense of not shedding much light on the etiology of the disease.
@@
 First, they are all equally bad at discriminating cases from controls.
-Yet, the classification accuracy of network methods is similar to that of a machine learning classifier, which suggests that cases and controls are difficult to separate in the GENESIS dataset.
-This might be due to unaccounted for environmental factors, and limited statistical power, which reduces the ability to identify relevant SNPs.
+Yet, the classification accuracy of network methods is similar to that of a classifier trained on the entire genome, which suggests that cases and controls are difficult to separate in the GENESIS dataset.
+This may be due to limited statistical power, which reduces the ability to identify relevant SNPs; but in any event, we do not expect to be able to separate people who have or will developed cancer from others on the sole basis of their genomes, ignoring all environmental factors and chance events.
 Second, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
@@
 Another source of instability might be the redundancy inherent to biological networks, since they are subject to an evolutionary pressure to avoid single points of failure \cite{wagnerAlternativeRoutesMutational2007}.
-Hence, often network methods will have multiple paths connecting two high-scoring nodes.
-
-To overcome the limitations of the individual methods while exploiting their strengths, we proposed combining them into a consensus solution.
+Hence, biological networks will often have multiple paths connecting two high-scoring nodes.
+
+To overcome these limitations while exploiting the strengths of the individual methods, we proposed combining them into a consensus solution.
 We use a straightforward strategy of including any node that was recovered by multiple methods.
-We proposed two networks: a consensus solution that tackled the heterogeneity of the solutions in the full dataset, and a stable consensus solution, that addressed the instability of the methods.
-They both synthesized the altered mechanisms: they both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer.
-Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but who are related to cancer at the expression level and whose neighborhood has consistent high association scores.
+We thus proposed two networks: a consensus solution, meant to address the heterogeneity of the solutions in the full dataset, and a stable consensus solution, which in addition addressed the instability of the methods.
+They both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer, thus synthetizing the mechanisms altered in breast cancer cases.
+Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but are related to cancer at the expression level and interact with genes with consistently high association scores.
 Importantly, while the consensus approach was as unstable as the individual network-guided methods, the stable consensus network retained the ability to provide compelling hypotheses and had better stability.
@@
 While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in explotatory settings.
-Nonetheless, BCAC is not an ideal counterfactual in which GENESIS had more samples.
+Nonetheless, BCAC is not an ideal ground truth.% in which GENESIS had more samples% .
 Yet, there are caveats which make it a non-optimal comparison.
@@
 Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
-Lastly GWAS are unlikely to identify all genes relevant for the disease: some might only show up in rare-variant studies; others might have too low effect sizes.
+Finally, and this is indeed the motivation for this study, GWAS are unlikely to identify all genes relevant for the disease: some might only show up in rare-variant studies; others might have too low effect sizes.
 Network methods account for this by including genes with low association scores but with relevant topological properties.
-Hence, network methods and GWAS are unlikely to capture exactly the same sets of genes.
+Hence, network methods and GWAS, even well-powered, are unlikely to capture exactly the same sets of genes.
 This might partly excuse the low precisions displayed in Section \ref{results:boost} and the low AUC displayed in Section \ref{results:benchmark}.
@@
 In consequence, they show their shortcomings on understudied genes, especially those not in the network.
-Out of the 32\,767 genes that we can map the genotyped SNPs to, 60.7\% (19\,887) are not in the PPIN.
+Out of the 32\,767 genes to which we can map the genotyped SNPs, 60.7\% (19\,887) are not in the PPIN.
 The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (\nameref{sfig:biotypes_excluded}).
@@
 On the other hand, Huang et al.
-\cite{huang_systematic_2018} found that network methods performed better on larger methods than in smaller networks at identifying disease genes.
+\cite{huang_systematic_2018} found that larger networks were more useful than smaller networks to identify disease genes.
 This would support using the largest networks in our experiments.
-When we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
+However, when we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
 This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best-known genes are artificially pushed into the solutions.
@@
 PEGASUS \cite{nakka_gene_2016}, SKAT \cite{ionita-laza_sequence_2013} or COMBAT \cite{wang_combat:_2017}) might have more statistical power.
-Another important decision is how to handle LD in a GWAS.
+
+How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners. % Another important decision is how to handle LD in a GWAS.
 VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
 With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
-However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem paramount to SConES' solutions, and an LD pruning step could potentially alter them.
-
-How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners.
+However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem to drive SConES' solutions, and an LD pruning step could potentially alter them.
+
 In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
@@
 Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the PPIN.
-But the opposite case is also true: when nearby genes are not functionally related (and hence disconnected in the PPIN), network methods might disregard them even if they have high association scores.
+But the opposite case is also true: when genes are not functionally related (and hence disconnected in the PPIN), network methods might disregard them even if they have high association scores.
 LD also affects SConES, since it penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Equation~\ref{eq:scones}.
-Due to linkage disequilibrium, nearby SNPs' P-values are be correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
+Due to linkage disequilibrium, nearby SNPs' P-values are correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
 Hence, SConES will tend to select LD-blocks formed by low P-value SNPs.
@@
 
-To produce the two consensus solutions, we faced the different interfaces, preprocessing steps, and unexpected behaviors of the various methods.
-To facilitate that other authors apply them to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
+To produce the two consensus solutions, we had to face practical challenges due to the differences in interfaces, preprocessing steps, and unexpected behaviors of the various methods.
+To make it easier for other to apply these methods to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
 They are available on GitHub: \url{https://github.com/hclimente/gwas-tools} (Section \ref{methods:data}).
@@
 Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
-Specifically, for each gene we only used the 10\% of SNPs mapped to it with lowest P\=/values.
-We mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
+More specifically, we mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
+For each gene, we computed VEGAS2 scores using only the 10\% of SNPs with lowest P\=/values among all those that were mapped to it.
 We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
@@
 
-\subsection{High-score subnetwork search algorithms}
+\subsection{Network methods}
 
@@
 
-In addition to weights, nodes have other properties, provided by the topology of the graph.
+Nodes can be described by properties provided by the topology of the graph.
 We focus on two of those: degree centrality, and betweenness centrality.
@@
 We also use two matrices that describe two different properties of a graph.
-The described matrices are square, and have as many rows and columns as nodes are in the network.
-The element $(i,j)$ represents a  selected relationship between $v_i$ and $v_j$.
-The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; the diagonal is zero.
+These matrices are square, and have as many rows and columns as nodes in the network.
+The element $(i,j)$ hence represents a relationship between $v_i$ and $v_j$.
+The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; its diagonal is zero.
 The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
@@
 \paragraph{Gene network}
-The statistical frameworks of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Yet, we used protein-protein interaction networks (PPIN) for all them except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
+The mathematicl formulations of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
+Here, we used protein-protein interaction networks (PPIN) for all them except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 
-Additionally, we compared the results of the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
+Additionally, we compared the results obtained on the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
 In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
@@
 
-\subsubsection{Network methods}
+\subsubsection{High-score subnetwork search algorithms}
 \label{methods:methods}
@@
 Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-But this is not the only aspect to model when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+Several aspects have to be taken into consideration when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
 Unsurprisingly, multiple solutions have been proposed.
@@
 \item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
-\item Add the neighboring nodes whose inclusion increases the Z\textsubscript{m+1} more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
+\item Add the neighboring nodes whose inclusion increases Z\textsubscript{m+1} by more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
 \item Repeat 2-4 until no further enlargement is possible.
-\item Add $S_i$ to the list of subnetworks to return. Its Z-score is normalized as
+\item Add $S_i$ to the list of subnetworks to return. Normalize its Z-score as
 
@@
 We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-We used the suggested hyperparameters $d = 2$ and $r = 0.1$.
+Unless otherwise specified, we used the suggested hyperparameters $d = 2$ and $r = 0.1$.
 We used the function \texttt{simpleChoose} to select the solution, which aggregates the top 1\% subnetworks.
@@
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P\=value into a score that is negative under weak association with the phenotype, and positive under a strong one.
+The authors propose a transformation of the genes' P\=/value into a score that is negative under weak association with the phenotype, and positive under a strong one.
 This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
@@
 
-As SConES, this optimization problem can also be solved by a graph min-cut approach.
+As for SConES, this optimization problem can also be solved by a graph min-cut approach.
 
 SigMod presents three important differences with SConES.
-First it is designed for PPINs.
+First, it is designed for PPINs.
 Second, it favors solutions containing many edges.
@@
 \begin{description}
-\item[{dmGWAS}] We essayed multiple values for $r$ (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and $d$ (1, 2, and 3). 
-\item[{heinz}] We essayed multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
-\item[{HotNet2}] We essayed different thresholds to decide which genes would receive a score of 0, and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
+\item[{dmGWAS}] We assayed multiple values for $r$ (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and $d$ (1, 2, and 3). 
+\item[{heinz}] We assayed multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
+\item[{HotNet2}] We assayed different thresholds to decide which genes would receive a score of 0, and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
 \item[{LEAN}] We used the following significance cutoffs for LEAN's P-values (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, and 1).
 \item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
-\item[{SigMod}] We used essayed multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
+\item[{SigMod}] We used multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
 \end{description}
@@
 
-In multiple steps of this article, we needed to compare the outcome of a methods that works on genes with the outcome of another methods that works on SNPs, or vice versa.
+In multiple steps of this article, we needed to compare the outcome of a method that works on genes with the outcome of another method that works on SNPs.
 For this purpose, we used the SNP-gene correspondance described in Section \ref{methods:conventional}.
@@
 A desirable solution offers good predictive power on unseen \emph{test} samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, a machine learning algorithm that searches a small subset of SNPs which provides good classification accuracy.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, which searches for a small subset of SNPs which provides good classification accuracy.
 We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
@@
 In this case, we did use the summary statistics of all 13\,250\,641 available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
-Since the these genotypes did not include chromosome X, we excluded that chromosome in this analysis.
-All comparisons included only the common genes between GENESIS and BCAC, so we used a less  stringent Bonferroni threshold ($1.66 \times 10^{-6}$) to call gene significance.
+Since the these genotypes did not include chromosome X, we excluded it from this analysis.
+All comparisons included only the genes common to GENESIS and BCAC, so we used the corresponding Bonferroni threshold ($1.66 \times 10^{-6}$) to call gene significance.
 
@@
 Rewiring the PPIN while preserving the number of edges of each gene allows to study the impact of the topology on the output of network methods.
+Indeed, the edges lose their biological meaning but the topology of the network is conserved.
 We produced 100 such rewirings by swapping edges in the PPIN.

### 2020-10-16 manuscript/plos/review_1.tex
added parameter space exploration and others

@@
 \section*{Abstract}
-Systems biology provides a comprehensive approach to uncovering the genetics of complex diseases and building hypotheses.
-Indeed, it allows to jointly consider the statistical association between genetic variation and a phenotype, and the biological context of each gene, represented as a network.
-In this work, we study six network methods which identify subnetworks with high scores of association with a phenotype.
-More specifically, we examine their use to discover new breast cancer susceptibility genes by interrogating a genome-wide association study (GWAS) data set focused on French women with a family history of breast cancer and tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
-We perform an in-depth benchmarking of these six methods with respect to the size of the solution, its predictive power, and the stability and the runtime of the methods.
-Network methods include prior knowledge in the analysis: they might tolerate selecting a gene with a high P-value, as long as it connects other two, low P-value genes.
-This alows them to give more compelling results than standard SNP- and gene-level analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-For instance, at least two network methods selected 14 neighbors of \emph{COPS5}, a gene related to multiple hallmarks of cancer.
-Importantly, we find significant overlaps between the genes in five of the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
-Yet, network methods are notably unstable: the average Pearson correlation across all the methods between the genes selected on different subsamples of the data is 0.45, versus 0.57 simply using Bonferroni correction.
-To compensate this instability, we produce a stable consensus solution, formed by the most consistently selected genes.
-This stable consensus is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than the average gene.
-The network seems organized around \emph{CUL3}, which encodes a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
-In conclusion, this article shows the pertinence of network-based analyses to address known limitations of GWAS, namely their lack of statistical power and the difficulty of their interpretation.
+Network methods provide a comprehensive approach to uncovering the genetics of complex diseases and building hypotheses.
+They discover susceptibility genes by jointly consider the statistical association between genetic variation and a phenotype, measured in a genome-wide association study (GWAS), and the biological context of each gene, represented as a network.
+Hence, a network method could select a gene with a high P-value of association if it connects multiple low P-value genes in the network.
+In this work, we studied six network methods which identify subnetworks with high scores of GWAS association with a phenotype.
+This allows them to give more compelling results than standard SNP- and gene-level GWAS analyses, recovering causal subnetworks tightly related to cancer susceptibility.
+We applied them to GENESIS, a GWAS on French women with a family history of breast cancer and who tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
+We critically compared these six methods, discussing the impact of their different mathematical frameworks, and the parameter choices.
+Additionally, we performed an in-depth benchmarking with respect to the size and predictive power of their solutions, as well as their stability and runtimes.
+Importantly, we found significant overlaps between the genes in five of the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
+Since most of the methods produced reasonable solutions, we proposed to combine them into a consensus solution, containing the genes selected by at least two methods.
+This aggregation brought further insights.
+For instance, it contained \emph{COPS5}, a gene related to multiple hallmarks of cancer, and 14 of its neighbors.
+The main drawback of network methods with regards to conventional $\chi^2$-based GWAS was instability i.e.
+network methods' outputs changed more in front of small perturbations in the input.
+To compensate for this instability, we proposed a stable consensus solution, formed by the most consistently selected genes in different subsamples of the data.
+This stable consensus was composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than most genes.
+The network was organized around \emph{CUL3}, which encodes a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
+In conclusion, we showed how network address limitations of GWAS, namely their lack of statistical power and the difficulty of their interpretation.
 Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
@@
 Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
-The omnigenic model offers an elegant explanation: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated so that many other genes can alter the function of the core genes.
+The omnigenic model offers an appealing explanation: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated so that many other genes can alter the function of the core genes.
 These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
@@
 However, Bonferroni correction is overly conservative when the statistical tests are correlated, as it is the case in GWAS \cite{wang_statistical_2018}.
-Another open issue is the interpretation of the results, as the functional consequences of most common variants are not well understood.
+Another open issue is the interpretation of the results, as the functional consequences of most common variants are unknown.
 On top of that, recent large-sampled studies suggest that numerous loci spread all along the genome contribute to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are strongly functionally inter-related and influence each other's behavior, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
+The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are strongly inter-related and influence each other's function, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
 Hence, a comprehensive statistical framework which includes the structure of biological data might help alleviate the aforementioned issues.
@@
 In this work, we analyze the application of these six network methods on GWAS data.
-They do so by using different interpretations of the omnigenic model, providing a representative view of the field.
+They use different interpretations of the omnigenic model, and provide a representative view of the field.
 We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
@@
   \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-} As LEAN did not produce any significant network (BH adjusted P\=/value~$<$~0.05), it was excluded.
+} As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it was excluded.
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
@@
 \textbf{(B)}~VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods.
-\textbf{(C)}~Manhattan plots of SNPs/genes; in black, the method's solution.
-The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{(D)}~Solution networks produced by the different methods.}
+\textbf{(C)}~Solution networks produced by the different methods.
+\textbf{(D)}~Manhattan plots of SNPs/genes; in black, the method's solution.
+The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
   \label{fig:solution_overview}
@@
   \centering
-  \includegraphics[width=.7\linewidth]{./figures/figure_3.pdf}
+  \includegraphics[width=.8\linewidth]{./figures/figure_3.pdf}
   \caption{ {\bf Consensus solution on GENESIS (Section~\ref{methods:methods}) }.
-Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
-The latter ones are also colored in pink.
-All gene names are indicated in \nameref{sfig:consensus-names}.}
+\textbf{A} ~Manhattan plots of genes; in black, the ones in the consensus solution.
+The Bonferroni threshold is indicated by a red line (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
+\textbf{B} Consensus network.
+Each gene is represented by a pie chart, which shows the methods that selected it.
+We enlarged the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
+\textbf{C} The nodes are in the same disposition as in panel B, but every gene name is indicated.
+We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset}
   \label{fig:consensus}
@@
 LEAN did not produce any solution in any of the subsamples.
+Using different combinations of parameters (Section \ref{methods:parameters}), we computed how good each of the methods was at recovering the results of a conventional GWAS on BCAC (Section \ref{methods:bcac}, Fig~\ref{fig:benchmark}B).
+SConES exhibits the largest area under the curve, since, when $\lambda = 0$ (network topology is disregarded), it is equivalent to a Bonferroni correction.
+Between the remaining network methods, they have similar areas under the curve, with heinz being the one with the largest area.
+
 To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (\nameref{sfig:additional-benchmarks}B).
@@
   \centering
-  \includegraphics[width=\linewidth]{./figures/figure_4.pdf}
+  \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
   \caption{{ \bf Comparison of network-based GWAS methods on GENESIS.
@@
 For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs.
+\textbf{(B)} True positive rate and true negative rate, using significant SNPs (for SConES and $\chi^2$ + Bonferroni) and genes (for the remaining methods) in BCAC (Section~\ref{methods:bcac}) as true positives, of multiple parameter combinations for different methods (Section \ref{results:parameters}).
 \textbf{(C)}~Pairwise Pearson correlations of the solutions produced by different methods.
@@
   \centering
-  \includegraphics[width=\linewidth]{./figures/figure_2.pdf}
+  \includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
   \caption{{\bf Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
@@
 This is despite the fact that highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
-This was especially the case of SigMod, which selected three highly central, unassociated genes: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
+This was especially the case of SigMod, which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
 As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
@@
 
+\subsection{Different parameters produce similarly-sized solutions}
+\label{results:parameters}
+
+We explored the parameter space of the different methods by running them under different combinations of parameters (Section \ref{methods:parameters}).
+In agreement with their formulations (Section \ref{methods:methods}), larger values of certain parameters produce less astringent solutions (\nameref{sfig:parameters}A): for HotNet2 and heinz, we examined the threshold to decide each gene has a positive score or zero; for dmGWAS, the \texttt{d} parameter controls how far neighbors could be added; node involved how far the search from the explored subnetwork; for SigMod \texttt{nmax} specifies the maximum size of the solution; and for LEAN, it was the P-value threshold to consider a solution significant.
+Two parameters had the opposite effect (larger is more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES $\eta$, where higher values produce more stringent solutions.
+However, two of the parameters did not have the expected effect: dmGWAS' \texttt{r}, which controls the minimum increment in the score required to add an additional gene; and SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''.
+In both cases, the size of the solution was very similar across the different values.
+Despite the differences in size, the size of the solutions was relatively robust to the choice of parameters (\nameref{sfig:parameters}B).
+
+We computed the Pearson correlation between the different solutions as in Section \ref{methods:algorithm_comparison} to study how the parameters affected which genes and SNPs were selected (\nameref{sfig:parameters}C).
+This showed that dmGWAS and SigMod are robust to certain parameters: dmGWAS' output was mostly determined by the parameter $d$, rather than $r$; SigMod's by $nmax$, rather than $maxjump$.
+
+SConES presented an interesting case in terms of feature selection: most of the explored combinations of parameters led to trivial solutions (either all the SNPs, or none of them were included) (\nameref{sfig:parameters}A).
+To explore a more meaningful parameter space, we selected the parameters in two rounds.
+First, we explored the whole sample space.
+Then, we focused in a range of $\eta$ and $\lambda$ 1.5 orders of magnitude above and below the best parameters, respectively.
+This second parameter space was more diverse, and allowed to find more interesting solutions.
+
 \subsection{Building a stable consensus network preserves global network properties}
@@
 Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
-We built a new, stable consensus solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
+We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
 Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
@@
   \centering
-  \includegraphics[width=.7\linewidth]{./figures/figure_5.pdf}
-  \caption{{\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}).} Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
-The latter ones are also colored in pink.
-All gene names are indicated in \nameref{sfig:stable-consensus-names}.}
+  \includegraphics[width=.8\linewidth]{./figures/figure_5.pdf}
+  \caption{ {\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}) }.
+\textbf{A} ~Manhattan plots of genes; in black, the ones in the stable consensus solution.
+The Bonferroni threshold is indicated by a red line (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
+\textbf{B} Stable consensus network.
+Each gene is represented by a pie chart, which shows the methods that selected it.
+We enlarged the most central gene (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
+\textbf{C} The nodes are in the same disposition as in panel B, but every gene name is indicated.
+We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset}
   \label{fig:stable-consensus}
@@
 On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While this deepens our understanding of the disease and provides biological hypotheses, they require further analyses, which might deter unexperienced practitioners.
-For instance, examining the network properties of the genes in the solution (e.g.
-centrality and neighbors).
-Those might drive the selection of genes, especially of those with low association scores.
+While this deepens our understanding of the disease and provides biological hypotheses, they require further analyses.
+For instance, a user might need to examine the centrality of the selected genes, and discern the extent to which that property is driving the selection of each gene.
 HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
@@
 Network methods show modest precision, but much higher recall at recovering BCAC biomarkers (Section \ref{results:boost}).
-While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in explotatory settings.
+While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in exploratory settings.
 Nonetheless, BCAC is not an ideal ground truth.% in which GENESIS had more samples% .
@@
 In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms (e.g.
-PEGASUS \cite{nakka_gene_2016}, SKAT \cite{ionita-laza_sequence_2013} or COMBAT \cite{wang_combat:_2017}) might have more statistical power.
+However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms like PEGASUS \cite{nakka_gene_2016}, SKAT \cite{ionita-laza_sequence_2013} or COMBAT \cite{wang_combat:_2017} might have more statistical power.
 
@@
 However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem to drive SConES' solutions, and an LD pruning step could potentially alter them.
-
-In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in linkage disequilibrium.
+In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in LD.
 In fact, the presented case is paradigmatic, since all three genes are located in the HLA region, the most gene-dense region of the genome \cite{Xie2003}.
@@
 LD also affects SConES, since it penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Equation~\ref{eq:scones}.
-Due to linkage disequilibrium, nearby SNPs' P-values are correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
+Due to LD, nearby SNPs' P-values are correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
 Hence, SConES will tend to select LD-blocks formed by low P-value SNPs.
-This might explain why SConES produces similar results on the GS and GM networks, heavily affected by linkage disequilibrium (Section \ref{results:drawbacks}).
+This might explain why SConES produces similar results on the GS and GM networks, heavily affected by LD (Section \ref{results:drawbacks}).
 However, this same behavior raises the burden of proof required to select SNPs with many interactions, like those mapped to hub genes in the PPIN.
@@
 In order to obtain significant SNPs, we performed Bonferroni correction, to keep family-wise error rate below 5\%.
-The threshold used was [$]\frac {0.05} {197083} = 2.54 \times 10^{-7}[/$].
+The threshold used was $\frac {0.05} {197083} = 2.54 \times 10^{-7}$.
 
@@
 Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
-We also used Bonferroni correction to obtain significant genes; in this case, the threshold of signigicance was [$]\frac {0.05} {32767} = 1.53 \times 10^{-6}[/$].
+We also used Bonferroni correction to obtain significant genes; in this case, the threshold of signigicance was $\frac {0.05} {32767} = 1.53 \times 10^{-6}$.
 
@@
 \paragraph{Gene network}
-The mathematicl formulations of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Here, we used protein-protein interaction networks (PPIN) for all them except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
+The mathematical formulations of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
+Here, we used protein-protein interaction networks (PPIN) for all of them except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
-However, at different stages we also used GS and GM for comparison.
+However, at different stages we also compared GI to GS and GM to understand how considering the PPIN affects SConES' output.
 For the GM network, we used the mapping described in Section~\ref{methods:snp2gene}.
-In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2}).
+In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2} statistic).
 The properties of these three subnetworks are available in \nameref{stab:snp_solutions}.
@@
 Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
+Additionally, that change needs to be larger than a user specified threshold \texttt{maxjump}.
 Such a large change implies that the network is densely interconnected.
@@
 A score is assigned to each candidate solution by summing their node scores and normalizing by size.
-The candidate solution with the highest standardized score is the chosen solution.
+The candidate solution with the highest standardized score and that is not larger than a user-specified threshold (\texttt{nmax}) is the chosen solution.
 SigMod is implemented in an R package \cite{sigmod}.
@@
 \begin{description}
-\item[{dmGWAS}] We assayed multiple values for $r$ (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and $d$ (1, 2, and 3). 
-\item[{heinz}] We assayed multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
-\item[{HotNet2}] We assayed different thresholds to decide which genes would receive a score of 0, and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
+\item[{dmGWAS}] We tested multiple values for $r$ (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and $d$ (1, 2, and 3). 
+\item[{heinz}] We tested multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
+\item[{HotNet2}] We tested different thresholds to decide which genes would receive a score of 0, and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
 \item[{LEAN}] We used the following significance cutoffs for LEAN's P-values (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, and 1).
 \item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
-\item[{SigMod}] We used multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
+\item[{SigMod}] We tested multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
 \end{description}
@@
 In multiple steps of this article, we needed to compare the outcome of a method that works on genes with the outcome of another method that works on SNPs.
-For this purpose, we used the SNP-gene correspondance described in Section \ref{methods:conventional}.
+For this purpose, we used the SNP-gene correspondence described in Section \ref{methods:conventional}.
 To convert a list of SNPs into a list of genes, we included all the genes that can be mapped to any of the SNPs.
@@
 This penalty was set by cross-validation, choosing the value that minimized misclassification error.
-We applied each network method to each \emph{train} set, and trained the classifier on it as well using only on the selected SNPs.
+We applied each network method to each \emph{train} set, and trained the classifier on the same train set using only the selected SNPs.
 When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
@@
 We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
-Indeed, the lifetime incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
+Indeed, the lifetime cumulative incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
 However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
@@
 In this case, we did use the summary statistics of all 13\,250\,641 available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
-Since the these genotypes did not include chromosome X, we excluded it from this analysis.
+Since these genotypes did not include chromosome X, we excluded it from this analysis.
 All comparisons included only the genes common to GENESIS and BCAC, so we used the corresponding Bonferroni threshold ($1.66 \times 10^{-6}$) to call gene significance.
@@
 We only applied only four methods on the rewirings: heinz, dmGWAS, LEAN and SigMod.
-We excluded HotNet2, since it takes notably longer to run than the other methods, while taking up more computational resources.
+We excluded HotNet2 and SConES, since they take notably longer to run than the other methods, while taking up more computational resources.
 
@@
 \paragraph*{S4 Fig.}
-\label{sfig:consensus-names}
-{\bf Consensus solution on GENESIS (Section~\ref{methods:methods}).
-} \textbf{(A)}~Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-This panel is identical to Fig~\ref{fig:consensus}.
-\textbf{(B)}~The same network, but every gene name is indicated.
-We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-% \includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf}
-
-\paragraph*{S5 Fig.}
 \label{sfig:consensus_stats}
@@
 
-\paragraph*{S6 Fig.}
+\paragraph*{S5 Fig.}
 \label{sfig:additional-benchmarks}
@@
 
+\paragraph*{S6 Fig.}
+\label{sfig:parameters}
+{\bf Parameter space of the network methods.} \textbf{(A)} Boxplot of the solution sizes of the methods under the explored parameters (Section \ref{methods:parameters}).
+\textbf{(B)} Size of SConES's with regards to each pair of parameters.
+\textbf{(C)} Pearson correlation between the solutions of the different runs.
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
+
 \paragraph*{S7 Fig.}
-\label{sfig:stable-consensus-names}
-{\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}).
-} \textbf{(A)}~Each node is represented by a pie chart, which shows the methods that selected it.
-We labeled (and enlarged) the two most central genes (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-The latter ones are also colored in pink.
-This panel is identical to Fig~\ref{fig:stable-consensus}.
-\textbf{(B)}~The same network, but every gene name is indicated.
-We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.
-% \includegraphics[width=.675\linewidth]{./figures/sfigure_7.pdf}
-
-\paragraph*{S8 Fig.}
 \label{sfig:biotypes_excluded}
@@
 
-\paragraph*{S9 Fig.}
+\paragraph*{S8 Fig.}
 \label{sfig:lc_ht_comparison}
@@
 
-\paragraph*{S10 Fig.}
+\paragraph*{S9 Fig.}
 \label{sfig:scones_gsm}

### 2020-10-16 manuscript/plos/review_1.tex
shorter abstract

@@
 \section*{Abstract}
-Network methods provide a comprehensive approach to uncovering the genetics of complex diseases and building hypotheses.
-They discover susceptibility genes by jointly consider the statistical association between genetic variation and a phenotype, measured in a genome-wide association study (GWAS), and the biological context of each gene, represented as a network.
-Hence, a network method could select a gene with a high P-value of association if it connects multiple low P-value genes in the network.
+% Network methods provide a comprehensive approach to uncovering the genetics of complex diseases and building hypotheses.
+% They discover susceptibility genes by jointly consider the statistical association between genetic variation and a phenotype, measured in a genome-wide association study (GWAS), and the biological context of each gene, represented as a network.
+Network approaches to disease use molecular networks, which model functional relationships between the molecules in a cell, to generate hypotheses about the genetics of complex diseases.
+Among them, several approaches jointly consider gene scores, representing the association between each gene and the disease, and the biological context of each gene, modeled by a network.
+%These approaches can select low-scoring genes if they connect multiple high-scoring genes.
+Here, we obtain gene scores from genome-wide association studies (GWAS) and study six network approaches on 
+% Hence, a network method could select a gene with a high P-value of association if it connects multiple low P-value genes in the network.
 In this work, we studied six network methods which identify subnetworks with high scores of GWAS association with a phenotype.
 This allows them to give more compelling results than standard SNP- and gene-level GWAS analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-We applied them to GENESIS, a GWAS on French women with a family history of breast cancer and who tested negative for pathogenic variants in \emph{BRCA1} and \emph{BRCA2}.
-We critically compared these six methods, discussing the impact of their different mathematical frameworks, and the parameter choices.
-Additionally, we performed an in-depth benchmarking with respect to the size and predictive power of their solutions, as well as their stability and runtimes.
-Importantly, we found significant overlaps between the genes in five of the solution networks and the genes significantly associated in the largest GWAS on susceptibility to breast cancer.
-Since most of the methods produced reasonable solutions, we proposed to combine them into a consensus solution, containing the genes selected by at least two methods.
-This aggregation brought further insights.
-For instance, it contained \emph{COPS5}, a gene related to multiple hallmarks of cancer, and 14 of its neighbors.
-The main drawback of network methods with regards to conventional $\chi^2$-based GWAS was instability i.e.
-network methods' outputs changed more in front of small perturbations in the input.
-To compensate for this instability, we proposed a stable consensus solution, formed by the most consistently selected genes in different subsamples of the data.
-This stable consensus was composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, Fisher's exact test P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than most genes.
-The network was organized around \emph{CUL3}, which encodes a ubiquitin ligase related protein that regulates the protein levels of several genes involved in cancer progression.
-In conclusion, we showed how network address limitations of GWAS, namely their lack of statistical power and the difficulty of their interpretation.
-Project-agnostic implementations of each of the network methods are available at \url{https://github.com/hclimente/gwas-tools} to facilitate their application to other GWAS datasets.
+% We apply them to
+GENESIS, a GWAS on French women with non-BRCA familial breast cancer% and no pathogenic variants in \emph{BRCA1} and \emph{BRCA2}
+.
+We provide a critical comparison of these six methods, discussing the impact of their mathematical formulation and of their parameters.
+% on % .
+Additionally, we performed an in-depth benchmarking with respect to
+% the size and predictive power of their solutions, as well as on their stability and runtimes.
+% Importantly,
+Using a molecular network yields more compelling results than standard % SNP- and gene-level
+GWAS analyses. Indeed, %recovering causal subnetworks tightly related to cancer susceptibility.
+we find significant overlaps between our solutions and the genes identified in the largest GWAS on breast cancer susceptibility. % Since most of the methods produce reasonable solutions, we
+We further
+propose to combine these solutions into a consensus network, %containing the genes selected by at least two methods. This aggregation
+which
+brings further insights.
+The consensus network contains \emph{COPS5}, a gene related to multiple hallmarks of cancer, and 14 of its neighbors.
+The main drawback of network methods % with respect to conventional $\chi^2$-based GWAS
+is that they are not robust to small perturbations in their inputs. %To compensate for this instability, we
+We therefore propose a stable consensus solution, formed by the most consistently selected genes in multiple subsamples of the data.
+In GENESIS, it is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than most genes.
+The network is organized around \emph{CUL3}, which is involved in the regulation of several genes linked to cancer progression.
+In conclusion, we showed how network methods help overcome the lack of statistical power of GWAS and improve their interpretation.
+Project-agnostic implementations of all methods are available at \url{https://github.com/hclimente/gwas-tools}.
+% to facilitate their application to other GWAS datasets.
 

### 2020-10-16 manuscript/plos/review_1.tex
addressed comments from fabienne and christine

--- 2020-10-16 manuscript/plos/review_1.tex
@@
 Hence, a network method could select a gene with a high P-value of association if it connects multiple low P-value genes in the network.
-In this work, we studied six network methods which identify subnetworks with high scores of GWAS association with a phenotype.
+In this work, we evaluate six network methods which identify subnetworks with high scores of GWAS association with a phenotype.
 This allows them to give more compelling results than standard SNP- and gene-level GWAS analyses, recovering causal subnetworks tightly related to cancer susceptibility.
@@
 These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
-In this study, we characterize six such network methods on GENESIS, a GWAS dataset for familial breast cancer in the French population.
+In this study, we apply and compare six such network methods on GENESIS, a GWAS dataset for familial breast cancer in the French population.
 Combining these approaches allows us to identify potentially novel breast cancer susceptibility genes, and provides a mechanistic explanation for their role in the development of the disease.
@@
 We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
-After a classical GWAS approach, we use these network methods to recover additional breast cancer susceptibility genes.
+After a classical GWAS approach, we use these network methods to identify additional breast cancer susceptibility genes.
 Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus solution of predisposition to familial breast cancer.
@@
 
-\subsection{Conventional SNP- and gene-based analyses confirm that \emph{FGFR2} locus is associated with familial breast cancer}
+\subsection{Conventional SNP- and gene-based analyses retrieve the \emph{FGFR2} locus in the GENESIS dataset}
 \label{results:conventional}
@@
 Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression using all the GENESIS genotypes as input, and the phenotype (cancer/healthy) as outcome (Section~\ref{methods:classifier}).
+We applied an L1-penalized logistic regression using all GENESIS genotypes as input, and the phenotype (cancer/healthy) as outcome (Section~\ref{methods:classifier}).
 The algorithm selected 100 SNPs, both from all aforementioned regions and new ones (\nameref{sfig:snp_gene_manhattan}C).
@@
 \emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS or BCAC (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus solution have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
+Despite its lack of association in GENESIS or in studies conducted by the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus solution have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
 
@@
 
-We compared the results obtained with different network methods to the European sample of the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017}, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
+We compared the results obtained with different network methods to the European sample of BCAC, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
 Although BCAC case-control studies do not necessarily target cases with a familial history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
@@
 
-Across this article, we have compared our results to significant genes and SNPs in the BCAC study \cite{Michailidou2017}.
-Network methods show modest precision, but much higher recall at recovering BCAC biomarkers (Section \ref{results:boost}).
+In this work, we have compared our results to significant genes and SNPs in the BCAC study \cite{Michailidou2017}.
+Network methods show modest precision, but much higher recall at recovering BCAC hits (Section \ref{results:boost}).
 While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in exploratory settings.

### 2020-10-16 manuscript/plos/review_1.tex
minor typos fixed

@@
 We explored the parameter space of the different methods by running them under different combinations of parameters (Section \ref{methods:parameters}).
-In agreement with their formulations (Section \ref{methods:methods}), larger values of certain parameters produce less astringent solutions (\nameref{sfig:parameters}A): for HotNet2 and heinz, we examined the threshold to decide each gene has a positive score or zero; for dmGWAS, the \texttt{d} parameter controls how far neighbors could be added; node involved how far the search from the explored subnetwork; for SigMod \texttt{nmax} specifies the maximum size of the solution; and for LEAN, it was the P-value threshold to consider a solution significant.
-Two parameters had the opposite effect (larger is more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES $\eta$, where higher values produce more stringent solutions.
+In agreement with their formulations (Section \ref{methods:methods}), larger values of certain parameters produce less stringent solutions (\nameref{sfig:parameters}A): for HotNet2 and heinz, this is the threshold above which genes are assigned a non-zero score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; % node involved how far the search from the explored subnetwork;
+for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
+Two parameters had the opposite effect (larger is more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES's $\eta$, where higher values produce smaller solutions.
 However, two of the parameters did not have the expected effect: dmGWAS' \texttt{r}, which controls the minimum increment in the score required to add an additional gene; and SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''.

### 2020-10-16 manuscript/plos/review_1.tex
address more comments

--- 2020-10-16 manuscript/plos/review_1.tex
@@
 Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
-\textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors.
+\textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors (Sections \ref{methods:algorithm_comparison} and \ref{methods:snp2gene}).
 \textbf{(B)}~VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods.
@@
 
-Most of the network methods, including the consensus, were highly unstable, raising questions about the reliability of the results.
+Most of the network methods, including the consensus, were highly unstable (Fig~\ref{fig:benchmark}), raising questions about the reliability of the results.
 We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
@@
 
+As not all databases compile the same interactions, the choice of the PPIN determines the final output.
+In this work we used exclusively interactions from HINT from high-throughput experiments.
+This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
+On the other hand, Huang et al.
+\cite{huang_systematic_2018} found that larger networks were more useful than smaller networks to identify disease genes.
+This would support using the largest networks in our experiments.
+However, when we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
+This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best-known genes are artificially pushed into the solutions.
+
 The strength of network-based analyses comes from leveraging prior knowledge to boost discovery.
@@
 eQTLs, SNPs associated to the expression of a gene) might lead to different results.
-
-As not all databases compile the same interactions, the choice of the PPIN determines the final output.
-In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
-On the other hand, Huang et al.
-\cite{huang_systematic_2018} found that larger networks were more useful than smaller networks to identify disease genes.
-This would support using the largest networks in our experiments.
-However, when we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
-This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best-known genes are artificially pushed into the solutions.
 

### 2020-10-17 manuscript/plos/review_1.tex
changed to past tense

@@
 Chloé-Agathe Azencott\textsuperscript{3,1,2},
-with the GENESIS study group\textsuperscript{\textpilcrow}
+GENESIS study group\textsuperscript{\textpilcrow}
 \\
@@
 % They discover susceptibility genes by jointly consider the statistical association between genetic variation and a phenotype, measured in a genome-wide association study (GWAS), and the biological context of each gene, represented as a network.
-Network approaches to disease use molecular networks, which model functional relationships between the molecules in a cell, to generate hypotheses about the genetics of complex diseases.
+Network approaches to disease use biological networks, which model functional relationships between the molecules in a cell, to generate hypotheses about the genetics of complex diseases.
 Among them, several approaches jointly consider gene scores, representing the association between each gene and the disease, and the biological context of each gene, modeled by a network.
@@
 % Importantly,
-Using a molecular network yields more compelling results than standard % SNP- and gene-level
+Using a biological network yields more compelling results than standard % SNP- and gene-level
 GWAS analyses. Indeed, %recovering causal subnetworks tightly related to cancer susceptibility.
@@
 Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
-The omnigenic model offers an appealing explanation: only a limited number of core genes are directly involved in the disease; but gene functions are deeply interrelated so that many other genes can alter the function of the core genes.
+The omnigenic model offers an appealing explanation: only a limited number of \emph{core} genes are directly involved in the disease; but gene functions are deeply interrelated, and so many other genes can alter the function of the core genes.
 These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
@@
 As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them.
-These relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expression is correlated, or their products establish a protein-protein interaction.
+These relationships come from evidence that the genes jointly contribute to a biological function; for instance, their expressions are correlated, or their products establish a protein-protein interaction.
 Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}.
-In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,pinero_uncovering_2016}: they tend to occupy central positions in the network, interconecting different modules TODO.
+In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,cai_broker_2010}: they tend to occupy central positions in the network (although not the most central ones); disease genes for the same pathology tend to be connected in modules; and often they are bottlenecks that interconect modules.
 Therefore, studying the neighborhood of disease-associated genes is effective at identifying new ones that are involved in the same biological functions \cite{huang_systematic_2018}.
@@
 
-In this work, we analyze the application of these six network methods on GWAS data.
+In this work, we studied the relevance of network-based approaches to genetics by applying these six network methods to GWAS data.
 They use different interpretations of the omnigenic model, and provide a representative view of the field.
 We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
-After a classical GWAS approach, we use these network methods to identify additional breast cancer susceptibility genes.
-Lastly, we carry out a comparison of the solutions obtained by the different methods, and aggregate them to obtain a consensus solution of predisposition to familial breast cancer.
+After a classical GWAS approach, we used these network methods to identify additional breast cancer susceptibility genes.
+Lastly, we carried out a comparison of the solutions obtained by the different methods, and aggregated them to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
 
@@
 Two genomic regions had a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
-The former overlaps with gene \emph{FGFR2}; the latter with \emph{CASC16} the protein-coding gene \emph{TOX3}.
+The former overlaps with the gene \emph{FGFR2}; the latter with \emph{CASC16} and the protein-coding gene \emph{TOX3}.
 Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
@@
 We applied six network methods to the GENESIS dataset (Section~\ref{methods:methods}).
-As none of the networks examined by LEAN was significant (Benjamini-Hochberg [BH] correction adjusted P\=/value $<$ 0.05), we obtained six solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, one for SConES GI (which works at the SNP level), and the consensus.
+As none of the networks examined by LEAN was significant (Benjamini-Hochberg [BH] correction adjusted P\=/value $<$ 0.05), we obtained five solutions (Fig~\ref{fig:solution_overview}): one for each of the remaining four gene-based methods, and one for SConES GI (which works at the SNP level).
+
+\begin{figure}[!ht]
+  \centering
+  \includegraphics[width=\textwidth]{./figures/figure_1.png}
+  \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
+} As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it was excluded.
+Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
+\textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors (Sections \ref{methods:algorithm_comparison} and \ref{methods:snp2gene}).
+\textbf{(B)}~Distribution of VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods.
+\textbf{(C)}~Solution networks produced by the different methods.
+\textbf{(D)}~Manhattan plots of SNPs/genes; in black, the method's solution.
+The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
+  \label{fig:solution_overview}
+\end{figure}
 
@@
 
-The topologies of the six solutions differ as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality  and the number of connected components (Table~\ref{tab:gene_solutions}).
-Only two methods yield more than one connected component: SConES, as described above, and HotNet2.
+The topologies of the five solutions differ as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality  and the number of connected components (Table~\ref{tab:gene_solutions}).
+Three methods yield more than one connected component: SConES, as described above, SigMod, and HotNet2.
 HotNet2 produced 135 subnetworks, 115 of which have fewer than five genes.
-The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM}.
-
-Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences between the solutions of the different methods.
-It linked different parts of SigMod's solution to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
-Interestingly, the dmGWAS solution (\nameref{stab:dmgwas_pwy}) is also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-However, it additionally includes submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
-As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) are involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~$<$~0.04).
+The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM} (Section \ref{methods:bcac}).
 
@@
 {\bf \# components }: number of connected components.
-\textbf{\mean{Betweenness}}: mean betweenness of the selected genes in the PPIN.
+\textbf{\mean{Betweenness}}: median betweenness of the selected genes in the PPIN.
 \textbf{\median{P}\textsubscript{gene}}: median VEGAS2 P\=/value of the selected genes.
-\textbf{\# genes in consensus}: Number of genes in common between the method's solution, and the 93 genes in the consensus solution.
+\textbf{\# genes in consensus}: number of genes in common between the method's solution, and the 93 genes in the consensus solution.
   \end{flushleft}
@@
 
+Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences between the solutions of the different methods.
+It linked different parts of SigMod's solution to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
+Interestingly, the dmGWAS solution (\nameref{stab:dmgwas_pwy}) was also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
+However, it additionally included submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) were involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~$<$~0.04).
+
+Despite their differences, there were additional common themes.
+All obtained solutions had lower association P\=/values than the whole PPIN (median VEGAS2 P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Fig~\ref{fig:solution_overview}D).
+This illustrates the trade-off between controlling for type I error and biological relevance.
+However, there were nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+
+Additionally, the solutions presented other desirable properties.
+First, four of them were enriched in known breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value$<$0.03).
+Second, the genes in three solutions displayed on average a significantly higher betweenness centrality than the rest of the genes (dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value $<$ 1.4 \texttimes{} 10\textsuperscript{-21}).
+This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
+Interestingly, SConES selected SNPs that are also more central than the average SNP (\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
+
+\subsection{A case study: the consensus solution}
+\label{results:consensus}
+
+Despite their shared properties, the differences between the solutions of the different methods suggested that each of them captured different aspects of cancer susceptibility.
+Indeed, out of the 668 genes that were selected by at least one method, only 93 were selected by at least two, 20 by three, and none by four or more.
+Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig \ref{fig:solution_overview}B), a relationship that plateaus at 2.
+Hence, to leverage on their strengths and compensate their respective weaknesses, we built a consensus solution that captured the genes shared among at least two solutions (Section~\ref{methods:methods}).
+This solution (Fig~\ref{fig:consensus}) contained 93 genes and exhibited the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
+
 \begin{figure}[!ht]
   \centering
-  \includegraphics[width=\textwidth]{./figures/figure_1.png}
-  \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-} As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it was excluded.
-Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
-\textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors (Sections \ref{methods:algorithm_comparison} and \ref{methods:snp2gene}).
-\textbf{(B)}~VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods.
-\textbf{(C)}~Solution networks produced by the different methods.
-\textbf{(D)}~Manhattan plots of SNPs/genes; in black, the method's solution.
-The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
-  \label{fig:solution_overview}
-\end{figure}
-
-Despite their differences, there are additional common themes.
-All obtained solutions have lower association P\=/values than the whole PPIN (median VEGAS2 P\=/value $\ll 0.46$, Table~\ref{tab:gene_solutions}), despite containing genes with higher P\=/values as well (Fig~\ref{fig:solution_overview}D).
-This illustrates the trade-off between controlling for type I error and biological relevance.
-However, there are nuances between solutions in this regard: heinz strongly favors genes with lower P\=/values, while dmGWAS is less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tends to select whole LD-blocks; and HotNet2 and SigMod are less likely to select low scoring genes.
-
-Additionally, the solutions presented other desirable properties.
-First, five of them were enriched in known breast cancer susceptibility genes (consensus, dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value$<$0.03).
-Second, the genes in four solutions displayed on average a higher betweenness centrality than the rest of the genes, a difference that is significant in four solutions (consensus, dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value$<$1.4 \texttimes{} 10\textsuperscript{-21}).
-This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
-
-\subsection{A case study: the consensus solution}
-\label{results:consensus}
-
-Despite their shared properties, the differences between the solutions of the different methods suggest that each of them captures different aspects of cancer susceptibility.
-Indeed, out of the 668 genes that are selected by at least one method, only 93 are selected by at least two, 20 by three, and none by four or more.
-Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig \ref{fig:solution_overview}B), a relationship that plateaus at 2.
-Hence, to leverage on their strengths and compensate their respective weaknesses, we built a consensus solution that captures the genes shared among at least two solutions (Section~\ref{methods:methods}).
-This solution (Fig~\ref{fig:consensus}) contains 93 genes and exhibits the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
-
-\begin{figure}[!ht]
-  \centering
-  \includegraphics[width=.8\linewidth]{./figures/figure_3.pdf}
+  \includegraphics[width=.9\linewidth]{./figures/figure_3.pdf}
   \caption{ {\bf Consensus solution on GENESIS (Section~\ref{methods:methods}) }.
-\textbf{A} ~Manhattan plots of genes; in black, the ones in the consensus solution.
+\textbf{(A)} ~Manhattan plot of genes; in black, the ones in the consensus solution.
 The Bonferroni threshold is indicated by a red line (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{B} Consensus network.
+\textbf{(B)} Consensus network.
 Each gene is represented by a pie chart, which shows the methods that selected it.
 We enlarged the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
-\textbf{C} The nodes are in the same disposition as in panel B, but every gene name is indicated.
-We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset}
+\textbf{(C)} The nodes are in the same disposition as in panel B, but every gene name is indicated.
+We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.}
   \label{fig:consensus}
@@
 
-A pathway enrichment analysis of the genes in the consensus solution also shows similar pathways as the individual solutions (\nameref{stab:consensus_pwy}).
+A pathway enrichment analysis of the genes in the consensus solution also showed similar pathways as the individual solutions (\nameref{stab:consensus_pwy}).
 We found two involved mechanisms: \emph{mitochondrial translation} and \emph{attenuation phase}.
@@
 Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
-With regards to the attenuation phase of heat shock response, it involves three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
+With regards to the attenuation phase of heat shock response, it involved three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
 The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 map to all of them, and 4 to two, making it hard to disentangle their effects.
+In fact, out of the 22 SNPs that map to any of these three genes, 9 mapped to all of them, and 4 to two, making it hard to disentangle their effects.
 \emph{HSPA1A} was the most strongly associated gene (VEGAS2 P\=/value~=~8.37~\texttimes{}~10\textsuperscript{-4}).
 
-Topologically, the consensus consists of a connected component composed of 49 genes, and multiple smaller subnetworks.
-Among the latter, 19 genes are in subnetworks containing a single gene or two connected nodes, implying that they do not have a consistently altered neighborhood, but are strongly associated themselves and hence picked by at least two methods.
-The large connected component contains genes that are highly central in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
-This suggests that these genes were selected because they were on the shortest path between another two highly associated genes.
+Topologically, the consensus consisted of a connected component composed of 49 genes, and multiple smaller subnetworks (Fig~\ref{fig:consensus}B and C).
+Among the latter, 19 genes were in subnetworks containing a single gene or two connected nodes, implying that they did not have a consistently altered neighborhood, but were strongly associated themselves and hence picked by at least two methods.
+The large connected component contained genes that are highly central in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
+This suggested that these genes were selected because they were on the shortest path between another two highly associated genes.
 In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus solution is COPS5, a component of the COP9 signalosome which regulates multiple signaling pathways.
+For instance, the most central node in the consensus solution is \emph{COPS5}, a component of the COP9 signalosome which regulates multiple signaling pathways.
 \emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS or in studies conducted by the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus solution have consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
+Despite its lack of association in GENESIS or in studies conducted by the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus solution had consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
 
@@
 Although BCAC case-control studies do not necessarily target cases with a familial history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
-This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provides a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
+This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provided a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
 We computed a gene association score on BCAC (Section~\ref{methods:bcac}).
-The solutions provided by the different  network methods overlap significantly with BCAC findings (Fisher's exact test P\=/value \textless~0.019).
-The gene-based methods achieve comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (\nameref{sfig:additional-benchmarks}A).
-Interestingly, while SConES GI, at the SNP-level, achieves a similar recall (8.6\%), it shows a much higher precision (47.3\%).
+The solutions provided by the different  network methods overlapped significantly with BCAC hits (Fisher's exact test P\=/value \textless~0.019).
+The gene-based methods achieved comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (\nameref{sfig:additional-benchmarks}A).
+Interestingly, while SConES GI, at the SNP-level, achieved a similar recall (8.6\%), it showed a much higher precision (47.3\%).
 
@@
 
-To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on the \emph{test dataset} (\nameref{sfig:additional-benchmarks}B).
+To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on \emph{test sets} (\nameref{sfig:additional-benchmarks}B).
 The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
 Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
-Of course, although these performances are low, we do not expect to separate cases from controls well using exclusively genetic data \cite{deloscamposComplexTraitPredictionEra2018}.
+Of course, although these performances were low, we did not expect to separate cases from controls well using exclusively genetic data \cite{deloscamposComplexTraitPredictionEra2018}.
 
@@
 HotNet2 was the slowest (3 days and 14 hours on average).
-Including the time required to compute the gene scores, however, slows down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
+Including the time required to compute the gene scores, however, slowed down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
 Including this first step, it therefore took 5 days on average for HotNet2 to produce a result.
 
-\subsection{Network topology matters, and might lead to ambiguous results}
+\subsection{Network topology and association scores matter, and might lead to ambiguous results}
 \label{results:drawbacks}
@@
 This is due to the fact that each of them models the problem differently.
-Importantly, understanding which assumptions they make allows to understand the results more in depth.
-For instance, the fact that LEAN did not return any gene implies that there is no gene such that both itself and its environment are on average strongly associated with the disease.
-
-In the GENESIS dataset, heinz's solution is very conservative, providing a small solution with the lowest median P\=/value (Table~\ref{tab:gene_solutions}).
+Importantly, understanding which assumptions they made allowed to understand the results more in depth.
+For instance, the fact that LEAN did not return any gene implied that there was no gene such that both itself and its environment were on average strongly associated with the disease.
+
+In the GENESIS dataset, heinz's solution was very conservative, providing a small solution with the lowest median P\=/value (Table~\ref{tab:gene_solutions}).
 By repeatedly selecting this compact solution, heinz was the most stable method (Fig~\ref{fig:benchmark}C).
-Its conservativeness stems from its preprocessing step, which models the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (\nameref{sfig:snp_gene_manhattan}B), only 36 of all the genes retain a positive score after that transformation.
-Yet, this small solution does not provide much insight into the susceptibility mechanisms to cancer.
-Importantly, it ignores genes that are associated to cancer in this dataset like \emph{FGFR2}.
+Its conservativeness stemmed from its preprocessing step, which modeled the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
+Due to the limited signal at the gene level in this dataset (\nameref{sfig:snp_gene_manhattan}B), only 36 of all the genes retained a positive score after that transformation.
+Yet, this small solution did not provide much insight into the susceptibility mechanisms to cancer.
+Importantly, it ignored genes that were associated to cancer in this dataset like \emph{FGFR2}.
 
 On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
-dmGWAS' solution is the lowest scoring solution on average.
-This is due to the greedy framework it uses, which has a bias for larger solutions \cite{nikolayeva_network_2018}.
+dmGWAS' solution was the lowest scoring solution on average.
+This was due to the greedy framework it uses, which had a bias for larger solutions \cite{nikolayeva_network_2018}.
 It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated gene if it was linked to another, high scoring one.
-This is exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
+This was exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
 This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
-In the same line, HotNet2's solution is even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
-Lastly, SigMod misses some of the highest scoring, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
+In the same line, HotNet2's solution was even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
+Lastly, SigMod missed some of the highest scoring, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
 
@@
   \caption{{\bf Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
-\textbf{(B)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
+\textbf{(B)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its centrality degree.
+\textbf{(C)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
 We highlighted the genes selected by each method, and the ones selected by more than one (``Consensus'').
 We labeled the three most central genes that were picked by any method.
-\textbf{(C)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its centrality degree.
 \textbf{(D)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
@@
 
-Another peculiarity of network methods is their relationship to degree centrality.
-We studied random rewirings of the PPIN while preserving node centrality (Section~\ref{methods:rewiring}).
+Another peculiarity of network methods was their relationship to degree centrality.
+We studied random rewirings of the PPIN that preserved node centrality (Section~\ref{methods:rewiring}).
 In this setting, network methods favored central genes, as they often connect high scoring nodes (Fig~\ref{fig:issues}B).
-This is despite the fact that highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
-This was especially the case of SigMod, which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
+This was despite the fact that highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
+This was especially the case of SigMod (\nameref{sfig:rewirings}), which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
 As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
@@
 
-By virtue of using a SNP subnetwork, SConES analyzes each SNP in their functional context.
-It can therefore select SNPs located in genes not included in the PPIN, as well as SNPs in non-coding regions or in non-interacting genes.
+By virtue of using a SNP subnetwork, SConES analyzed each SNP in their functional context.
+It could therefore select SNPs located in genes not included in the PPIN, as well as SNPs in non-coding regions or in non-interacting genes.
 We compared the solution of SConES in the GI network (using PPIN information), to the one using only positional information (GS network), or positional and gene annotations (GM network).
-Importantly, SConES produces similar results on the GS and GM networks (\nameref{sfig:pearson_methods}).
-While the solutions on those two greatly overlap with SConES GI's, they contain additional gene-coding segments (Fig~\ref{fig:issues}C).
-In fact the formers' solutions are chromosome regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
-On top of those SConES GS selects region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
+Importantly, SConES produced similar results on the GS and GM networks (\nameref{sfig:pearson_methods}).
+While the solutions on those two greatly overlap with SConES GI's, they contained additional gene-coding segments (Fig~\ref{fig:issues}C).
+In fact the formers' solutions were chromosome regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
+On top of those SConES GS selected region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
 
@@
 We explored the parameter space of the different methods by running them under different combinations of parameters (Section \ref{methods:parameters}).
-In agreement with their formulations (Section \ref{methods:methods}), larger values of certain parameters produce less stringent solutions (\nameref{sfig:parameters}A): for HotNet2 and heinz, this is the threshold above which genes are assigned a non-zero score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; % node involved how far the search from the explored subnetwork;
-for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
+In agreement with their formulations (Section \ref{methods:methods}), larger values of certain parameters produced less stringent solutions (\nameref{sfig:parameters}A): for HotNet2 and heinz, this is the threshold above which genes are assigned a non-zero score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
 Two parameters had the opposite effect (larger is more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES's $\eta$, where higher values produce smaller solutions.
@@
 We computed the Pearson correlation between the different solutions as in Section \ref{methods:algorithm_comparison} to study how the parameters affected which genes and SNPs were selected (\nameref{sfig:parameters}C).
-This showed that dmGWAS and SigMod are robust to certain parameters: dmGWAS' output was mostly determined by the parameter $d$, rather than $r$; SigMod's by $nmax$, rather than $maxjump$.
+This showed that dmGWAS and SigMod were robust to certain parameters: dmGWAS' output was mostly determined by the parameter $d$, rather than $r$; SigMod's by $nmax$, rather than $maxjump$.
 
 SConES presented an interesting case in terms of feature selection: most of the explored combinations of parameters led to trivial solutions (either all the SNPs, or none of them were included) (\nameref{sfig:parameters}A).
-To explore a more meaningful parameter space, we selected the parameters in two rounds.
+To explore a more meaningful parameter space, in our experiments we selected the parameters in two rounds.
 First, we explored the whole sample space.
@@
 The resulting stability-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}).
-This network shares most of the properties of the consensus: breast cancer susceptibility genes are overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes are more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) are isolated.
+This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig \ref{fig:stable-consensus}B and C).
 
@@
   \caption{ {\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}) }.
-\textbf{A} ~Manhattan plots of genes; in black, the ones in the stable consensus solution.
+\textbf{(A)} ~Manhattan plot of genes; in black, the ones in the stable consensus solution.
 The Bonferroni threshold is indicated by a red line (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{B} Stable consensus network.
+\textbf{(B)} Stable consensus network.
 Each gene is represented by a pie chart, which shows the methods that selected it.
 We enlarged the most central gene (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
-\textbf{C} The nodes are in the same disposition as in panel B, but every gene name is indicated.
-We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset}
+\textbf{(C)} The nodes are in the same disposition as in panel B, but every gene name is indicated.
+We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.}
   \label{fig:stable-consensus}
@@
 
-Despite these similarities, the consensus and the stable consensus solutions include different genes.
-In the stable consensus network, the most central gene is \emph{CUL3}, which is absent from the previous consensus solution and has a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
-This gene is a component of Cullin-RING ubiquitin ligases.
+Despite these similarities, the consensus and the stable consensus solutions included different genes.
+In the stable consensus network, the most central gene was \emph{CUL3}, which was absent from the previous consensus solution and had a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
+This gene was a component of Cullin-RING ubiquitin ligases.
 Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
@@
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided discovery have been proposed in recent years.
-In this article we evaluated the relevance of six of these methods by applying them to the study of GENESIS, a GWAS dataset on familial breast cancer.
-
-DmGWAS, Heinz, HotNet2, SConES and SigMod all yield interesting solutions, which include (but are not limited to) known breast cancer susceptibility genes.
-In general, the selected genes and SNPs were more central than average, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
+In this article we evaluated six of these methods (Section \ref{methods:methods}) by applying them to the GENESIS study, a GWAS dataset on familial breast cancer (Section \ref{methods:data}).
+
+DmGWAS, Heinz, HotNet2, SConES and SigMod all yielded interesting solutions, which include (but are not limited to) known breast cancer susceptibility genes (Section \ref{results:separate_networks}).
+In general, the selected genes and SNPs were more central than most other genes and SNPs, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
 However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by network methods (Section \ref{results:drawbacks}).
-Yet, across this article we show that highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
+Yet, across this article we showed that the highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
 Despite these similarities, the solutions obtained were notably different.
-At one end of the spectrum, SConES and heinz preferred small, highly associated solutions, at the expense of not shedding much light on the etiology of the disease.
-On the other end, SigMod and dmGWAS gravitate towards larger, less associated solutions which provide a wide overview of the biological context.
-While this deepens our understanding of the disease and provides biological hypotheses, they require further analyses.
-For instance, a user might need to examine the centrality of the selected genes, and discern the extent to which that property is driving the selection of each gene.
-HotNet2 balances both approaches at the expense of producing the largest solution: a constellation of many, highly associated, small subnetworks.
-Additionally, all solutions share two drawbacks.
-First, they are all equally bad at discriminating cases from controls.
-Yet, the classification accuracy of network methods is similar to that of a classifier trained on the entire genome, which suggests that cases and controls are difficult to separate in the GENESIS dataset.
+At one end of the spectrum, SConES and heinz preferred highly associated solutions, which in turn were small and did not shed much light on the etiology of the disease.
+On the other end, SigMod and dmGWAS gravitated towards larger, less associated solutions which provided a wide overview of the biological context.
+While this deepened our understanding of breast cancer susceptibility and provided biological hypotheses, they required further analyses.
+For instance, we had to examine the centrality of the selected genes to discern the extent to which that property is driving the selection of each gene (Section \ref{results:drawbacks}).
+Lastly, HotNet2 produced a third kind of solution: a large constellation of many, high-scoring, small subnetworks.
+However, all solutions shared two drawbacks.
+First, they were all equally bad at discriminating cases from controls.
+Yet, the classification accuracy of network methods was similar to that of a classifier trained on the entire genome (Section \ref{results:benchmark}), which suggests that cases and controls are difficult to separate in the GENESIS dataset.
 This may be due to limited statistical power, which reduces the ability to identify relevant SNPs; but in any event, we do not expect to be able to separate people who have or will developed cancer from others on the sole basis of their genomes, ignoring all environmental factors and chance events.
-Second, all methods are remarkably unstable, yielding different solutions for slightly different inputs.
-This might partly be caused by the instability of the P\=/values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
-Hence, heinz's conservative transformation of P\=/values, which favors only the most extreme ones, leads to improved stability.
-Another source of instability might be the redundancy inherent to biological networks, since they are subject to an evolutionary pressure to avoid single points of failure \cite{wagnerAlternativeRoutesMutational2007}.
+Hence, in this case network methods were preferable to the logistic regression classifier, since they did ``no worse'' while providing an interpretable solution.
+Second, all methods were remarkably unstable, yielding different solutions for slightly different inputs.
+This might partly have been caused by the instability of the P\=/values themselves in low statistical power settings \cite{halseyFickleValueGenerates2015}.
+Hence, heinz's conservative transformation of P\=/values, which favored only the most extreme ones, led to improved stability.
+Another source of instability might have been the redundancy inherent to biological networks, since they are subject to an evolutionary pressure to avoid single points of failure \cite{wagnerAlternativeRoutesMutational2007}.
 Hence, biological networks will often have multiple paths connecting two high-scoring nodes.
@@
 To overcome these limitations while exploiting the strengths of the individual methods, we proposed combining them into a consensus solution.
-We use a straightforward strategy of including any node that was recovered by multiple methods.
-We thus proposed two networks: a consensus solution, meant to address the heterogeneity of the solutions in the full dataset, and a stable consensus solution, which in addition addressed the instability of the methods.
+We used a straightforward strategy of including any node that was recovered by multiple methods.
+We thus proposed two networks (Sections \ref{results:consensus} and \ref{results:stable-consensus}): a consensus solution, meant to address the heterogeneity of the solutions in the full dataset, and a stable consensus solution, which in addition addressed the instability of the methods.
 They both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer, thus synthetizing the mechanisms altered in breast cancer cases.
@@
 In this work, we have compared our results to significant genes and SNPs in the BCAC study \cite{Michailidou2017}.
-Network methods show modest precision, but much higher recall at recovering BCAC hits (Section \ref{results:boost}).
+Network methods showed modest precision, but much higher recall at recovering BCAC hits (Section \ref{results:boost}).
 While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in exploratory settings.
-Nonetheless, BCAC is not an ideal ground truth.% in which GENESIS had more samples% .
-Yet, there are caveats which make it a non-optimal comparison.
-First, the studied populations are non-overlapping: BCAC focuses on a pan-European cohort, while GENESIS targets the French population specifically.
-Second, the study designs differ: a high proportion of breast cancer cases investigated in BCAC are sporadic (not selected according to family history), while GENESIS is a homogeneous dataset not included in BCAC and which focuses on the French high-risk population attending the family cancer clinics.
-Despite these differences, we expect some degree of shared genetic architecture, especially at the gene level.
+Nonetheless, BCAC was not an ideal ground truth.
+First, the studied populations are not completely overlapping: BCAC focused on a pan-European cohort, while GENESIS targeted the French population specifically.
+Second, the study designs differed: a high proportion of breast cancer cases investigated in BCAC were sporadic (not selected according to family history), while GENESIS was a homogeneous dataset not included in BCAC and which focused on the French high-risk population attending the family cancer clinics.
+Despite these differences, we expected some degree of shared genetic architecture, especially at the gene level.
 Finally, and this is indeed the motivation for this study, GWAS are unlikely to identify all genes relevant for the disease: some might only show up in rare-variant studies; others might have too low effect sizes.
@@
 
-As not all databases compile the same interactions, the choice of the PPIN determines the final output.
+As not all PPIN databases compiled the same interactions, the choice of the PPIN determines the final output.
 In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responds to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
+This responded to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
 On the other hand, Huang et al.
@@
 However, when we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
-This supports using only interactions from high-throughput experiments, which produces apparently similar solutions and avoids falling into ``circular reasonings'', where the best-known genes are artificially pushed into the solutions.
+This supports using only interactions from high-throughput experiments, which produced apparently similar solutions and avoided falling into ``circular reasonings'', where the best-known genes were artificially pushed into the solutions as we observed in Section \ref{results:drawbacks}.
 
@@
 In consequence, they show their shortcomings on understudied genes, especially those not in the network.
-Out of the 32\,767 genes to which we can map the genotyped SNPs, 60.7\% (19\,887) are not in the PPIN.
+Out of the 32\,767 genes to which we mapped the genotyped SNPs, 60.7\% (19\,887) were not in the PPIN.
 The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (\nameref{sfig:biotypes_excluded}).
-Yet, RNA genes like \emph{CASC16} are associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
+Yet, RNA genes like \emph{CASC16} were associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
 In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, overlapping with \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the PPIN.
-However, on average protein-coding genes absent from the PPIN are less associated with breast cancer susceptibility (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
-This cannot be due to well-known genes having more known interactions because we are only using interactions from high-throughput experiments.
-As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it is due to interactions between central genes being more likely.
+However, on average protein-coding genes absent from the PPIN were less associated with breast cancer susceptibility (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
+This could not be due to well-known genes having more known interactions because we only used interactions from high-throughput experiments.
+As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it was due to interactions between central genes being more likely.
 It is worth noting that network methods that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
@@
 
-How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners. % Another important decision is how to handle LD in a GWAS.
+How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners.
 VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
-With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
-However, as mentioned in Section~\ref{results:drawbacks}, LD patterns seem to drive SConES' solutions, and an LD pruning step could potentially alter them.
-In Section~\ref{results:consensus} we highlight ambiguities that appear when genes overlap or are in LD.
+With regards to gene-based network methods, in Section~\ref{results:consensus} we highlighted ambiguities that appear when genes overlap or are in LD.
 In fact, the presented case is paradigmatic, since all three genes are located in the HLA region, the most gene-dense region of the genome \cite{Xie2003}.
@@
 But the opposite case is also true: when genes are not functionally related (and hence disconnected in the PPIN), network methods might disregard them even if they have high association scores.
-LD also affects SConES, since it penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Equation~\ref{eq:scones}.
+With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
+However, LD patterns also affect SConES' in other ways, since its formulation penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Equation~\ref{eq:scones}.
 Due to LD, nearby SNPs' P-values are correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
-Hence, SConES will tend to select LD-blocks formed by low P-value SNPs.
+Hence, SConES tends to select LD-blocks formed by low P-value SNPs.
 This might explain why SConES produces similar results on the GS and GM networks, heavily affected by LD (Section \ref{results:drawbacks}).
@@
 
-To produce the two consensus solutions, we had to face practical challenges due to the differences in interfaces, preprocessing steps, and unexpected behaviors of the various methods.
-To make it easier for other to apply these methods to new datasets and aggregate their solutions, we built six nextflow pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
-They are available on GitHub: \url{https://github.com/hclimente/gwas-tools} (Section \ref{methods:data}).
-Importantly, those methods that had a permissive license were compiled into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
+To produce the two consensus solutions, we faced practical challenges due to the differences in interfaces, preprocessing steps, and unexpected behaviors of the various methods.
+To make it easier for other to apply these methods to new datasets and aggregate their solutions, we built six \texttt{nextflow} pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
+They are available on GitHub: \href{https://github.com/hclimente/gwas-tools}{hclimente/gwas-tools} (Section \ref{methods:code}).
+Importantly, we compiled those methods that had a permissive license into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
 
@@
 The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
-Index cases are patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
-Controls are unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
+Index cases were patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
+Controls were unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
 We focused on the 2\,577 samples of European ancestry, of which 1\,279 were controls and 1\,298 were cases.
 The genotyping was performed using the iCOGS array, a custom Illumina array designed to study the genetic susceptibility to hormone-related cancers \cite{sakoda_turning_2013}.
-It contains 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
+It contained 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
 
@@
 \label{methods:notation}
-In this article, we use undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
+In this article, we used undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
 $V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
@@
 Nodes can be described by properties provided by the topology of the graph.
-We focus on two of those: degree centrality, and betweenness centrality.
+We focused on two of those: degree centrality, and betweenness centrality.
 The degree centrality, or degree, is the number of edges that a node has.
@@
 
-We also use two matrices that describe two different properties of a graph.
+We also used two matrices that describe two different properties of a graph.
 These matrices are square, and have as many rows and columns as nodes in the network.
@@
 \paragraph{Gene network}
-The mathematical formulations of the different network methods are compatible with any type of network (protein interactions, gene coexpression, regulatory, etc.).
-Here, we used protein-protein interaction networks (PPIN) for all of them except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
+The mathematical formulations of the different network methods are compatible with any type of biological network (protein interactions, gene coexpression, regulatory, etc.).
+Here, we used protein-protein interaction networks (PPIN) for all of the network methods except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 \paragraph{SNP networks}
-SConES \cite{azencott_efficient_2013} is the only network method designed to handle SNP networks.
-As in gene networks, two SNPs are connected in a SNP network when there is evidence of shared functionality between two SNPs.
+SConES \cite{azencott_efficient_2013} was the only network method designed to handle SNP networks.
+As in gene networks, two SNPs were connected in a SNP network when there was evidence of shared functionality between them.
 Azencott et al.
 \cite{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
-We focused on the GI network, as it fits the scope of this work better, using the PPIN described above.
+We focused on the GI network, as it fitted the scope of this work better, using the PPIN described above.
 However, at different stages we also compared GI to GS and GM to understand how considering the PPIN affects SConES' output.
 For the GM network, we used the mapping described in Section~\ref{methods:snp2gene}.
-In all three the node scores are the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2} statistic).
+In all three the node scores were the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2} statistic).
 The properties of these three subnetworks are available in \nameref{stab:snp_solutions}.
@@
 \textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
-\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected hyperparameters are explored.
-\textbf{\emph{Solution}}: additional properties are enforced on the solution, other than containing high scoring, connected nodes.
+\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected parameters are explored.
+\textbf{\emph{Solution}}: additional properties that are enforced on the solution, other than containing high scoring, connected nodes.
 \textbf{\emph{Comp.}}: number of connected components in the solution.
 \textbf{\emph{Input}}: genotype data or GWAS summary statistics.
-\textbf{\emph{Scoring}}: how SNP/gene P\=/values are transformed into node scores.
+\textbf{\emph{Scoring}}: how SNP/gene P\=/values were transformed into node scores.
 In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the cumulative distribution function of the standard Normal distribution.
@@
 where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) d.
-\item Add the neighboring nodes whose inclusion increases Z\textsubscript{m+1} by more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + r).
+\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) \texttt{d}.
+\item Add the neighboring nodes whose inclusion increases Z\textsubscript{m+1} by more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + \texttt{r}).
 \item Repeat 2-4 until no further enlargement is possible.
@@
 We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-Unless otherwise specified, we used the suggested hyperparameters $d = 2$ and $r = 0.1$.
+Unless otherwise specified, we used the suggested parameters \texttt{d} $ = 2$ and \texttt{r} $ = 0.1$.
 We used the function \texttt{simpleChoose} to select the solution, which aggregates the top 1\% subnetworks.
@@
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors propose a transformation of the genes' P\=/value into a score that is negative under weak association with the phenotype, and positive under a strong one.
+The authors proposed a transformation of the genes' P\=/value into a score that is negative under weak association with the phenotype, and positive under a strong one.
 This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
@@
 Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
-We used the implementation of LEAN from the LEANR R package \cite{leanr}.
+We used the implementation of LEAN from the \texttt{LEANR} R package \cite{leanr}.
 \item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem
@@
 
-where \(\lambda\) and \(\eta\) are hyperparameters that control the sparsity and the connectivity of the model.
+where \(\lambda\) and \(\eta\) are parameters that control the sparsity and the connectivity of the model.
 The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
@@
 \cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
-In this case, the selected hyperparameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
-We used the version on SConES implemented in the R package martini \cite{martini}.
+In this case, the selected parameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
+We used the version on SConES implemented in the R package \texttt{martini} \cite{martini}.
 
 \item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones.
+It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones:
 
@@
 SConES, instead, penalizes connections between the selected and unselected nodes.
-Third, it explores the grid of hyperparameters differently, and processes their respective solutions.
+Third, it explores the grid of parameters differently, and processes their respective solutions.
 Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
@@
 The candidate solution with the highest standardized score and that is not larger than a user-specified threshold (\texttt{nmax}) is the chosen solution.
+We used the default parameters \texttt{maxjump} $ = 10$ and \texttt{nmax} $= 300$.
 SigMod is implemented in an R package \cite{sigmod}.
@@
 \begin{description}
-\item[{dmGWAS}] We tested multiple values for $r$ (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and $d$ (1, 2, and 3). 
+\item[{dmGWAS}] We tested multiple values for \texttt{r} (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and \texttt{d} (1, 2, and 3). 
 \item[{heinz}] We tested multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
@@
 \item[{LEAN}] We used the following significance cutoffs for LEAN's P-values (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, and 1).
-\item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
+\item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores by default (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
 \item[{SigMod}] We tested multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
@@
 We searched for pathways enriched in the gene solutions produced by the above methods.
-We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the ReactomePA R package \cite{Yu2016}.
+We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the \texttt{ReactomePA} R package \cite{Yu2016}.
 The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:conventional}).
-We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH).
-Pathways with a BH adjusted P\=/value \textless~0.05 were deemed significant.
+We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH): pathways with a BH adjusted P\=/value \textless~0.05 were deemed significant.
 
@@
 
-We evaluated multiple properties (described below) of the different methods through a 5-fold subsampling setting.
+We evaluated multiple properties (described in Sections \ref{methods:algorithm_comparison} and \ref{methods:classifier}) of the different methods through a 5-fold subsampling setting.
 We applied each method to 5 random subsets of the original GENESIS dataset containing 80\% of the samples (\emph{train set}).
@@
 This penalty was set by cross-validation, choosing the value that minimized misclassification error.
-We applied each network method to each \emph{train} set, and trained the classifier on the same train set using only the selected SNPs.
+We applied each network method to each train set, and trained the classifier on the same train set using only the selected SNPs.
 When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
-Then we evaluated the sensitivity and the specificity on the \emph{test set}.
+Then we evaluated the sensitivity and the specificity of the classifier on the test set.
 The active set gave an estimate of a plausible, more sparse solution with comparable predictive power to the original solution.
-To obtain a baseline, we also trained the classifier on all the SNPs.
+To obtain a baseline, we also trained the classifier on all the SNPs of the train set.
 We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
@@
 
-\subsubsection{Comparison to state-of-the-art}
+\subsection{Comparison to state-of-the-art}
 \label{methods:bcac}
@@
 
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries which included 13\,250\,641 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries.
+BCAC included 13\,250\,641 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
 Through imputation, BCAC includes more SNPs than the iCOGS array used for GENESIS (Section~\ref{methods:data}).
-Yet, in all the comparisons in this paper, we focused on the subset of the GENESIS SNPs that passed quality control (Section~\ref{methods:data}).
+Yet, in all the comparisons in this paper, we focused on the subset of the GENESIS SNPs that passed quality control.
 Hence, we used the same Bonferroni threshold as in Section~\ref{methods:conventional} to determine the significant SNPs in BCAC.
@@
 
-\subsubsection{Network rewirings}
+\subsection{Network rewirings}
 \label{methods:rewiring}
 
-Rewiring the PPIN while preserving the number of edges of each gene allows to study the impact of the topology on the output of network methods.
+Rewiring the PPIN while preserving the number of edges of each gene allowed to study the impact of the topology on the output of network methods.
 Indeed, the edges lose their biological meaning but the topology of the network is conserved.
@@
 We only applied only four methods on the rewirings: heinz, dmGWAS, LEAN and SigMod.
-We excluded HotNet2 and SConES, since they take notably longer to run than the other methods, while taking up more computational resources.
+We excluded HotNet2 and SConES, since they take notably longer to run than the other methods.
 
@@
 They are compiled in \url{https://github.com/hclimente/gwas-tools}.
-Although the GENESIS data is not public, the code to apply the pipelines to this data, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+The code to apply the pipelines to GENESIS, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
 We deposited all the produced gene solutions on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
@@
 \paragraph*{S6 Fig.}
+\label{sfig:rewirings}
+{\bf Number of times a gene was selected by either dmGWAS, heinz, LEAN,or SigMod in 100 rewirings of the PPIN (Section \ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
+
+\paragraph*{S7 Fig.}
 \label{sfig:parameters}
-{\bf Parameter space of the network methods.} \textbf{(A)} Boxplot of the solution sizes of the methods under the explored parameters (Section \ref{methods:parameters}).
-\textbf{(B)} Size of SConES's with regards to each pair of parameters.
+{\bf Parameter space of the network methods.} \textbf{(A)} Size of the solution produced by different parameter values, expressed as a percentage of the maximum solution size for the method, or the highest tested value for the parameter, respectively.
+The size of the solution is the median among all the solution sizes for the same parameter.
+\textbf{(B)} Boxplot of the solution sizes of the methods under the explored parameters (Section \ref{methods:parameters}).
 \textbf{(C)} Pearson correlation between the solutions of the different runs.
@@
 
-\paragraph*{S7 Fig.}
+\paragraph*{S8 Fig.}
 \label{sfig:biotypes_excluded}
@@
 
-\paragraph*{S8 Fig.}
+\paragraph*{S9 Fig.}
 \label{sfig:lc_ht_comparison}
@@
 
-\paragraph*{S9 Fig.}
+\paragraph*{S10 Fig.}
 \label{sfig:scones_gsm}

### 2020-10-18 manuscript/plos/original.tex
updated latest version of the paper

@@
 Nadine Andrieu\textsuperscript{1,2,3}, 
-Chloé-Agathe Azencott\textsuperscript{3,1,2},
-with the GENESIS study group\textsuperscript{\textpilcrow}
+Chloé-Agathe Azencott\textsuperscript{3,1,2}
 \\
@@
 We used the function \emph{simpleChoose} to select the solution subnetwork, which aggregates the top 1\% subnetworks.
-\end{description}
-
-\begin{description}
+
 \item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
@@
 We used the version on SConES implemented in the R package martini \cite{martini}.
-\end{description}
-
-\begin{description}
+
 \item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected gene subnetwork \cite{liu_sigmod:_2017}.
@@
 SigMod is implemented in an R package \cite{sigmod}.
-\end{description}
-
-\begin{description}
+
 \item[{Consensus}] We built a consensus network by retaining the nodes that were selected by at least two of the six methods (using SConES GI for SConES).

### 2020-10-18 manuscript/plos/review_1.tex
updated latest version of the paper

@@
 \section*{Abstract}
-% Network methods provide a comprehensive approach to uncovering the genetics of complex diseases and building hypotheses.
-% They discover susceptibility genes by jointly consider the statistical association between genetic variation and a phenotype, measured in a genome-wide association study (GWAS), and the biological context of each gene, represented as a network.
 Network approaches to disease use biological networks, which model functional relationships between the molecules in a cell, to generate hypotheses about the genetics of complex diseases.
-Among them, several approaches jointly consider gene scores, representing the association between each gene and the disease, and the biological context of each gene, modeled by a network.
-%These approaches can select low-scoring genes if they connect multiple high-scoring genes.
-Here, we obtain gene scores from genome-wide association studies (GWAS) and study six network approaches on 
-% Hence, a network method could select a gene with a high P-value of association if it connects multiple low P-value genes in the network.
-In this work, we studied six network methods which identify subnetworks with high scores of GWAS association with a phenotype.
-This allows them to give more compelling results than standard SNP- and gene-level GWAS analyses, recovering causal subnetworks tightly related to cancer susceptibility.
-% We apply them to
-GENESIS, a GWAS on French women with non-BRCA familial breast cancer% and no pathogenic variants in \emph{BRCA1} and \emph{BRCA2}
-.
-We provide a critical comparison of these six methods, discussing the impact of their mathematical formulation and of their parameters.
-% on % .
-Additionally, we performed an in-depth benchmarking with respect to
-% the size and predictive power of their solutions, as well as on their stability and runtimes.
-% Importantly,
-Using a biological network yields more compelling results than standard % SNP- and gene-level
-GWAS analyses. Indeed, %recovering causal subnetworks tightly related to cancer susceptibility.
-we find significant overlaps between our solutions and the genes identified in the largest GWAS on breast cancer susceptibility. % Since most of the methods produce reasonable solutions, we
-We further
-propose to combine these solutions into a consensus network, %containing the genes selected by at least two methods. This aggregation
-which
-brings further insights.
+Several among them jointly consider gene scores, representing the association between each gene and the disease, and the biological context of each gene, modeled by a network.
+Here, we study six such network methods using gene scores from GENESIS, a genome-wide association study (GWAS) on French women with non-BRCA familial breast cancer.
+We provide a critical comparison of these six methods, discussing the impact of their mathematical formulation and parameters.
+Using a biological network yields more compelling results than standard GWAS analyses.
+Indeed, we find significant overlaps between our solutions and the genes identified in the largest GWAS on breast cancer susceptibility.
+We further propose to combine these solutions into a consensus network, which brings further insights.
 The consensus network contains \emph{COPS5}, a gene related to multiple hallmarks of cancer, and 14 of its neighbors.
-The main drawback of network methods % with respect to conventional $\chi^2$-based GWAS
-is that they are not robust to small perturbations in their inputs. %To compensate for this instability, we
-We therefore propose a stable consensus solution, formed by the most consistently selected genes in multiple subsamples of the data.
+The main drawback of network methods is that they are not robust to small perturbations in their inputs.
+Therefore, we propose a stable consensus solution, formed by the most consistently selected genes in multiple subsamples of the data.
 In GENESIS, it is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than most genes.
@@
 Project-agnostic implementations of all methods are available at \url{https://github.com/hclimente/gwas-tools}.
-% to facilitate their application to other GWAS datasets.
 
@@
 \section*{Author summary}
-In genome-wide association studies (GWAS), thousands of genomes are scanned to identify variants associated with a complex trait.
+Genome-wide association studies (GWAS) scan thousands of genomes to identify variants associated with a complex trait.
 Over the last 15 years, GWAS have advanced our understanding of the genetics of complex diseases, and in particular of hereditary cancers.
-Yet, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
-The omnigenic model offers an appealing explanation: only a limited number of \emph{core} genes are directly involved in the disease; but gene functions are deeply interrelated, and so many other genes can alter the function of the core genes.
+However, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
+The omnigenic model offers an appealing explanation: only a limited number of \emph{core} genes are directly involved in the disease, but gene functions are deeply interrelated, and so many other genes can alter the function of the core genes.
 These interrelations are often modeled as networks, and multiple algorithms have been proposed to use these networks to identify the subset of core genes involved in a specific trait.
-In this study, we apply and compare six such network methods on GENESIS, a GWAS dataset for familial breast cancer in the French population.
-Combining these approaches allows us to identify potentially novel breast cancer susceptibility genes, and provides a mechanistic explanation for their role in the development of the disease.
+This study applies and compares six such network methods on GENESIS, a GWAS dataset for familial breast cancer in the French population.
+Combining these approaches allows us to identify potentially novel breast cancer susceptibility genes and provides a mechanistic explanation for their role in the development of the disease.
 We provide ready-to-use implementations of all the examined methods.
@@
 In human health, genome-wide association studies (GWAS) aim at quantifying how single-nucleotide polymorphisms (SNPs) predispose to complex diseases, like diabetes or some forms of cancer \cite{bush_chapter_2012}.
-To that end, in a typical GWAS thousands of unrelated samples are genotyped: the cases, suffering from the disease of interest, and the controls, taken from the general population.
-Then, a statistical test of association (e.g.
-based on logistic regression) is conducted between each individual SNP and the phenotype.
+To that end, in a typical GWAS, thousands of unrelated samples are genotyped: the cases, suffering from the disease of interest, and the controls, taken from the general population.
+Then, a statistical test of association (e.g., based on logistic regression) is conducted between each SNP and the phenotype.
 Those SNPs with a P\=/value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
-Once the risk SNPs have been discovered, they can be used for risk assessment, and to deepen our understanding of the disease.
+Once the risk SNPs have been discovered, they can be used for risk assessment and deepening our understanding of the disease.
 
 GWAS have successfully identified thousands of variants underlying many common diseases \cite{buniello_nhgri-ebi_2019}.
-However, this experimental setting also presents intrinsic challenges.
+However, this experimental setting also presents inherent challenges.
 Some of them stem from the high dimensionality of the problem, as every GWAS to date studies more variants than samples are genotyped.
-This limits the statistical power of the experiment, as only variants with larger effects can be detected \cite{visscher_10_2017}.
+This limits the statistical power of the experiment, as it can only detect variants with larger effects \cite{visscher_10_2017}.
 This is particularly problematic since the prevailing view is that most genetic architectures involve many variants with small effects \cite{visscher_10_2017}.
-Additionally, to avoid false positives, a conservative multiple test correction is applied, typically the previously mentioned Bonferroni correction.
-However, Bonferroni correction is overly conservative when the statistical tests are correlated, as it is the case in GWAS \cite{wang_statistical_2018}.
+Additionally, to avoid false positives, most GWAS apply a conservative multiple test correction, typically the previously mentioned Bonferroni correction.
+However, Bonferroni correction is overly conservative when the statistical tests correlate, as happens in GWAS \cite{wang_statistical_2018}.
 Another open issue is the interpretation of the results, as the functional consequences of most common variants are unknown.
 On top of that, recent large-sampled studies suggest that numerous loci spread all along the genome contribute to a degree to any complex trait, in accordance with the infinitesimal model \cite{barton_infinitesimal_2017}.
-The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are strongly inter-related and influence each other's function, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
-Hence, a comprehensive statistical framework which includes the structure of biological data might help alleviate the aforementioned issues.
-
-For this reason, many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013}.
+The recently proposed omnigenic model \cite{boyle_expanded_2017} offers an explanation: genes are strongly interrelated and influence each other's function, which allows alterations in most genes to impact the subset of ``core'' genes directly involved in the disease's mechanism.
+Hence, a comprehensive statistical framework that includes the structure of biological data might help alleviate the issues above.
+
+For this reason, many authors turn to network biology to handle the complex interplay of biomolecules that lead to disease \cite{furlong_human_2013,leisersonNetworkAnalysisGWAS2013}.
 As its name suggests, network biology models biology as a network, where the biomolecules under study, often genes, are nodes, and selected functional relationships are edges that link them.
@@
 Under this view, complex diseases are not the consequence of a single altered gene, but of the interaction of multiple interdependent molecules \cite{barabasi_network_2011}.
-In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,cai_broker_2010}: they tend to occupy central positions in the network (although not the most central ones); disease genes for the same pathology tend to be connected in modules; and often they are bottlenecks that interconect modules.
-Therefore, studying the neighborhood of disease-associated genes is effective at identifying new ones that are involved in the same biological functions \cite{huang_systematic_2018}.
-
-Network-based discovery methods exploit the differential properties described above to identify disease genes on GWAS data \cite{azencott_network-guided_2016}.
-In essence, each gene is assigned a score of association with the disease, computed from the GWAS data, and biological relationships, given by a network built on prior knowledge.
+In fact, an examination of biological networks shows that disease genes have differential properties \cite{barabasi_network_2011,cai_broker_2010}: they tend to occupy central positions in the network (although not the most central ones); disease genes for the same pathology tend to cluster in modules; and often they are bottlenecks that interconnect modules.
+
+Network-based discovery methods exploit the differential properties described above to identify disease genes on GWAS data \cite{azencott_network-guided_2016,huang_systematic_2018}.
+In essence, each gene receives a score of association with the disease, computed from the GWAS data, and a set of biological relationships, given by a network built on prior knowledge.
 Then, the problem becomes finding a functionally-related set of highly-scoring genes.
@@
 For example, some methods restrict the problem to specific types of subnetworks.
-Such is the case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e.
-instances were both a gene and its direct interactors are associated with the disease.
-Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, do not impose such strong constraints, and search for subnetworks interconnecting genes with high association scores.
-However, they differ in their tolerance to the inclusion of low-scoring nodes, and the topology of the solution.
-Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring, but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
+Such is the case of LEAN \cite{gwinner_network-based_2016}, which focuses on ``star'' subnetworks, i.e., instances where both a gene and its direct interactors are associated with the disease.
+Other algorithms, like dmGWAS \cite{jia_dmgwas:_2011} and heinz \cite{dittrich_identifying_2008}, do not impose such strong constraints and search for subnetworks interconnecting genes with high association scores.
+However, they differ in their tolerance to the inclusion of low-scoring nodes and the topology of the solution.
+Lastly, other methods also consider the topology of the network, favoring groups of nodes that are not only high-scoring but also densely interconnected; such is the case of HotNet2 \cite{leiserson_pan-cancer_2015}, SConES \cite{azencott_efficient_2013}, and SigMod \cite{liu_sigmod:_2017}.
 
 In this work, we studied the relevance of network-based approaches to genetics by applying these six network methods to GWAS data.
-They use different interpretations of the omnigenic model, and provide a representative view of the field.
+They use different interpretations of the omnigenic model and provide a representative view of the field.
 We worked on the GENESIS dataset \cite{sinilnikova_genesis:_2016}, a study on familial breast cancer conducted in the French population.
 After a classical GWAS approach, we used these network methods to identify additional breast cancer susceptibility genes.
-Lastly, we carried out a comparison of the solutions obtained by the different methods, and aggregated them to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
+Lastly, we compared the solutions obtained by the different methods and aggregated them to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
+
+\section{Materials and methods}
+
+\subsection{GENESIS dataset, preprocessing, and quality control}
+\label{methods:data}
+
+The GENE Sisters (GENESIS) study investigated risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
+Index cases were patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
+Controls were unaffected colleagues or friends of the cases born around the year of birth of their corresponding case (\textpm{} 3 years).
+We focused on the 2\,577 samples of European ancestry, of which 1\,279 were controls, and 1\,298 were cases.
+The genotyping platform was the iCOGS array, a custom Illumina array designed to study the genetic susceptibility to hormone-related cancers \cite{sakoda_turning_2013}.
+It contained 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
+
+We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
+We also removed a subset of 20 duplicated SNPs in \emph{FGFR2}.
+We excluded the samples with more than 10\% missing genotypes.
+After controlling for relatedness, we excluded 17 additional samples (6 for sample identity error, 6 controls related to other samples, 2 cases related to an index case, and 3 additional controls having a high relatedness score).
+Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
+The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
+
+We looked for population structure that could produce spurious associations.
+A principal component analysis revealed no visual differential population structure between cases and controls (\nameref{sfig:pcs}).
+Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
+
+\subsection{SNP- and gene-based GWAS}
+\label{methods:conventional}
+
+To measure the association between genotype and susceptibility to breast cancer, we performed a per-SNP 1~d.f.~$\chi^2$ allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
+To obtain significant SNPs, we performed a Bonferroni correction to keep the family-wise error rate below 5\%.
+The threshold used was $\frac {0.05} {197083} = 2.54 \times 10^{-7}$.
+
+Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
+More specifically, we mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
+We computed VEGAS2 scores for each gene using only the 10\% of SNPs with the lowest P\=/values among all those mapped to it.
+We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 mapped to at least one SNP.
+Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
+Out of the 197\,083 SNPs remaining after quality control, 164\,037 mapped to at least one of these genes.
+We also performed a Bonferroni correction to obtain significant genes; in this case, the threshold of significance was $\frac {0.05} {32767} = 1.53 \times 10^{-6}$.
+
+\subsection{Network methods}
+
+\subsubsection{Mathematical notations}
+\label{methods:notation}
+In this article, we used undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
+$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
+Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
+When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
+A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
+
+Nodes can be described by properties provided by the topology of the graph.
+We focused on two of those: degree centrality and betweenness centrality.
+The degree centrality, or simply emph{degree}, is the number of edges that a node has.
+The betweenness centrality, or \emph{betweenness}, is the number of times a node participates in shortest paths between two other nodes.
+
+We also used two matrices that describe two different properties of a graph.
+Both matrices are square and have as many rows and columns as nodes are in the network.
+The element $(i,j)$ hence represents a relationship between $v_i$ and $v_j$.
+The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; its diagonal is zero.
+The \emph{degree matrix} $D_G$ is a diagonal matrix that contains the degree of the different nodes.
+
+\subsubsection{Networks}
+\label{methods:networks}
+
+\paragraph{Gene network}
+The mathematical formulations of the different network methods are compatible with any type of biological network (e.g., from protein interactions or gene co-expression).
+Here, we used protein-protein interaction networks (PPIN) for all network methods except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
+We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
+Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the PPIN.
+Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we could map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
+12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
+The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
+
+Additionally, we compared the results obtained on this PPIN with those obtained on a PPIN built using interactions coming from both high-throughput and targeted studies.
+In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
+Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
+
+\paragraph{SNP networks}
+SConES \cite{azencott_efficient_2013} was the only network method designed to handle SNP networks.
+As in gene networks, two SNPs were connected in a SNP network when there was evidence of shared functionality between them.
+Azencott et al.
+\cite{azencott_efficient_2013} proposed three ways of building such networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
+We focused on the GI network using the PPIN described above, as it fitted the scope of this work better.
+However, at different stages, we also compared GI to GS and GM to understand how including the PPIN affects SConES' output.
+For the GM network, we used the mapping described in Section~\ref{methods:snp2gene}.
+In all three, we scored the nodes using the 1~d.f.~\(\chi\)\textsuperscript{2} statistic of association.
+The properties of these three subnetworks are available in \nameref{stab:snp_solutions}.
+
+\subsubsection{High-score subnetwork search algorithms}
+\label{methods:methods}
+
+Genes that contribute to the same function are nearby in the PPIN and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
+Several aspects have to be considered when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
+Unsurprisingly, multiple solutions have been proposed.
+We examined six of them: five that explore the PPIN, and one which explores SNP networks.
+We selected open-source methods that had an implementation available and accessible documentation.
+We summarize their main differences in Table~\ref{tab:method_comparison}.
+We scored both SNPs and genes with the P-values (or transformations) computed in Section~\ref{methods:conventional}.
+
+\begin{table}[!ht]
+  \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
+  \centering
+  \caption{
+  {\bf Summary of the differences between the network methods.}}
+  \begin{tabular}{l|llllllll}
+  {\bf Method} & {\bf Field } & {\bf Nodes} & {\bf Exhaustive} & {\bf Solution} & {\bf Comp.} & {\bf Input} & {\bf Scoring} & {\bf Ref.}\\ \thickhline
+  dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
+  heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
+  HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
+  LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
+  SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
+  SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \cite{liu_sigmod:_2017}\\
+  \end{tabular}
+  \begin{flushleft} \textbf{\emph{Field}}: field in which the algorithm was developed.
+\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
+\textbf{\emph{Exhaustive}}: whether the method explores all the possible solutions given the selected parameters.
+\textbf{\emph{Solution}}: additional properties enforced on the solution, other than containing high scoring, connected nodes.
+\textbf{\emph{Comp.}}: number of connected components in the solution.
+\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
+\textbf{\emph{Scoring}}: how SNP/gene P\=/values were transformed into node scores.
+In the case of heinz, BUM stands for beta-uniform model; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the cumulative distribution function of the standard Normal distribution.
+\textbf{\emph{Ref.}}: original publication featuring the algorithm.
+  \end{flushleft}
+  \label{tab:method_comparison}
+  \end{adjustwidth}
+\end{table}
+
+\begin{description}
+\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
+To that end, it searches candidate solutions using a greedy, ``seed and extend'', heuristic:
+
+\begin{enumerate}
+\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
+\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
+
+\begin{eqnarray} 
+Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
+\end{eqnarray}
+
+where \emph{k} is the number of genes in $S_i$, z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}), and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
+\item Identify neighboring nodes of $S_i$, i.e., nodes at distance \(\le\) \texttt{d}.
+\item Add the neighboring nodes whose inclusion increases Z\textsubscript{m+1} by more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + \texttt{r}).
+\item Repeat 2-4 until no further enlargement is possible.
+\item Add $S_i$ to the list of subnetworks to return. Normalize its Z-score as
+
+\begin{eqnarray}
+Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
+\end{eqnarray}
+
+where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
+\end{enumerate}
+
+DmGWAS carries out this process on every gene in the PPIN.
+We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
+Unless otherwise specified, we used the suggested parameters \texttt{d} $ = 2$ and \texttt{r} $ = 0.1$.
+We used the function \texttt{simpleChoose} to select the solution, which aggregates the top 1\% subnetworks.
+
+\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
+The authors proposed a transformation of the genes' P\=/value into a score that is negative under weak association with the phenotype, and positive under a strong one.
+This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
+Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
+To solve it efficiently, it is re-cast as the Prize-Collecting Steiner Tree Problem, which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
+
+\begin{eqnarray}
+p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
+\end{eqnarray}
+
+were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
+All three are positive quantities.
+Heinz implements the algorithm from Ljubić et al.
+\cite{ljubic_algorithmic_2006} which, in practice, is often fast and optimal, although neither is guaranteed.
+We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
+
+\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}.
+To that end, it considers both the local topology of the PPIN and the nodes' scores.
+An insulated heat diffusion process captures the former: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors and receives heat from its ``hotter'' neighbors while retaining part of its own (hence, \emph{insulated}).
+This process continues until a stationary state is reached, in which the temperature of the nodes does not change anymore, and results in a diffusion matrix $F$.
+$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
+\begin{eqnarray} 
+E = F \operatorname{diag}(w(V)), 
+\end{eqnarray}
+where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
+For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$.
+Hence, $E_{ij}$ can be interpreted as a (non-symmetric) similarity between those two nodes.
+To obtain densely connected solutions, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
+Lastly, HotNet2 evaluates the statistical significance of the solutions by comparing their size to the size of PPINs obtained by permuting the node scores.
+We assigned the initial node scores as in Nakka et al.
+\cite{nakka_gene_2016}, giving a 0 to the genes with a VEGAS2 P-values of association with the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+In the GENESIS dataset, the threshold separating both was a P\=/value of 0.125, which we obtained using a local FDR approach \cite{scheid_twilight;_2005}.
+HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
+Both parameters are set automatically by the algorithm, which is robust to their values \cite{leiserson_pan-cancer_2015}.
+HotNet2 is implemented in Python \cite{hotnet2}.
+
+\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed of one central node and all its interactors \cite{gwinner_network-based_2016}.
+By imposing this restriction, LEAN can exhaustively test all such subnetworks (one per node).
+For a particular star subnetwork of size \emph{m} LEAN performs three steps:
+
+\begin{enumerate}
+\item Rank the P\=/values of the involved nodes as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}. 
+\item Conduct \emph{k} binomial tests to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
+The minimum of these \emph{k} P\=/values is the score of the subnetwork.
+\item Transform this score into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed. 
+\end{enumerate}
+
+We adjust these P\=/values for multiple testing through a Benjamini-Hochberg correction. We used the implementation of LEAN from the \texttt{LEANR} R package \cite{leanr}.
+
+\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem
+
+\begin{eqnarray}
+\label{eq:scones}
+\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }},
+\end{eqnarray}
+
+where \(\lambda\) and \(\eta\) are parameters that control the sparsity and the connectivity of the model.
+The connectivity term penalizes disconnected solutions, with many edges between selected and unselected nodes.
+Given a $\lambda$ and an $\eta$, Eq~\ref{eq:scones} has a unique solution that SConES finds using a graph min-cut procedure.
+As in Azencott et al.
+\cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
+In this case, the selected parameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
+We used the version on SConES implemented in the R package \texttt{martini} \cite{martini}.
+
+\item[{SigMod}] SigMod searches the highest-scoring, most densely connected subnetwork \cite{liu_sigmod:_2017}.
+It addresses an optimization problem similar to that of SConES (Eq~\ref{eq:scones}), but with a different connectivity term that favors solutions containing many edges:
+
+\begin{eqnarray}
+\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
+\end{eqnarray}
+
+As for SConES, this optimization problem can also be solved by a graph min-cut approach.
+
+SigMod presents three important differences with SConES.
+First, it was designed for PPINs.
+Second, it favors solutions containing many edges between the selected nodes.
+SConES, instead, penalizes connections between selected and unselected nodes.
+Third, it explores the grid of parameters differently, and processes their respective solutions.
+Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
+Additionally, that change needs to be larger than a user-specified threshold \texttt{maxjump}.
+Such a large change implies that the network is densely interconnected.
+This results in one candidate solution for each \(\eta\), which is processed by removing any node not connected to any other.
+A score is assigned to each candidate solution by summing their node scores and normalizing by size.
+Finally, SigMod chooses the candidate solution with the highest standardized score, and that is not larger than a user-specified threshold (\texttt{nmax}).
+We used the default parameters \texttt{maxjump} $ = 10$ and \texttt{nmax} $= 300$.
+SigMod is implemented in an R package \cite{sigmod}.
+
+\item[{Consensus}] We built a consensus solution by retaining the genes selected by at least two of the six methods (using SConES GI for SConES).
+It includes any edge between the selected genes in the PPIN.
+\end{description}
+
+We performed all the computations in the cluster described in Section~\ref{methods:computation}.
+
+\subsubsection{Parameter space}
+\label{methods:parameters}
+
+We used the network methods with the parameters recommended by their authors, or with the default parameters in their absence.
+Additionally, we explored the parameter space of the different methods to study how they alter the output.
+
+\begin{description}
+\item[{dmGWAS}] We tested multiple values for \texttt{r} (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and \texttt{d} (1, 2, and 3). 
+\item[{heinz}] We tested multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
+\item[{HotNet2}] We tested different thresholds to decide which genes would receive a score of 0 and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
+\item[{LEAN}] We used the following significance cutoffs for LEAN's P-values (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, and 1).
+\item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores by default (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
+\item[{SigMod}] We tested multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
+\end{description}
+
+\subsubsection{Comparing SNP-methods to gene-methods and vice versa}
+\label{methods:snp2gene}
+
+In multiple steps of this article, we compared the outcome of a method that works on genes with the outcome of one that works on SNPs.
+For this purpose, we used the SNP-gene correspondence described in Section~\ref{methods:conventional}.
+To convert a list of SNPs into a list of genes, we included all the genes mapped to any of those SNPs.
+Conversely, to convert a list of genes into a list of SNPs, we included all the SNPs mapped to any of those genes.
+
+\subsection{Pathway enrichment analysis}
+\label{methods:pathway_enrichment}
+
+We searched for pathways enriched in the gene solutions produced by the above methods.
+We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the \texttt{ReactomePA} R package \cite{Yu2016}.
+The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section~\ref{methods:conventional}).
+We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH): pathways with a BH adjusted P\=/value \textless~0.05 were deemed significant.
+
+\subsection{Benchmark of methods}
+\label{methods:benchmark}
+
+We evaluated multiple properties of the different methods (described in Sections~\ref{methods:algorithm_comparison} and \ref{methods:classifier}) through a 5-fold subsampling setting.
+We applied each method to 5 random subsets of the original GENESIS dataset containing 80\% of the samples (\emph{train set}).
+When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
+We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
+Every method and repetition ran in the same computational settings (Section~\ref{methods:computation}).
+
+\subsubsection{Properties of the solution}
+\label{methods:algorithm_comparison}
+
+We compared the runtime, the number of selected features (genes or SNPs), and the stability (sensitivity to the choice of train set) of the different network methods.
+Nogueira and Brown \cite{nogueira_measuring_2016} proposed quantifying a method's stability using the Pearson correlation between the genes selected on different subsamples.
+This correlation was calculated between vectors with the length of the total number of features, containing a 0 at position $i$ if feature $i$ was not selected and a 1 if it was.
+
+\subsubsection{Classification accuracy of selected SNPs}
+\label{methods:classifier}
+
+A desirable solution offers good predictive power on the unseen test samples.
+We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, which searches for a small subset of SNPs that provides good classification accuracy at predicting the outcome (case/control).
+The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}).
+The active set was a plausible, more sparse solution with comparable predictive power to the original solution.
+The L1 penalty was set by cross-validation, choosing the value that minimized misclassification error.
+
+We applied each network method to each train set.
+Then, we trained the classifier on the same train set using only the selected SNPs.
+When the method retrieved a list of genes, we proceeded as explained in Section~\ref{methods:snp2gene}.
+Lastly we evaluated the sensitivity and the specificity of the classifier on the test set.
+To obtain a baseline, we also trained the classifier on all the SNPs of the train set.
+
+We did not expect a linear model on selected SNPs to separate cases from controls well.
+Indeed, the lifetime cumulative incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
+However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
+
+\subsection{Comparison to state-of-the-art}
+\label{methods:bcac}
+
+An alternative way to evaluate the methods is by comparing their solutions to an external dataset.
+For that purpose, we used the 153 genes associated to familial breast cancer on DisGeNET \cite{pinero_disgenet:_2017}.
+Across this article, we refer to these genes as \emph{breast cancer susceptibility genes}.
+
+Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries.
+BCAC included 13\,250\,641 SNPs genotyped or imputed on 228\,951 women of European ancestry, mostly from the general population \cite{Michailidou2017}.
+Through imputation, BCAC includes more SNPs than the iCOGS array used for GENESIS (Section~\ref{methods:data}).
+However, in all the comparisons in this paper we focused on the SNPs that passed quality control in GENESIS.
+Hence, we used the same Bonferroni threshold as in Section~\ref{methods:conventional} to determine the significant SNPs in BCAC.
+We also computed gene-scores in the BCAC data using VEGAS2, as in Section~\ref{methods:data}.
+In this case, we did use the summary statistics of all 13\,250\,641 available SNPs and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
+Since these genotypes did not include chromosome X, we excluded it from this analysis.
+All comparisons included only the genes in common between GENESIS and BCAC, so we used a different Bonferroni threshold ($1.66 \times 10^{-6}$) to call gene significance.
+
+\subsection{Network rewirings}
+\label{methods:rewiring}
+
+Rewiring the PPIN while preserving the number of edges of each gene allowed to study the impact of the topology on the output of network methods.
+Indeed, the edges lose their biological meaning while the topology of the network is conserved.
+We produced 100 such rewirings by randomly swapping edges in the PPIN.
+We still scored the genes as described in Section~\ref{methods:methods}.
+We only applied only four methods on the rewirings: heinz, dmGWAS, LEAN, and SigMod.
+We excluded HotNet2 and SConES since they took notably longer to run.
+
+\subsection{Computational resources}
+\label{methods:computation}
+
+We ran all the computations on a Slurm cluster, running Ubuntu 16.04.2 on the nodes.
+The CPU models on the nodes were Intel Xeon CPU E5-2450 v2 at 2.50GHz and Intel Xeon E5-2440 at 2.40GHz.
+The nodes running heinz and HotNet2 had 20GB of memory; the ones running dmGWAS, LEAN, SConES, and SigMod, 60GB.
+For the benchmark (Section~\ref{methods:benchmark}), we ran each of the methods on the same Ubuntu 16.04.2 node, with a CPU Intel Xeon E5-2450 v2 at 2.50GHz, and 60GB of memory.
+
+\subsection{Code and data availability}
+\label{methods:code}
+
+We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and running six different network methods.
+We created a pipeline with a clear interface that should work on any GWAS dataset for each of those processes.
+They are compiled in \url{https://github.com/hclimente/gwas-tools}.
+The code that applies them to GENESIS, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
+We deposited all the produced gene solutions on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
+
+Summary statistics for SNPs and genes are available at \url{https://github.com/hclimente/genewa}.
+We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact \href{mailto:nadine.andrieu@curie.fr}{nadine.andrieu(at)curie.fr}.
 
@@
 We conducted association analyses in the GENESIS dataset (Section~\ref{methods:data}) at both SNP and gene levels (Section~\ref{methods:conventional}).
-Two genomic regions had a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
-The former overlaps with the gene \emph{FGFR2}; the latter with \emph{CASC16} and the protein-coding gene \emph{TOX3}.
-Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
-In GENESIS only \emph{FGFR2} was significantly associated with breast cancer at the gene-level (\nameref{sfig:snp_gene_manhattan}B).
-
-Closer examination reveals two other regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
-Both of them have been associated to breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
-We applied an L1-penalized logistic regression using all GENESIS genotypes as input, and the phenotype (cancer/healthy) as outcome (Section~\ref{methods:classifier}).
-The algorithm selected 100 SNPs, both from all aforementioned regions and new ones (\nameref{sfig:snp_gene_manhattan}C).
-Yet, it is unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
-Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map to (Section~\ref{methods:snp2gene}) are not interconnected in the protein-protein interaction network (PPIN, Section~\ref{methods:networks}).
-Moreover, the classification performance of the model is low (sensitivity = 55\%, specificity = 55\%, Section \ref{methods:benchmark}).
-Together, these issues motivate exploring network methods, which consider not only statistical association, but also the location of each gene in a PPIN to find susceptibility genes.
+At the SNP level, two genomic regions had a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
+The former overlaps with the gene \emph{FGFR2}, the latter with \emph{CASC16} and the protein-coding gene \emph{TOX3}.
+Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
+At the gene level, only \emph{FGFR2} was significantly associated with breast cancer (\nameref{sfig:snp_gene_manhattan}B).
+
+Closer examination revealed two other regions (3p24 and 8q24) having low, albeit not genome-wide significant, P\=/values.
+Both of them have been associated with breast cancer susceptibility in the past \cite{brisbin_meta-analysis_2011,search_newly_2009}.
+We applied an L1-penalized logistic regression using all GENESIS genotypes as input and the phenotype (cancer/healthy) as the outcome (Section~\ref{methods:classifier}).
+The algorithm selected 100 SNPs, both from all regions mentioned above and new ones (\nameref{sfig:snp_gene_manhattan}C).
+However, it was unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
+Moreover, and in opposition to what would be expected under the omnigenic model, the genes to which these SNPs map (Section~\ref{methods:snp2gene}) were not interconnected in the protein-protein interaction network (PPIN, Section~\ref{methods:networks}).
+Moreover, the classification performance of the model was low (sensitivity = 55\%, specificity = 55\%, Section~\ref{methods:benchmark}).
+Together, these issues motivated exploring network methods, which consider not only statistical association but also the location of each gene in a PPIN to find susceptibility genes.
 
@@
   \includegraphics[width=\textwidth]{./figures/figure_1.png}
-  \caption{{\bf Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.
-} As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it was excluded.
-Unless indicated otherwise, results refer to genes, except for SConES GI which are at the SNP-level.
-\textbf{(A)}~Overlap between the genes selected by each of the methods, measured by Pearson correlation between indicator vectors (Sections \ref{methods:algorithm_comparison} and \ref{methods:snp2gene}).
+  \caption{\textbf{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.} As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it is not shown.
+Unless indicated otherwise, results refer to SNPs for SConES GI, and to genes for the other methods.
+\textbf{(A)}~Overlap between the genes selected by each method, measured by Pearson correlation between indicator vectors (Sections~\ref{methods:algorithm_comparison} and \ref{methods:snp2gene}).
 \textbf{(B)}~Distribution of VEGAS2 P-values of the genes in the PPIN not selected by any network method (12\,213), and of those selected by 1 (575), 2 (73), or 3 (20) methods.
@@
 \textbf{(D)}~Manhattan plots of SNPs/genes; in black, the method's solution.
-The Bonferroni threshold is indicated by a red line (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
+The red line indicates the Bonferroni threshold (2.54~\texttimes{}~10\textsuperscript{-7} for SNPs, 1.53~\texttimes{}~10\textsuperscript{-6} for genes).}
   \label{fig:solution_overview}
@@
 
-These solutions differ in many aspects, making it hard to draw joint conclusions.
-For starters, the overlap between the genes featured in each solution is quite small (Fig~\ref{fig:solution_overview}A).
-However, the methods tend to agree on the genes with the strongest signal: genes selected by more methods tended to have lower P-value of association (Fig~\ref{fig:solution_overview}B).
-
-Another prominent difference is the solution size: the largest solution, produced by HotNet2, contains 440 genes, while heinz's contains only 4 genes.
-While SConES GI did not recover any protein coding gene, working with SNP networks rather than gene networks allowed it to retriev four subnetworks in intergenic regions, and another one overlapping an RNA gene (\emph{RNU6-420P}).
-
-The topologies of the five solutions differ as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality  and the number of connected components (Table~\ref{tab:gene_solutions}).
-Three methods yield more than one connected component: SConES, as described above, SigMod, and HotNet2.
+These solutions differed in many aspects, making it hard to draw joint conclusions.
+For starters, the overlap between the genes featured in each solution was relatively small (Fig~\ref{fig:solution_overview}A).
+However, the methods tended to agree on the genes with the strongest signal: genes selected by more methods tended to have lower P-value of association (Fig~\ref{fig:solution_overview}B).
+
+Another major difference was the solution size: the largest solution, produced by HotNet2, contained 440 genes, while heinz's contained only 4 genes.
+While SConES GI did not recover any protein coding gene, working with SNP networks rather than gene networks allowed it to retrieve four subnetworks in intergenic regions and another subnetwork overlapping an RNA gene (\emph{RNU6-420P}).
+
+The topologies of the five solutions differed as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality and the number of connected components (Table~\ref{tab:gene_solutions}).
+Three methods yielded more than one connected component: SConES, as described above, SigMod, and HotNet2.
 HotNet2 produced 135 subnetworks, 115 of which have fewer than five genes.
-The second largest subnetwork (13 nodes) contains the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM} (Section \ref{methods:bcac}).
+The second largest subnetwork (13 nodes) contained the two breast cancer susceptibility genes \emph{CASP8} and \emph{BLM} (Section~\ref{methods:bcac}).
 
@@
   \end{tabular}
-  \begin{flushleft} \textbf{\# genes}: number of genes selected out of those that are part of the PPIN; for SConES GI the total number of genes, including RNA genes, was added in parentheses.
-{\bf \# components }: number of connected components.
-\textbf{\mean{Betweenness}}: median betweenness of the selected genes in the PPIN.
-\textbf{\median{P}\textsubscript{gene}}: median VEGAS2 P\=/value of the selected genes.
-\textbf{\# genes in consensus}: number of genes in common between the method's solution, and the 93 genes in the consensus solution.
+  \begin{flushleft} \textbf{\# genes}:~number of genes selected out of those that are part of the PPIN; for SConES GI, the total number of genes, including RNA genes, was added in parentheses.
+\textbf{\# components}:~number of connected components.
+\textbf{\mean{Betweenness}}:~median betweenness of the selected genes in the PPIN.
+\textbf{\median{P}\textsubscript{gene}}:~median VEGAS2 P\=/value of the selected genes.
+\textbf{\# genes in consensus}:~number of genes in common between the method's solution and the 93 genes in the consensus solution.
   \end{flushleft}
@@
 
-Lastly, a pathway enrichment analysis (Section \ref{methods:pathway_enrichment}) also showed similarities and differences between the solutions of the different methods.
+Lastly, a pathway enrichment analysis (Section~\ref{methods:pathway_enrichment}) also showed similarities and differences between the methods' solutions.
 It linked different parts of SigMod's solution to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
@@
 This illustrates the trade-off between controlling for type I error and biological relevance.
-However, there were nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
+However, there are nuances between solutions in this regard: heinz strongly favored genes with lower P\=/values, while dmGWAS was less conservative (median VEGAS2 P\=/values 0.0012 and 0.19, respectively); SConES tended to select whole LD-blocks; and HotNet2 and SigMod were less likely to select low scoring genes.
 
 Additionally, the solutions presented other desirable properties.
-First, four of them were enriched in known breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value$<$0.03).
-Second, the genes in three solutions displayed on average a significantly higher betweenness centrality than the rest of the genes (dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value $<$ 1.4 \texttimes{} 10\textsuperscript{-21}).
+First, four of them were enriched in known breast cancer susceptibility genes (dmGWAS, heinz, HotNet2, and SigMod, Fisher's exact test one-sided P\=/value~$<$~0.03).
+Second, the genes in three solutions displayed, on average, a significantly higher betweenness centrality than the rest of the genes (dmGWAS, HotNet2, and SigMod, Wilcoxon rank-sum test P\=/value~$<$~1.4~\texttimes{}~10\textsuperscript{-21}).
 This agrees with the notion that disease genes are more central than other non-essential genes \cite{pinero_uncovering_2016}, an observation that holds in breast cancer (one-tailed Wilcoxon rank-sum test P\=/value~=~2.64 \texttimes{} 10\textsuperscript{-5} when comparing the betweenness of known susceptibility genes versus the rest).
-Interestingly, SConES selected SNPs that are also more central than the average SNP (\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
+Interestingly, the SNPs in SConES' solution were also more central than the average SNP (\nameref{stab:snp_solutions}), suggesting that causal SNPs are also more central than non-associated SNPs.
 
@@
 
-Despite their shared properties, the differences between the solutions of the different methods suggested that each of them captured different aspects of cancer susceptibility.
+Despite their shared properties, the differences between the solutions suggested that each of them captured different aspects of cancer susceptibility.
 Indeed, out of the 668 genes that were selected by at least one method, only 93 were selected by at least two, 20 by three, and none by four or more.
-Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig \ref{fig:solution_overview}B), a relationship that plateaus at 2.
-Hence, to leverage on their strengths and compensate their respective weaknesses, we built a consensus solution that captured the genes shared among at least two solutions (Section~\ref{methods:methods}).
+Encouragingly, the more methods selected a gene, the higher its association score to the phenotype (Fig~\ref{fig:solution_overview}B), a relationship that plateaued at 2.
+Hence, to leverage their strengths and compensate for their respective weaknesses, we built a consensus solution using the genes shared among at least two solutions (Section~\ref{methods:methods}).
 This solution (Fig~\ref{fig:consensus}) contained 93 genes and exhibited the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
@@
   \includegraphics[width=.9\linewidth]{./figures/figure_3.pdf}
-  \caption{ {\bf Consensus solution on GENESIS (Section~\ref{methods:methods}) }.
-\textbf{(A)} ~Manhattan plot of genes; in black, the ones in the consensus solution.
-The Bonferroni threshold is indicated by a red line (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{(B)} Consensus network.
+  \caption{ \textbf{Consensus solution on GENESIS (Section~\ref{methods:methods}).} \textbf{(A)}~Manhattan plot of genes; in black, the ones in the consensus solution.
+The red line indicates the Bonferroni threshold (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
+\textbf{(B)}~Consensus network.
 Each gene is represented by a pie chart, which shows the methods that selected it.
-We enlarged the two most central genes (\emph{COPS5} and \emph{OFD1}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
-\textbf{(C)} The nodes are in the same disposition as in panel B, but every gene name is indicated.
-We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.}
+We enlarged the two most central genes (\emph{COPS5} and \emph{OFD1}), the known breast cancer susceptibility genes, and the BCAC-significant genes (Section~\ref{methods:bcac}).
+\textbf{(C)}~The nodes are in the same disposition as in panel B, but we indicated every gene name.
+We colored in pink the names of known breast cancer susceptibility genes and BCAC-significant genes.}
   \label{fig:consensus}
@@
 The former is supported by genes like \emph{MRPS30} (VEGAS2 P\=/value~=~0.001), which encode a mitochondrial ribosomal protein and was also linked to breast cancer susceptibility \cite{quigley_5p12_2014}.
-Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition proposed as a therapeutic target.
+Interestingly, increased mitochondrial translation has been found in cancer cells \cite{Yu2016Repositioning}, and its inhibition was proposed as a therapeutic target.
 With regards to the attenuation phase of heat shock response, it involved three Hsp70 chaperones: HSPA1A, HSPA1B, and HSPA1L.
 The genes encoding these proteins are all near each other at 6p21, in the region known as HLA.
-In fact, out of the 22 SNPs that map to any of these three genes, 9 mapped to all of them, and 4 to two, making it hard to disentangle their effects.
+In fact, out of the 22 SNPs mapped to any of these three genes, 9 mapped to all three, and 4 to two, which made it hard to disentangle their effects.
 \emph{HSPA1A} was the most strongly associated gene (VEGAS2 P\=/value~=~8.37~\texttimes{}~10\textsuperscript{-4}).
 
-Topologically, the consensus consisted of a connected component composed of 49 genes, and multiple smaller subnetworks (Fig~\ref{fig:consensus}B and C).
-Among the latter, 19 genes were in subnetworks containing a single gene or two connected nodes, implying that they did not have a consistently altered neighborhood, but were strongly associated themselves and hence picked by at least two methods.
-The large connected component contained genes that are highly central in the PPIN, a property which is weakly anti-correlated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
-This suggested that these genes were selected because they were on the shortest path between another two highly associated genes.
-In view of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistently with the omnigenic model of disease \cite{boyle_expanded_2017}.
-For instance, the most central node in the consensus solution is \emph{COPS5}, a component of the COP9 signalosome which regulates multiple signaling pathways.
+Topologically, the consensus consisted of a connected component composed of 49 genes and multiple smaller subnetworks (Fig~\ref{fig:consensus}B and C).
+Among the latter, 19 genes were in subnetworks containing a single gene or two connected nodes.
+This implied that they did not have a consistently altered neighborhood but were strongly associated themselves and hence picked by at least two methods.
+The large connected component contained genes that are highly central in the PPIN.
+This property weakly anticorrelated with the P\=/value of association to the disease (Pearson correlation coefficient~=~-0.26, \nameref{sfig:consensus_stats}).
+This anticorrelation suggested that these genes were selected because they were on the shortest path between two high scoring genes.
+Because of this, we hypothesize that highly central genes might contribute to the heritability through alterations of their neighborhood, consistent with the omnigenic model of disease \cite{boyle_expanded_2017}.
+For instance, the most central node in the consensus solution was \emph{COPS5}, a component of the COP9 signalosome that regulates multiple signaling pathways.
 \emph{COPS5} is related to multiple hallmarks of cancer and is overexpressed in multiple tumors, including breast and ovarian cancer \cite{liu_jab1_cops5_2018}.
-Despite its lack of association in GENESIS or in studies conducted by the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (VEGAS2 P\=/value of 0.22 and 0.14 respectively), its neighbors in the consensus solution had consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
+Despite its lack of association in GENESIS or in studies conducted by the Breast Cancer Association Consortium (BCAC) \cite{Michailidou2017} (VEGAS2 P\=/value of 0.22 and 0.14, respectively), its neighbors in the consensus solution had consistently low P\=/values (median VEGAS2 P\=/value~=~0.006).
 
@@
 We compared the results obtained with different network methods to the European sample of BCAC, the largest GWAS to date on breast cancer (Section~\ref{methods:bcac}).
-Although BCAC case-control studies do not necessarily target cases with a familial history of breast cancer, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
-This shared genetic architecture, together with BCAC's scale (90 times more samples than GENESIS) provided a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
+Although BCAC case-control studies do not necessarily target cases with a familial history of breast cancer like GENESIS does, this comparison is pertinent since we expect a shared genetic architecture at the gene level, at which most network methods operate.
+Together with BCAC's scale (90 times more samples than GENESIS), this shared genetic architecture provided a reasonable counterfactual of what we would expect if GENESIS had a larger sample size.
 We computed a gene association score on BCAC (Section~\ref{methods:bcac}).
@@
 The gene-based methods achieved comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (\nameref{sfig:additional-benchmarks}A).
-Interestingly, while SConES GI, at the SNP-level, achieved a similar recall (8.6\%), it showed a much higher precision (47.3\%).
+Interestingly, while SConES GI achieved a similar recall at the SNP-level (8.6\%), it showed a much higher precision (47.3\%).
 
@@
 
-We compared the six network methods in a 5-fold subsampling setting (Section \ref{methods:benchmark}).
-This allowed us to measure five properties (Fig~\ref{fig:benchmark}): size of the solution; sensitivity and specificity of an L1-penalized logistic regression classifier on the selected SNPs; stability; and computational runtime.
+We compared the six network methods in a 5-fold subsampling setting (Section~\ref{methods:benchmark}).
+In this comparison we measured properties (Fig~\ref{fig:benchmark} and \nameref{sfig:additional-benchmarks}): the size of the solution; the sensitivity and the specificity of an L1-penalized logistic regression classifier on the selected SNPs; the stability of the methods; and their computational runtime.
 The solution size varies greatly between the different methods (Fig~\ref{fig:benchmark}A).
-Heinz produced the smallest solutions, with an average of 182 selected SNPs (Section~\ref{methods:snp2gene}).
-The largest solutions came from SConES GI (6\,256.6 SNPs), and dmGWAS (4\,255.0 SNPs).
+Heinz produced the smallest solutions, with an average of 182 selected SNPs (Section~\ref{methods:snp2gene}) while the largest ones came from SConES GI (6\,256.6 SNPs) and dmGWAS (4\,255.0 SNPs).
 LEAN did not produce any solution in any of the subsamples.
-Using different combinations of parameters (Section \ref{methods:parameters}), we computed how good each of the methods was at recovering the results of a conventional GWAS on BCAC (Section \ref{methods:bcac}, Fig~\ref{fig:benchmark}B).
-SConES exhibits the largest area under the curve, since, when $\lambda = 0$ (network topology is disregarded), it is equivalent to a Bonferroni correction.
-Between the remaining network methods, they have similar areas under the curve, with heinz being the one with the largest area.
-
-To determine whether the selected SNPs could be used for patient classification, we computed the performance of the classifier on \emph{test sets} (\nameref{sfig:additional-benchmarks}B).
-The different classifiers displayed similarly poor sensitivities and specificities, all in the 0.52 -- 0.56 range.
-Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming only at minimizing prediction error.
+
+To determine whether the selected SNPs could predict cancer susceptibility, we computed the classifiers' performances on \emph{test sets} (\nameref{sfig:additional-benchmarks}B).
+The different classifiers displayed similarly low sensitivities and specificities, all in the 0.52 -- 0.56 range.
+Interestingly, the classifier trained on all the SNPs had a similar performance, despite being the only method aiming to minimize prediction error.
 Of course, although these performances were low, we did not expect to separate cases from controls well using exclusively genetic data \cite{deloscamposComplexTraitPredictionEra2018}.
@@
   \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
-  \caption{{ \bf Comparison of network-based GWAS methods on GENESIS.
-} Each method was run 5 times on a random subset containing 80\% of the samples, and tested on the remaining samples (Section \ref{methods:benchmark}).
-As LEAN did not select any gene, it was excluded from all panels except \textbf{D}.
-\textbf{(A)}~Number of SNPs selected by each method and number of SNPs in the active set found by the classifier(Section~\ref{methods:classifier}).
+  \caption{\textbf{Comparison of network methods on GENESIS.} Each method was run 5 times on a random subset containing 80\% of the samples and tested on the remaining samples (Section~\ref{methods:benchmark}).
+As LEAN did not select any gene, we excluded it from panels A and B.
+\textbf{(A)}~Number of SNPs selected by each method and number of SNPs in the active set (i.e., the number of SNPs selected by the classifier, Section~\ref{methods:classifier}).
 Points are the average over the 5 runs; the error bars represent the standard error of the mean.
-A grey diagonal line with slope 1 is added for comparison, indicating the upper bound of the active set (i.e.
-the number of SNPs in the solution).
-For reference, the active set of Lasso using all the SNPs included, on average, 154\,117.4 SNPs.
-\textbf{(B)} True positive rate and true negative rate, using significant SNPs (for SConES and $\chi^2$ + Bonferroni) and genes (for the remaining methods) in BCAC (Section~\ref{methods:bcac}) as true positives, of multiple parameter combinations for different methods (Section \ref{results:parameters}).
-\textbf{(C)}~Pairwise Pearson correlations of the solutions produced by different methods.
+A grey diagonal line with slope 1 is added for comparison, indicating the upper bound of the active set.
+For reference, the active set of the classifier using all the SNPs as input included, on average, 154\,117.4 SNPs.
+\textbf{(B)}~Pairwise Pearson correlations of the solutions produced by different methods.
 A Pearson correlation of 1 means the two solutions are the same.
-A Pearson correlation of 0 means that there is no SNP in common between the two solutions.
-\textbf{(D)}~Runtime of the evaluated methods, by type of network used (PPIN or SNP).
-For gene-based methods, inverted triangles represent the runtime of the algorithm itself, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.}
+\textbf{(C)}~Runtime of the evaluated methods, by type of network used (PPIN or SNP).
+For gene-based methods, inverted triangles represent the runtime of the algorithm alone, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
+\textbf{(D)}~True positive rate and true negative rate of the methods, obtained using different parameter combinations (Section~\ref{results:parameters}).
+We used as true positives BCAC-significant SNPs (for SConES and $\chi^2$ + Bonferroni) and genes (for the remaining methods, Section~\ref{methods:bcac}).
+We used the whole dataset in this panel.}
   \label{fig:benchmark}
@@
 Another desirable quality of a selection algorithm is the stability of the solution with respect to small changes in the input (Section~\ref{methods:algorithm_comparison}).
-Heinz was highly stable in our benchmark, while the other methods displayed similarly low stabilities (Fig~\ref{fig:benchmark}C).
-
-In terms of computational runtime, the fastest method was heinz (Fig~\ref{fig:benchmark}D), which returned a solution in a few seconds.
+Heinz was highly stable in our benchmark, while the other methods displayed similarly low stabilities (Fig~\ref{fig:benchmark}B).
+
+In terms of computational runtime, the fastest method was heinz (Fig~\ref{fig:benchmark}C), which returned a solution in a few seconds.
 HotNet2 was the slowest (3 days and 14 hours on average).
 Including the time required to compute the gene scores, however, slowed down considerably gene-based methods; on this benchmark, that step took on average 1 day and 9.33 hours.
-Including this first step, it therefore took 5 days on average for HotNet2 to produce a result.
-
-\subsection{Network topology and association scores matter, and might lead to ambiguous results}
+Including this first step, it took 5 days on average for HotNet2 to produce a result.
+
+Using different combinations of parameters (Section~\ref{methods:parameters}), we computed how good each of the methods was at recovering the results of a conventional GWAS on BCAC (Section~\ref{methods:bcac}, Fig~\ref{fig:benchmark}D).
+SConES exhibits the largest area under the curve since, when $\lambda = 0$ (i.e., network topology is disregarded), it is equivalent to a Bonferroni correction.
+The remaining network methods have similar areas under the curve, with heinz having the largest one.
+
+\subsection{Network topology and association scores matter and might lead to ambiguous results}
 \label{results:drawbacks}
 
-As shown above, and despite their similarities, the network methods produced remarkably different solutions.
-This is due to the fact that each of them models the problem differently.
-Importantly, understanding which assumptions they made allowed to understand the results more in depth.
-For instance, the fact that LEAN did not return any gene implied that there was no gene such that both itself and its environment were on average strongly associated with the disease.
+As shown above, and despite their similarities, the different ways of modeling the problem led to remarkably different solutions.
+Importantly, understanding which assumptions the methods made allowed us to understand the results more in depth.
+For instance, the fact that LEAN did not return any gene implied that there was no gene such that both itself and its environment were, on average, strongly associated with the disease.
 
 In the GENESIS dataset, heinz's solution was very conservative, providing a small solution with the lowest median P\=/value (Table~\ref{tab:gene_solutions}).
-By repeatedly selecting this compact solution, heinz was the most stable method (Fig~\ref{fig:benchmark}C).
+By repeatedly selecting this compact solution, heinz was the most stable method (Fig~\ref{fig:benchmark}B).
 Its conservativeness stemmed from its preprocessing step, which modeled the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (\nameref{sfig:snp_gene_manhattan}B), only 36 of all the genes retained a positive score after that transformation.
-Yet, this small solution did not provide much insight into the susceptibility mechanisms to cancer.
-Importantly, it ignored genes that were associated to cancer in this dataset like \emph{FGFR2}.
+Due to the limited signal at the gene level in this dataset (\nameref{sfig:snp_gene_manhattan}B), only 36 genes retained a positive score after that transformation.
+However, this small solution did not provide much insight into the susceptibility mechanisms to cancer.
+Importantly, it ignored genes that were associated with cancer in this dataset, like \emph{FGFR2}.
 
 On the other end of the spectrum, dmGWAS, HotNet2, and SigMod produced large solutions.
-dmGWAS' solution was the lowest scoring solution on average.
-This was due to the greedy framework it uses, which had a bias for larger solutions \cite{nikolayeva_network_2018}.
-It considered all nodes at distance 2 of the examined subnetwork, and accepted a weakly associated gene if it was linked to another, high scoring one.
-This was exacerbated when the results of successive greedy searches are aggregated, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
+DmGWAS' solution was the lowest scoring solution on average because of its greedy framework, which is biased towards larger solutions \cite{nikolayeva_network_2018}.
+It considered all nodes at distance 2 of the examined subnetwork and accepted a weakly associated gene if it was linked to another, high scoring one.
+Aggregating the results of successive greedy searches exacerbates this bias, leading to a large, tightly connected cluster of unassociated genes (Fig~\ref{fig:issues}A).
 This relatively low signal-to-noise ratio combined with the large solution requires additional analyses to draw conclusions, such as enrichment analyses.
 In the same line, HotNet2's solution was even harder to interpret, being composed of 440 genes divided into 135 subnetworks.
-Lastly, SigMod missed some of the highest scoring, breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
+Lastly, SigMod missed some of the highest scoring breast cancer susceptibility genes in the dataset, like \emph{FGFR2} and \emph{TOX3}.
 
@@
   \includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
-  \caption{{\bf Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
+  \caption{\textbf{Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
 \textbf{(B)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its centrality degree.
 \textbf{(C)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
-We highlighted the genes selected by each method, and the ones selected by more than one (``Consensus'').
+We highlighted the genes selected by each method and the ones selected by more than one (``Consensus'').
 We labeled the three most central genes that were picked by any method.
-\textbf{(D)}~Overlap between the solutions of SConES GS, GM or GI in the different genomic regions.
-SNPs that were not selected in the studied network, but were selected in another one, are displayed in background color.
-}
+\textbf{(D)}~Overlap between the solutions of SConES GS, GM, or GI.
+Barplots are colored based on whether the SNPs map to a gene or not (Section~\ref{methods:snp2gene}).}
   \label{fig:issues}
@@
 We studied random rewirings of the PPIN that preserved node centrality (Section~\ref{methods:rewiring}).
-In this setting, network methods favored central genes, as they often connect high scoring nodes (Fig~\ref{fig:issues}B).
-This was despite the fact that highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
-This was especially the case of SigMod (\nameref{sfig:rewirings}), which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3} and \emph{FN1}.
-As we showed in Section \ref{results:consensus}, and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
+In this setting, network methods favored central genes (Fig~\ref{fig:issues}B) even though highly central genes often had no association to breast cancer susceptibility (Fig~\ref{fig:issues}C).
+We found this bias especially in SigMod (\nameref{sfig:rewirings}), which selected three highly central, unassociated genes in both the PPIN and in many of the random rewirings: \emph{COPS5}, \emph{CUL3}, and \emph{FN1}.
+However, as we showed in Section~\ref{results:consensus} and will show in \ref{results:stable-consensus}, there is evidence in the literature of the contribution of the first two to breast cancer susceptibility.
 With regards to \emph{FN1}, it encodes a fibronectin, a protein of the extracellular matrix involved in cell adhesion and migration.
-Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it is negatively correlated with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.
-
-By virtue of using a SNP subnetwork, SConES analyzed each SNP in their functional context.
-It could therefore select SNPs located in genes not included in the PPIN, as well as SNPs in non-coding regions or in non-interacting genes.
-We compared the solution of SConES in the GI network (using PPIN information), to the one using only positional information (GS network), or positional and gene annotations (GM network).
+Overexpression of \emph{FN1} has been observed in breast cancer \cite{Ioachim2002}, and it anticorrelates with poor prognosis in other cancer types \cite{Yi2016,Sponziello2016}.
+
+By using a SNP subnetwork, SConES analyzed each SNP in its functional context.
+Therefore, it could select SNPs located in genes not included in the PPIN and in non-coding regions.
+We compared the solution of SConES in the GI network (using PPIN information), to the one using only positional information (GS network) and to the one using positional and gene annotations (GM network).
 Importantly, SConES produced similar results on the GS and GM networks (\nameref{sfig:pearson_methods}).
-While the solutions on those two greatly overlap with SConES GI's, they contained additional gene-coding segments (Fig~\ref{fig:issues}C).
-In fact the formers' solutions were chromosome regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
-On top of those SConES GS selected region 8q24 (\emph{POU5F1B} \cite{breyer_expressed_2014}).
+While the solutions on those two considerably overlap with SConES GI's, they contained additional gene-coding segments (Fig~\ref{fig:issues}C).
+In fact, both SConES GS and GM selected chromosome regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
+In addition to those, SConES GS selected region 8q24, also linked to breast cancer (\emph{POU5F1B} \cite{breyer_expressed_2014}).
 
@@
 
-We explored the parameter space of the different methods by running them under different combinations of parameters (Section \ref{methods:parameters}).
-In agreement with their formulations (Section \ref{methods:methods}), larger values of certain parameters produced less stringent solutions (\nameref{sfig:parameters}A): for HotNet2 and heinz, this is the threshold above which genes are assigned a non-zero score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
-Two parameters had the opposite effect (larger is more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES's $\eta$, where higher values produce smaller solutions.
-However, two of the parameters did not have the expected effect: dmGWAS' \texttt{r}, which controls the minimum increment in the score required to add an additional gene; and SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''.
+We explored methods' parameter space by running them under different combinations of parameters (Section~\ref{methods:parameters}).
+In agreement with their formulations (Section~\ref{methods:methods}), larger values of specific parameters produced less stringent solutions (\nameref{sfig:params_size}A): for HotNet2 and heinz, this is the threshold above which genes receive a positive score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
+Two parameters had the opposite effect (the larger, the more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES' $\eta$, where higher values produce smaller solutions.
+However, two of the parameters did not have the expected effect: dmGWAS' \texttt{r}, which controls the minimum increment in the score required to add a gene; and SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''.
 In both cases, the size of the solution was very similar across the different values.
-Despite the differences in size, the size of the solutions was relatively robust to the choice of parameters (\nameref{sfig:parameters}B).
-
-We computed the Pearson correlation between the different solutions as in Section \ref{methods:algorithm_comparison} to study how the parameters affected which genes and SNPs were selected (\nameref{sfig:parameters}C).
-This showed that dmGWAS and SigMod were robust to certain parameters: dmGWAS' output was mostly determined by the parameter $d$, rather than $r$; SigMod's by $nmax$, rather than $maxjump$.
-
-SConES presented an interesting case in terms of feature selection: most of the explored combinations of parameters led to trivial solutions (either all the SNPs, or none of them were included) (\nameref{sfig:parameters}A).
-To explore a more meaningful parameter space, in our experiments we selected the parameters in two rounds.
+Despite the differences in size, the solutions' size was relatively robust to the choice of parameters (\nameref{sfig:params_size}B).
+
+We computed the Pearson correlation between the different solutions as in Section~\ref{methods:algorithm_comparison} to study how the parameters affected which genes and SNPs were selected (\nameref{sfig:params_pearson}).
+This analysis showed that dmGWAS and SigMod were robust to two parameters: the parameter \texttt{d} determined dmGWAS' output more than \texttt{r}; for SigMod, it was $nmax$ rather than $maxjump$.
+
+SConES presented an interesting case in terms of feature selection: most of the explored combinations of parameters led to trivial solutions (they included either all the SNPs or none of them) (\nameref{sfig:params_pearson}).
+To explore a more meaningful parameter space, we selected the parameters in two rounds in our experiments.
 First, we explored the whole sample space.
-Then, we focused in a range of $\eta$ and $\lambda$ 1.5 orders of magnitude above and below the best parameters, respectively.
-This second parameter space was more diverse, and allowed to find more interesting solutions.
+Then, we focused on a range of $\eta$ and $\lambda$ 1.5 orders of magnitude above and below the best parameters, respectively.
+This second parameter space was more diverse, which allowed to find more interesting solutions.
 
@@
 
-Most of the network methods, including the consensus, were highly unstable (Fig~\ref{fig:benchmark}), raising questions about the reliability of the results.
-We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section \ref{methods:benchmark}).
-Such a network is expected to capture the subnetworks more often found altered, and hence should be more resistant to noise.
-We used only genes selected in at least 7 of the solutions, which corresponded to 1\% of all genes selected at least once.
-The resulting stability-based consensus was composed of 68 genes (Fig \ref{fig:stable-consensus}).
-This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig \ref{fig:stable-consensus}B and C).
+Most network methods, including the consensus, were highly unstable (Fig~\ref{fig:benchmark}B), raising questions about the results' reliability.
+We built a new, \emph{stable consensus} solution using the genes selected most often across the 30 solutions obtained by running the 6 methods on 5 different splits of the data (Section~\ref{methods:benchmark}).
+Such a network should capture the subnetworks more often found altered, and hence should be more resistant to noise.
+We used only genes selected in at least 7 solutions, which corresponded to 1\% of all genes selected at least once.
+The resulting stability-based consensus was composed of 68 genes (Fig~\ref{fig:stable-consensus}).
+This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig~\ref{fig:stable-consensus}B and C).
 
@@
   \includegraphics[width=.8\linewidth]{./figures/figure_5.pdf}
-  \caption{ {\bf Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus}) }.
-\textbf{(A)} ~Manhattan plot of genes; in black, the ones in the stable consensus solution.
-The Bonferroni threshold is indicated by a red line (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
-\textbf{(B)} Stable consensus network.
+  \caption{ \textbf{Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus})}.
+\textbf{(A)}~Manhattan plot of genes; in black, the ones in the stable consensus solution.
+The red line indicates the Bonferroni threshold (1.53~\texttimes{}~10\textsuperscript{-6} for genes).
+\textbf{(B)}~Stable consensus network.
 Each gene is represented by a pie chart, which shows the methods that selected it.
-We enlarged the most central gene (\emph{CUL3}) and those genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset (Section~\ref{methods:bcac}).
-\textbf{(C)} The nodes are in the same disposition as in panel B, but every gene name is indicated.
-We colored in pink the names of genes that are known breast cancer susceptibility genes and/or significantly associated with breast cancer susceptibility in the BCAC dataset.}
+We enlarged the most central gene (\emph{CUL3}), the known breast cancer susceptibility genes, and the BCAC-significant genes (Section~\ref{methods:bcac}).
+\textbf{(C)}~The nodes are in the same disposition as in panel B, but we indicated every gene name.
+We colored in pink the names of known breast cancer susceptibility genes and BCAC-significant genes.}
   \label{fig:stable-consensus}
@@
 
-Despite these similarities, the consensus and the stable consensus solutions included different genes.
+Despite these similarities, the consensus and the stable consensus included different genes.
 In the stable consensus network, the most central gene was \emph{CUL3}, which was absent from the previous consensus solution and had a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
@@
 
-In recent years, the ability of GWAS to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
-First, the omnigenic model proposes that gene functions are interwoven with each other in a dense co-function network.
-The practical consequence is that larger and larger GWAS will lead to the discovery of an uninformative wide-spread pleiotropy.
-Second, discovery in GWAS is hindered by a conservative statistical framework.
-Network methods elegantly address these two issues by using both association scores and an interaction network to take into consideration the biological context of each of the genes and SNPs.
+In recent years, the GWAS' ability to unravel the mechanisms leading to complex diseases has been called into question \cite{boyle_expanded_2017}.
+First, the omnigenic model proposes that gene functions are interwoven in a dense co-function network.
+The practical consequence is that larger and larger GWAS will lead to discovering an uninformative wide-spread pleiotropy.
+Second, its conservative statistical framework hinders GWAS discovery.
+Network methods elegantly address these two issues by using both association scores and an interaction network to consider the biological context of each of the genes and SNPs.
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided discovery have been proposed in recent years.
-In this article we evaluated six of these methods (Section \ref{methods:methods}) by applying them to the GENESIS study, a GWAS dataset on familial breast cancer (Section \ref{methods:data}).
-
-DmGWAS, Heinz, HotNet2, SConES and SigMod all yielded interesting solutions, which include (but are not limited to) known breast cancer susceptibility genes (Section \ref{results:separate_networks}).
-In general, the selected genes and SNPs were more central than most other genes and SNPs, in accordance with the observation that disease genes are more relatively central \cite{pinero_uncovering_2016}.
-However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by network methods (Section \ref{results:drawbacks}).
-Yet, across this article we showed that the highly central genes that were selected (\emph{COPS5}, \emph{CUL3} and \emph{FN1}) could plausibly be involved in breast cancer susceptibility.
-Despite these similarities, the solutions obtained were notably different.
-At one end of the spectrum, SConES and heinz preferred highly associated solutions, which in turn were small and did not shed much light on the etiology of the disease.
-On the other end, SigMod and dmGWAS gravitated towards larger, less associated solutions which provided a wide overview of the biological context.
+In this article we evaluated six of these methods (Section~\ref{methods:methods}) by applying them to the GENESIS study, a GWAS dataset on familial breast cancer (Section~\ref{methods:data}).
+
+DmGWAS, Heinz, HotNet2, SConES, and SigMod all yielded compelling solutions, which include (but are not limited to) known breast cancer susceptibility genes (Section~\ref{results:separate_networks}).
+In general, the selected genes and SNPs were more central than most other genes and SNPs, agreeing with the observation that disease genes are more central \cite{pinero_uncovering_2016}.
+However, very central nodes are also more likely to be connecting any given random pair of nodes, making them more likely to be selected by network methods (Section~\ref{results:drawbacks}).
+However, we found support in the literature for the involvement of the selected highly central genes (\emph{COPS5}, \emph{FN1}, and \emph{CUL3}) in breast cancer susceptibility (Sections~\ref{results:consensus}, \ref{results:drawbacks}, and \ref{results:stable-consensus}).
+Despite these similarities, the methods' solutions were notably different.
+At one end of the spectrum, SConES and heinz preferred high scoring solutions, which were also small and hence did not shed much light on the disease's etiology.
+On the other end, dmGWAS, HotNet2 and SigMod gravitated towards lower scoring but larger solutions, which provided a wide overview of the biological context.
 While this deepened our understanding of breast cancer susceptibility and provided biological hypotheses, they required further analyses.
-For instance, we had to examine the centrality of the selected genes to discern the extent to which that property is driving the selection of each gene (Section \ref{results:drawbacks}).
-Lastly, HotNet2 produced a third kind of solution: a large constellation of many, high-scoring, small subnetworks.
+For instance, we examined the centrality of the selected genes to understand how much that property was driving their selection (Section~\ref{results:drawbacks}).
 However, all solutions shared two drawbacks.
 First, they were all equally bad at discriminating cases from controls.
-Yet, the classification accuracy of network methods was similar to that of a classifier trained on the entire genome (Section \ref{results:benchmark}), which suggests that cases and controls are difficult to separate in the GENESIS dataset.
-This may be due to limited statistical power, which reduces the ability to identify relevant SNPs; but in any event, we do not expect to be able to separate people who have or will developed cancer from others on the sole basis of their genomes, ignoring all environmental factors and chance events.
-Hence, in this case network methods were preferable to the logistic regression classifier, since they did ``no worse'' while providing an interpretable solution.
+However, the classification accuracy of network methods was similar to that of a classifier trained on the entire genome (Section~\ref{results:benchmark}), which suggests that cases and controls are difficult to separate in the GENESIS dataset.
+This may be due to limited statistical power, which reduces the ability to identify relevant SNPs.
+However, in any event, we do not expect to separate people who have or will develop cancer from others on the sole basis of their genomes, ignoring all environmental factors and chance events.
+Hence, network methods were preferable to the logistic regression classifier since they did ``no worse'' at classification while providing an interpretable solution.
 Second, all methods were remarkably unstable, yielding different solutions for slightly different inputs.
@@
 Hence, heinz's conservative transformation of P\=/values, which favored only the most extreme ones, led to improved stability.
-Another source of instability might have been the redundancy inherent to biological networks, since they are subject to an evolutionary pressure to avoid single points of failure \cite{wagnerAlternativeRoutesMutational2007}.
+Another source of instability might have been the redundancy inherent to biological networks, a consequence of an evolutionary pressure to avoid single points of failure \cite{wagnerAlternativeRoutesMutational2007}.
 Hence, biological networks will often have multiple paths connecting two high-scoring nodes.
 
-To overcome these limitations while exploiting the strengths of the individual methods, we proposed combining them into a consensus solution.
-We used a straightforward strategy of including any node that was recovered by multiple methods.
-We thus proposed two networks (Sections \ref{results:consensus} and \ref{results:stable-consensus}): a consensus solution, meant to address the heterogeneity of the solutions in the full dataset, and a stable consensus solution, which in addition addressed the instability of the methods.
-They both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer, thus synthetizing the mechanisms altered in breast cancer cases.
-Thanks to their smaller size and their network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease, but are related to cancer at the expression level and interact with genes with consistently high association scores.
-Importantly, while the consensus approach was as unstable as the individual network-guided methods, the stable consensus network retained the ability to provide compelling hypotheses and had better stability.
-This supports that instability might be caused by redundant but equivalent biological mechanisms, and hence validates the conclusions obtained on the individual solutions and the consensus.
+To overcome these limitations while exploiting the each method's strengths, we proposed combining them into a consensus solution.
+We used the straightforward strategy of including any node that was recovered by at least two methods.
+We thus proposed two solutions (Sections~\ref{results:consensus} and \ref{results:stable-consensus}): a consensus solution, which addressed the heterogeneity of the solutions, and a stable consensus solution, which addresseded the instability of the methods.
+They both included the majority of the strongly associated smaller solutions and captured genes and broader mechanisms related to cancer, thus synthesizing the mechanisms altered in breast cancer cases.
+Thanks to their smaller size and network structure, they provided compelling hypotheses on genes like \emph{COPS5} and \emph{CUL3}, which lack genome-wide association with the disease but are related to cancer at the expression level and consistently interact with high scoring genes.
+Notably, while the consensus approach was as unstable as the individual network-guided methods, the stable consensus network retained the ability to provide compelling hypotheses and had better stability.
+This supported that redundant but equivalent mechanisms might cause instability and supported the conclusions obtained on the individual solutions.
 
 In this work, we have compared our results to significant genes and SNPs in the BCAC study \cite{Michailidou2017}.
-Network methods showed modest precision, but much higher recall at recovering BCAC hits (Section \ref{results:boost}).
-While precision might be desirable when a subset of good markers is required (for instance, for diagnosis), higher recall is desirable in exploratory settings.
+Network methods showed modest precision but much higher recall at recovering BCAC hits (Section~\ref{results:boost}).
+While precision might be desirable when a subset of useful markers is required (for instance, for diagnosis), higher recall is desirable in exploratory settings.
 Nonetheless, BCAC was not an ideal ground truth.
-First, the studied populations are not completely overlapping: BCAC focused on a pan-European cohort, while GENESIS targeted the French population specifically.
-Second, the study designs differed: a high proportion of breast cancer cases investigated in BCAC were sporadic (not selected according to family history), while GENESIS was a homogeneous dataset not included in BCAC and which focused on the French high-risk population attending the family cancer clinics.
-Despite these differences, we expected some degree of shared genetic architecture, especially at the gene level.
-Finally, and this is indeed the motivation for this study, GWAS are unlikely to identify all genes relevant for the disease: some might only show up in rare-variant studies; others might have too low effect sizes.
+First, the studied populations are not entirely overlapping: BCAC focused on a pan-European cohort, while GENESIS targeted the French population.
+Second, the study designs differed: a high proportion of breast cancer cases investigated in BCAC were sporadic (not selected according to family history), while GENESIS was a homogeneous dataset not included in BCAC focused on the French high-risk population attending the family cancer clinics.
+Finally, and this is indeed the motivation for this study, GWAS are unlikely to identify all genes relevant for the disease: some might only show up in rare-variant studies; others might have too small effect sizes.
 Network methods account for this by including genes with low association scores but with relevant topological properties.
-Hence, network methods and GWAS, even well-powered, are unlikely to capture exactly the same sets of genes.
-This might partly excuse the low precisions displayed in Section \ref{results:boost} and the low AUC displayed in Section \ref{results:benchmark}.
+Hence, network methods and GWAS, even when well-powered, are unlikely to capture exactly the same sets of genes.
+This might partly excuse the low precisions displayed in Section~\ref{results:boost} and the low AUC displayed in Section~\ref{results:benchmark}.
 
 As not all PPIN databases compiled the same interactions, the choice of the PPIN determines the final output.
-In this work we used exclusively interactions from HINT from high-throughput experiments.
-This responded to concerns about adding interactions identified in targeted studies and prone to a ``rich getting richer'' phenomenon: popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, and they might bias discovery by reducing the average shortest path length between two random nodes.
+In this work, we used only interactions from HINT from high-throughput experiments (Section~\ref{methods:networks}).
+This responded to concerns about adding interactions identified in targeted studies and falling into ``rich getting richer'' problems: since popular genes have a higher proportion of their interactions described \cite{cai_broker_2010,das_hint:_2012}, they might bias discovery towards themselves by reducing the average shortest path length between two random nodes.
 On the other hand, Huang et al.
@@
 This would support using the largest networks in our experiments.
-However, when we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), we found that for most of the methods it did not greatly change the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
-This supports using only interactions from high-throughput experiments, which produced apparently similar solutions and avoided falling into ``circular reasonings'', where the best-known genes were artificially pushed into the solutions as we observed in Section \ref{results:drawbacks}.
+However, when we compared the impact of using a larger PPIN containing interactions from both high-throughput experiments and the literature (Section~\ref{methods:networks}), for most of the methods it did not change much the size or the stability of the solution, the classification accuracy, or the runtime (\nameref{sfig:lc_ht_comparison}).
+This supports using only interactions from high-throughput experiments, which produced similar solutions and avoided falling into ``circular reasonings'', where the best-known genes were artificially pushed into the solutions, as we observed in Section~\ref{results:drawbacks}.
 
@@
 The majority of those (14\,660) are non-coding genes, mainly lncRNA, miRNA, and snRNA (\nameref{sfig:biotypes_excluded}).
-Yet, RNA genes like \emph{CASC16} were associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
-In addition, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, overlapping with \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the PPIN.
+Nevertheless, RNA genes like \emph{CASC16} were associated to breast cancer (Section~\ref{results:conventional}), reminding us of the importance of using networks beyond coding genes.
+Besides, even protein-coding genes linked to breast cancer susceptibility \cite{ahmed_newly_2009}, like \emph{NEK10} (P\=/value 1.6 \texttimes{} 10\textsuperscript{-5}, overlapping with \emph{SLC4A7}) or \emph{POU5F1B}, were absent from the PPIN.
 However, on average protein-coding genes absent from the PPIN were less associated with breast cancer susceptibility (Wilcoxon rank-sum P\=/value~=~2.79 \texttimes{} 10\textsuperscript{-8}, median P\=/values of 0.43 and 0.47).
@@
 As disease genes tend to be more central \cite{pinero_uncovering_2016}, we hypothesize that it was due to interactions between central genes being more likely.
-It is worth noting that network methods that do not use PPIs, like SConES GS and GM, did recover SNPs in \emph{NEK10} and \emph{CASC16}.
-Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018} and, yet, are excluded from gene-centric approaches.
+It is worth noting that network methods that do not use PPIs, like SConES GS and GM, recovered SNPs in \emph{NEK10} and \emph{CASC16}.
+Moreover, both SConES GM and GI recovered intergenic regions, which might contain key regulatory elements \cite{gallagherPostGWASEraAssociation2018}, but are excluded from gene-centric approaches.
 This shows the potential of SNP networks, in which SNPs are linked when there is evidence of co-function, to perform network-guided GWAS even in the absence of gene-level interactions.
-Lastly, all the methods are heavily affected by how SNPs are mapped to genes, and other strategies (e.g.
-eQTLs, SNPs associated to the expression of a gene) might lead to different results.
-
-A crucial step for the gene-based methods is the computation of the gene score.
-In this work we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
-However, it presents known problems (selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}), and other algorithms like PEGASUS \cite{nakka_gene_2016}, SKAT \cite{ionita-laza_sequence_2013} or COMBAT \cite{wang_combat:_2017} might have more statistical power.
+Lastly, all the methods are heavily affected by how SNPs are mapped to genes, and other strategies (e.g., eQTLs, SNPs associated to gene expression) might lead to different results.
+
+A crucial step for the gene-based methods is the computation of gene scores.
+In this work, we used VEGAS2 \cite{mishra_vegas2:_2015} due to the flexibility it offers to use user-specified gene annotations.
+However, it presents known problems: selection of an appropriate percentage of top SNPs, long runtimes and P\=/value precision limited to the number of permutations \cite{nakka_gene_2016}).
+Additionally, other algorithms like PEGASUS \cite{nakka_gene_2016}, SKAT \cite{ionita-laza_sequence_2013} or COMBAT \cite{wang_combat:_2017} might have more statistical power.
 
 How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners.
+Often, the question is whether an LD-based pruning of the genotypes will improve the results.
 VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
-With regards to gene-based network methods, in Section~\ref{results:consensus} we highlighted ambiguities that appear when genes overlap or are in LD.
-In fact, the presented case is paradigmatic, since all three genes are located in the HLA region, the most gene-dense region of the genome \cite{Xie2003}.
+In Section~\ref{results:consensus} we highlighted ambiguities that appear when genes overlap or are in LD.
+The presented case is paradigmatic since all three genes are in the HLA region, the most gene-dense region of the genome \cite{Xie2003}.
 Network methods are prone to selecting such genes when they are functionally related, and hence interconnected in the PPIN.
@@
 With regards to SConES, fewer SNPs would lead to simpler SNP networks and, possibly, shorter runtimes.
-However, LD patterns also affect SConES' in other ways, since its formulation penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Equation~\ref{eq:scones}.
-Due to LD, nearby SNPs' P-values are correlated; and since SNP networks are determined by positional information, nearby SNPs are likely to be linked.
+However, LD patterns also affect SConES' in other ways, since its formulation penalizes selecting a SNP and not its neighbors, via a nonzero parameter $\eta$ in Eq~\ref{eq:scones}.
+Due to LD, nearby SNPs' P-values correlate; since positional information determines SNP networks, nearby SNPs are likely to be connected.
 Hence, SConES tends to select LD-blocks formed by low P-value SNPs.
-This might explain why SConES produces similar results on the GS and GM networks, heavily affected by LD (Section \ref{results:drawbacks}).
+This might explain why SConES produced similar results on the GS and GM networks, heavily affected by LD (Section~\ref{results:drawbacks}).
 However, this same behavior raises the burden of proof required to select SNPs with many interactions, like those mapped to hub genes in the PPIN.
 For this reason, SConES GI did not select any protein coding gene.
-We hypothesize that this is caused by the absence of joint association of a gene and a majority of its neighbors.
-This is supported by the lack of results from LEAN as well.
-Yet, a different combination of parameters could lead to a more informative SConES' solution (e.g.
-a lower \(\lambda\) in Equation~\ref{eq:scones}), although it is unclear how to find it.
-In addition, due to the design of the iCOGS array (Section~\ref{methods:data}), the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are under studied --- hindering the accuracy with which the GS network captures the genome structure.
-A stringent LD pruning might address such problems.
+This could be caused by the absence of joint association of a gene and most of its neighbors, a hypothesis supported by LEAN's lack of results.
+Yet, a different combination of parameters could lead to a more informative SConES' solution (e.g., a lower \(\lambda\) in Eq~\ref{eq:scones}), although it is unclear how to find it.
+In addition, due to the design of the iCOGS array (Section~\ref{methods:data}), the genome of GENESIS participants has not been unbiasedly surveyed: some regions are fine-mapped --- which might distort gene structure in GM and GI networks --- while others are understudied --- hindering the accuracy with which the GS network captures the genome structure.
+A strong LD pruning might address such problems.
 
 To produce the two consensus solutions, we faced practical challenges due to the differences in interfaces, preprocessing steps, and unexpected behaviors of the various methods.
-To make it easier for other to apply these methods to new datasets and aggregate their solutions, we built six \texttt{nextflow} pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
-They are available on GitHub: \href{https://github.com/hclimente/gwas-tools}{hclimente/gwas-tools} (Section \ref{methods:code}).
-Importantly, we compiled those methods that had a permissive license into a Docker image for easier use, which is available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
-
-\section{Materials and methods}
-
-\subsection{GENESIS dataset, preprocessing and quality control}
-\label{methods:data}
-
-The GENE Sisters (GENESIS) study was designed to investigate risk factors for familial breast cancer in the French population \cite{sinilnikova_genesis:_2016}.
-Index cases were patients with infiltrating mammary or ductal adenocarcinoma, who had a sister with breast cancer, and who have been tested negative for \emph{BRCA1} and \emph{BRCA2} pathogenic variants.
-Controls were unaffected colleagues and/or friends of the cases, born around the year of birth of their corresponding case (\textpm{} 3 years).
-We focused on the 2\,577 samples of European ancestry, of which 1\,279 were controls and 1\,298 were cases.
-The genotyping was performed using the iCOGS array, a custom Illumina array designed to study the genetic susceptibility to hormone-related cancers \cite{sakoda_turning_2013}.
-It contained 211\,155 SNPs, including SNPs putatively associated with breast, ovarian, and prostate cancers, SNPs associated with survival after diagnosis, and SNPs associated to other cancer-related traits, as well as candidate functional variants in selected genes and pathways.
-
-We discarded SNPs with a minor allele frequency lower than 0.1\%, those not in Hardy--Weinberg equilibrium in controls (P\=/value \textless~0.001), and those with genotyping data missing on more than 10\% of the samples.
-A subset of 20 duplicated SNPs in \emph{FGFR2} were also removed.
-In addition, we removed the samples with more than 10\% missing genotypes.
-After controlling for relatedness, 17 additional samples were removed (6 for sample identity error, 6 controls related to other samples, 2 cases being related to an index case, and 3 additional controls having a high relatedness score).
-Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
-The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
-
-We looked for population structure that could produce spurious associations.
-A principal component analysis revealed no visual differential population structure between cases and controls (\nameref{sfig:pcs}).
-Independently, we did not find evidence of genomic inflation (\(\lambda\)~=~1.05) either, further confirming the absence of confounding population structure.
-
-\subsection{SNP- and gene-based GWAS}
-\label{methods:conventional}
-
-To measure association between a genotype and susceptibility to breast cancer, we performed a per-SNP 1~d.f.~$\chi^2$ allelic test using PLINK v1.90 \cite{chang_second-generation_2015}.
-In order to obtain significant SNPs, we performed Bonferroni correction, to keep family-wise error rate below 5\%.
-The threshold used was $\frac {0.05} {197083} = 2.54 \times 10^{-7}$.
-
-Then, we used VEGAS2 \cite{mishra_vegas2:_2015} to compute the gene-level association score from the P\=/values of the SNPs mapped to them.
-More specifically, we mapped SNPs to genes through their genomic coordinates: all SNPs located within the boundaries of a gene, \textpm 50 kb, were mapped to that gene.
-For each gene, we computed VEGAS2 scores using only the 10\% of SNPs with lowest P\=/values among all those that were mapped to it.
-We used the 62\,193 genes described in GENCODE 31 \cite{frankish_gencode_2019}, although only 54\,612 could be mapped to at least one SNP.
-Out of those, we focused exclusively on the 32\,767 that had a gene symbol.
-Out of the 197\,083 SNPs remaining after quality control, 164\,037 were mapped to at least one of these genes.
-We also used Bonferroni correction to obtain significant genes; in this case, the threshold of signigicance was $\frac {0.05} {32767} = 1.53 \times 10^{-6}$.
-
-\subsection{Network methods}
-
-\subsubsection{Mathematical notations}
-\label{methods:notation}
-In this article, we used undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
-$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
-Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
-When referring to a subnetwork $S$, $V_{S}$ is the set of nodes in $S$ and $E_{S}$ is the set of edges in $S$.
-A special case of subgraphs are \emph{connected} subgraphs, which occur when every node in the subgraph can be reached from any other node.
-
-Nodes can be described by properties provided by the topology of the graph.
-We focused on two of those: degree centrality, and betweenness centrality.
-The degree centrality, or degree, is the number of edges that a node has.
-The betweenness centrality, or betweenness, is the number of times a node participates in the shortest paths between two other nodes.
-
-We also used two matrices that describe two different properties of a graph.
-These matrices are square, and have as many rows and columns as nodes in the network.
-The element $(i,j)$ hence represents a relationship between $v_i$ and $v_j$.
-The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; its diagonal is zero.
-The \emph{degree matrix} $D_G$ is a diagonal matrix which contains the degree of the different nodes.
-
-\subsubsection{Networks}
-\label{methods:networks}
-
-\paragraph{Gene network}
-The mathematical formulations of the different network methods are compatible with any type of biological network (protein interactions, gene coexpression, regulatory, etc.).
-Here, we used protein-protein interaction networks (PPIN) for all of the network methods except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
-We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
-Unless otherwise specified, we used only interactions coming from high-throughput experiments, leaving out targeted studies that might bias the topology of the PPIN.
-Out of the 146\,722 interactions from high-throughput experiments that HINT stores, we were able to map 142\,541 to a pair of gene symbols, involving 13\,619 genes.
-12\,880 of those mapped to a genotyped SNP after quality control, involving 127\,604 interactions.
-The scoring function for the nodes changed from method to method (Section~\ref{methods:methods}).
-
-Additionally, we compared the results obtained on the aforementioned PPIN with those obtained on another PPIN built using interactions coming from both high-throughput and targeted studies.
-In that case, out of the 179\,332 interactions in HINT, 173\,797 mapped to a pair of gene symbols.
-Out of those, 13\,735 mapped to a genotyped SNP after quality control, involving 156\,190 interactions.
-
-\paragraph{SNP networks}
-SConES \cite{azencott_efficient_2013} was the only network method designed to handle SNP networks.
-As in gene networks, two SNPs were connected in a SNP network when there was evidence of shared functionality between them.
-Azencott et al.
-\cite{azencott_efficient_2013} proposed three ways of building these networks: connecting the SNPs consecutive in the genomic sequence (``GS network''); interconnecting all the SNPs mapped to the same gene, on top of GS (``GM network''); and interconnecting all SNPs mapped to two genes for which a protein-protein interaction exists, on top of GM (``GI network'').
-We focused on the GI network, as it fitted the scope of this work better, using the PPIN described above.
-However, at different stages we also compared GI to GS and GM to understand how considering the PPIN affects SConES' output.
-For the GM network, we used the mapping described in Section~\ref{methods:snp2gene}.
-In all three the node scores were the association scores of the individual SNPs with the phenotype (1~d.f.~\(\chi\)\textsuperscript{2} statistic).
-The properties of these three subnetworks are available in \nameref{stab:snp_solutions}.
-
-\subsubsection{High-score subnetwork search algorithms}
-\label{methods:methods}
-
-Genes that contribute to the same function are nearby in the PPIN, and can be topologically related to each other in diverse ways (densely interconnected modules, nodes around a hub, a path, etc.).
-Several aspects have to be taken into consideration when developing a network method: how to score the nodes, whether the affected mechanisms form a single connected component or several, how to frame the problem in a computationally efficient fashion, which network to use, etc.
-Unsurprisingly, multiple solutions have been proposed.
-We examined six of them: five that explore the PPIN, and one which explores SNP networks.
-We selected methods that were open-source, and had an implementation available and accessible documentation.
-Their main differences are summarized in Table~\ref{tab:method_comparison}.
-We scored both SNPs and genes with the P-values (or transformations of them) computed in Section~\ref{methods:conventional}.
-
-\begin{table}[!ht]
-  \begin{adjustwidth}{-2.25in}{0in} % Comment out/remove adjustwidth environment if table fits in text column.
-  \centering
-  \caption{
-  {\bf Summary of the differences between the network methods.}}
-  \begin{tabular}{l|llllllll}
-  {\bf Method} & {\bf Field } & {\bf Nodes} & {\bf Exhaustive} & {\bf Solution} & {\bf Comp.} & {\bf Input} & {\bf Scoring} & {\bf Ref.}\\ \thickhline
-  dmGWAS & GWAS & Genes & No & - & 1 & Summary & -log\textsubscript{10}(P) & \cite{jia_dmgwas:_2011}\\
-  heinz & Omics & Genes & Yes & - & 1 & Summary & BUM & \cite{dittrich_identifying_2008}\\
-  HotNet2 & Omics & Genes & Yes & Module & \(\ge\) 1 & Summary & Local FDR & \cite{leiserson_pan-cancer_2015}\\
-  LEAN & Omics & Genes & Yes & Star & \(\ge\) 1 & Summary & -log\textsubscript{10}(P) & \cite{gwinner_network-based_2016}\\
-  SConES & GWAS & SNPs & Yes & Module & \(\ge\) 1 & Genotypes & 1 d.f. \(\chi\)\textsuperscript{2} & \cite{azencott_efficient_2013}\\
-  SigMod & GWAS & Genes & Yes & Module & \(\ge\) 1 & Summary & $\Phi$\textsuperscript{-1}(1 - P) & \cite{liu_sigmod:_2017}\\
-  \end{tabular}
-  \begin{flushleft} \textbf{\emph{Field}}: field in which the algorithm was developed.
-\textbf{\emph{Nodes}}: the type of nodes in the network, either genes (PPIN) or SNPs.
-\textbf{\emph{Exhaustive}}: whether all the possible solutions given the selected parameters are explored.
-\textbf{\emph{Solution}}: additional properties that are enforced on the solution, other than containing high scoring, connected nodes.
-\textbf{\emph{Comp.}}: number of connected components in the solution.
-\textbf{\emph{Input}}: genotype data or GWAS summary statistics.
-\textbf{\emph{Scoring}}: how SNP/gene P\=/values were transformed into node scores.
-In the case of heinz, BUM stands for beta-uniform model, used to transform the P-values; for SigMod, $\Phi$\textsuperscript{-1} represents the inverse of the cumulative distribution function of the standard Normal distribution.
-\textbf{\emph{Ref.}}: original publication featuring the algorithm.
-  \end{flushleft}
-  \label{tab:method_comparison}
-  \end{adjustwidth}
-\end{table}
-
-\begin{description}
-\item[{dmGWAS}] dmGWAS seeks the subgraph with the highest local density in low P\=/values \cite{jia_dmgwas:_2011}.
-To that end it searches candidate solutions using a greedy, ``seed and extend'', heuristic:
-
-\begin{enumerate}
-\item Select a seed node $i$ and form the subnetwork $S_i=\{i\}$.
-\item Compute Stouffer's Z-score Z\textsubscript{m} for $S_i$ as
-
-\begin{eqnarray} 
-Z_m = \frac{1}{\sqrt{k}} \sum_{j \in S_i} z_j,
-\end{eqnarray}
-
-where \emph{k} is the number of genes in $S_i$; z\textsubscript{j} is the Z score of gene $j$, computed as \(\phi\)\textsuperscript{-1}(1 - P\=/value\textsubscript{j}); and \(\phi\)\textsuperscript{-1} is the inverse normal distribution function.
-\item Identify neighboring nodes of $S_i$,  i.e. nodes at distance \(\le\) \texttt{d}.
-\item Add the neighboring nodes whose inclusion increases Z\textsubscript{m+1} by more than a threshold Z\textsubscript{m}~\texttimes{}~(1 + \texttt{r}).
-\item Repeat 2-4 until no further enlargement is possible.
-\item Add $S_i$ to the list of subnetworks to return. Normalize its Z-score as
-
-\begin{eqnarray}
-Z_{N}=\frac{Z_{m}-\operatorname{mean}\left(Z_{m}(\pi)\right)}{\operatorname{SD}\left(Z_{m}(\pi)\right)},
-\end{eqnarray}
-
-where Z\textsubscript{m}(\(\pi\)) represents a vector containing 100\,000 random subsets of the same number of genes.
-\end{enumerate}
-
-DmGWAS carries out this process on every gene in the PPIN.
-We used the implementation of dmGWAS in the dmGWAS 3.0 R package \cite{dmgwas}.
-Unless otherwise specified, we used the suggested parameters \texttt{d} $ = 2$ and \texttt{r} $ = 0.1$.
-We used the function \texttt{simpleChoose} to select the solution, which aggregates the top 1\% subnetworks.
-
-\item[{heinz}] The goal of heinz is to identify the highest-scored connected subnetwork \cite{dittrich_identifying_2008}.
-The authors proposed a transformation of the genes' P\=/value into a score that is negative under weak association with the phenotype, and positive under a strong one.
-This transformation is achieved by modeling the distribution of P\=/values by a beta-uniform model (BUM) parameterized by the desired false discovery rate (FDR).
-Thus formulated, the problem is NP-complete, and hence solving it would require a prohibitively long computational time.
-To solve it efficiently it is re-cast as the Prize-Collecting Steiner Tree Problem (PCST), which seeks to select the connected subnetwork S that maximizes the \emph{profit} p(S), defined as:
-
-\begin{eqnarray}
-p(S) = \sum_{v \in V_S} p(v) - \sum_{e \in E_S} c(e). 
-\end{eqnarray}
-
-were $p(v)~=~w(v) - w'$ is the \emph{profit} of adding a node, $c(e)~=~w'$ is the \emph{cost} of adding an edge, and $w' = min_{v \in V_{G}} w(v)$ is the smallest node weight of $G$.
-All three are positive quantities.
-Heinz implements the algorithm from Ljubić et al.
-\cite{ljubic_algorithmic_2006} which, in practice is often fast and optimal, although neither is guaranteed.
-We used BioNet's implementation of heinz \cite{beisser_bionet:_2010,heinz}.
-
-\item[{HotNet2}] HotNet2 was developed to find connected subgraphs of genes frequently mutated in cancer \cite{leiserson_pan-cancer_2015}.
-To that end, it considers both the local topology of the PPIN and the scores of the nodes.
-The former is captured by an insulated heat diffusion process: at initialization, the score of the node determines its initial heat; iteratively each node yields heat to its ``colder'' neighbors, and receives heat from its ``hotter'' neighbors while retaining part of its own (hence, \emph{insulated}).
-This process continues until a stationary state is reached, in which the temperature of nodes does not change anymore, and results in a diffusion matrix $F$.
-$F$ is used to compute the similarity matrix $E$ that models exchanged heat as
-\begin{eqnarray} 
-E = F \operatorname{diag}(w(V)), 
-\end{eqnarray}
-where $\operatorname{diag}(w(V))$ is a diagonal matrix with the node scores in its diagonal.
-For any two nodes $i$ and $j$, $E_{ij}$ models the amount of heat that diffuses from node $j$ to node $i$, which can be interpreted as a (non-symmetric) similarity between those two nodes.
-To obtain densely connected solutions, HotNet2 prunes $E$, only preserving edges such that $w(E) > \delta$.
-Lastly, HotNet2 evaluates the statistical significance of the solutions by comparing their size to the size of PPINs obtained by permuting the node scores.
-We assigned initial node scores as in Nakka et al.
-\cite{nakka_gene_2016}, assigning a score of 0 for the genes with a low probability of being associated to the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
-In the GENESIS dataset, the threshold separating both was a P\=/value of 0.125, which was obtained using a local FDR approach \cite{scheid_twilight;_2005}.
-HotNet2 has two parameters: the restart probability \(\beta\), and the threshold heat \(\delta\).
-Both parameters are set automatically by the algorithm, which is robust to their values \cite{leiserson_pan-cancer_2015}.
-HotNet2 is implemented in Python \cite{hotnet2}.
-
-\item[{LEAN}] LEAN searches altered ``star'' subnetworks, that is, subnetworks composed by one central node and all its interactors \cite{gwinner_network-based_2016}.
-By imposing this restriction, LEAN can exhaustively test all such subnetworks (one per node).
-For a particular subnetwork of size \emph{m}, the P\=/values corresponding to the involved nodes are ranked as p\textsubscript{1} \(\le\) \dots{} \(\le\) p\textsubscript{m}.
-Then, \emph{k} binomial tests are conducted, to compute the probability of having \emph{k} out of \emph{m} P\=/values lower or equal to p\textsubscript{k} under the null hypothesis.
-The minimum of these \emph{k} P\=/values is the score of the subnetwork.
-This score is transformed into a P\=/value through an empirical distribution obtained via a subsampling scheme, where gene sets of the same size are selected randomly, and their score computed.
-Lastly, P\=/values are corrected for multiple testing through a Benjamini-Hochberg correction.
-We used the implementation of LEAN from the \texttt{LEANR} R package \cite{leanr}.
-\item[{SConES}] SConES searches the minimal, modular, and maximally associated subnetwork in a SNP graph \cite{azencott_efficient_2013}. Specifically, it solves the problem
-
-\begin{eqnarray}
-\label{eq:scones}
-\underset{S \subseteq G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} - \underbrace{\lambda \sum_{v \in V_S} \sum_{u \not\in V_S} {W}_{vu} }_{\text { connectivity }}-\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}
-\end{eqnarray}
-
-where \(\lambda\) and \(\eta\) are parameters that control the sparsity and the connectivity of the model.
-The connectivity term penalizes disconnected solutions, with many edges between nodes that are selected and nodes that are not.
-Given a $\lambda$ and an $\eta$, the aforementioned problem has a unique solution, that SConES finds using a graph min-cut procedure.
-As in Azencott et al.
-\cite{azencott_efficient_2013}, we selected \(\lambda\) and \(\eta\) by cross-validation, choosing the values that produce the most stable solution across folds.
-In this case, the selected parameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
-We used the version on SConES implemented in the R package \texttt{martini} \cite{martini}.
-
-\item[{SigMod}] SigMod aims at identifying the highest-scoring, most densely connected subnetwork \cite{liu_sigmod:_2017}.
-It addresses an optimization problem similar to that of SConES (Equation \ref{eq:scones}), but the connectivity term encourages connected solutions by favoring solutions where many edges connect two selected nodes, rather than penalizing disconnected ones:
-
-\begin{eqnarray}
-\underset{S \in G}{\arg \max } \underbrace{\sum_{v \in V_S} w(v)}_{\text { association }} + \underbrace{\lambda \sum_{v \in V_S} \sum_{u \in V_S} W_{vu} }_{\text { connectivity }} -\underbrace{\eta \lvert V_S \rvert }_{\text { sparsity }}.
-\end{eqnarray}
-
-As for SConES, this optimization problem can also be solved by a graph min-cut approach.
-
-SigMod presents three important differences with SConES.
-First, it is designed for PPINs.
-Second, it favors solutions containing many edges.
-SConES, instead, penalizes connections between the selected and unselected nodes.
-Third, it explores the grid of parameters differently, and processes their respective solutions.
-Specifically, for the range of \(\lambda\)~=~\(\lambda\)\textsubscript{min}, \dots{}, \(\lambda\)\textsubscript{max} for the same \(\eta\), it prioritizes the solution with the largest change in size from \(\lambda\)\textsubscript{n} to \(\lambda\)\textsubscript{n+1}.
-Additionally, that change needs to be larger than a user specified threshold \texttt{maxjump}.
-Such a large change implies that the network is densely interconnected.
-This results in one candidate solution for each \(\eta\), which are processed by removing any node not connected to any other.
-A score is assigned to each candidate solution by summing their node scores and normalizing by size.
-The candidate solution with the highest standardized score and that is not larger than a user-specified threshold (\texttt{nmax}) is the chosen solution.
-We used the default parameters \texttt{maxjump} $ = 10$ and \texttt{nmax} $= 300$.
-SigMod is implemented in an R package \cite{sigmod}.
-
-\item[{Consensus}] We built a consensus solution by retaining the genes that were selected by at least two of the six methods (using SConES GI for SConES).
-Any edge between the selected genes in the PPIN was included.
-\end{description}
-
-We performed all the computations in the cluster described in Section \ref{methods:computation}.
-
-\subsubsection{Parameter space}
-\label{methods:parameters}
-
-We used the network methods with the parameters recommended by their authors, or with the default parameters in their absence.
-Additionally, we explored the parameter space of the different methods to study how they alter the output.
-
-\begin{description}
-\item[{dmGWAS}] We tested multiple values for \texttt{r} (0.0001, 0.001, 0.01, 0.05, 0.1, 0.25, 0.5, and 1) and \texttt{d} (1, 2, and 3). 
-\item[{heinz}] We tested multiple FDR thresholds (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, 1).
-\item[{HotNet2}] We tested different thresholds to decide which genes would receive a score of 0, and which ones a score of $-log_{10}$(P-value): 0.001, 0.01, 0.05, 0.125, 0.25, and 0.5.
-\item[{LEAN}] We used the following significance cutoffs for LEAN's P-values (0.05, 0.1, 0.15, 0.2, 0.25, 0.3, 0.35, 0.4, 0.45, 0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.9, 0.95, and 1).
-\item[{SConES}] We used the values of $\lambda$ and $\eta$ that \texttt{martini} explores by default (35.54, 5.40, 0.82, 0.12, 0.02, 0.01, 4.39e-4, 6.68e-5, 1.02e-5, and 1.55e-6 in both cases)
-\item[{SigMod}] We tested multiple values for the parameters \texttt{nmax} (10, 50, 100, 300, 700, 1000, and 10\,000) and \texttt{maxjump} (5, 10, 20, 30, and 50).
-\end{description}
-
-\subsubsection{Comparing SNP-methods to gene-methods, and vice versa}
-\label{methods:snp2gene}
-
-In multiple steps of this article, we needed to compare the outcome of a method that works on genes with the outcome of another method that works on SNPs.
-For this purpose, we used the SNP-gene correspondence described in Section \ref{methods:conventional}.
-To convert a list of SNPs into a list of genes, we included all the genes that can be mapped to any of the SNPs.
-Conversely, to convert a list of genes into a list of SNPs, we included all the SNPs that can be mapped to any of the genes.
-
-\subsection{Pathway enrichment analysis}
-\label{methods:pathway_enrichment}
-
-We searched for pathways enriched in the gene solutions produced by the above methods.
-We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the \texttt{ReactomePA} R package \cite{Yu2016}.
-The universe of genes included any gene that we could map to a SNP in the iCOGS array (Section \ref{methods:conventional}).
-We adjusted the P-values for multiple testing as in Benjamini and Hochberg \cite{Benjamini1995} (BH): pathways with a BH adjusted P\=/value \textless~0.05 were deemed significant.
-
-\subsection{Benchmark of methods}
-\label{methods:benchmark}
-
-We evaluated multiple properties (described in Sections \ref{methods:algorithm_comparison} and \ref{methods:classifier}) of the different methods through a 5-fold subsampling setting.
-We applied each method to 5 random subsets of the original GENESIS dataset containing 80\% of the samples (\emph{train set}).
-When pertinent, we evaluated the solution on the remaining 20\% (\emph{test set}).
-We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
-Every method and repetition ran on the same computational settings (Section \ref{methods:computation}).
-
-\subsubsection{Properties of the solution}
-\label{methods:algorithm_comparison}
-
-We compared the runtime, the number of selected features (genes or SNPs), and the stability (sensitivity of the result to small changes in the input, here, using different \emph{train} sets) of the different network methods.
-Nogueira and Brown \cite{nogueira_measuring_2016} proposed quantifying the stability of a method using the Pearson correlation between the genes selected on different subsamples.
-This correlation was calculated between vectors with the length of the total number of features, containing a 0 at position $i$ if feature $i$ was not selected, and 1 if it was.
-
-\subsubsection{Classification accuracy of selected SNPs}
-\label{methods:classifier}
-
-A desirable solution offers good predictive power on unseen \emph{test} samples.
-We evaluated the predicting power of the SNPs selected by the different methods through the performance of an L1-penalized logistic regression classifier, which searches for a small subset of SNPs which provides good classification accuracy.
-We trained the classifier exclusively on those selected SNPs to predict the outcome (case/control).
-The L1 penalty helps to account for linkage disequilibrium by reducing the number of SNPs included in the model (\emph{active set}) while improving the generalization of the classifier.
-This penalty was set by cross-validation, choosing the value that minimized misclassification error.
-We applied each network method to each train set, and trained the classifier on the same train set using only the selected SNPs.
-When the method retrieved a list of genes (all of them except SConES), we considered as selected all the SNPs mapped to any of those genes.
-Then we evaluated the sensitivity and the specificity of the classifier on the test set.
-The active set gave an estimate of a plausible, more sparse solution with comparable predictive power to the original solution.
-To obtain a baseline, we also trained the classifier on all the SNPs of the train set.
-We do not expect a linear model on selected SNPs to be able to separate cases from controls well.
-Indeed, the lifetime cumulative incidence of breast cancer among women with a family history of breast or ovarian cancer, and no \emph{BRCA1/2} mutations, is only 3.9 times more than in the general population \cite{Metcalfe2008}.
-However, classification accuracy may be one additional informative criterion on which to evaluate solutions.
-
-\subsection{Comparison to state-of-the-art}
-\label{methods:bcac}
-
-An alternative way to evaluate the results is by comparing our results to an external dataset.
-For that purpose, we used the 153 genes associated to familial breast cancer on DisGeNET \cite{pinero_disgenet:_2017}.
-Across this article we refer to these genes as \emph{breast cancer susceptibility genes}.
-
-Additionally, we used the summary statistics from the Breast Cancer Association Consortium (BCAC), a meta-analysis of case-control studies conducted in multiple countries.
-BCAC included 13\,250\,641 SNPs genotyped or imputed on 228\,951 women of European ancestry mostly from the general population \cite{Michailidou2017}.
-Through imputation, BCAC includes more SNPs than the iCOGS array used for GENESIS (Section~\ref{methods:data}).
-Yet, in all the comparisons in this paper, we focused on the subset of the GENESIS SNPs that passed quality control.
-Hence, we used the same Bonferroni threshold as in Section~\ref{methods:conventional} to determine the significant SNPs in BCAC.
-We also computed gene-scores in the BCAC data using VEGAS2, as in Section~\ref{methods:data}.
-In this case, we did use the summary statistics of all 13\,250\,641 available SNPs, and the genotypes from European samples from the 1000 Genomes Project \cite{the_1000_genomes_project_consortium_global_2015} to compute the LD patterns.
-Since these genotypes did not include chromosome X, we excluded it from this analysis.
-All comparisons included only the genes common to GENESIS and BCAC, so we used the corresponding Bonferroni threshold ($1.66 \times 10^{-6}$) to call gene significance.
-
-\subsection{Network rewirings}
-\label{methods:rewiring}
-
-Rewiring the PPIN while preserving the number of edges of each gene allowed to study the impact of the topology on the output of network methods.
-Indeed, the edges lose their biological meaning but the topology of the network is conserved.
-We produced 100 such rewirings by swapping edges in the PPIN.
-We scored the genes as described in section \ref{methods:methods}.
-We only applied only four methods on the rewirings: heinz, dmGWAS, LEAN and SigMod.
-We excluded HotNet2 and SConES, since they take notably longer to run than the other methods.
-
-\subsection{Computational resources}
-\label{methods:computation}
-
-We ran all the computations on a Slurm cluster, running Ubuntu 16.04.2 on the nodes.
-The CPU models on the nodes were Intel Xeon CPU E5-2450 v2 at 2.50GHz and Intel Xeon E5-2440 at 2.40GHz.
-The nodes running heinz and HotNet2 had 20GB of memory; the ones running dmGWAS, LEAN, SConES, and SigMod, 60GB.
-For the benchmark (Section \ref{methods:benchmark}), we ran each of the methods on the same Ubuntu 16.04.2 node, with a CPU Intel Xeon E5-2450 v2 at 2.50GHz, and 60GB of memory.
-
-\subsection{Code and data availability}
-\label{methods:code}
-
-We developed computational pipelines for several steps of GWAS analyses, such as physically mapping SNPs to genes, computing gene scores, and performing six different network analyses.
-For each of those processes, we created a pipeline with a clear interface that should work on any GWAS dataset.
-They are compiled in \url{https://github.com/hclimente/gwas-tools}.
-The code to apply the pipelines to GENESIS, as well as the code that reproduces all the analyses in this article are available at \url{https://github.com/hclimente/genewa}.
-We deposited all the produced gene solutions on NDEx (\url{http://www.ndexbio.org}), under the UUID e9b0e22a-e9b0-11e9-bb65-0ac135e8bacf.
-
-Summary statistics for SNPs and genes are available at \url{https://github.com/hclimente/genewa}.
-We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact \href{mailto:nadine.andrieu@curie.fr}{nadine.andrieu(at)curie.fr}.
+To make it easier for others to apply these methods to new datasets and aggregate their solutions, we built six \texttt{nextflow} pipelines \cite{di_tommaso_nextflow_2017} with a consistent interface and, whenever possible, parallelized computation.
+They are available on GitHub: \href{https://github.com/hclimente/gwas-tools}{hclimente/gwas-tools} (Section~\ref{methods:code}).
+Importantly, we compiled those methods with a permissive license into a Docker image for easier use, available on Docker Hub \href{https://hub.docker.com/r/hclimente/gwas-tools}{hclimente/gwas-tools}.
 
@@
 \label{stab:snp_solutions}
-{\bf Summary statistics on the results of SConES on the three SNP-SNP interaction networks.} The first row within each block contains the summary statistics on the whole network.
+\textbf{Summary statistics on the results of SConES on the three SNP networks (Section~\ref{methods:networks}).} The first row within each block contains the summary statistics on the whole network.
 
 \paragraph*{S2 Table.}
-\label{stab:scones_gene_solutions}
-{\bf Summary statistics on the results of multiple network methods on the PPIN.} The first row contains the summary statistics on the whole PPIN.
+\label{stab:sigmod_pwy}
+\textbf{Pathway enrichment analyses of the genes in SigMod's solution.}
 
 \paragraph*{S3 Table.}
-\label{stab:sigmod_pwy}
-{\bf Pathway enrichment analyses of the genes in SigMod solution.}
+\label{stab:dmgwas_pwy}
+\textbf{Pathway enrichment analyses of the genes in dmGWAS' solution.}
 
 \paragraph*{S4 Table.}
-\label{stab:dmgwas_pwy}
-{\bf Pathway enrichment analyses of the genes in dmGWAS solution.}
+\label{stab:hotnet2_pwy}
+\textbf{Pathway enrichment analyses of the genes in HotNet2's solution.}
 
 \paragraph*{S5 Table.}
-\label{stab:hotnet2_pwy}
-{\bf Pathway enrichment analyses of the genes in HotNet2 solution.}
-
-\paragraph*{S6 Table.}
 \label{stab:consensus_pwy}
-{\bf Pathway enrichment analyses of the genes in the consensus solution.}
+\textbf{Pathway enrichment analyses of the genes in the consensus' solution.}
 
@@
 \label{sfig:pcs}
-{\bf GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)}~Eight main principal components computed on the genotypes of GENESIS.
+\textbf{GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)}~Eight main principal components, computed on the genotypes of GENESIS.
 Cases are colored in green, controls in orange.
@@
 \label{sfig:snp_gene_manhattan}
-{\bf Association in GENESIS.
-The red line represents the Bonferroni threshold.} \textbf{(A)}~SNP association, measured from the outcome of a 1 d.f.
+\textbf{Association in GENESIS.
+The red lines represent the Bonferroni thresholds.} \textbf{(A)}~SNP association, measured from the outcome of a 1 d.f.
 $\chi^2$ allelic test (Section~\ref{methods:conventional}).
-Significant SNPs that are within a coding gene, or within 50 kilobases of its boundaries, are annotated.
+Significant SNPs within a coding gene, or within 50 kilobases of its boundaries, are annotated.
 The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
@@
 \textbf{(C)}~SNP association as in panel (A).
-The SNPs in black are selected by a L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).
+The SNPs in black were selected by an L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).
 %\includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
@@
 \paragraph*{S3 Fig.}
-\label{sfig:pearson_methods}
-{\bf Pearson correlation between the different solutions.
-} \textbf{(A)}~Correlation between selected SNPs.
-\textbf{(B)}~Correlation between selected genes.
-In general, the solutions display a very low overlap.
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_3.pdf}
+\label{sfig:consensus_stats}
+\textbf{Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus solution.} The blue line represents a fitted generalized linear model.
+% \includegraphics[width=.65\textwidth]{./figures/sfigure_3.pdf}
 
 \paragraph*{S4 Fig.}
-\label{sfig:consensus_stats}
-{\bf Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus solution.} The blue line represents a fitted generalized linear model.
-% \includegraphics[width=.65\textwidth]{./figures/sfigure_5.pdf}
-
-\paragraph*{S5 Fig.}
 \label{sfig:additional-benchmarks}
-{\bf Additional benchmarks of the network methods} \textbf{(A)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
+\textbf{Additional benchmarks of the network methods.} \textbf{(A)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
 For reference, we added a gray line with a slope of 1.
@@
 Points are the average over the 5 runs; the error bars represent the standard error of the mean.
-% \includegraphics[width=.675\linewidth]{./figures/sfigure_6.pdf
+% \includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf
+
+\paragraph*{S5 Fig.}
+\label{sfig:pearson_methods}
+\textbf{Pearson correlation between the different solutions.} \textbf{(A)}~Correlation between selected SNPs.
+\textbf{(B)}~Correlation between selected genes.
+In general, the solutions display a very low overlap.
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
 
@@
 \label{sfig:rewirings}
-{\bf Number of times a gene was selected by either dmGWAS, heinz, LEAN,or SigMod in 100 rewirings of the PPIN (Section \ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
+\textbf{Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf
+
+\paragraph*{S7 Fig.}
+\label{sfig:params_size}
+\textbf{Size of the solutions obtained under different parameters.} \textbf{(A)}~Size of the solution produced by different parameter values, expressed as a percentage of the maximum solution size for the method, or the highest tested value for the parameter, respectively.
+The size of the solution is the median among all the solution sizes for the same parameter.
+\textbf{(B)}~Boxplot of the solution sizes of the methods under the explored parameters (Section~\ref{methods:parameters}).
+% \includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf
+
+\paragraph*{S8 Fig.}
+\label{sfig:params_pearson}
+\textbf{Pearson correlation between the solutions obtained under different parameters, computed as in Section~\ref{methods:algorithm_comparison}.} Grey tiles represent the cases where we could not compute the Pearson correlation because the two vectors were either all ones (all genes/SNPs were selected) or zeros (no genes/SNPs were selected).
 % \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
 
-\paragraph*{S7 Fig.}
-\label{sfig:parameters}
-{\bf Parameter space of the network methods.} \textbf{(A)} Size of the solution produced by different parameter values, expressed as a percentage of the maximum solution size for the method, or the highest tested value for the parameter, respectively.
-The size of the solution is the median among all the solution sizes for the same parameter.
-\textbf{(B)} Boxplot of the solution sizes of the methods under the explored parameters (Section \ref{methods:parameters}).
-\textbf{(C)} Pearson correlation between the solutions of the different runs.
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
-
-\paragraph*{S8 Fig.}
+\paragraph*{S9 Fig.}
 \label{sfig:biotypes_excluded}
-{\bf Biotypes of genes from the annotation that are not present in the HINT PPIN.}
+\textbf{Biotypes of genes from the annotation that are not present in the HINT PPIN.}
 % \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf}
 
-\paragraph*{S9 Fig.}
+\paragraph*{S10 Fig.}
 \label{sfig:lc_ht_comparison}
-{\bf Comparison of the benchmark on high-throughput (HT) interactions to the benchmark on both high-throughput and literature curated interactions (HT+LC).
-} Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
+\textbf{Comparison of the benchmark on high-throughput (HT) interactions to the benchmark on both high-throughput and literature curated interactions (HT+LC).} Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
 \textbf{(A)}~Ratios of the selected features between both benchmarks and of the active set (Section~\ref{methods:classifier}).
@@
 For gene-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
-In general, adding additional interactions slightly improves the stability of the solution, but increases the solution size, has mixed effects on the sensitivity and specificity, and impacts negatively the required runtime of the algorithms.
+In general, adding additional interactions slightly improved the stability of the solution.
+However, it increased the solution size and the required runtime, and had mixed effects on the sensitivity and specificity.
 % \includegraphics[width=.85\linewidth]{./figures/sfigure_9.pdf}
 
-\paragraph*{S10 Fig.}
+\paragraph*{S11 Fig.}
 \label{sfig:scones_gsm}
-{\bf Overview of the solutions produced by the SConES on the GS and GM networks (Section \ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)}~Manhattan plots of SNPs (Section \ref{methods:conventional}); in black, the method’s solution.
-The Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}) is indicated by a red line.
+\textbf{Overview of the solutions produced by the SConES on the GS and GM networks (Section~\ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)}~Manhattan plots of SNPs (Section~\ref{methods:conventional}); in black, the method’s solution.
+The red line indicates the Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}).
 \textbf{(B)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
@@
 Mebirouk, E.
-Cavaciuti), the biological resource centre (S.
+Cavaciuti), the biological resource center (S.
 Mazoyer, F.
@@
 Sornin).
-We wish to pay a tribute to Olga M.
-Sinilnikova, who was one of the initiators and principal investigators of GENESIS and who died prematurely on June 30, 2014.
+We wish to pay tribute to Olga M.
+Sinilnikova, one of the initiators and principal investigators of GENESIS, and who died prematurely on June 30, 2014.
 

### 2021-01-14 manuscript/plos/review_2.tex
revisions 2: added manuscript and some changes

--- 2020-10-18 manuscript/plos/review_1.tex
@@
 Fabienne Lesueur\textsuperscript{1,2,3},
+GENESIS study group\textsuperscript{\textpilcrow},
 Dominique Stoppa-Lyonnet\textsuperscript{5,6,7\textcurrency}, 
 Nadine Andrieu\textsuperscript{1,2,3}, 
-Chloé-Agathe Azencott\textsuperscript{3,1,2},
-GENESIS study group\textsuperscript{\textpilcrow}
+Chloé-Agathe Azencott\textsuperscript{3,1,2}
 \\
@@
 After a classical GWAS approach, we used these network methods to identify additional breast cancer susceptibility genes.
-Lastly, we compared the solutions obtained by the different methods and aggregated them to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
+Lastly, we compared the solutions obtained by the different methods and computed their intersection to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
 
@@
 The threshold used was $\frac {0.05} {197083} = 2.54 \times 10^{-7}$.
+The summary statistics of the analyzed SNPs are available in \nameref{stab:snp_summary}.
 
@@
 We also performed a Bonferroni correction to obtain significant genes; in this case, the threshold of significance was $\frac {0.05} {32767} = 1.53 \times 10^{-6}$.
+The summary statistics of the analyzed genes are available in \nameref{stab:gene_summary}.
 
@@
 
-We searched for pathways enriched in the gene solutions produced by the above methods.
+We searched for pathways enriched in the solution of each network method.
 We conducted a hypergeometric test on pathways from Reactome \cite{Jassal2019} using the function \texttt{enrichPathway} from the \texttt{ReactomePA} R package \cite{Yu2016}.
@@
 We still scored the genes as described in Section~\ref{methods:methods}.
-We only applied only four methods on the rewirings: heinz, dmGWAS, LEAN, and SigMod.
+We applied only four methods on the rewirings: heinz, dmGWAS, LEAN, and SigMod.
 We excluded HotNet2 and SConES since they took notably longer to run.
+As on the real PPIN, LEAN did not produce significant results in the rewirings.
 
@@
 
-Lastly, a pathway enrichment analysis (Section~\ref{methods:pathway_enrichment}) also showed similarities and differences between the methods' solutions.
+Lastly, pathway enrichment analyses on each method's solution (Section~\ref{methods:pathway_enrichment}) also revealed similarities and differences between them.
 It linked different parts of SigMod's solution to four processes (\nameref{stab:sigmod_pwy}): protein translation (including mitochondrial), mRNA splicing, protein misfolding, and keratinization (BH adjusted P\=/values~$<$~0.03).
 Interestingly, the dmGWAS solution (\nameref{stab:dmgwas_pwy}) was also related to protein misfolding (\emph{attenuation phase}, BH adjusted P\=/value~=~0.01).
-However, it additionally included submodules of proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
+However, it additionally included proteins related to mitosis, DNA damage, and regulation of TP53 (BH adjusted P\=/values~$<$~0.05), which match previously known mechanisms of breast cancer susceptibility \cite{nielsen_hereditary_2016}.
 As with SigMod, the genes in HotNet2's solution (\nameref{stab:hotnet2_pwy}) were involved in mitochondrial translation (BH adjusted P\=/value~=~1.87 \texttimes{} 10\textsuperscript{-4}), but also in glycogen metabolism and transcription of nuclear receptors (BH adjusted P\=/value~$<$~0.04).
@@
   \centering
-  \includegraphics[width=.9\linewidth]{./figures/figure_3.pdf}
+  \includegraphics[width=.9\linewidth]{./figures/figure_2.pdf}
   \caption{ \textbf{Consensus solution on GENESIS (Section~\ref{methods:methods}).} \textbf{(A)}~Manhattan plot of genes; in black, the ones in the consensus solution.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
+  \includegraphics[width=.8\linewidth]{./figures/figure_3.pdf}
   \caption{\textbf{Comparison of network methods on GENESIS.} Each method was run 5 times on a random subset containing 80\% of the samples and tested on the remaining samples (Section~\ref{methods:benchmark}).
@@
 For gene-based methods, inverted triangles represent the runtime of the algorithm alone, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
-\textbf{(D)}~True positive rate and true negative rate of the methods, obtained using different parameter combinations (Section~\ref{results:parameters}).
+\textbf{(D)}~True positive rate and false positive rate of the methods, obtained using different parameter combinations (Section~\ref{results:parameters}).
 We used as true positives BCAC-significant SNPs (for SConES and $\chi^2$ + Bonferroni) and genes (for the remaining methods, Section~\ref{methods:bcac}).
@@
   \centering
-  \includegraphics[width=.8\linewidth]{./figures/figure_2.pdf}
+  \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
   \caption{\textbf{Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
@@
 Importantly, SConES produced similar results on the GS and GM networks (\nameref{sfig:pearson_methods}).
-While the solutions on those two considerably overlap with SConES GI's, they contained additional gene-coding segments (Fig~\ref{fig:issues}C).
+While the solutions on those two considerably overlap with SConES GI's, they contained additional gene-coding segments (Fig~\ref{fig:issues}D).
 In fact, both SConES GS and GM selected chromosome regions related to breast cancer, like 3p24 (\emph{SLC4A7}/\emph{NEK10} \cite{ahmed_newly_2009}), 5p12 (\emph{FGF10}, \emph{MRPS30} \cite{quigley_5p12_2014}), 10q26 (\emph{FGFR2}), and 16q12 (\emph{TOX3}).
@@
 
-To overcome these limitations while exploiting the each method's strengths, we proposed combining them into a consensus solution.
+To overcome these limitations while exploiting each method's strengths, we proposed combining them into a consensus solution.
 We used the straightforward strategy of including any node that was recovered by at least two methods.
@@
 \paragraph*{S1 Table.}
+\label{stab:snp_summary}
+\textbf{SNP summary statistics on GENESIS.}
+
+\paragraph*{S2 Table.}
+\label{stab:gene_summary}
+\textbf{Gene summary statistics on GENESIS.}
+
+\paragraph*{S3 Table.}
 \label{stab:snp_solutions}
@@
 
-\paragraph*{S2 Table.}
+\paragraph*{S4 Table.}
 \label{stab:sigmod_pwy}
@@
 
-\paragraph*{S3 Table.}
+\paragraph*{S5 Table.}
 \label{stab:dmgwas_pwy}
@@
 
-\paragraph*{S4 Table.}
+\paragraph*{S6 Table.}
 \label{stab:hotnet2_pwy}
@@
 
-\paragraph*{S5 Table.}
+\paragraph*{S7 Table.}
 \label{stab:consensus_pwy}
@@
 Cases are colored in green, controls in orange.
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
 
@@
 \textbf{Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus solution.} The blue line represents a fitted generalized linear model.
-% \includegraphics[width=.65\textwidth]{./figures/sfigure_3.pdf}
+%\includegraphics[width=.65\textwidth]{./figures/sfigure_3.pdf}
 
@@
 Points are the average over the 5 runs; the error bars represent the standard error of the mean.
-% \includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf
+%\includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf
 
 \paragraph*{S5 Fig.}
+\label{sfig:rewirings}
+\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf
+
+\paragraph*{S6 Fig.}
 \label{sfig:pearson_methods}
@@
 
-\paragraph*{S6 Fig.}
-\label{sfig:rewirings}
-\textbf{Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf
-
 \paragraph*{S7 Fig.}
@@
 \textbf{(B)}~Boxplot of the solution sizes of the methods under the explored parameters (Section~\ref{methods:parameters}).
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf
 
@@
 \textbf{Pearson correlation between the solutions obtained under different parameters, computed as in Section~\ref{methods:algorithm_comparison}.} Grey tiles represent the cases where we could not compute the Pearson correlation because the two vectors were either all ones (all genes/SNPs were selected) or zeros (no genes/SNPs were selected).
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
 
 \paragraph*{S9 Fig.}
-\label{sfig:biotypes_excluded}
-\textbf{Biotypes of genes from the annotation that are not present in the HINT PPIN.}
-% \includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf}
-
-\paragraph*{S10 Fig.}
 \label{sfig:lc_ht_comparison}
@@
 However, it increased the solution size and the required runtime, and had mixed effects on the sensitivity and specificity.
-% \includegraphics[width=.85\linewidth]{./figures/sfigure_9.pdf}
+%\includegraphics[width=.85\linewidth]{./figures/sfigure_10.pdf}
+
+\paragraph*{S10 Fig.}
+\label{sfig:biotypes_excluded}
+\textbf{Biotypes of genes from the annotation that are not present in the HINT PPIN.}
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_9.pdf}
 
@@
 \textbf{(C)}~Solution networks.
-% \includegraphics[width=\linewidth]{./figures/sfigure_10.png}
+%\includegraphics[width=\linewidth]{./figures/sfigure_11.png}
 
@@
 
-We thank all the GENESIS collaborating cancer clinics clinics (Clinique Sainte Catherine, Avignon: H.
+We thank all the GENESIS collaborating cancer clinics (Clinique Sainte Catherine, Avignon: H.
 Dreyfus; Hôpital Saint Jacques, Besançon: M-A.

### 2021-01-15 manuscript/plos/review_2.tex
updated figures with normalized betweenness
reworked abstract and title

@@
 {\Large
-\textbf\newline{Biological networks and GWAS: comparing and combining network methods to understand the genetics of familial breast cancer susceptibility in the GENESIS study} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{How can biological networks boost GWAS?
+A case study on familial breast cancer susceptibility} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 \section*{Abstract}
-Network approaches to disease use biological networks, which model functional relationships between the molecules in a cell, to generate hypotheses about the genetics of complex diseases.
-Several among them jointly consider gene scores, representing the association between each gene and the disease, and the biological context of each gene, modeled by a network.
-Here, we study six such network methods using gene scores from GENESIS, a genome-wide association study (GWAS) on French women with non-BRCA familial breast cancer.
-We provide a critical comparison of these six methods, discussing the impact of their mathematical formulation and parameters.
-Using a biological network yields more compelling results than standard GWAS analyses.
-Indeed, we find significant overlaps between our solutions and the genes identified in the largest GWAS on breast cancer susceptibility.
-We further propose to combine these solutions into a consensus network, which brings further insights.
-The consensus network contains \emph{COPS5}, a gene related to multiple hallmarks of cancer, and 14 of its neighbors.
-The main drawback of network methods is that they are not robust to small perturbations in their inputs.
-Therefore, we propose a stable consensus solution, formed by the most consistently selected genes in multiple subsamples of the data.
-In GENESIS, it is composed of 68 genes, enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$) and occupying more central positions in the network than most genes.
-The network is organized around \emph{CUL3}, which is involved in the regulation of several genes linked to cancer progression.
-In conclusion, we showed how network methods help overcome the lack of statistical power of GWAS and improve their interpretation.
-Project-agnostic implementations of all methods are available at \url{https://github.com/hclimente/gwas-tools}.
+Genome-wide association studies (GWAS) explore the genetic causes of complex diseases.
+However, they ignore the biological context of the SNPs and genes under study.
+To address this, one can seek susceptibility genes that are functionally related, using networks to integrate this functional information.
+Many such network methods exist, stemming from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism.
+Unsurprisingly, their results are also very different.
+To sort through this diversity, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer.
+First, we verified that network methods recovered more interesting and interpretable genes than standard GWAS.
+But, as expected, these results were notably different.
+To address this heterogeneity, we combined them into a \emph{consensus}.
+The key gene in the consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks.
+Another drawback was that network methods were unstable, selecting very different genes on different subsamples of GENESIS.
+Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected in multiple subsamples.
+Importantly, this solution was also enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
+The most central gene was \emph{CUL3}, which regulates several genes linked to cancer progression.
+Lastly, we evaluated the biases of each method and the impact of the parameters on the outcome.
+In general, network methods preferred highly central genes, even in random rewirings that stripped networks of any biological meaning.
+In conclusion, we showed the advantages of network-guided GWAS, characterized their shortcomings, and provided strategies to address them.
+To compute the consensus networks, implementations of all six methods are available at \url{https://github.com/hclimente/gwas-tools}.
 

### 2021-01-18 manuscript/plos/review_2.tex
replaced betweenness by normalized betweenness

@@
 \textbf\newline{How can biological networks boost GWAS?
-A case study on familial breast cancer susceptibility} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+A case study on susceptibility to familial breast cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 However, they ignore the biological context of the SNPs and genes under study.
-To address this, one can seek susceptibility genes that are functionally related, using networks to integrate this functional information.
-Many such network methods exist, stemming from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism.
-Unsurprisingly, their results are also very different.
+To address this shortcoming, one can seek susceptibility genes that are functionally related, employing networks to integrate such functional information.
+Many such network methods exist, each stemming from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism.
+Unsurprisingly, their results are also disparate.
 To sort through this diversity, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer.
-First, we verified that network methods recovered more interesting and interpretable genes than standard GWAS.
-But, as expected, these results were notably different.
-To address this heterogeneity, we combined them into a \emph{consensus}.
-The key gene in the consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks.
-Another drawback was that network methods were unstable, selecting very different genes on different subsamples of GENESIS.
-Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected in multiple subsamples.
-Importantly, this solution was also enriched in known breast cancer susceptibility genes (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
-The most central gene was \emph{CUL3}, which regulates several genes linked to cancer progression.
+First, we verified that network methods recovered more interesting and interpretable genes than a standard GWAS.
+Nevertheless, as expected, these results were notably different.
+We addressed this heterogeneity studying their overlap, computing what we called the \emph{consensus}.
+The key gene in this consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks.
+Another drawback of network methods was their instability, selecting very different genes on different subsamples of GENESIS.
+Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected across multiple subsamples.
+Importantly, this solution was also enriched in genes associated with breast cancer susceptibility (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
+The most connected gene was \emph{CUL3}, a regulator of several genes linked to cancer progression.
 Lastly, we evaluated the biases of each method and the impact of the parameters on the outcome.
-In general, network methods preferred highly central genes, even in random rewirings that stripped networks of any biological meaning.
-In conclusion, we showed the advantages of network-guided GWAS, characterized their shortcomings, and provided strategies to address them.
+In general, network methods preferred highly connected genes, even after random rewirings that stripped the connections of any biological meaning.
+In conclusion, we presented the advantages of network-guided GWAS, characterized their shortcomings, and provided strategies to address them.
 To compute the consensus networks, implementations of all six methods are available at \url{https://github.com/hclimente/gwas-tools}.
@@
 After a classical GWAS approach, we used these network methods to identify additional breast cancer susceptibility genes.
-Lastly, we compared the solutions obtained by the different methods and computed their intersection to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
+Lastly, we compared the solutions obtained by the different methods and studied their intersection to obtain consensus solutions of predisposition to familial breast cancer that addressed their shortcomings.
 
@@
 In this article, we used undirected, vertex-weighted networks, or graphs, $G = (V,E,w)$.
-$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices, with weights $w: V \rightarrow \mathbb{R}$.
+$V = \{v_{1}, \dots{}, v_{n}\}$ refers to the vertices (or nodes), with weights $w: V \rightarrow \mathbb{R}$.
 Equivalently, $E \subseteq \{\{x,y\} | x,y \in V \wedge x \neq y\}$ refers to the edges.
@@
 Nodes can be described by properties provided by the topology of the graph.
-We focused on two of those: degree centrality and betweenness centrality.
-The degree centrality, or simply emph{degree}, is the number of edges that a node has.
-The betweenness centrality, or \emph{betweenness}, is the number of times a node participates in shortest paths between two other nodes.
+We focused on the betweenness centrality, or \emph{betweenness}: the number of times a node participates in a shortest path between two other nodes.
+We normalized the betweenness by dividing it by $\frac {(N-1)(N-2)} {2}$, where $N$ is the number of nodes.
 
@@
 The \emph{adjacency matrix} $W_G$ contains a 1 when the corresponding nodes are connected, and 0 otherwise; its diagonal is zero.
-The \emph{degree matrix} $D_G$ is a diagonal matrix that contains the degree of the different nodes.
+The \emph{degree matrix} $D_G$ is a diagonal matrix that contains the number of edges of each node.
 
@@
 We excluded HotNet2 and SConES since they took notably longer to run.
-As on the real PPIN, LEAN did not produce significant results in the rewirings.
+As on the real PPIN, LEAN did not produce significant results on the rewirings either.
 
@@
 
-The topologies of the five solutions differed as well (Fig~\ref{fig:solution_overview}C), as measured by the median centrality and the number of connected components (Table~\ref{tab:gene_solutions}).
+The topologies of the five solutions differed as well (Fig~\ref{fig:solution_overview}C), as measured by the median betweenness and the number of connected components (Table~\ref{tab:gene_solutions}).
 Three methods yielded more than one connected component: SConES, as described above, SigMod, and HotNet2.
@@
   \thickhline
-  HINT HT           & 13\,619 & 142\,541  & 15  & 16\,706   & 0.46  & 93/93 \\
+  HINT HT           & 13\,619 & 142\,541  & 15  & $1.8 \times 10^{-4}$ & 0.46  & 93/93 \\
   \hline
-  dmGWAS            & 194     & 450       & 1   & 49\,115   & 0.19  & 55/93 \\
-  heinz             & 4       & 3         & 1   & 113\,633  & 0.001 & 4/93  \\
-  HotNet2           & 440     & 374       & 130 & 7\,739    & 0.048 & 63/93 \\
+  dmGWAS            & 194     & 450       & 1   & $5.3 \times 10^{-4}$ & 0.19  & 55/93 \\
+  heinz             & 4       & 3         & 1   & $1.2 \times 10^{-3}$ & 0.001 & 4/93  \\
+  HotNet2           & 440     & 374       & 130 & $8.3 \times 10^{-5}$ & 0.048 & 63/93 \\
   LEAN              & 0       & 0         & 0   & -         & -     & 0/93  \\
   SConES GI         & 0 (1)   & 0         & 0   & -         & -     & 0/93  \\
-  SigMod            & 142     & 249       & 11  & 92\,603   & 0.008 & 84/93 \\
+  SigMod            & 142     & 249       & 11  & $1 \times 10^{-3}$   & 0.008 & 84/93 \\
   \hline
-  Consensus         & 93      & 186       & 21  & 50\,737   & 0.006 & 93/93 \\
-  Stable consensus  & 68      & 49        & 32  & 94\,854   & 0.005 & 43/93 \\
+  Consensus         & 93      & 186       & 21  & $5.5 \times 10^{-4}$ & 0.006 & 93/93 \\
+  Stable consensus  & 68      & 49        & 32  & $1 \times 10^{-3}$   & 0.005 & 43/93 \\
   \end{tabular}
@@
 \textbf{\# components}:~number of connected components.
-\textbf{\mean{Betweenness}}:~median betweenness of the selected genes in the PPIN.
+\textbf{\mean{Betweenness}}:~median (normalized) betweenness of the selected genes in the PPIN.
 \textbf{\median{P}\textsubscript{gene}}:~median VEGAS2 P\=/value of the selected genes.
@@
 Hence, to leverage their strengths and compensate for their respective weaknesses, we built a consensus solution using the genes shared among at least two solutions (Section~\ref{methods:methods}).
-This solution (Fig~\ref{fig:consensus}) contained 93 genes and exhibited the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
+This solution (Fig~\ref{fig:consensus}A) contained 93 genes and exhibited the aforementioned properties of the individual solutions: enrichment in breast cancer susceptibility genes and higher betweenness centrality than the rest of the genes.
 
@@
   \caption{\textbf{Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
-\textbf{(B)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its centrality degree.
-\textbf{(C)}~Centrality degree and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
+\textbf{(B)}~Number of times a gene was selected by either dmGWAS, heinz, LEAN, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:pathway_enrichment}) and its betweenness.
+\textbf{(C)}~Betweenness and -log\textsubscript{10} of the VEGAS2 P\=/value in BCAC for each of the nodes in the PPIN.
 We highlighted the genes selected by each method and the ones selected by more than one (``Consensus'').
@@
 
-Another peculiarity of network methods was their relationship to degree centrality.
+Another peculiarity of network methods was their relationship to betweenness centrality.
 We studied random rewirings of the PPIN that preserved node centrality (Section~\ref{methods:rewiring}).
@@
 We used only genes selected in at least 7 solutions, which corresponded to 1\% of all genes selected at least once.
-The resulting stability-based consensus was composed of 68 genes (Fig~\ref{fig:stable-consensus}).
+The resulting stability-based consensus was composed of 68 genes (Fig~\ref{fig:stable-consensus}A).
 This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig~\ref{fig:stable-consensus}B and C).
@@
 \label{sfig:rewirings}
-\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its centrality degree.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
+\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its betweenness.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
 %\includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf
@@
 
-We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S.
+We wish to thank Om Kulkarni for helpful discussion on gene-based GWAS and PPIN databases, Vivien Goepp for his constructive comments of the manuscript, and the genetic epidemiology platform (the PIGE, Plateforme d'Investigation en Génétique et Epidemiologie: S.
 Eon-Marchais, M.

### 2021-01-19 manuscript/plos/review_2.tex
updated figures to refer only to normalized betweenness

@@
 Genome-wide association studies (GWAS) explore the genetic causes of complex diseases.
-However, they ignore the biological context of the SNPs and genes under study.
+However, they ignore the biological context of the genetic variants and genes under study.
 To address this shortcoming, one can seek susceptibility genes that are functionally related, employing networks to integrate such functional information.
@@
 The key gene in this consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks.
-Another drawback of network methods was their instability, selecting very different genes on different subsamples of GENESIS.
+Another issue of network methods was their instability, selecting very different genes on different subsamples of GENESIS.
 Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected across multiple subsamples.

### 2021-01-26 manuscript/plos/review_2.tex
added chloe's comments

@@
 However, they ignore the biological context of the genetic variants and genes under study.
-To address this shortcoming, one can seek susceptibility genes that are functionally related, employing networks to integrate such functional information.
-Many such network methods exist, each stemming from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism.
-Unsurprisingly, their results are also disparate.
-To sort through this diversity, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer.
-First, we verified that network methods recovered more interesting and interpretable genes than a standard GWAS.
-Nevertheless, as expected, these results were notably different.
-We addressed this heterogeneity studying their overlap, computing what we called the \emph{consensus}.
+To address this shortcoming, one can use biological networks, which model functional relationships, to search for functionally related susceptibility loci.
+Many such network methods exist, each arising from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism.
+Unsurprisingly, this results in disparate solutions.
+To explore how to exploit these heterogeneous approaches, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer.
+First, we verified that network methods recovered more interpretable genes than a standard GWAS.
+We addressed the heterogeneity of their solutions by studying their overlap, computing what we called the \emph{consensus}.
 The key gene in this consensus solution was \emph{COPS5}, a gene related to multiple cancer hallmarks.
-Another issue of network methods was their instability, selecting very different genes on different subsamples of GENESIS.
+Another issue we observed was that network methods were unstable, selecting very different genes on different subsamples of GENESIS.
 Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected across multiple subsamples.
-Importantly, this solution was also enriched in genes associated with breast cancer susceptibility (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
+This solution was also enriched in genes associated with breast cancer susceptibility (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
 The most connected gene was \emph{CUL3}, a regulator of several genes linked to cancer progression.
-Lastly, we evaluated the biases of each method and the impact of the parameters on the outcome.
+Lastly, we evaluated the biases of each method and the impact of their parameters on the outcome.
 In general, network methods preferred highly connected genes, even after random rewirings that stripped the connections of any biological meaning.

### 2021-01-28 manuscript/plos/review_2.tex
revision 2 submitted

@@
 {\Large
-\textbf\newline{How can biological networks boost GWAS?
-A case study on susceptibility to familial breast cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Boosting GWAS using biological networks: a study on susceptibility to familial breast cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 Genome-wide association studies (GWAS) explore the genetic causes of complex diseases.
-However, they ignore the biological context of the genetic variants and genes under study.
+However, classical approaches ignore the biological context of the genetic variants and genes under study.
 To address this shortcoming, one can use biological networks, which model functional relationships, to search for functionally related susceptibility loci.
-Many such network methods exist, each arising from different mathematical frameworks, pre-processing steps, and assumptions about the susceptibility mechanism.
+Many such network methods exist, each arising from different mathematical frameworks, pre-processing steps, and assumptions about the network properties of the susceptibility mechanism.
 Unsurprisingly, this results in disparate solutions.
-To explore how to exploit these heterogeneous approaches, we selected six network methods and applied them to GENESIS, a GWAS on French women with non-BRCA familial breast cancer.
-First, we verified that network methods recovered more interpretable genes than a standard GWAS.
+To explore how to exploit these heterogeneous approaches, we selected six network methods and applied them to GENESIS, a nationwide French study on familial breast cancer.
+First, we verified that network methods recovered more interpretable results than a standard GWAS.
 We addressed the heterogeneity of their solutions by studying their overlap, computing what we called the \emph{consensus}.
@@
 Therefore, we proposed a \emph{stable consensus} solution formed by the 68 genes most consistently selected across multiple subsamples.
-This solution was also enriched in genes associated with breast cancer susceptibility (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
+This solution was also enriched in genes known to be associated with breast cancer susceptibility (\emph{BLM}, \emph{CASP8}, \emph{CASP10}, \emph{DNAJC1}, \emph{FGFR2}, \emph{MRPS30}, and \emph{SLC4A7}, P\=/value~=~$3 \times 10^{-4}$).
 The most connected gene was \emph{CUL3}, a regulator of several genes linked to cancer progression.
@@
 In general, network methods preferred highly connected genes, even after random rewirings that stripped the connections of any biological meaning.
-In conclusion, we presented the advantages of network-guided GWAS, characterized their shortcomings, and provided strategies to address them.
+In conclusion, we present the advantages of network-guided GWAS, characterize their shortcomings, and provide strategies to address them.
 To compute the consensus networks, implementations of all six methods are available at \url{https://github.com/hclimente/gwas-tools}.
@@
 Genome-wide association studies (GWAS) scan thousands of genomes to identify variants associated with a complex trait.
-Over the last 15 years, GWAS have advanced our understanding of the genetics of complex diseases, and in particular of hereditary cancers.
+Over the last 15 years, GWAS have advanced our understanding of the genetics of complex diseases, and in particular of cancers.
 However, they have led to an apparent paradox: the more we perform such studies, the more it seems that the entire genome is involved in every disease.
@@
 Then, a statistical test of association (e.g., based on logistic regression) is conducted between each SNP and the phenotype.
-Those SNPs with a P\=/value lower than a conservative Bonferroni threshold are candidates to further studies in an independent cohort.
+Those SNPs with a P\=/value lower than a conservative Bonferroni threshold are candidates to further studies in independent cohorts.
 Once the risk SNPs have been discovered, they can be used for risk assessment and deepening our understanding of the disease.
@@
 
-Network-based discovery methods exploit the differential properties described above to identify disease genes on GWAS data \cite{azencott_network-guided_2016,huang_systematic_2018}.
+Network-based discovery methods exploit the differential properties described above to identify disease genes using GWAS data \cite{azencott_network-guided_2016,huang_systematic_2018}.
 In essence, each gene receives a score of association with the disease, computed from the GWAS data, and a set of biological relationships, given by a network built on prior knowledge.
@@
 Lastly, based on study selection criteria, 11 other samples were removed (1 control having cancer, 4 index cases with no affected sister, 3 half-sisters, 1 sister with lobular carcinoma \emph{in situ}, 1 with a \emph{BRCA1} or \emph{BRCA2} pathogenic variant detected in the family, 1 with unknown molecular diagnosis).
-The final dataset included 1\,271 controls and 1\,280 cases, genotyped over 197\,083 SNPs.
+The final dataset included 1\,271 controls, 1\,280 cases, and 197\,083 SNPs.
 
@@
 
-\paragraph{Gene network}
+\paragraph{Gene networks}
 The mathematical formulations of the different network methods are compatible with any type of biological network (e.g., from protein interactions or gene co-expression).
-Here, we used protein-protein interaction networks (PPIN) for all network methods except SConES, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
+Here, we used protein-protein interaction networks (PPIN) for all gene-centric network methods, as PPINs are interpretable, well-characterized, and the methods were designed to run efficiently on them.
 We built our PPIN from both binary and co-complex interactions stored in the HINT database (release April 2019) \cite{das_hint:_2012}.
@@
 We assigned the initial node scores as in Nakka et al.
-\cite{nakka_gene_2016}, giving a 0 to the genes with a VEGAS2 P-values of association with the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
+\cite{nakka_gene_2016}, giving a 0 to the genes unlikely to be truly associated with the disease, and -log\textsubscript{10}(P\=/value) to those likely to be.
 In the GENESIS dataset, the threshold separating both was a P\=/value of 0.125, which we obtained using a local FDR approach \cite{scheid_twilight;_2005}.
@@
 In this case, the selected parameters were \(\eta\)~=~3.51, \(\lambda\)~=~210.29 for SConES GS; \(\eta\)~=~3.51, \(\lambda\)~=~97.61 for SConES GM; and \(\eta\)~=~3.51, \(\lambda\)~=~45.31 for SConES GI.
-We used the version on SConES implemented in the R package \texttt{martini} \cite{martini}.
+We used the version on SConES implemented in the R package \texttt{martini} \cite{martini,climente-gonzalezMartiniPackageGenomewide2021}.
 
@@
 
-\subsubsection{Comparing SNP-methods to gene-methods and vice versa}
+\subsubsection{Comparing SNP-based methods to gene-based methods and vice versa}
 \label{methods:snp2gene}
@@
 We used the 5 repetitions to estimate the average and the standard deviation of the different measures.
-Every method and repetition ran in the same computational settings (Section~\ref{methods:computation}).
+Every method and repetition was run in the same computational settings (Section~\ref{methods:computation}).
 
@@
 BCAC included 13\,250\,641 SNPs genotyped or imputed on 228\,951 women of European ancestry, mostly from the general population \cite{Michailidou2017}.
-Through imputation, BCAC includes more SNPs than the iCOGS array used for GENESIS (Section~\ref{methods:data}).
+Through imputations, BCAC includes more SNPs than the iCOGS array used for GENESIS (Section~\ref{methods:data}).
 However, in all the comparisons in this paper we focused on the SNPs that passed quality control in GENESIS.
@@
 We excluded HotNet2 and SConES since they took notably longer to run.
-As on the real PPIN, LEAN did not produce significant results on the rewirings either.
+As on the real PPIN, LEAN did not produce significant results on any of the rewirings either.
 
@@
 Summary statistics for SNPs and genes are available at \url{https://github.com/hclimente/genewa}.
-We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact \href{mailto:nadine.andrieu@curie.fr}{nadine.andrieu(at)curie.fr}.
+We cannot share genotype data publicly for confidentiality reasons, but are available from GENESIS. Interested researchers can contact Séverine Eon-Marchais (severine.eon-marchais(at)curie.fr).
 
@@
 The former overlaps with the gene \emph{FGFR2}, the latter with \emph{CASC16} and the protein-coding gene \emph{TOX3}.
-Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, \emph{BRCA1} and \emph{BRCA2} carrier studies \cite{Mulligan2011}, and hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
+Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, in studies on \emph{BRCA1} and \emph{BRCA2} carriers \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
 At the gene level, only \emph{FGFR2} was significantly associated with breast cancer (\nameref{sfig:snp_gene_manhattan}B).
@@
 In the stable consensus network, the most central gene was \emph{CUL3}, which was absent from the previous consensus solution and had a low association score in both GENESIS and BCAC (P\=/values of 0.04 and 0.26, respectively).
-This gene was a component of Cullin-RING ubiquitin ligases.
-Encouragingly, it impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression was also linked to increased sensitivity to carcinogens \cite{Loignon2009}.
+This gene is a component of Cullin-RING ubiquitin ligases.
+It impacts the protein levels of multiple genes relevant for cancer progression \cite{Chen2016}, and its overexpression has also been linked to increased sensitivity to carcinogens \cite{Loignon2009}.
 
@@
 Based on what could be considered diverse interpretations of the omnigenic model, several methods for network-guided discovery have been proposed in recent years.
-In this article we evaluated six of these methods (Section~\ref{methods:methods}) by applying them to the GENESIS study, a GWAS dataset on familial breast cancer (Section~\ref{methods:data}).
+In this article we evaluated six of these methods (Section~\ref{methods:methods}) by applying them to the GENESIS GWAS dataset on familial breast cancer (Section~\ref{methods:data}).
 
@@
 On the other end, dmGWAS, HotNet2 and SigMod gravitated towards lower scoring but larger solutions, which provided a wide overview of the biological context.
-While this deepened our understanding of breast cancer susceptibility and provided biological hypotheses, they required further analyses.
+While this deepens our understanding of breast cancer susceptibility and provides biological hypotheses, interpreting their solutions required further analyses.
 For instance, we examined the centrality of the selected genes to understand how much that property was driving their selection (Section~\ref{results:drawbacks}).
@@
 Hence, network methods were preferable to the logistic regression classifier since they did ``no worse'' at classification while providing an interpretable solution.
+Using an algorithm other than L1-penalized logistic regression might improve classification accuracy.
+Specifically, PRS-based strategies seem to perform slightly better on GENESIS \cite{lonjouGenePathwayLevel2021}.
 Second, all methods were remarkably unstable, yielding different solutions for slightly different inputs.
@@
 How to handle linkage disequilibrium (LD) is often a concern among GWAS practitioners.
-Often, the question is whether an LD-based pruning of the genotypes will improve the results.
+Often, the question is whether an LD-based pruning of the analyzed SNPs will improve the results.
 VEGAS2 accounts for LD patterns, and hence an LD pruning step would not impact gene-based network methods, although it would speed up VEGAS2's computation time.
-In Section~\ref{results:consensus} we highlighted ambiguities that appear when genes overlap or are in LD.
+In Section~\ref{results:consensus} we highlighted ambiguities that appear when genes overlap or contain SNPs that are in LD.
 The presented case is paradigmatic since all three genes are in the HLA region, the most gene-dense region of the genome \cite{Xie2003}.
@@
 % Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
-\paragraph*{S1 Table.}
+\paragraph*{S1 Table}
 \label{stab:snp_summary}
@@
 
-\paragraph*{S2 Table.}
+\paragraph*{S2 Table}
 \label{stab:gene_summary}
@@
 
-\paragraph*{S3 Table.}
+\paragraph*{S3 Table}
 \label{stab:snp_solutions}
@@
 
-\paragraph*{S4 Table.}
+\paragraph*{S4 Table}
 \label{stab:sigmod_pwy}
@@
 
-\paragraph*{S5 Table.}
+\paragraph*{S5 Table}
 \label{stab:dmgwas_pwy}
@@
 
-\paragraph*{S6 Table.}
+\paragraph*{S6 Table}
 \label{stab:hotnet2_pwy}
@@
 
-\paragraph*{S7 Table.}
+\paragraph*{S7 Table}
 \label{stab:consensus_pwy}
@@
 
-\paragraph*{S1 Fig.}
+\paragraph*{S1 Fig}
 \label{sfig:pcs}
@@
 
-\paragraph*{S2 Fig.}
+\paragraph*{S2 Fig}
 \label{sfig:snp_gene_manhattan}
@@
 
-\paragraph*{S3 Fig.}
+\paragraph*{S3 Fig}
 \label{sfig:consensus_stats}
@@
 
-\paragraph*{S4 Fig.}
+\paragraph*{S4 Fig}
 \label{sfig:additional-benchmarks}
@@
 For reference, we added a gray line with a slope of 1.
-This panel is identical to Fig~\ref{fig:consensus}.
 \textbf{(B)}~Sensitivity and specificity on the test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
@@
 
-\paragraph*{S5 Fig.}
+\paragraph*{S5 Fig}
 \label{sfig:rewirings}
@@
 
-\paragraph*{S6 Fig.}
+\paragraph*{S6 Fig}
 \label{sfig:pearson_methods}
@@
 
-\paragraph*{S7 Fig.}
+\paragraph*{S7 Fig}
 \label{sfig:params_size}
@@
 
-\paragraph*{S8 Fig.}
+\paragraph*{S8 Fig}
 \label{sfig:params_pearson}
@@
 
-\paragraph*{S9 Fig.}
+\paragraph*{S9 Fig}
 \label{sfig:lc_ht_comparison}
@@
 
-\paragraph*{S10 Fig.}
+\paragraph*{S10 Fig}
 \label{sfig:biotypes_excluded}
@@
 
-\paragraph*{S11 Fig.}
+\paragraph*{S11 Fig}
 \label{sfig:scones_gsm}

### 2021-02-25 manuscript/plos/review_3.tex
added final (?) review

--- 2021-01-28 manuscript/plos/review_2.tex
@@
 \bigskip
-\textbf{1} Institut Curie, PSL Research University, F-75005 Paris, France;\\
-\textbf{2} INSERM, U900, F-75005 Paris, France;\\
-\textbf{3} MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, F-75006 Paris, France;\\
+\textbf{1} Institut Curie, PSL Research University, Paris, France;\\
+\textbf{2} INSERM, U900, Paris, France;\\
+\textbf{3} MINES ParisTech, PSL Research University, CBIO-Centre for Computational Biology, Paris, France;\\
 \textbf{4} RIKEN Center for Advanced Intelligence Project (AIP), Tokyo, Japan; \\
-\textbf{5} Service de Génétique, Institut Curie, F-75005 Paris, France;\\
-\textbf{6} INSERM, U830, F-75005 Paris, France;\\
-\textbf{7} Université Paris Descartes.\\
+\textbf{5} Service de Génétique, Institut Curie, Paris, France;\\
+\textbf{6} INSERM, U830, Paris, France;\\
+\textbf{7} Université Paris Descartes, Paris, France.\\
 \bigskip
@@
   \centering
-  \includegraphics[width=\textwidth]{./figures/figure_1.png}
+  %\includegraphics[width=\textwidth]{./figures/figure_1.png}
   \caption{\textbf{Overview of the solutions produced by the different network methods (Section~\ref{methods:methods}) on the GENESIS dataset.} As LEAN did not produce any significant solution (BH adjusted P\=/value~$<$~0.05), it is not shown.
@@
   \centering
-  \includegraphics[width=.9\linewidth]{./figures/figure_2.pdf}
+  %\includegraphics[width=.9\linewidth]{./figures/figure_2.pdf}
   \caption{ \textbf{Consensus solution on GENESIS (Section~\ref{methods:methods}).} \textbf{(A)}~Manhattan plot of genes; in black, the ones in the consensus solution.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{./figures/figure_3.pdf}
+  %\includegraphics[width=.8\linewidth]{./figures/figure_3.pdf}
   \caption{\textbf{Comparison of network methods on GENESIS.} Each method was run 5 times on a random subset containing 80\% of the samples and tested on the remaining samples (Section~\ref{methods:benchmark}).
@@
   \centering
-  \includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
+  %\includegraphics[width=.8\linewidth]{./figures/figure_4.pdf}
   \caption{\textbf{Drawbacks encountered when using network methods.} \textbf{(A)}~DmGWAS solution, with the genes colored according to the -log$_{10}$ of their P\=/value.
@@
   \centering
-  \includegraphics[width=.8\linewidth]{./figures/figure_5.pdf}
+  %\includegraphics[width=.8\linewidth]{./figures/figure_5.pdf}
   \caption{ \textbf{Stable consensus solution on GENESIS (Section~\ref{results:stable-consensus})}.
@@
 
-\section{Supporting information}
-
-% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
-\paragraph*{S1 Table}
-\label{stab:snp_summary}
-\textbf{SNP summary statistics on GENESIS.}
-
-\paragraph*{S2 Table}
-\label{stab:gene_summary}
-\textbf{Gene summary statistics on GENESIS.}
-
-\paragraph*{S3 Table}
-\label{stab:snp_solutions}
-\textbf{Summary statistics on the results of SConES on the three SNP networks (Section~\ref{methods:networks}).} The first row within each block contains the summary statistics on the whole network.
-
-\paragraph*{S4 Table}
-\label{stab:sigmod_pwy}
-\textbf{Pathway enrichment analyses of the genes in SigMod's solution.}
-
-\paragraph*{S5 Table}
-\label{stab:dmgwas_pwy}
-\textbf{Pathway enrichment analyses of the genes in dmGWAS' solution.}
-
-\paragraph*{S6 Table}
-\label{stab:hotnet2_pwy}
-\textbf{Pathway enrichment analyses of the genes in HotNet2's solution.}
-
-\paragraph*{S7 Table}
-\label{stab:consensus_pwy}
-\textbf{Pathway enrichment analyses of the genes in the consensus' solution.}
-
-\paragraph*{S1 Fig}
-\label{sfig:pcs}
-\textbf{GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)}~Eight main principal components, computed on the genotypes of GENESIS.
-Cases are colored in green, controls in orange.
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
-
-\paragraph*{S2 Fig}
-\label{sfig:snp_gene_manhattan}
-\textbf{Association in GENESIS.
-The red lines represent the Bonferroni thresholds.} \textbf{(A)}~SNP association, measured from the outcome of a 1 d.f.
-$\chi^2$ allelic test (Section~\ref{methods:conventional}).
-Significant SNPs within a coding gene, or within 50 kilobases of its boundaries, are annotated.
-The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
-\textbf{(B)}~Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values (Section~\ref{methods:conventional}).
-The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
-\textbf{(C)}~SNP association as in panel (A).
-The SNPs in black were selected by an L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).
-%\includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
-
-\paragraph*{S3 Fig}
-\label{sfig:consensus_stats}
-\textbf{Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus solution.} The blue line represents a fitted generalized linear model.
-%\includegraphics[width=.65\textwidth]{./figures/sfigure_3.pdf}
-
-\paragraph*{S4 Fig}
-\label{sfig:additional-benchmarks}
-\textbf{Additional benchmarks of the network methods.} \textbf{(A)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(B)}~Sensitivity and specificity on the test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
-The performance of the classifier trained on all SNPs is also displayed.
-Points are the average over the 5 runs; the error bars represent the standard error of the mean.
-%\includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf
-
-\paragraph*{S5 Fig}
-\label{sfig:rewirings}
-\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its betweenness.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf
-
-\paragraph*{S6 Fig}
-\label{sfig:pearson_methods}
-\textbf{Pearson correlation between the different solutions.} \textbf{(A)}~Correlation between selected SNPs.
-\textbf{(B)}~Correlation between selected genes.
-In general, the solutions display a very low overlap.
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
-
-\paragraph*{S7 Fig}
-\label{sfig:params_size}
-\textbf{Size of the solutions obtained under different parameters.} \textbf{(A)}~Size of the solution produced by different parameter values, expressed as a percentage of the maximum solution size for the method, or the highest tested value for the parameter, respectively.
-The size of the solution is the median among all the solution sizes for the same parameter.
-\textbf{(B)}~Boxplot of the solution sizes of the methods under the explored parameters (Section~\ref{methods:parameters}).
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf
-
-\paragraph*{S8 Fig}
-\label{sfig:params_pearson}
-\textbf{Pearson correlation between the solutions obtained under different parameters, computed as in Section~\ref{methods:algorithm_comparison}.} Grey tiles represent the cases where we could not compute the Pearson correlation because the two vectors were either all ones (all genes/SNPs were selected) or zeros (no genes/SNPs were selected).
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
-
-\paragraph*{S9 Fig}
-\label{sfig:lc_ht_comparison}
-\textbf{Comparison of the benchmark on high-throughput (HT) interactions to the benchmark on both high-throughput and literature curated interactions (HT+LC).} Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
-\textbf{(A)}~Ratios of the selected features between both benchmarks and of the active set (Section~\ref{methods:classifier}).
-\textbf{(B)}~Shifts in sensitivity and specificity.
-\textbf{(C)}~Shift in Pearson correlation between benchmarks.
-\textbf{(D)}~Ratio between the runtimes of the benchmarks.
-For gene-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
-In general, adding additional interactions slightly improved the stability of the solution.
-However, it increased the solution size and the required runtime, and had mixed effects on the sensitivity and specificity.
-%\includegraphics[width=.85\linewidth]{./figures/sfigure_10.pdf}
-
-\paragraph*{S10 Fig}
-\label{sfig:biotypes_excluded}
-\textbf{Biotypes of genes from the annotation that are not present in the HINT PPIN.}
-%\includegraphics[width=.9\linewidth]{./figures/sfigure_9.pdf}
-
-\paragraph*{S11 Fig}
-\label{sfig:scones_gsm}
-\textbf{Overview of the solutions produced by the SConES on the GS and GM networks (Section~\ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)}~Manhattan plots of SNPs (Section~\ref{methods:conventional}); in black, the method’s solution.
-The red line indicates the Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}).
-\textbf{(B)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(C)}~Solution networks.
-%\includegraphics[width=\linewidth]{./figures/sfigure_11.png}
-
 \section*{Acknowledgments}
@@
 Sornin).
-We wish to pay tribute to Olga M.
-Sinilnikova, one of the initiators and principal investigators of GENESIS, and who died prematurely on June 30, 2014.
 
@@
 
-\section*{Author contributions}
-
-\begin{description}
-  \item[Conceptualization] Héctor Climente-González, Christine Lonjou, Chloé-Agathe Azencott.
-  \item[Data curation] Christine Lonjou, GENESIS Study collaborators.
-  \item[Formal Analysis] Héctor Climente-González, Christine Lonjou.
-  \item[Funding acquisition] Dominique Stoppa-Lyonnet, Nadine Andrieu, Chloé-Agathe Azencott.
-  \item[Investigation] Héctor Climente-González, Christine Lonjou.
-  \item[Methodology] Héctor Climente-González, Christine Lonjou, Chloé-Agathe Azencott.
-  \item[Project administration] Chloé-Agathe Azencott.
-  \item[Resources] GENESIS Study collaborators, Dominique Stoppa-Lyonnet, Nadine Andrieu.
-  \item[Software] Héctor Climente-González, Christine Lonjou.
-  \item[Supervision] Christine Lonjou, Fabienne Lesueur, Nadine Andrieu, Chloé-Agathe Azencott.
-  \item[Validation] Christine Lonjou, Fabienne Lesueur.
-  \item[Visualization] Héctor Climente-González.
-  \item[Writing – original draft] Héctor Climente-González.
-  \item[Writing – review \& editing] Héctor Climente-González, Christine Lonjou, Fabienne Lesueur, Nadine Andrieu, Chloé-Agathe Azencott.
-\end{description}
-
 \nolinenumbers
 
-\bibliography{bibliography.bib}
+% \bibliography{bibliography.bib} 
+\begin{thebibliography}{10}
+
+  \bibitem{bush_chapter_2012}
+  Bush WS, Moore JH.
+  \newblock Chapter 11: {Genome}-{Wide} {Association} {Studies}.
+  \newblock PLoS Computational Biology. 2012;8(12):e1002822.
+  \newblock doi:{10.1371/journal.pcbi.1002822}.
+
+  \bibitem{buniello_nhgri-ebi_2019}
+  Buniello A, MacArthur JA, Cerezo M, Harris LW, Hayhurst J, Malangone C, et~al.
+  \newblock The {NHGRI}-{EBI} {GWAS} {Catalog} of published genome-wide
+    association studies, targeted arrays and summary statistics 2019.
+  \newblock Nucleic Acids Research. 2019;47(D1):D1005--D1012.
+  \newblock doi:{10.1093/nar/gky1120}.
+
+  \bibitem{visscher_10_2017}
+  Visscher PM, Wray NR, Zhang Q, Sklar P, McCarthy MI, Brown MA, et~al.
+  \newblock 10 {Years} of {GWAS} {Discovery}: {Biology}, {Function}, and
+    {Translation}.
+  \newblock The American Journal of Human Genetics. 2017;101(1):5--22.
+  \newblock doi:{10.1016/j.ajhg.2017.06.005}.
+
+  \bibitem{wang_statistical_2018}
+  Wang MH, Cordell HJ, Van~Steen K.
+  \newblock Statistical methods for genome-wide association studies.
+  \newblock Seminars in Cancer Biology.
+    2018;doi:{10.1016/j.semcancer.2018.04.008}.
+
+  \bibitem{barton_infinitesimal_2017}
+  Barton NH, Etheridge AM, Véber A.
+  \newblock The infinitesimal model: {Definition}, derivation, and implications.
+  \newblock Theoretical Population Biology. 2017;118:50--73.
+  \newblock doi:{10.1016/j.tpb.2017.06.001}.
+
+  \bibitem{boyle_expanded_2017}
+  Boyle EA, Li YI, Pritchard JK.
+  \newblock An {Expanded} {View} of {Complex} {Traits}: {From} {Polygenic} to
+    {Omnigenic}.
+  \newblock Cell. 2017;169(7):1177--1186.
+  \newblock doi:{10.1016/j.cell.2017.05.038}.
+
+  \bibitem{furlong_human_2013}
+  Furlong LI.
+  \newblock Human diseases through the lens of network biology.
+  \newblock Trends in Genetics. 2013;29(3):150--159.
+  \newblock doi:{10.1016/j.tig.2012.11.004}.
+
+  \bibitem{leisersonNetworkAnalysisGWAS2013}
+  Leiserson MD, {Eldridge, Jonathan V}, {Ramachandran, Sohini}, {Raphael,
+    Benjamin J}.
+  \newblock Network analysis of {GWAS} data.
+  \newblock Current Opinion in Genetics \& Development. 2013;23(6):602--610.
+
+  \bibitem{barabasi_network_2011}
+  Barabási AL, Gulbahce N, Loscalzo J.
+  \newblock Network medicine: a network-based approach to human disease.
+  \newblock Nature Reviews Genetics. 2011;12(1):56--68.
+  \newblock doi:{10.1038/nrg2918}.
+
+  \bibitem{cai_broker_2010}
+  Cai JJ, Borenstein E, Petrov DA.
+  \newblock Broker {Genes} in {Human} {Disease}.
+  \newblock Genome Biology and Evolution. 2010;2:815--825.
+  \newblock doi:{10.1093/gbe/evq064}.
+
+  \bibitem{azencott_network-guided_2016}
+  Azencott CA.
+  \newblock Network-{Guided} {Biomarker} {Discovery}.
+  \newblock In: Machine {Learning} for {Health} {Informatics}. vol. 9605. Cham:
+    Springer International Publishing; 2016. p. 319--336.
+  \newblock Available from:
+    \url{http://link.springer.com/10.1007/978-3-319-50478-0_16}.
+
+  \bibitem{huang_systematic_2018}
+  Huang JK, Carlin DE, Yu MK, Zhang W, Kreisberg JF, Tamayo P, et~al.
+  \newblock Systematic {Evaluation} of {Molecular} {Networks} for {Discovery} of
+    {Disease} {Genes}.
+  \newblock Cell Systems. 2018;6(4):484--495.e5.
+  \newblock doi:{10.1016/j.cels.2018.03.001}.
+
+  \bibitem{gwinner_network-based_2016}
+  Gwinner F, Boulday G, Vandiedonck C, Arnould M, Cardoso C, Nikolayeva I, et~al.
+  \newblock Network-based analysis of omics data: {The} {LEAN} method.
+  \newblock Bioinformatics. 2016; p. btw676.
+  \newblock doi:{10.1093/bioinformatics/btw676}.
+
+  \bibitem{jia_dmgwas:_2011}
+  Jia P, Zheng S, Long J, Zheng W, Zhao Z.
+  \newblock {dmGWAS}: dense module searching for genome-wide association studies
+    in protein–protein interaction networks.
+  \newblock Bioinformatics. 2011;27(1):95--102.
+  \newblock doi:{10.1093/bioinformatics/btq615}.
+
+  \bibitem{dittrich_identifying_2008}
+  Dittrich MT, Klau GW, Rosenwald A, Dandekar T, Muller T.
+  \newblock Identifying functional modules in protein-protein interaction
+    networks: an integrated exact approach.
+  \newblock Bioinformatics. 2008;24(13):i223--i231.
+  \newblock doi:{10.1093/bioinformatics/btn161}.
+
+  \bibitem{leiserson_pan-cancer_2015}
+  Leiserson MDM, Vandin F, Wu HT, Dobson JR, Eldridge JV, Thomas JL, et~al.
+  \newblock Pan-cancer network analysis identifies combinations of rare somatic
+    mutations across pathways and protein complexes.
+  \newblock Nature Genetics. 2015;47(2):106--114.
+  \newblock doi:{10.1038/ng.3168}.
+
+  \bibitem{azencott_efficient_2013}
+  Azencott CA, Grimm D, Sugiyama M, Kawahara Y, Borgwardt KM.
+  \newblock Efficient network-guided multi-locus association mapping with graph
+    cuts.
+  \newblock Bioinformatics. 2013;29(13):i171--i179.
+  \newblock doi:{10.1093/bioinformatics/btt238}.
+
+  \bibitem{liu_sigmod:_2017}
+  Liu Y, Brossard M, Roqueiro D, Margaritte-Jeannin P, Sarnowski C, Bouzigon E,
+    et~al.
+  \newblock {SigMod}: an exact and efficient method to identify a strongly
+    interconnected disease-associated module in a gene network.
+  \newblock Bioinformatics. 2017; p. btx004.
+  \newblock doi:{10.1093/bioinformatics/btx004}.
+
+  \bibitem{sinilnikova_genesis:_2016}
+  Sinilnikova OM, Dondon MG, Eon-Marchais S, Damiola F, Barjhoux L, Marcou M,
+    et~al.
+  \newblock {GENESIS}: a {French} national resource to study the missing
+    heritability of breast cancer.
+  \newblock BMC Cancer. 2016;16(1):13.
+  \newblock doi:{10.1186/s12885-015-2028-9}.
+
+  \bibitem{sakoda_turning_2013}
+  Sakoda LC, Jorgenson E, Witte JS.
+  \newblock Turning of {COGS} moves forward findings for hormonally mediated
+    cancers.
+  \newblock Nature Genetics. 2013;45(4):345--348.
+  \newblock doi:{10.1038/ng.2587}.
+
+  \bibitem{chang_second-generation_2015}
+  Chang CC, Chow CC, Tellier LC, Vattikuti S, Purcell SM, Lee JJ.
+  \newblock Second-generation {PLINK}: rising to the challenge of larger and
+    richer datasets.
+  \newblock GigaScience. 2015;4(1):7.
+  \newblock doi:{10.1186/s13742-015-0047-8}.
+
+  \bibitem{mishra_vegas2:_2015}
+  Mishra A, Macgregor S.
+  \newblock {VEGAS}2: {Software} for {More} {Flexible} {Gene}-{Based} {Testing}.
+  \newblock Twin Research and Human Genetics. 2015;18(1):86--91.
+  \newblock doi:{10.1017/thg.2014.79}.
+
+  \bibitem{frankish_gencode_2019}
+  Frankish A, Diekhans M, Ferreira AM, Johnson R, Jungreis I, Loveland J, et~al.
+  \newblock {GENCODE} reference annotation for the human and mouse genomes.
+  \newblock Nucleic Acids Research. 2019;47(D1):D766--D773.
+  \newblock doi:{10.1093/nar/gky955}.
+
+  \bibitem{das_hint:_2012}
+  Das J, Yu H.
+  \newblock {HINT}: {High}-quality protein interactomes and their applications in
+    understanding human disease.
+  \newblock BMC Systems Biology. 2012;6(1):92.
+  \newblock doi:{10.1186/1752-0509-6-92}.
+
+  \bibitem{dmgwas}
+  Wang Q, Jia P. dmGWAS 3.0; 2014.
+  \newblock \url{https://bioinfo.uth.edu/dmGWAS/}.
+
+  \bibitem{ljubic_algorithmic_2006}
+  Ljubić I, Weiskircher R, Pferschy U, Klau GW, Mutzel P, Fischetti M.
+  \newblock An {Algorithmic} {Framework} for the {Exact} {Solution} of the
+    {Prize}-{Collecting} {Steiner} {Tree} {Problem}.
+  \newblock Mathematical Programming. 2006;105(2-3):427--449.
+  \newblock doi:{10.1007/s10107-005-0660-x}.
+
+  \bibitem{beisser_bionet:_2010}
+  Beisser D, Klau GW, Dandekar T, Muller T, Dittrich MT.
+  \newblock {BioNet}: an {R}-{Package} for the functional analysis of biological
+    networks.
+  \newblock Bioinformatics. 2010;26(8):1129--1130.
+  \newblock doi:{10.1093/bioinformatics/btq089}.
+
+  \bibitem{heinz}
+  Dittrich M, Beisser D. BioNet; 2008.
+  \newblock \url{https://bioconductor.org/packages/BioNet/}.
+
+  \bibitem{nakka_gene_2016}
+  Nakka P, Raphael BJ, Ramachandran S.
+  \newblock Gene and {Network} {Analysis} of {Common} {Variants} {Reveals}
+    {Novel} {Associations} in {Multiple} {Complex} {Diseases}.
+  \newblock Genetics. 2016;204(2):783--798.
+  \newblock doi:{10.1534/genetics.116.188391}.
+
+  \bibitem{scheid_twilight;_2005}
+  Scheid S, Spang R.
+  \newblock twilight; a {Bioconductor} package for estimating the local false
+    discovery rate.
+  \newblock Bioinformatics. 2005;21(12):2921--2922.
+  \newblock doi:{10.1093/bioinformatics/bti436}.
+
+  \bibitem{hotnet2}
+  Leiserson MDM, Vandin F, Wu HT, Dobson JR, Eldridge JV, Thomas JL, et~al..
+    HotNet2; 2018.
+  \newblock \url{https://github.com/raphael-group/hotnet2}.
+
+  \bibitem{leanr}
+  Gwinner F. LEANR; 2016.
+  \newblock \url{https://cran.r-project.org/web/packages/LEANR/}.
+
+  \bibitem{martini}
+  Climente-González H, Azencott CA. martini; 2019.
+  \newblock \url{https://www.bioconductor.org/packages/martini/}.
+
+  \bibitem{climente-gonzalezMartiniPackageGenomewide2021}
+  Climente-González H, Azencott CA.
+  \newblock martini: an {R} package for genome-wide association studies using
+    {SNP} networks.
+  \newblock Bioinformatics; 2021.
+  \newblock Available from:
+    \url{http://biorxiv.org/lookup/doi/10.1101/2021.01.25.428047}.
+
+  \bibitem{sigmod}
+  Liu Y. SigMod v2; 2018.
+  \newblock \url{https://github.com/YuanlongLiu/SigMod}.
+
+  \bibitem{Jassal2019}
+  Jassal B, Matthews L, Viteri G, Gong C, Lorente P, Fabregat A, et~al.
+  \newblock The reactome pathway knowledgebase.
+  \newblock Nucleic Acids Research. 2019;doi:{10.1093/nar/gkz1031}.
+
+  \bibitem{Yu2016}
+  Yu G, He QY.
+  \newblock {ReactomePA}: an R/Bioconductor package for reactome pathway analysis
+    and visualization.
+  \newblock Molecular {BioSystems}. 2016;12(2):477--479.
+  \newblock doi:{10.1039/c5mb00663e}.
+
+  \bibitem{Benjamini1995}
+  Benjamini Y, Hochberg Y.
+  \newblock Controlling the False Discovery Rate: A Practical and Powerful
+    Approach to Multiple Testing.
+  \newblock Journal of the Royal Statistical Society: Series B (Methodological).
+    1995;57(1):289--300.
+  \newblock doi:{10.1111/j.2517-6161.1995.tb02031.x}.
+
+  \bibitem{nogueira_measuring_2016}
+  Nogueira S, Brown G.
+  \newblock Measuring the {Stability} of {Feature} {Selection}.
+  \newblock In: Machine {Learning} and {Knowledge} {Discovery} in {Databases}.
+    vol. 9852. Cham: Springer International Publishing; 2016. p. 442--457.
+  \newblock Available from:
+    \url{http://link.springer.com/10.1007/978-3-319-46227-1_28}.
+
+  \bibitem{Metcalfe2008}
+  Metcalfe KA, Finch A, Poll A, Horsman D, Kim-Sing C, Scott J, et~al.
+  \newblock Breast cancer risks in women with a family history of breast or
+    ovarian cancer who have tested negative for a {BRCA}1 or {BRCA}2 mutation.
+  \newblock British Journal of Cancer. 2008;100(2):421--425.
+  \newblock doi:{10.1038/sj.bjc.6604830}.
+
+  \bibitem{pinero_disgenet:_2017}
+  Piñero J, Bravo A, Queralt-Rosinach N, Gutiérrez-Sacristán A, Deu-Pons J,
+    Centeno E, et~al.
+  \newblock {DisGeNET}: a comprehensive platform integrating information on human
+    disease-associated genes and variants.
+  \newblock Nucleic Acids Research. 2017;45(D1):D833--D839.
+  \newblock doi:{10.1093/nar/gkw943}.
+
+  \bibitem{Michailidou2017}
+  Michailidou K, Lindstr\"{o}m S, Dennis J, Beesley J, Hui S, Kar S, et~al.
+  \newblock Association analysis identifies 65 new breast cancer risk loci.
+  \newblock Nature. 2017;551(7678):92--94.
+  \newblock doi:{10.1038/nature24284}.
+
+  \bibitem{the_1000_genomes_project_consortium_global_2015}
+  {The 1000 Genomes Project Consortium}, Gibbs RA, Boerwinkle E, Doddapaneni H,
+    Han Y, Korchina V, et~al.
+  \newblock A global reference for human genetic variation.
+  \newblock Nature. 2015;526(7571):68--74.
+  \newblock doi:{10.1038/nature15393}.
+
+  \bibitem{Mulligan2011}
+  Mulligan AM, , Couch FJ, Barrowdale D, Domchek SM, Eccles D, et~al.
+  \newblock Common breast cancer susceptibility alleles are associated with
+    tumour subtypes in {BRCA}1 and {BRCA}2 mutation carriers: results from the
+    Consortium of Investigators of Modifiers of {BRCA}1/2.
+  \newblock Breast Cancer Research. 2011;13(6).
+  \newblock doi:{10.1186/bcr3052}.
+
+  \bibitem{rinella_genetic_2013}
+  Rinella ES, Shao Y, Yackowski L, Pramanik S, Oratz R, Schnabel F, et~al.
+  \newblock Genetic variants associated with breast cancer risk for {Ashkenazi}
+    {Jewish} women with strong family histories but no identifiable {BRCA}1/2
+    mutation.
+  \newblock Human Genetics. 2013;132(5):523--536.
+  \newblock doi:{10.1007/s00439-013-1269-4}.
+
+  \bibitem{brisbin_meta-analysis_2011}
+  Brisbin AG, Asmann YW, Song H, Tsai YY, Aakre JA, Yang P, et~al.
+  \newblock Meta-analysis of 8q24 for seven cancers reveals a locus between {NOV}
+    and {ENPP}2 associated with cancer development.
+  \newblock BMC Medical Genetics. 2011;12(1):156.
+  \newblock doi:{10.1186/1471-2350-12-156}.
+
+  \bibitem{search_newly_2009}
+  {SEARCH}, {The GENICA Consortium}, {kConFab}, {Australian Ovarian Cancer Study
+    Group}, Ahmed S, Thomas G, et~al.
+  \newblock Newly discovered breast cancer susceptibility loci on 3p24 and
+    17q23.2.
+  \newblock Nature Genetics. 2009;41(5):585--590.
+  \newblock doi:{10.1038/ng.354}.
+
+  \bibitem{nielsen_hereditary_2016}
+  Nielsen FC, van Overeem~Hansen T, Sørensen CS.
+  \newblock Hereditary breast and ovarian cancer: new genes in confined pathways.
+  \newblock Nature Reviews Cancer. 2016;16(9):599--612.
+  \newblock doi:{10.1038/nrc.2016.72}.
+
+  \bibitem{pinero_uncovering_2016}
+  Piñero J, Berenstein A, Gonzalez-Perez A, Chernomoretz A, Furlong LI.
+  \newblock Uncovering disease mechanisms through network biology in the era of
+    {Next} {Generation} {Sequencing}.
+  \newblock Scientific Reports. 2016;6(1):24570.
+  \newblock doi:{10.1038/srep24570}.
+
+  \bibitem{quigley_5p12_2014}
+  Quigley DA, Fiorito E, Nord S, Van~Loo P, Alnaes GG, Fleischer T, et~al.
+  \newblock The 5p12 breast cancer susceptibility locus affects {MRPS}30
+    expression in estrogen-receptor positive tumors.
+  \newblock Molecular Oncology. 2014;8(2):273--284.
+  \newblock doi:{10.1016/j.molonc.2013.11.008}.
+
+  \bibitem{Yu2016Repositioning}
+  Yu M, Li R, Zhang J.
+  \newblock Repositioning of antibiotic levofloxacin as a mitochondrial
+    biogenesis inhibitor to target breast cancer.
+  \newblock Biochemical and Biophysical Research Communications.
+    2016;471(4):639--645.
+  \newblock doi:{10.1016/j.bbrc.2016.02.072}.
+
+  \bibitem{liu_jab1_cops5_2018}
+  Liu G, Claret FX, Zhou F, Pan Y.
+  \newblock Jab1/{COPS}5 as a {Novel} {Biomarker} for {Diagnosis}, {Prognosis},
+    {Therapy} {Prediction} and {Therapeutic} {Tools} for {Human} {Cancer}.
+  \newblock Frontiers in Pharmacology. 2018;9:135.
+  \newblock doi:{10.3389/fphar.2018.00135}.
+
+  \bibitem{deloscamposComplexTraitPredictionEra2018}
+  de~los Campos G, Vazquez AI, Hsu S, Lello L.
+  \newblock Complex-{Trait} {Prediction} in the {Era} of {Big} {Data}.
+  \newblock Trends in Genetics. 2018;34(10):746--754.
+  \newblock doi:{10.1016/j.tig.2018.07.004}.
+
+  \bibitem{nikolayeva_network_2018}
+  Nikolayeva I, Guitart~Pla O, Schwikowski B.
+  \newblock Network module identification—{A} widespread theoretical bias and
+    best practices.
+  \newblock Methods. 2018;132:19--25.
+  \newblock doi:{10.1016/j.ymeth.2017.08.008}.
+
+  \bibitem{Ioachim2002}
+  Ioachim E, Charchanti A, Briasoulis E, Karavasilis V, Tsanou H, Arvanitis DL,
+    et~al.
+  \newblock Immunohistochemical expression of extracellular matrix components
+    tenascin, fibronectin, collagen type {IV} and laminin in breast cancer: their
+    prognostic value and role in tumour invasion and progression.
+  \newblock European Journal of Cancer. 2002;38(18):2362--2370.
+  \newblock doi:{10.1016/s0959-8049(02)00210-1}.
+
+  \bibitem{Yi2016}
+  Yi W, Xiao E, Ding R, Luo P, Yang Y.
+  \newblock High expression of fibronectin is associated with poor prognosis,
+    cell proliferation and malignancy via the {NF}-$\kappa$B/p53-apoptosis
+    signaling pathway in colorectal cancer.
+  \newblock Oncology Reports. 2016;36(6):3145--3153.
+  \newblock doi:{10.3892/or.2016.5177}.
+
+  \bibitem{Sponziello2016}
+  Sponziello M, Rosignolo F, Celano M, Maggisano V, Pecce V, Rose RFD, et~al.
+  \newblock Fibronectin-1 expression is increased in aggressive thyroid cancer
+    and favors the migration and invasion of cancer cells.
+  \newblock Molecular and Cellular Endocrinology. 2016;431:123--132.
+  \newblock doi:{10.1016/j.mce.2016.05.007}.
+
+  \bibitem{ahmed_newly_2009}
+  Ahmed S, Thomas G, Ghoussaini M, Healey CS, Humphreys MK, Platte R, et~al.
+  \newblock Newly discovered breast cancer susceptibility loci on 3p24 and
+    17q23.2.
+  \newblock Nature Genetics. 2009;41(5):585--590.
+  \newblock doi:{10.1038/ng.354}.
+
+  \bibitem{breyer_expressed_2014}
+  Breyer J, Dorset D, Clark T, Bradley K, Wahlfors T, McReynolds K, et~al.
+  \newblock An {Expressed} {Retrogene} of the {Master} {Embryonic} {Stem} {Cell}
+    {Gene} {POU}5F1 {Is} {Associated} with {Prostate} {Cancer} {Susceptibility}.
+  \newblock The American Journal of Human Genetics. 2014;94(3):395--404.
+  \newblock doi:{10.1016/j.ajhg.2014.01.019}.
+
+  \bibitem{Chen2016}
+  Chen HY, Chen RH.
+  \newblock Cullin 3 Ubiquitin Ligases in Cancer Biology: Functions and
+    Therapeutic Implications.
+  \newblock Frontiers in Oncology. 2016;6.
+  \newblock doi:{10.3389/fonc.2016.00113}.
+
+  \bibitem{Loignon2009}
+  Loignon M, Miao W, Hu L, Bier A, Bismar TA, Scrivens PJ, et~al.
+  \newblock Cul3 overexpression depletes Nrf2 in breast cancer and is associated
+    with sensitivity to carcinogens, to oxidative stress, and to chemotherapy.
+  \newblock Molecular Cancer Therapeutics. 2009;8(8):2432--2440.
+  \newblock doi:{10.1158/1535-7163.mct-08-1186}.
+
+  \bibitem{lonjouGenePathwayLevel2021}
+  Lonjou C, Eon-Marchais S, Truong T, Dondon MG, Karimi M, Jiao Y, et~al.
+  \newblock Gene- and pathway-level analyses of {iCOGS} variants highlight novel
+    signaling pathways underlying familial breast cancer susceptibility.
+  \newblock International Journal of Cancer. 2021; p. ijc.33457.
+  \newblock doi:{10.1002/ijc.33457}.
+
+  \bibitem{halseyFickleValueGenerates2015}
+  Halsey LG, Curran-Everett D, Vowler SL, Drummond GB.
+  \newblock The fickle {P} value generates irreproducible results.
+  \newblock Nature Methods. 2015;12(3):179--185.
+  \newblock doi:{10.1038/nmeth.3288}.
+
+  \bibitem{wagnerAlternativeRoutesMutational2007}
+  Wagner A, Wright J.
+  \newblock Alternative routes and mutational robustness in complex regulatory
+    networks.
+  \newblock Biosystems. 2007;88(1-2):163--172.
+  \newblock doi:{10.1016/j.biosystems.2006.06.002}.
+
+  \bibitem{gallagherPostGWASEraAssociation2018}
+  Gallagher MD, Chen-Plotkin AS.
+  \newblock The {Post}-{GWAS} {Era}: {From} {Association} to {Function}.
+  \newblock The American Journal of Human Genetics. 2018;102(5):717--730.
+  \newblock doi:{10.1016/j.ajhg.2018.04.002}.
+
+  \bibitem{ionita-laza_sequence_2013}
+  Ionita-Laza I, Lee S, Makarov V, Buxbaum J, Lin X.
+  \newblock Sequence {Kernel} {Association} {Tests} for the {Combined} {Effect}
+    of {Rare} and {Common} {Variants}.
+  \newblock The American Journal of Human Genetics. 2013;92(6):841--853.
+  \newblock doi:{10.1016/j.ajhg.2013.04.015}.
+
+  \bibitem{wang_combat:_2017}
+  Wang M, Huang J, Liu Y, Ma L, Potash JB, Han S.
+  \newblock {COMBAT}: {A} {Combined} {Association} {Test} for {Genes} {Using}
+    {Summary} {Statistics}.
+  \newblock Genetics. 2017;207(3):883--891.
+  \newblock doi:{10.1534/genetics.117.300257}.
+
+  \bibitem{Xie2003}
+  Xie T.
+  \newblock Analysis of the Gene-Dense Major Histocompatibility Complex Class
+    {III} Region and Its Comparison to Mouse.
+  \newblock Genome Research. 2003;13(12):2621--2636.
+  \newblock doi:{10.1101/gr.1736803}.
+
+  \bibitem{di_tommaso_nextflow_2017}
+  Di~Tommaso P, Chatzou M, Floden EW, Barja PP, Palumbo E, Notredame C.
+  \newblock Nextflow enables reproducible computational workflows.
+  \newblock Nature Biotechnology. 2017;35(4):316--319.
+  \newblock doi:{10.1038/nbt.3820}.
+
+  \end{thebibliography}
 
@@
 
+\section{Supporting information}
+
+% Include only the SI item label in the paragraph heading. Use the \nameref{label} command to cite SI items in the text.
+\paragraph*{S1 Table}
+\label{stab:snp_summary}
+\textbf{SNP summary statistics on GENESIS.}
+
+\paragraph*{S2 Table}
+\label{stab:gene_summary}
+\textbf{Gene summary statistics on GENESIS.}
+
+\paragraph*{S3 Table}
+\label{stab:snp_solutions}
+\textbf{Summary statistics on the results of SConES on the three SNP networks (Section~\ref{methods:networks}).} The first row within each block contains the summary statistics on the whole network.
+
+\paragraph*{S4 Table}
+\label{stab:sigmod_pwy}
+\textbf{Pathway enrichment analyses of the genes in SigMod's solution.}
+
+\paragraph*{S5 Table}
+\label{stab:dmgwas_pwy}
+\textbf{Pathway enrichment analyses of the genes in dmGWAS' solution.}
+
+\paragraph*{S6 Table}
+\label{stab:hotnet2_pwy}
+\textbf{Pathway enrichment analyses of the genes in HotNet2's solution.}
+
+\paragraph*{S7 Table}
+\label{stab:consensus_pwy}
+\textbf{Pathway enrichment analyses of the genes in the consensus' solution.}
+
+\paragraph*{S1 Fig}
+\label{sfig:pcs}
+\textbf{GENESIS shows no differential population structure between cases and controls.} \textbf{(A,B,C,D)}~Eight main principal components, computed on the genotypes of GENESIS.
+Cases are colored in green, controls in orange.
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_1.pdf}
+
+\paragraph*{S2 Fig}
+\label{sfig:snp_gene_manhattan}
+\textbf{Association in GENESIS.
+The red lines represent the Bonferroni thresholds.} \textbf{(A)}~SNP association, measured from the outcome of a 1 d.f.
+$\chi^2$ allelic test (Section~\ref{methods:conventional}).
+Significant SNPs within a coding gene, or within 50 kilobases of its boundaries, are annotated.
+The Bonferroni threshold is 2.54 \texttimes{} 10\textsuperscript{-7}.
+\textbf{(B)}~Gene association, measured by P\=/value of VEGAS2 \cite{mishra_vegas2:_2015} using the 10\% of SNPs with the lowest P\=/values (Section~\ref{methods:conventional}).
+The Bonferroni threshold is 1.53 \texttimes{} 10\textsuperscript{-6}.
+\textbf{(C)}~SNP association as in panel (A).
+The SNPs in black were selected by an L1-penalized logistic regression (Section~\ref{methods:classifier}, $\lambda = 0.03$).
+%\includegraphics[width=.7\linewidth]{./figures/sfigure_2.png}
+
+\paragraph*{S3 Fig}
+\label{sfig:consensus_stats}
+\textbf{Relationship between the log\textsubscript{10} of the betweenness centrality and the -log\textsubscript{10} of the VEGAS2 P\=/value of the genes in the consensus solution.} The blue line represents a fitted generalized linear model.
+%\includegraphics[width=.65\textwidth]{./figures/sfigure_3.pdf}
+
+\paragraph*{S4 Fig}
+\label{sfig:additional-benchmarks}
+\textbf{Additional benchmarks of the network methods.} \textbf{(A)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs/genes in BCAC.
+For reference, we added a gray line with a slope of 1.
+\textbf{(B)}~Sensitivity and specificity on the test set of the L1-penalized logistic regression trained on the features selected by each of the methods.
+The performance of the classifier trained on all SNPs is also displayed.
+Points are the average over the 5 runs; the error bars represent the standard error of the mean.
+%\includegraphics[width=.675\linewidth]{./figures/sfigure_4.pdf
+
+\paragraph*{S5 Fig}
+\label{sfig:rewirings}
+\textbf{Number of times a gene was selected by either dmGWAS, heinz, or SigMod in 100 rewirings of the PPIN (Section~\ref{methods:rewiring}) and its betweenness.} This figure is equivalent to Fig~\ref{fig:issues}B, split by method.
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_5.pdf
+
+\paragraph*{S6 Fig}
+\label{sfig:pearson_methods}
+\textbf{Pearson correlation between the different solutions.} \textbf{(A)}~Correlation between selected SNPs.
+\textbf{(B)}~Correlation between selected genes.
+In general, the solutions display a very low overlap.
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_6.pdf}
+
+\paragraph*{S7 Fig}
+\label{sfig:params_size}
+\textbf{Size of the solutions obtained under different parameters.} \textbf{(A)}~Size of the solution produced by different parameter values, expressed as a percentage of the maximum solution size for the method, or the highest tested value for the parameter, respectively.
+The size of the solution is the median among all the solution sizes for the same parameter.
+\textbf{(B)}~Boxplot of the solution sizes of the methods under the explored parameters (Section~\ref{methods:parameters}).
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_7.pdf
+
+\paragraph*{S8 Fig}
+\label{sfig:params_pearson}
+\textbf{Pearson correlation between the solutions obtained under different parameters, computed as in Section~\ref{methods:algorithm_comparison}.} Grey tiles represent the cases where we could not compute the Pearson correlation because the two vectors were either all ones (all genes/SNPs were selected) or zeros (no genes/SNPs were selected).
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_8.pdf
+
+\paragraph*{S9 Fig}
+\label{sfig:lc_ht_comparison}
+\textbf{Comparison of the benchmark on high-throughput (HT) interactions to the benchmark on both high-throughput and literature curated interactions (HT+LC).} Grey lines represent no change in the statistic between the benchmarks (1 for ratios mean(HT) / mean(HT + LC), 0 for differences mean(HT) - mean(HT + LC)).
+\textbf{(A)}~Ratios of the selected features between both benchmarks and of the active set (Section~\ref{methods:classifier}).
+\textbf{(B)}~Shifts in sensitivity and specificity.
+\textbf{(C)}~Shift in Pearson correlation between benchmarks.
+\textbf{(D)}~Ratio between the runtimes of the benchmarks.
+For gene-based methods, inverted triangles represent the ratio of runtimes of the algorithms themselves, and circles the total time, which includes the algorithm themselves and the additional 119\,980 seconds (1 day and 9.33 hours) that VEGAS2 took on average to compute the gene scores from SNP summary statistics.
+In general, adding additional interactions slightly improved the stability of the solution.
+However, it increased the solution size and the required runtime, and had mixed effects on the sensitivity and specificity.
+%\includegraphics[width=.85\linewidth]{./figures/sfigure_10.pdf}
+
+\paragraph*{S10 Fig}
+\label{sfig:biotypes_excluded}
+\textbf{Biotypes of genes from the annotation that are not present in the HINT PPIN.}
+%\includegraphics[width=.9\linewidth]{./figures/sfigure_9.pdf}
+
+\paragraph*{S11 Fig}
+\label{sfig:scones_gsm}
+\textbf{Overview of the solutions produced by the SConES on the GS and GM networks (Section~\ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)}~Manhattan plots of SNPs (Section~\ref{methods:conventional}); in black, the method’s solution.
+The red line indicates the Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}).
+\textbf{(B)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
+For reference, we added a gray line with a slope of 1.
+\textbf{(C)}~Solution networks.
+%\includegraphics[width=\linewidth]{./figures/sfigure_11.png}
+
 \end{document}

### 2021-03-03 manuscript/plos/review_4.tex
minor modifications

--- 2021-02-25 manuscript/plos/review_3.tex
@@
 We conducted association analyses in the GENESIS dataset (Section~\ref{methods:data}) at both SNP and gene levels (Section~\ref{methods:conventional}).
-At the SNP level, two genomic regions had a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (\nameref{sfig:snp_gene_manhattan}A).
+At the SNP level, two genomic regions had a P\=/value lower than the Bonferroni threshold on chromosomes 10 and 16 (S2A Fig).
 The former overlaps with the gene \emph{FGFR2}, the latter with \emph{CASC16} and the protein-coding gene \emph{TOX3}.
 Variants in both \emph{FGFR2} and \emph{TOX3} have been repeatedly associated with breast cancer susceptibility in other case-control studies \cite{Michailidou2017}, in studies on \emph{BRCA1} and \emph{BRCA2} carriers \cite{Mulligan2011}, and in hereditary breast and ovarian cancer families negative for mutations in \emph{BRCA1} and \emph{BRCA2} \cite{rinella_genetic_2013}.
-At the gene level, only \emph{FGFR2} was significantly associated with breast cancer (\nameref{sfig:snp_gene_manhattan}B).
+At the gene level, only \emph{FGFR2} was significantly associated with breast cancer (S2B Fig).
 
@@
 We applied an L1-penalized logistic regression using all GENESIS genotypes as input and the phenotype (cancer/healthy) as the outcome (Section~\ref{methods:classifier}).
-The algorithm selected 100 SNPs, both from all regions mentioned above and new ones (\nameref{sfig:snp_gene_manhattan}C).
+The algorithm selected 100 SNPs, both from all regions mentioned above and new ones (S2C Fig).
 However, it was unclear why those SNPs were selected, as emphasized by the high P\=/value of some of them, which further complicates the biological interpretation.
@@
 
-Topologically, the consensus consisted of a connected component composed of 49 genes and multiple smaller subnetworks (Fig~\ref{fig:consensus}B and C).
+Topologically, the consensus consisted of a connected component composed of 49 genes and multiple smaller subnetworks (Fig~\ref{fig:consensus}B and \ref{fig:consensus}C).
 Among the latter, 19 genes were in subnetworks containing a single gene or two connected nodes.
@@
 The solutions provided by the different  network methods overlapped significantly with BCAC hits (Fisher's exact test P\=/value \textless~0.019).
-The gene-based methods achieved comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (\nameref{sfig:additional-benchmarks}A).
+The gene-based methods achieved comparable precision (2\%-25\%) and recall (1.3-12.1\%) at recovering BCAC-significant genes (S4A Fig).
 Interestingly, while SConES GI achieved a similar recall at the SNP-level (8.6\%), it showed a much higher precision (47.3\%).
@@
 We compared the six network methods in a 5-fold subsampling setting (Section~\ref{methods:benchmark}).
-In this comparison we measured properties (Fig~\ref{fig:benchmark} and \nameref{sfig:additional-benchmarks}): the size of the solution; the sensitivity and the specificity of an L1-penalized logistic regression classifier on the selected SNPs; the stability of the methods; and their computational runtime.
+In this comparison we measured properties (Figs~\ref{fig:benchmark} and S4): the size of the solution; the sensitivity and the specificity of an L1-penalized logistic regression classifier on the selected SNPs; the stability of the methods; and their computational runtime.
 The solution size varies greatly between the different methods (Fig~\ref{fig:benchmark}A).
@@
 
-To determine whether the selected SNPs could predict cancer susceptibility, we computed the classifiers' performances on \emph{test sets} (\nameref{sfig:additional-benchmarks}B).
+To determine whether the selected SNPs could predict cancer susceptibility, we computed the classifiers' performances on \emph{test sets} (S4B Fig).
 The different classifiers displayed similarly low sensitivities and specificities, all in the 0.52 -- 0.56 range.
@@
 Its conservativeness stemmed from its preprocessing step, which modeled the gene P\=/values as a mixture model of a beta distribution and a uniform distribution, controlled by an FDR parameter.
-Due to the limited signal at the gene level in this dataset (\nameref{sfig:snp_gene_manhattan}B), only 36 genes retained a positive score after that transformation.
+Due to the limited signal at the gene level in this dataset (S2B Fig), only 36 genes retained a positive score after that transformation.
 However, this small solution did not provide much insight into the susceptibility mechanisms to cancer.
@@
 We explored methods' parameter space by running them under different combinations of parameters (Section~\ref{methods:parameters}).
-In agreement with their formulations (Section~\ref{methods:methods}), larger values of specific parameters produced less stringent solutions (\nameref{sfig:params_size}A): for HotNet2 and heinz, this is the threshold above which genes receive a positive score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
+In agreement with their formulations (Section~\ref{methods:methods}), larger values of specific parameters produced less stringent solutions (S7A Fig): for HotNet2 and heinz, this is the threshold above which genes receive a positive score; for dmGWAS, it is the \texttt{d} parameter, which controls how far neighbors could be added; for SigMod, it is \texttt{nmax}, which specifies the maximum size of the solution; and for LEAN, it is the P-value threshold to consider a solution significant.
 Two parameters had the opposite effect (the larger, the more stringent): SigMod's \texttt{maxjump}, which sets the threshold to consider an increment in $\lambda$ ``large enough''; and SConES' $\eta$, where higher values produce smaller solutions.
@@
 In both cases, the size of the solution was very similar across the different values.
-Despite the differences in size, the solutions' size was relatively robust to the choice of parameters (\nameref{sfig:params_size}B).
+Despite the differences in size, the solutions' size was relatively robust to the choice of parameters (Fig S7B).
 
@@
 The resulting stability-based consensus was composed of 68 genes (Fig~\ref{fig:stable-consensus}A).
-This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig~\ref{fig:stable-consensus}B and C).
+This network shared most of the properties of the consensus: breast cancer susceptibility genes were overrepresented (P\=/value = 3 \texttimes{} 10\textsuperscript{-4}), as well as genes involved in mitochondrial translation and the attenuation phase (adjusted P\=/values 0.001 and 3 \texttimes{} 10\textsuperscript{-5} respectively); the selected genes were more central than average (P\=/value = 1.1 \texttimes{} 10\textsuperscript{-14}); and a considerable number of nodes (19) were isolated (Fig~\ref{fig:stable-consensus}B and \ref{fig:stable-consensus}C).
 
@@
 
-\paragraph*{S11 Fig}
-\label{sfig:scones_gsm}
-\textbf{Overview of the solutions produced by the SConES on the GS and GM networks (Section~\ref{methods:networks}) on the GENESIS dataset.} \textbf{(A)}~Manhattan plots of SNPs (Section~\ref{methods:conventional}); in black, the method’s solution.
-The red line indicates the Bonferroni threshold (2.54 \texttimes{} 10\textsuperscript{-7}).
-\textbf{(B)}~Precision and recall of the evaluated methods with respect to Bonferroni-significant SNPs (SConES) or genes (other methods) in BCAC.
-For reference, we added a gray line with a slope of 1.
-\textbf{(C)}~Solution networks.
-%\includegraphics[width=\linewidth]{./figures/sfigure_11.png}
-
 \end{document}
